### 2019-07-10 resolution.tex
new repo

% qjrms4doc.tex V1.10, 4 October 2013

\documentclass[times]{qjrms4}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
\usepackage{moreverb}

%\def\volumeyear{2013}
\def\volumenumber{00}

\begin{document}

\runningheads{A.~R.~Herrington and K.~A.~Reed}{CAM resolution sensitivity}

\title{Parameterized convection, grid-scale clouds and resolution sensitivity in the Community Atmosphere Model}

\author{Adam R. Herrington\corrauth, Kevin A. Reed}
\address{School of Marine and Atmospheric Sciences, Stony Brook University, Stony Brook, NY 11794}

\corraddr{\url{adam.herrington@stonybrook.edu}}

\begin{abstract}
This paper describes...
\end{abstract}

\keywords{Climate models, physical parameterizations, physics-dynamics coupling}

\maketitle

\section{Introduction}

An increasing number of atmospheric dynamical cores are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution ($\Delta x = 50$ km or less), or even globally high-resolution weather and climate simulations \citep{MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}. Incorporating these advances into Atmospheric General Circulation Models (AGCMs) requires the development of physical parameterizations ({\em{physics}}) appropriate for the diversity of grid configurations that these dynamical cores are now able to support, referred to as scale-aware physics. The most common approach towards developing scale-aware physics has been through the lens of limited area, cloud resolving simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}. Through subsequently filtering large-eddy simulation solutions to lower-resolution grids, a relationship between resolved and unresolved moments may be understood and ultimately parameterized. While this approach is likely necessary for developing scale-aware physics, it is not sufficient. Since the equations of motions have inherent scale dependencies, resolved dynamical modes are a function of the native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS,J2017JAMES}. Scale-aware physics should also recognize native grid dependencies.

The sensitivity of the Community Atmosphere Model (CAM), and its predecessor, the Community Climate Model (CCM) to horizontal resolution is well documented. Despite thirty years of continual model development, there are robust sensitivities to resolution (hereafter {\em{resolution}} refers to {\em{horizontal resolution}}) which have persisted in all versions of the model. Total precipitable water decreases with resolution \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}, which typically, but not always \citep[see][]{WETAL1995CD,ZetAl2014JCb}, coincides with a reduction in cloud cover. \cite{KW1991JGR,WETAL1995CD} suggested that the drying of the atmosphere is consistent with the observed greater magnitude resolved vertical velocities with resolution, with greater subsiding motion increasing the export of dry air from the upper troposphere. This mechanism is consistent with an analysis of moisture budgets in CAM, version 4 (CAM4) across multiple resolutions \citep{YETAL2014JCLIM,HR2017JCLIM}.

It is well known that the magnitude of vertical velocities increase with resolution in atmospheric models. While the cause of this sensitivity is established in the cloud-resolving and large-eddy simulation literature \citep[see][and references therein]{J2017JAMES}, only recently has the vertical velocity field in AGCMs and there sensitivity to resolution received attention, albeit with conflicting explanations \citep{DETALA2016ACP,RETAL2016CD,HR2018JAMES}. In this paper, it is shown that the vertical velocities in CAM, version 6 (CAM6) scale like $\Delta x^{-1}$, and that the inadequacy of this scaling to explain prior CAM behavior \citep{HR2017JCLIM} is due to time-truncation errors arising from too large a physics time-step ($\Delta t_{phys}$) at higher-resolutions \citep{HR2018JAMES,HETAL2019JAMES}. Furthermore, the $\Delta x^{-1}$ scaling is entirely consistent with the mechanism of resolution sensitivity in cloud-resolving and large-eddy models, i.e., the inherent sensitivities of resolved dynamical modes to native grid resolution.

Another robust response of the CAM-lineage to resolution is an increase in grid-scale precipitation rates (also referred to as stratiform precipitation in the literature), at the expense of parameterized convective precipitation rates. This behavior is summarized in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological grid-scale and convective precipitation rates in prior CAM/CCM convergence studies. The studies of \cite{KW1991JGR,WETAL1995CD,W2013QJRMS} indicate that the tendency to reduce $\Delta t_{phys}$ with resolution would by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convection precipitation rates with resolution.

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=20pc,angle=0]{figs/cam-history.pdf}\\
\end{center}
\caption{Bar-graph of the convective (solid) and grid-scale (white) climatological precipitation rates in prior CAM/CCM convergence studies. Each window contains a single convergence study, with identical x-axis; the approximate grid resolution. Colors indicate the model configuration; January ensemble (black) and aqua-planet configurations with SST profiles $QOBS$ (blue) and $CNTL$ (red) after \cite{NH2000ASL}. Studies included in this figure are \cite{KW1991JGR} (CCM1), \cite{WETAL1995CD} (CCM2), \cite{W2008TELLUS} (CAM3), \cite{RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM} (CAM4), \cite{ZetAl2014JCb} (CAM5) and this study (CAM6). CCM2* refers to the modified parameter experiment of \cite{WETAL1995CD}, where parameters vary with resolution to reduce the dependence of cloud fraction on resolution.}
\label{fig:cam-history}
\end{figure}

In this study, the reduction in convective precipitation rates in CAM6 is shown to result from the greater subsiding motion with resolution, leading to a more stable atmosphere in which the criterion for parameterized convection occurs less often. The feedback of the resolved vertical motion on the physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution. Section 2 describes CAM6 and the simulations used in this study. Section 3 contains a thorough analysis of the CAM6 simulations and Section 4 provides some discussion and conclusions.

%\input{results.tex}

\ack This class file was developed by Sunrise Setting Ltd,
Paignton, Devon, UK. Website:\\
\href{http://www.sunrise-setting.co.uk}{\texttt{www.sunrise-setting.co.uk}}

\bibliographystyle{wileyqj}
\bibliography{bib}
\end{document}

### 2019-07-10 results.tex
new repo

\section{Methods}

\subsection{Dynamical Core}

This study uses the spectral-element dynamical core option of Community Atmosphere Model \citep[CAM-SE;][]{TF2010JCP,DetAl2012IJHPCA}, coupled with the mass conserving, semi-Lagrangian advection method for accelerated multi-tracer transport \citep[CSLAM;][]{LTOUNGK2017MWR}, and dry-mass vertical coordinate with comprehensive treatment of moisture and energy \citep{LetAl2018JAMES}. Dry dynamics are solved using the spectral element method in which the vector-invariant form of the hydrostatic equations of motion are solved with a degree 3 polynomial basis in each element, and the solutions are patched together at the boundaries using the direct stiffness summation to form a global basis. The elements are defined by a cubed-sphere grid, and the notation for horizontal grid resolution is an `$ne$' followed by the number of elements contained on an edge of one cubed-sphere face, e.g., $ne30$. Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds. The advection of tracers uses a semi-Lagrangian method on a finite volume grid constructed the cubed-sphere grid, and contains the same degrees of freedom and the dry dynamics. The coupled system conserves energy, mass and preserves linear correlations between two reactive species to within machine precision \citep{HL2018MWR}.

The physics are evaluated on the tracer advection grid, and the tendencies mapped back to the spectral element grid \citep[CAM-SE-CSLAM;][]{HL2018MWR}. A coarser physics grid, containing $\frac{5}{9}$ fewer degrees of freedom than the dynamical core grid is also available as part of the CAM-SE-CSLAM package (Chapter~\ref{sec:chapter5}). This configuration is also used in this study, but it is considered a perturbed configuration from the default configuration in this study. The dynamics time-step is subcycled within a longer physics time-step $\Delta t_{phys}$, and the temperature and momentum increments from the physics are then divided by the number of subcycles and added to the dynamical core at the beginning of each subcycle; the full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LW2019JAMES}.

\subsection{Moist Physics}

The simulations use Community Atmosphere Model, version 6 physics (CAM6; \url{https://ncar.github.io/CAM/doc/build/html/users_guide/index.html}). The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed PDF higher order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics. The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with a three mode Modular Aerosol Model \citep{MAM}. Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} using a dilute-plume calculation as the convective trigger and to close the mass fluxes \citep{NRJ2008JC}. The deep convection scheme includes convective momentum transport \citep{RR2008JC}.

\subsection{Experimental Design}

This convergence study uses an aqua-planet configuration \citep{NH2000ASL,MWO2016JAMES}, an all ocean planet with fixed, zonally symmetric sea surface temperatures modeled after the present day Earth \citep[$QOBS$ in][]{NH2000ASL}. The aqua-planets are in a perpetual equinox, and aerosols are largely absent from the simulations. Each simulation is ran for one simulated year. Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table6-1}. The horizontal hyper-viscosity operators $\nu$ vary with resolution after Chapter~\ref{sec:chapter5}, and are also provided in the table. The values of $\nu$ are a factor 2.5 greater for divergence damping and not shown in the Table. $\Delta t_{phys}$ is chosen to scale with resolution, in proportion to the grid spacing,
\begin{equation}
\Delta t_{phys} = \Delta t_{phys,0} \times \frac{n_{e,0}}{n_e}~s,\label{eq:dt-scale}
\end{equation}
where $\Delta t_{phys,0}$ is taken to be the standard $1800 s$ used in CAM-SE-CSLAM at the standard resolution, $n_{e,0} = ne30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km). This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test (Appendix~\ref{sec:app1}), and it is understood that it will likely lead to greater resolution sensitivity \citep{W2008TELLUS,HR2018JAMES}. The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.

\section{Results}

 \begin{table}
 \caption{Experimental design and global means.}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   Variable & $ne20$ & $ne30$ & $ne40$ & $ne60$ & $ne80$ & $ne120$ \\
   \hline
   $\nu$ ($m^4/s$) & $1.5 \times 10^{15}$ & $4.0 \times 10^{14}$ & $1.5 \times 10^{14}$ & $4.0 \times 10^{13}$  & $1.5 \times 10^{13}$ & $4.0 \times 10^{12}$\\
    $\Delta t_{phys}$ (s) & 2700 & 1800 & 1350 & 900 & 675 & 450 \\
   Total Cloud Fraction & 0.844 & 0.835 & 0.824 & 0.810 & 0.804 & 0.800 \\ 
   Total Precipitable Water (mm) & 23.31& 23.01 & 22.62 & 22.25 & 21.93 & 21.72 \\
   Convective Precipitation (mm/day) & 1.91 & 1.83 & 1.68 & 1.47 & 1.29 & 1.08 \\
   Stratiform Precipitation (mm/day) & 1.26 & 1.42 & 1.60 & 1.85 & 2.05 & 2.22 \\      
 \hline
 \end{tabular}
 \label{tbl:table6-1}
 \end{table}

Table~\ref{tbl:table6-1} provides a list of some relevant globally averaged metrics that are typically published in CAM-lineage convergence studies. Total precipitable water, total cloud fraction and deep convective precipitation rate decreases, while stratiform precipitation increases, monotonically with resolution. Resolution sensitivity in CAM6 is similar to all prior model versions within the models' lineage.

The probability density function (PDF) of negative, or upward vertical pressure velocities $\omega$ in the aqua-planets is shown in Figure~\ref{fig:2pdf}a. The magnitude of upward $\omega$ increases in a monotonic way with resolution, with positive, or downward $\omega$ behaving similarly (not shown). The PDF's may be scaled to the high-resolution $ne120$ resolution through $P(\omega)_s = \alpha P (\omega / \alpha)$, where $\alpha$ is the scale factor equation~\ref{eq:eq6-1}. In general, the scaled PDF's all line up on top of the high-resolution reference (Figure~\ref{fig:2pdf}b); equation~\ref{eq:eq6-1} explains the variation in vertical velocity across resolutions to first order.

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=33pc,angle=0]{figs/temp_2pdf.pdf}\\
\end{center}
\caption{Probability density distribution of the upward vertical pressure velocities $\omega$ computed everywhere in the model from 1 year of 6-hourly data (a) raw values (b) scaled to the $ne120$ resolution using $P(\omega)_s = \alpha P (\omega / \alpha)$, where $\alpha$ is from equation~\ref{eq:eq6-1}.}
\label{fig:2pdf}
\end{figure}

Changes to the time-mean vertical velocity field can be understood through decomposing the global mean $\omega$ into upward and downward components,
\begin{equation}
\overline{\langle \omega \rangle} = \overline{\langle f_{u} \rangle} \, \overline{\langle \omega_{u} \rangle} + \overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}, \label{eq:eq6-2}
\end{equation}
where $\langle f_x \rangle$ and $\langle \omega_x \rangle$ refers to the vertical mass fraction and contribution to mass weighted vertical mean $\omega$, respectively, subscript $u$ refers to upward motion and $d$, downward motion; overbars indicate time-mean.

The components of equation~\ref{eq:eq6-2} are shown in Figure~\ref{fig:8panel}a,b,e,f for the six aqua-planet simulations. The magnitude of both $\overline{\langle \omega_{u} \rangle}$ and $\overline{\langle \omega_{d} \rangle}$ increase monotonically with resolution, the upward component increasing more than the downward component. In contrast, $\overline{\langle f_{u} \rangle}$ decreases with resolution while $\overline{\langle f_{d} \rangle}$ increases, although in both cases there is a reversal in this trend in the highest resolution simulation, $ne120$. Note that the magnitude of the change in $\overline{\langle f_{u} \rangle}$ and $\overline{\langle f_{d} \rangle}$ is small, a range of about $0.015$ across all simulations. Figure~\ref{fig:8panel}c,g shows that the products  $\overline{\langle f_{u} \rangle} \, \overline{\langle \omega_{u} \rangle}$ and $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ are equal and opposite, which is a requirement of mass conservation in the model and a convenient check of the calculation. Besides the six aqua-planets described in Section 6.2, 23 additional year-long aqua-planet experiments were carried out at various resolutions, with modified parameters in the dynamical core and the physics. The spread in the components of equation~\ref{eq:eq6-2} in the simulations with perturbed parameters is relatively large for $\overline{\langle f_{x} \rangle}$, compared with $\overline{\langle \omega_{x} \rangle}$.

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=33pc,angle=0]{figs/temp_diags_8panel.pdf}\\
\end{center}
\caption{(a,b,c,e,f,g) Components of the global mean vertical pressure velocity $\overline{\langle \omega \rangle}$ (d) ratio of global mean ZM precipitation rate to the total precipitation rate versus grid spacing $\Delta x$ and (e) scatter plot of $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ and FREQZM, and the fitted linear regression which has a Pearson's R-value = 0.99. (a) $\overline{\langle \omega_{u} \rangle}$ (b) $\overline{\langle f_{u} \rangle}$ (c) $\overline{\langle f_{u} \rangle} \, \overline{\langle \omega_{u} \rangle}$ (e) $\overline{\langle \omega_{d} \rangle}$ (f) $ \overline{\langle f_{d} \rangle}$ (g) $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$. Colors are as in Figure~\ref{fig:2pdf}. Grey crosses are for the perturbed parameter runs.}
\label{fig:8panel}
\end{figure}

The reduction in the \cite{ZM1995AO} deep convective precipitation rate (hereafter referred to as the {\em{ZM precipitation}}) is depicted in Figure~\ref{fig:8panel}d, which shows the global mean fraction of total precipitation arising from ZM precipitation decreasing from about $0.60$ at low resolution ($ne20$) to about $0.32$ at high resolution ($ne120$). This reduction in ZM precipitation manifest through triggering less frequently (not shown). The space-time fraction the ZM precipitation scheme is triggered (hereafter referred to as the {\em{FREQZM}}) is highly negatively correlated with $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ in the 29 simulations (Pearson's R-value = 0.99; Figure~\ref{fig:8panel}h), more then with any individual component of equation~\ref{eq:eq6-2}.

Figure~\ref{fig:vomg}a,b shows the time mean zonal mean $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and FREQZM. $\langle f_{d} \rangle \langle \omega_{d} \rangle$ is smallest in the deep tropics (equatorward of about $10^{\circ}$), increasing sharply poleward throughout the subtropics and then begins to decrease in the middle latitudes ($\geq 40^{\circ}$) and conitnuing into the polar regions. FREQZM is largely opposite of $\langle f_{d} \rangle \langle \omega_{d} \rangle$; where greater subsiding motion occurs FREQZM is small, except in the mid-latitude region between $30^{\circ}$ and $60^{\circ}$ where the spatial variations are quite different. The resolution sensitivity of these two quantities has no zonal dependence; $\langle f_{d} \rangle \langle \omega_{d} \rangle$ increases and FREQZM decreases just about everywhere with resolution, in similar proportions.

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=17pc,angle=0]{figs/temp_zonal_fracd*vomgd.pdf}\\
\end{center}
\caption{Time mean zonal mean (a) $\langle f_{d} \rangle \langle \omega_{d} \rangle$, (b) FREQZM. Colors are as in Figure~\ref{fig:2pdf}.}
\label{fig:vomg}
\end{figure}

The ZM trigger uses a dilute form of CAPE, which is consistent with the negative relation between $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ and FREQZM in space, and across resolutions. In the classical, non-dilute case, CAPE can be seperated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through advection of dry static energy and moisture by the environment, i.e., the resolved flow. The latter term generally contributes positively to CAPE in regions of ascent and negatively in regions of subsidence \citep{SZ2018JCLIM}, consistent with the negative correlation of $\langle f_{d} \rangle \langle \omega_{d} \rangle$ in space, and across resolutions.

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=25pc,angle=0]{figs/temp_cape.pdf}\\
\end{center}
\caption{(a) time mean fraction of the deep tropics in the simulations with upward $\langle \omega \rangle$ (red) and downward $\langle \omega \rangle$ (blue), (b) CAPE computed from the mean temperature and moisture profiles of upward regions and downward regions. Black is for CAPE computed from the mean temperature and moisture profiles for the entire deep tropics, grey is the approximate discussed in the text.}
\label{fig:cape}
\end{figure}

To further support a negative relation between subsiding motion and CAPE, temperature and moisture profiles are collected from the deep tropics in the simulations, and conditionally sampled depending on whether $\langle \omega \rangle$ is positive or negative, indicating predominantly subsiding or ascending grid columns. The mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute-CAPE used by the ZM scheme, offline. Figure~\ref{fig:cape}b indicates that the mean profile of ascending regions are associated with large values of CAPE ($>170$ J/kg), and low values in subsiding regions ($<110$ J/kg). The mean profiles of subsiding regions have an anomalous warming layer in the $600-800$ hPa layer and an anomalous moisture deficit throughout the entire column, relative to the non-conditionally sampled mean (not shown). Figure~\ref{fig:cape}a shows that fractional area of air columns in the deep tropics that are predominantly subsiding (ascending) changes drastically with resolution, from $0.32$ ($0.68$) in the $ne20$ ($\Delta x = 166.8$ km) run, and monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in the $ne120$ ($\Delta x = 27.8$ km) run. Interestingly, the sum of the product of the fractional areas with their corresponding CAPE values gives approximately the same CAPE values computed from the non-conditionally sampled mean temperature and moisture profiles (Figure~\ref{fig:cape}b). This provides strong evidence that CAPE values in the deep tropics are decreasing with resolution because a larger (smaller) fraction of the deep tropics are made up of subsiding (ascending) grid columns with resolution.

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=14pc,angle=0]{figs/temp_zonal_4reg_dwn.pdf}\\
\end{center}
\caption{Zonal mean (a) R-values and (b) sensitivity parameter in the logistic regression, (c) drizzle fraction, (d) time mean surface latent heat fluxes. Colors are as in Figure~\ref{fig:2pdf}.}
\label{fig:4reg}
\end{figure}

To understand how the relationship between subsiding motion and the trigger in the ZM scheme manifests within the simulations themselves, a logistic regression is performed for each grid column in the simulations. Logistic regression uses an iterative method to fit a continuous variable predictor, $x$ to a binary predictand $p$ \citep{WILKSBOOK},
\begin{equation}
p = \frac{exp{[b_0 + b_1 x]}}{1 + exp{[b_0 + b_1 x]}}, \label{eq:eq6-3}
\end{equation}
where $b_0$ and $b_1$ are the shape parameters of the exponential. The predictor is chosen as the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand is whether or not the ZM scheme is active, $1$ for yes and $0$ for no. Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and parameter $b_1$ (hereafter referred to as the sensitivity parameter). Figure~\ref{fig:4reg}a shows the zonal mean R-values indicates the goodness of fit is greatest in the deep tropics. Figure~\ref{fig:4reg}d shows the time mean zonal mean latent heat flux in the simulations, which is expected to contribute positively to CAPE through the component associated with the thermodynamic state of boundary layer parcels. In the deep tropics, the latent heat flux is small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4reg}b), indicating that subsiding motion is successfully depressing CAPE, and the activity of the ZM scheme in that region.

Poleward of $10^{\circ}$, the R-value decreases to a local minimum between $15^{\circ} - 20^{\circ}$ latitude, and the magnitude of the sensitivity parameter steeply declines. Even though the subsiding motion is increasing polewards of $10^{\circ}$, this motion is less effective in depressing the convection scheme. The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes, indicating that the subsiding motion is less skillful a predictor of the ZM scheme when the boundary layer is being driven unstable by large surface latent heat fluxes. The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4reg}c). The predominance of drizzle in this region is probably a result of the large subsiding motion (Figure~\ref{fig:vomg}) constraining CAPE from becoming too large. AGCMs are known to suffer from an excess drizzle bias in precisely this approximate region \citep{D2006JCLIM}, and this analysis indicates that this bias is due to the use of a CAPE trigger function. In the $30^{\circ} - 60^{\circ}$ latitude region the shape parameter reaches it's global minimum and then begins to increase in the polar regions. This is consistent with lack of spatial correspondence between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and FREQZM in the mid-latitudes, and the reemergence of this correspondence in the polar regions (Figure~\ref{fig:vomg}).

The sensitivity parameter becomes less negative in the deep tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ in that region with resolution, which requires a lower sensitivity parameter to predict the binary of whether the ZM scheme is active. The R-values generally decrease with resolution indicating that there is degradation in the relationship with resolution. All values shown are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK}.

\section{Conclusions}

A convergence study is carried out in order to unravel the relationships between grid-scale clouds, parameterized convection and resolution sensitivity in an aqua-planet configuration of the Community Atmosphere Model (CAM) with version 6 physics (CAM6), the spectral-element dynamical core coupled to the Conservative Semi-Lagrangian Advection Method for advection of tracers and dry-mass vertical coordinate (CAM-SE-CSLAM). Common with just about all versions in the CAM-lineage, total cloud fraction, total precipitable water and convective precipitation decrease with resolution, while stratiform precipitation increases with resolution. The magnitude of vertical velocities everywhere in the model increase, and scale like the inverse of the grid spacing $\Delta x^{-1}$. The climatological area of regions of ascent decrease with resolution, the area of subsiding motion increases.

 The mass weighted vertical mean of downward vertical pressure velocities ($\langle f_{d} \rangle \langle \omega_{d} \rangle$) is highly negatively correlated with the frequency the \cite{ZM1995AO} deep convection scheme is triggered (FREQZM). The correlation of the climatological quantities is linear using 29 simulations containing a large range of grid resolutions, and has a goodness of fit Pearson's R-value of 0.99. A logistic regression between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and FREQZM was performed at each grid column in the simulations, and indicates that $\langle f_{d} \rangle \langle \omega_{d} \rangle$ is skillful at predicting FREQZM, with greater downward motion leading to less frequent convective triggering. This relationship is particularly strong in the deep tropics.

The \cite{ZM1995AO} deep convection scheme (ZM scheme) uses a Convective Available Potential Energy (CAPE) threshold as the convective trigger, and the high negative correlation between subsiding motion and the trigger is consistent with the dependence of CAPE on vertical advection of dry static energy and moisture \citep{Z2002JGR}. The precipitation rates in the $15^{\circ} - 20^{\circ}$ latitude region consist primarily of drizzle from the ZM scheme. The convective trigger is weakly correlated with rather strong subsiding motion in this region, but the latent heat fluxes into the boundary layer are at a global maximum. It is suggested that the downward motion is opposing the generation of CAPE by the latent heat fluxes of this region, but due to the overwhelming magnitude of the latent heat fluxes, cannot depress CAPE below the trigger threshold. The resultant CAPE values are small and ZM scheme produces drizzle most of the time. The excess simulation of drizzle in climate models is a known model bias \citep{D2006JCLIM}, and it is suggested here that this bias arises from the use of CAPE as the sole criterion for convective triggering.

The conclusion of this study builds on previous work \citep[][Chapter~\ref{sec:chapter5}]{HR2017JCLIM,HR2018JAMES}, and it is of the following relationships between parameterized convection, grid-scale clouds and resolution sensitivity in CAM. When the ZM scheme is triggered, it moistens the middle-to-upper troposphere through detrainment \citep{ZM1995AO}, moisture which is then condensed by the stratiform cloud scheme to produce grid-scale clouds. The buoyancy produced by grid-scale clouds leads to large magnitude upward motion by the dynamical core, and to conserve mass, downward motion radiating away from the grid-scale clouds as horizontally propagating gravity waves. The grid-scale clouds are grid-limited; their horizontal size is set by the effective resolution of the model. The vertical velocities associated with the buoyancy of stratiform clouds are inversely proportional to their horizontal extent, and therefore scale like $\Delta x^{-1}$. Due to the greater efficiency of vertical mass transport with increasing resolution, the area of ascending motion decreases with resolution. The area and magnitude of the subsiding regions increase with resolution in order to conserve mass, which has the effect of stabilizing and drying the atmosphere with increasing resolution. The ZM scheme, with its stability dependent CAPE trigger, responds through becoming less active with increasing resolution.

### 2019-08-06 resolution.tex
intro stuff

@@
 
-An increasing number of atmospheric dynamical cores are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution ($\Delta x = 50$ km or less), or even globally high-resolution weather and climate simulations \citep{MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
-Incorporating these advances into Atmospheric General Circulation Models (AGCMs) requires the development of physical parameterizations ({\em{physics}}) appropriate for the diversity of grid configurations that these dynamical cores are now able to support, referred to as scale-aware physics.
-The most common approach towards developing scale-aware physics has been through the lens of limited area, cloud resolving simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
-Through subsequently filtering large-eddy simulation solutions to lower-resolution grids, a relationship between resolved and unresolved moments may be understood and ultimately parameterized.
+An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x$ on the order of kilometers) and climate ($\Delta x = 50$ km or less) simulations \citep{MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
+These models are built using unstructured meshes that while allow for substantial grid flexibility, requires physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
+The most common approach towards developing scale-aware physics has been through the lens of limited area, large-eddy simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
+Through subsequently filtering large-eddy solutions to lower-resolution grids, a relationship between first- and higher-order moments \citep{G1992JFM} may be understood and ultimately parameterized.
 While this approach is likely necessary for developing scale-aware physics, it is not sufficient.
-Since the equations of motions have inherent scale dependencies, resolved dynamical modes are a function of the native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS,J2017JAMES}.
-Scale-aware physics should also recognize native grid dependencies.
+Since the equations of motions have inherent scale dependencies, resolved dynamical modes vary with the native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS,J2017JAMES}.
+Scale-aware physics should also recognize these native grid dependencies.
 
-The sensitivity of the Community Atmosphere Model (CAM), and its predecessor, the Community Climate Model (CCM) to horizontal resolution is well documented.
+The sensitivity of the Community Atmosphere Model \citep[CAM;][]{CAM5}, and its predecessor, the Community Climate Model (CCM) to horizontal resolution is well documented.
 Despite thirty years of continual model development, there are robust sensitivities to resolution (hereafter {\em{resolution}} refers to {\em{horizontal resolution}}) which have persisted in all versions of the model.
-Total precipitable water decreases with resolution \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}, which typically, but not always \citep[see][]{WETAL1995CD,ZetAl2014JCb}, coincides with a reduction in cloud cover.
-\cite{KW1991JGR,WETAL1995CD} suggested that the drying of the atmosphere is consistent with the observed greater magnitude resolved vertical velocities with resolution, with greater subsiding motion increasing the export of dry air from the upper troposphere.
-This mechanism is consistent with an analysis of moisture budgets in CAM, version 4 (CAM4) across multiple resolutions \citep{YETAL2014JCLIM,HR2017JCLIM}.
+The atmosphere dries with increasing resolution, seen through the reduction in total precipitable water \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}, which typically, but not always \citep[see][]{WETAL1995CD,ZetAl2014JCb}, coincides with a reduction in cloud cover.
+\cite{KW1991JGR} and \cite{WETAL1995CD} suggested that the drying of the atmosphere is consistent with the greater magnitude resolved vertical velocities with increasing resolution, with greater subsiding motion increasing the export of dry air from the upper troposphere.
+This mechanism is consistent with an analysis of moisture budgets in CAM, version 4 \citep[CAM4;][]{CAM4} across multiple resolutions \citep{YETAL2014JCLIM,HR2017JCLIM}.
 
 It is well known that the magnitude of vertical velocities increase with resolution in atmospheric models.
-While the cause of this sensitivity is established in the cloud-resolving and large-eddy simulation literature \citep[see][and references therein]{J2017JAMES}, only recently has the vertical velocity field in AGCMs and there sensitivity to resolution received attention, albeit with conflicting explanations \citep{DETALA2016ACP,RETAL2016CD,HR2018JAMES}.
-In this paper, it is shown that the vertical velocities in CAM, version 6 (CAM6) scale like $\Delta x^{-1}$, and that the inadequacy of this scaling to explain prior CAM behavior \citep{HR2017JCLIM} is due to time-truncation errors arising from too large a physics time-step ($\Delta t_{phys}$) at higher-resolutions \citep{HR2018JAMES,HETAL2019JAMES}.
-Furthermore, the $\Delta x^{-1}$ scaling is entirely consistent with the mechanism of resolution sensitivity in cloud-resolving and large-eddy models, i.e., the inherent sensitivities of resolved dynamical modes to native grid resolution.
+While the cause of this sensitivity has been established in the large-eddy simulation literature \citep[see][and references therein]{J2017JAMES}, only recently has the vertical velocity field in AGCMs and their sensitivity to resolution received attention, albeit with conflicting explanations \citep{DETALA2016ACP,RETAL2016CD,HR2018JAMES}.
+To generalize the relationship between vertical velocity and resolution, let $\alpha$ refer be the ratio of $W_0$, the vertical velocity scale of some reference grid spacing $\Delta x_0$, to $W$, the vertical velocity scale of any $\Delta x$.
+A power-law for $\alpha$ in $\Delta x$ is then,
+\begin{equation}
+\alpha^{-1} = \frac{W}{W_0} = \left( \frac{\Delta x}{\Delta x_0} \right)^n, \label{eq:alpha}
+\end{equation}
+where $n$ is the power-law exponent.
+
+\cite{RETAL2016CD} derive an estimate $n= b-1$ by combining a scale analysis of the continuity equation with a power law representation $\Delta x^{2b}$ of the second-order structure function of the horizontal wind.
+Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem, is equivalent to the observed $-\frac{5}{3}$ slope of the kinetic energy spectrum  \citep{NG1985JAS} through the relation $-\left( 2b+1 \right) $.
+\cite{RETAL2016CD} argue that the emergent constraint on the slope of the kinetic energy spectrum in both models and observations indicates that $n= -\frac{2}{3}$.
+
+For large-eddy simulations, the sensitivity of vertical velocities to resolution is explained through a scale analysis of the dynamical equations \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
+For hydrostatic scales relevant to AGCMs, a scale analysis of the Poisson yields $W \propto D^{-1}$, where $D$ is the horizontal scale of the buoyancy perturbation \citep{HR2018JAMES}.
+In CAM aqua-planet simulations, the largest source of buoyancy is associated with grid-scale cloud formation, whose horizontal extent are set by the effective resolution of the model, about $5-10 \Delta x$, and so this gives $n=-1$ \citep{HR2018JAMES}.
+In this paper, it is shown that the resolution sensitivity of vertical velocities in CAM, version 6 (CAM6) are well described with $n=-1$, and that the inadequacy of this scaling to explain prior CAM behavior \citep{HR2017JCLIM} is due to time-truncation errors arising from too large a physics time-step ($\Delta t_{phys}$) at higher-resolutions \citep{HR2018JAMES,HETAL2019JAMES}.
 
@@
 This behavior is summarized in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological grid-scale and convective precipitation rates in prior CAM/CCM convergence studies.
-The studies of \cite{KW1991JGR,WETAL1995CD,W2013QJRMS} indicate that the tendency to reduce $\Delta t_{phys}$ with resolution would by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convection precipitation rates with resolution.
+The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the tendency to reduce $\Delta t_{phys}$ with resolution would by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convection precipitation rates with resolution.
 

### 2019-08-07 resolution.tex
diags fig

--- 2019-07-10 resolution.tex
@@
 
-An increasing number of atmospheric dynamical cores are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution ($\Delta x = 50$ km or less), or even globally high-resolution weather and climate simulations \citep{MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
-Incorporating these advances into Atmospheric General Circulation Models (AGCMs) requires the development of physical parameterizations ({\em{physics}}) appropriate for the diversity of grid configurations that these dynamical cores are now able to support, referred to as scale-aware physics.
-The most common approach towards developing scale-aware physics has been through the lens of limited area, cloud resolving simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
-Through subsequently filtering large-eddy simulation solutions to lower-resolution grids, a relationship between resolved and unresolved moments may be understood and ultimately parameterized.
+An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x = 5$ km and less) and climate ($\Delta x = 50$ km and less) simulations \citep{MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
+These models are built using unstructured meshes that while allow for substantial grid flexibility, require physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
+The most common approach towards developing scale-aware physics is through the lens of limited area, large-eddy simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
+Through subsequently filtering large-eddy solutions to lower-resolution grids, a relationship between first- and higher-order moments \citep{G1992JFM} may be understood and ultimately parameterized as a function of grid resolution.
 While this approach is likely necessary for developing scale-aware physics, it is not sufficient.
-Since the equations of motions have inherent scale dependencies, resolved dynamical modes are a function of the native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS,J2017JAMES}.
-Scale-aware physics should also recognize native grid dependencies.
+The equations of motions have inherent scale dependencies, and the properties of dynamical modes also vary with native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS,J2017JAMES}.
+Scale-aware physics should also recognize these native grid dependencies.
 
-The sensitivity of the Community Atmosphere Model (CAM), and its predecessor, the Community Climate Model (CCM) to horizontal resolution is well documented.
-Despite thirty years of continual model development, there are robust sensitivities to resolution (hereafter {\em{resolution}} refers to {\em{horizontal resolution}}) which have persisted in all versions of the model.
-Total precipitable water decreases with resolution \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}, which typically, but not always \citep[see][]{WETAL1995CD,ZetAl2014JCb}, coincides with a reduction in cloud cover.
-\cite{KW1991JGR,WETAL1995CD} suggested that the drying of the atmosphere is consistent with the observed greater magnitude resolved vertical velocities with resolution, with greater subsiding motion increasing the export of dry air from the upper troposphere.
-This mechanism is consistent with an analysis of moisture budgets in CAM, version 4 (CAM4) across multiple resolutions \citep{YETAL2014JCLIM,HR2017JCLIM}.
+The sensitivity of the Community Atmosphere Model \citep[CAM;][]{CAM5}, and its predecessor, the Community Climate Model (CCM) to resolution ({\em{resolution}} refers to {\em{horizontal resolution}}, hereafter) is well documented through convergence studies \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}.
+Despite thirty years of continual model development, there are robust sensitivities to resolution that have persisted in all versions of the model.
+This study argues that a unifying cause, the inherent scale sensitivities of the underlying dynamical equations, can explain the robust responses to resolution that occur in CAM/CCM, {\color{red}{since it is difficult to conceive that inevitable responses to native grid resolution could be ignored in the pursuit of scale-aware physics.}}
+
+In CAM/CCM, the atmosphere progressively dries with increasing resolution, seen through a reduction in simulated total precipitable water \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}, which typically, but not always \citep[see][]{WETAL1995CD,ZetAl2014JCb}, coincides with a reduction in cloud cover.
+\cite{KW1991JGR} and \cite{WETAL1995CD} suggested that the drying of the atmosphere is due to greater magnitude resolved vertical velocities with increasing resolution, with greater subsiding motion increasing the export of dry air from the upper troposphere.
+This mechanism is consistent with an analysis of moisture budgets in CAM, version 4 \citep[CAM4;][]{CAM4} across multiple resolutions \citep{YETAL2014JCLIM,HR2017JCLIM}.
 
 It is well known that the magnitude of vertical velocities increase with resolution in atmospheric models.
-While the cause of this sensitivity is established in the cloud-resolving and large-eddy simulation literature \citep[see][and references therein]{J2017JAMES}, only recently has the vertical velocity field in AGCMs and there sensitivity to resolution received attention, albeit with conflicting explanations \citep{DETALA2016ACP,RETAL2016CD,HR2018JAMES}.
-In this paper, it is shown that the vertical velocities in CAM, version 6 (CAM6) scale like $\Delta x^{-1}$, and that the inadequacy of this scaling to explain prior CAM behavior \citep{HR2017JCLIM} is due to time-truncation errors arising from too large a physics time-step ($\Delta t_{phys}$) at higher-resolutions \citep{HR2018JAMES,HETAL2019JAMES}.
-Furthermore, the $\Delta x^{-1}$ scaling is entirely consistent with the mechanism of resolution sensitivity in cloud-resolving and large-eddy models, i.e., the inherent sensitivities of resolved dynamical modes to native grid resolution.
+While the cause of this sensitivity has been established for large-eddy simulations \citep[see][and references therein]{J2017JAMES}, only recently has the vertical velocity field in AGCMs and their sensitivity to resolution received attention \citep{DETALA2016ACP,OETAL2016JAMES}, albeit with conflicting explanations \citep{RETAL2016CD,HR2018JAMES}.
+To generalize the relationship between vertical velocity and resolution, let $\alpha$ refer to the ratio of $W_0$, the vertical velocity scale of some reference grid spacing $\Delta x_0$, to $W$, the vertical velocity scale of any $\Delta x$.
+A power law for $\alpha$ in $\Delta x$ is then,
+\begin{equation}
+\alpha = \frac{W_0}{W} = \left( \frac{\Delta x_0}{\Delta x} \right)^n, \label{eq:alpha}
+\end{equation}
+where $n$ is the power law exponent.
 
-Another robust response of the CAM-lineage to resolution is an increase in grid-scale precipitation rates (also referred to as stratiform precipitation in the literature), at the expense of parameterized convective precipitation rates.
-This behavior is summarized in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological grid-scale and convective precipitation rates in prior CAM/CCM convergence studies.
-The studies of \cite{KW1991JGR,WETAL1995CD,W2013QJRMS} indicate that the tendency to reduce $\Delta t_{phys}$ with resolution would by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convection precipitation rates with resolution.
+\cite{RETAL2016CD} derive an estimate $n= b-1$ by combining a scale analysis of the continuity equation with a power law representation $\Delta x^{2b}$ of the second-order structure function of the horizontal wind.
+Strictly speaking, $\Delta x$ here refers to the distance between two points for which the velocity increment is computed in the structure function, but with this distance set to the model grid-spacing.
+Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum for mesoscale motions \citep{NG1985JAS}.
+\cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope is common in both observations and models, and provides an emergent constraint for $n= -\frac{2}{3}$.
+
+In large-eddy simulations, the sensitivity of vertical velocities to resolution is adequately explained by a scale analysis of the dynamical equations \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
+For hydrostatic scales relevant to AGCMs, a scale analysis of the Poisson equation gives $W \propto D^{-1}$, where $D$ is the horizontal scale of a buoyancy perturbation driving vertical motion \citep{HR2018JAMES}.
+In CAM aqua-planet simulations, the largest source of buoyancy is from grid-scale cloud formation, whose horizontal extents are set by the effective resolution of the model (i.e., some multiple of $\Delta x$), indicating $n=-1$ \citep{HR2018JAMES}.
+\cite{HR2017JCLIM} has shown that the $n=-1$ scaling does not explain the behavior of CAM4 in a convergence experiment, but follow-up studies \citep{HR2018JAMES,HETAL2019JAMES} indicate that the inadequacy of the $n=-1$ scaling is not definitive due to time-truncation errors associated with fixing the physics time-step ($\Delta t_{phys}$) across resolutions in that study.
+
+Another robust response of the CAM/CCM lineage to resolution is an increase in stratiform precipitation rates (i.e., the precipitation from grid-scale clouds), at the expense of parameterized convective precipitation rates.
+This behavior is summarized in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies.
+The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the tendency to reduce $\Delta t_{phys}$ with resolution would by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convective precipitation rates with resolution.
 
@@
 
-In this study, the reduction in convective precipitation rates in CAM6 is shown to result from the greater subsiding motion with resolution, leading to a more stable atmosphere in which the criterion for parameterized convection occurs less often.
+In this study, a typical convergence experiment using CAM, version 6 (CAM6; \url{https://ncar.github.io/CAM/doc/build/html/users_guide/index.html}) is carried out and analyzed in detail.
+It is shown that the resolution sensitivity of vertical velocities are well described with $n=-1$ in equation~\eqref{eq:alpha}, provided $\Delta t_{phys}$ is defined in a way to avoid large truncation errors across resolutions.
+The reduction in convective precipitation rates with resolution in CAM6 is shown to result from the greater magnitude subsiding motion, creating a more stable atmosphere in which the criterion for parameterized convection occurs less often.
 The feedback of the resolved vertical motion on the physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution.
-Section 2 describes CAM6 and the simulations used in this study.
+Section 2 describes CAM6 and details the convergence experiment.
 Section 3 contains a thorough analysis of the CAM6 simulations and Section 4 provides some discussion and conclusions.
 
-%\input{results.tex}
+\input{results.tex}
 

### 2019-08-07 results.tex
diags fig

@@
 
-This study uses the spectral-element dynamical core option of Community Atmosphere Model \citep[CAM-SE;][]{TF2010JCP,DetAl2012IJHPCA}, coupled with the mass conserving, semi-Lagrangian advection method for accelerated multi-tracer transport \citep[CSLAM;][]{LTOUNGK2017MWR}, and dry-mass vertical coordinate with comprehensive treatment of moisture and energy \citep{LetAl2018JAMES}.
-Dry dynamics are solved using the spectral element method in which the vector-invariant form of the hydrostatic equations of motion are solved with a degree 3 polynomial basis in each element, and the solutions are patched together at the boundaries using the direct stiffness summation to form a global basis.
-The elements are defined by a cubed-sphere grid, and the notation for horizontal grid resolution is an `$ne$' followed by the number of elements contained on an edge of one cubed-sphere face, e.g., $ne30$.
+This study uses the spectral-element dynamical core option of Community Atmosphere Model \citep[CAM-SE;][]{DetAl2012IJHPCA}, coupled with a mass conserving, semi-Lagrangian advection method for accelerated multi-tracer transport \citep[CSLAM;][]{LTOUNGK2017MWR}, and dry-mass vertical coordinate with comprehensive treatment of moisture and energy \citep{LetAl2018JAMES}.
+The dry dynamics are solved using the high-order, momentum, mass and energy conserving spectral element method \citep{TF2010JCP}, with the elements defined by a cubed-sphere grid.
+The notation for the horizontal grid resolution is an `$ne$' followed by the number of elements making up an edge of one cubed-sphere face, e.g., $ne30$.
 Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds.
-The advection of tracers uses a semi-Lagrangian method on a finite volume grid constructed the cubed-sphere grid, and contains the same degrees of freedom and the dry dynamics.
+The advection of tracers by CSLAM uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.
 The coupled system conserves energy, mass and preserves linear correlations between two reactive species to within machine precision \citep{HL2018MWR}.
 
-The physics are evaluated on the tracer advection grid, and the tendencies mapped back to the spectral element grid \citep[CAM-SE-CSLAM;][]{HL2018MWR}.
-A coarser physics grid, containing $\frac{5}{9}$ fewer degrees of freedom than the dynamical core grid is also available as part of the CAM-SE-CSLAM package (Chapter~\ref{sec:chapter5}).
-This configuration is also used in this study, but it is considered a perturbed configuration from the default configuration in this study.
-The dynamics time-step is subcycled within a longer physics time-step $\Delta t_{phys}$, and the temperature and momentum increments from the physics are then divided by the number of subcycles and added to the dynamical core at the beginning of each subcycle; the full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LW2019JAMES}.
+The physics are evaluated on the finite-volume CSLAM grid, and the tendencies mapped back to the spectral element grid \citep[CAM-SE-CSLAM;][]{HL2018MWR}.
+A coarser physics grid, containing $\frac{5}{9}$ fewer degrees of freedom than the dynamical core grid is also available as part of the CAM-SE-CSLAM package \citep{HETAL2019JAMES}.
+This lower-resolution physics grid is used in this study, but only as a member of a perturbed parameter ensemble and not in the default convergence experiment.
+The dynamics time-step is subcycled within a longer physics time-step $\Delta t_{phys}$, and the temperature and momentum increments from the physics are divided by the number of subcycles and added to the dynamical core at the beginning of each subcycle; the full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LetAl2018JAMES}.
 
@@
 
-The simulations use Community Atmosphere Model, version 6 physics (CAM6; \url{https://ncar.github.io/CAM/doc/build/html/users_guide/index.html}).
-The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed PDF higher order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
+The simulations use the CAM6 physics package.
+The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed PDF high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
 The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with a three mode Modular Aerosol Model \citep{MAM}.
-Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} using a dilute-plume calculation as the convective trigger and to close the mass fluxes \citep{NRJ2008JC}.
-The deep convection scheme includes convective momentum transport \citep{RR2008JC}.
+The macrophysics/microphysics combine to produce stratiform precipitation from grid-scale clouds.
+Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} and a dilute form of the convective available potential energy (CAPE) is computed to assess the convective trigger ($\geq 70$ J/kg), and for closing the mass fluxes in the cloud ensemble \citep{NRJ2008JC}.
+The deep convection scheme also parameterizes convective momentum transport \citep{RR2008JC}.
 
@@
 
-This convergence study uses an aqua-planet configuration \citep{NH2000ASL,MWO2016JAMES}, an all ocean planet with fixed, zonally symmetric sea surface temperatures modeled after the present day Earth \citep[$QOBS$ in][]{NH2000ASL}.
+The convergence experiment is performed in an aqua-planet configuration \citep{NH2000ASL,MWO2016JAMES}, an all ocean planet with fixed, zonally symmetric sea surface temperatures modeled after present day Earth \citep[$QOBS$ in][]{NH2000ASL}.
 The aqua-planets are in a perpetual equinox, and aerosols are largely absent from the simulations.
 Each simulation is ran for one simulated year.
-Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table6-1}.
-The horizontal hyper-viscosity operators $\nu$ vary with resolution after Chapter~\ref{sec:chapter5}, and are also provided in the table.
+Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
+The horizontal hyper-viscosity operators $\nu$ vary with resolution after \cite{HETAL2019JAMES}, and are also provided in the table.
 The values of $\nu$ are a factor 2.5 greater for divergence damping and not shown in the Table.
@@
 where $\Delta t_{phys,0}$ is taken to be the standard $1800 s$ used in CAM-SE-CSLAM at the standard resolution, $n_{e,0} = ne30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km).
-This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test (Appendix~\ref{sec:app1}), and it is understood that it will likely lead to greater resolution sensitivity \citep{W2008TELLUS,HR2018JAMES}.
+This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test (Appendix~\ref{sec:app1}), and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS,HR2018JAMES}.
 The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.
@@
  \end{tabular}
- \label{tbl:table6-1}
+ \label{tbl:table1}
  \end{table}
@@
 \begin{center}
-\noindent\includegraphics[width=33pc,angle=0]{figs/temp_2pdf.pdf}\\
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_2pdf.pdf}\\
 \end{center}
@@
 \begin{center}
-\noindent\includegraphics[width=33pc,angle=0]{figs/temp_diags_8panel.pdf}\\
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_diags_4panel.pdf}\\
 \end{center}

### 2019-08-07 resolution.tex
edits

@@
 An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x = 5$ km and less) and climate ($\Delta x = 50$ km and less) simulations \citep{MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
-These models are built using unstructured meshes that while allow for substantial grid flexibility, require physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
+These models are built using unstructured meshes that while allows for substantial grid flexibility, require physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
 The most common approach towards developing scale-aware physics is through the lens of limited area, large-eddy simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
@@
 While this approach is likely necessary for developing scale-aware physics, it is not sufficient.
-The equations of motions have inherent scale dependencies, and the properties of dynamical modes also vary with native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS,J2017JAMES}.
+The equations of motions have inherent scale dependencies, with the properties of dynamical modes a function of native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
 Scale-aware physics should also recognize these native grid dependencies.
@@
 
-In this study, a typical convergence experiment using CAM, version 6 (CAM6; \url{https://ncar.github.io/CAM/doc/build/html/users_guide/index.html}) is carried out and analyzed in detail.
-It is shown that the resolution sensitivity of vertical velocities are well described with $n=-1$ in equation~\eqref{eq:alpha}, provided $\Delta t_{phys}$ is defined in a way to avoid large truncation errors across resolutions.
+In this study, a convergence experiment using CAM, version 6 (CAM6; \url{https://ncar.github.io/CAM/doc/build/html/users_guide/index.html}) is carried out and analyzed in detail.
+It is shown that the resolution sensitivity of vertical velocities are well described with $n=-1$ in equation~\eqref{eq:alpha}, provided $\Delta t_{phys}$ is defined in a way that avoids large truncation errors across resolutions.
 The reduction in convective precipitation rates with resolution in CAM6 is shown to result from the greater magnitude subsiding motion, creating a more stable atmosphere in which the criterion for parameterized convection occurs less often.

### 2019-08-07 results.tex
edits

@@
 Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds.
-The advection of tracers by CSLAM uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.
-The coupled system conserves energy, mass and preserves linear correlations between two reactive species to within machine precision \citep{HL2018MWR}.
+CSLAM tracer transport uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.
 
-The physics are evaluated on the finite-volume CSLAM grid, and the tendencies mapped back to the spectral element grid \citep[CAM-SE-CSLAM;][]{HL2018MWR}.
+\subsection{Physical Parameterizations}
+
+The physics are evaluated on the finite-volume CSLAM grid, and the tendencies mapped back to the spectral element grid.
+The coupled system, referred to as CAM-SE-CSLAM, conserves energy, mass and preserves linear correlations between two reactive species to within machine precision \citep{HL2018MWR}.
 A coarser physics grid, containing $\frac{5}{9}$ fewer degrees of freedom than the dynamical core grid is also available as part of the CAM-SE-CSLAM package \citep{HETAL2019JAMES}.
 This lower-resolution physics grid is used in this study, but only as a member of a perturbed parameter ensemble and not in the default convergence experiment.
-The dynamics time-step is subcycled within a longer physics time-step $\Delta t_{phys}$, and the temperature and momentum increments from the physics are divided by the number of subcycles and added to the dynamical core at the beginning of each subcycle; the full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LetAl2018JAMES}.
-
-\subsection{Moist Physics}
+The dynamics time-step is subcycled within a longer physics time-step $\Delta t_{phys}$, and the temperature and momentum increments from the physics are divided by the number of subcycles and added to the dynamical core at the beginning of each subcycle.
+The full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LetAl2018JAMES}.
 
@@
 The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with a three mode Modular Aerosol Model \citep{MAM}.
-The macrophysics/microphysics combine to produce stratiform precipitation from grid-scale clouds.
+The combined macrophysics/microphysics routines generate grid-scale clouds with stratiform precipitation.
 Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} and a dilute form of the convective available potential energy (CAPE) is computed to assess the convective trigger ($\geq 70$ J/kg), and for closing the mass fluxes in the cloud ensemble \citep{NRJ2008JC}.
@@
 Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
-The horizontal hyper-viscosity operators $\nu$ vary with resolution after \cite{HETAL2019JAMES}, and are also provided in the table.
-The values of $\nu$ are a factor 2.5 greater for divergence damping and not shown in the Table.
+The horizontal hyper-viscosity operators $\nu$ vary with resolution after \cite{HETAL2019JAMES}, also provided in Table~\ref{tbl:table1}.
+The values of $\nu$ are a factor 2.5 greater for divergence damping and are not shown.
 $\Delta t_{phys}$ is chosen to scale with resolution, in proportion to the grid spacing,
@@
 where $\Delta t_{phys,0}$ is taken to be the standard $1800 s$ used in CAM-SE-CSLAM at the standard resolution, $n_{e,0} = ne30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km).
-This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test (Appendix~\ref{sec:app1}), and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS,HR2018JAMES}.
+This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test \citep[Appendix A in][]{HETAL2019JAMES}, and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS}.
 The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.
@@
 \begin{center}
-\noindent\includegraphics[width=25pc,angle=0]{figs/temp_cape.pdf}\\
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_cape.pdf}\\
 \end{center}

### 2019-08-08 resolution.tex
edits

@@
 
-\input{results.tex}
+\section{Methods}
+
+\subsection{Dynamical Core}
+
+This study uses the spectral-element dynamical core option of Community Atmosphere Model \citep[CAM-SE;][]{DetAl2012IJHPCA}, coupled with a mass conserving, semi-Lagrangian advection method for accelerated multi-tracer transport \citep[CSLAM;][]{LTOUNGK2017MWR}, and dry-mass vertical coordinate with comprehensive treatment of moisture and energy \citep{LetAl2018JAMES}.
+The dry dynamics are solved using the high-order, momentum, mass and energy conserving spectral element method \citep{TF2010JCP}, with the elements defined by a cubed-sphere grid.
+The notation for the horizontal grid resolution is an `$ne$' followed by the number of elements making up an edge of one cubed-sphere face, e.g., $ne30$.
+Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds.
+CSLAM tracer transport uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.
+
+\subsection{Physical Parameterizations}
+
+The physics are evaluated on the finite-volume CSLAM grid, and the tendencies mapped back to the spectral element grid.
+The coupled system, referred to as CAM-SE-CSLAM, conserves energy, mass and preserves linear correlations between two reactive species to within machine precision \citep{HL2018MWR}.
+A coarser physics grid, containing $\frac{5}{9}$ fewer degrees of freedom than the dynamical core grid is also available as part of the CAM-SE-CSLAM package \citep{HETAL2019JAMES}.
+This lower-resolution physics grid is used in this study, but only as a member of a perturbed parameter ensemble and not in the default convergence experiment.
+The dynamics time-step is subcycled within a longer physics time-step $\Delta t_{phys}$, and the temperature and momentum increments from the physics are divided by the number of subcycles and added to the dynamical core at the beginning of each subcycle.
+The full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LetAl2018JAMES}.
+
+The simulations use the CAM6 physics package.
+The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed probability distribution function (PDF) high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
+The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with a three mode Modular Aerosol Model \citep{MAM}.
+The combined macrophysics/microphysics routines generate grid-scale clouds with stratiform precipitation.
+Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} and a dilute form of the convective available potential energy (CAPE) is computed and used as the convective trigger (convection occurs if dilute CAPE $\geq 70$ J/kg), and for closing the mass fluxes in the cloud ensemble \citep{NRJ2008JC}.
+The deep convection scheme also parameterizes convective momentum transport \citep{RR2008JC}.
+
+\subsection{Experimental Design}
+
+The convergence experiment is performed in an aqua-planet configuration \citep{NH2000ASL,MWO2016JAMES}, an all ocean planet with fixed, zonally symmetric sea surface temperatures modeled after present day Earth \citep[$QOBS$ in][]{NH2000ASL}.
+The aqua-planets are in a perpetual equinox, and aerosols are largely absent from the simulations.
+Each simulation is ran for one simulated year.
+Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
+The horizontal hyper-viscosity operators $\nu$ vary with resolution after \cite{HETAL2019JAMES}, also provided in Table~\ref{tbl:table1}.
+The values of $\nu$ are a factor 2.5 greater for divergence damping and are not shown.
+$\Delta t_{phys}$ is chosen to scale with resolution, in proportion to the grid spacing,
+\begin{equation}
+\Delta t_{phys} = \Delta t_{phys,0} \times \frac{n_{e,0}}{n_e}~s,\label{eq:dt-scale}
+\end{equation}
+where $\Delta t_{phys,0}$ is taken to be the standard $1800 s$ used in CAM-SE-CSLAM at the standard resolution, $n_{e,0} = ne30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km).
+This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test \citep[Appendix A in][]{HETAL2019JAMES}, and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS}.
+The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.
+
+\section{Results}
+
+ \begin{table}
+ \caption{Experimental design and global means.}
+ \centering
+ \scriptsize
+ \begin{tabular}{lcccccc}
+   \hline
+   Variable & $ne20$ & $ne30$ & $ne40$ & $ne60$ & $ne80$ & $ne120$ \\ 
+   \hline
+   $\Delta x$ (km) & 166.8 & 111.2 & 83.4 & 55.6 & 41.7 & 27.8 \\
+   $\nu$ ($m^4/s$) & $1.5 \times 10^{15}$ & $4.0 \times 10^{14}$ & $1.5 \times 10^{14}$ & $4.0 \times 10^{13}$  & $1.5 \times 10^{13}$ & $4.0 \times 10^{12}$\\
+    $\Delta t_{phys}$ (s) & 2700 & 1800 & 1350 & 900 & 675 & 450 \\
+   Total Cloud Fraction & 0.844 & 0.835 & 0.824 & 0.810 & 0.804 & 0.800 \\ 
+   Total Precipitable Water (mm) & 23.31& 23.01 & 22.62 & 22.25 & 21.93 & 21.72 \\
+   Convective Precipitation (mm/day) & 1.91 & 1.83 & 1.68 & 1.47 & 1.29 & 1.08 \\
+   Stratiform Precipitation (mm/day) & 1.26 & 1.42 & 1.60 & 1.85 & 2.05 & 2.22 \\      
+ \hline
+ \end{tabular}
+ \label{tbl:table1}
+ \end{table}
+
+Table~\ref{tbl:table1} provides globally averaged, climatological (omitting the first month) metrics for the CAM6 simulations which are typically published in CAM/CCM convergence studies.
+Total precipitable water, total cloud fraction and deep convective precipitation rate decreases, while stratiform precipitation increases, monotonically with resolution.
+Resolution sensitivity in CAM6 is similar to all prior versions of the model.
+
+\subsection{Vertical Velocities and Resolution}
+
+\begin{figure}[t]
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_2pdf.pdf}\\
+\end{center}
+\caption{Probability density distribution of the upward vertical pressure velocities $\omega$ computed everywhere in the model from 1 year of 6-hourly data (a) raw values (solid) and values remapped to the $ne20$ grid (dotted), (b) values scaled to the $ne120$ resolution.}
+\label{fig:2pdf}
+\end{figure}
+
+The PDF of negative, or upward vertical pressure velocities $\omega$ in the aqua-planets is shown in Figure~\ref{fig:2pdf}a.
+The magnitude of upward $\omega$ increases monotonicaly with resolution, with positive, or downward $\omega$ behaving similarly (not shown).
+This monotonic increase in the magnitude of $\omega$ is evident even after remapping all the model output to a common grid ($ne20$; dotted curves Figure~\ref{fig:2pdf}a).
+
+The PDF's may be scaled to the highest-resolution resolution grid through $P(\omega)_s = \alpha P (\omega / \alpha)$, where $\alpha$ is the scale factor from equation~\ref{eq:alpha}, and setting $\Delta x_0$ to the $ne120$ grid-spacing.
+Figure~\ref{fig:2pdf}b shows the scaled PDFs using $n=-1$ in equation~\ref{eq:alpha}.
+The scaled PDF's all collapse onto the high-resolution reference, indicating the a power-law with $n=-1$ explains to first-order the variation in vertical velocity with resolution in the aqua-planet simulations.
+
+Changes to the vertical velocity field can be further understood through decomposing the mass weighted vertical mean $\omega$ into upward and downward components,
+\begin{equation}
+\langle \omega \rangle =\langle f_{u} \rangle \, \langle \omega_{u} \rangle + \langle f_{d} \rangle \, \langle \omega_{d} \rangle, \label{eq:omega}
+\end{equation}
+where $\langle f_x \rangle$ and $\langle \omega_x \rangle$ refers to the vertical mass fraction $ \left( \frac{\int dp_x}{\int dp} \right)$ and the $x$ component of the mass weighted vertical mean of $\omega$ $ \left( \frac{\int \omega_x dp_x}{\int dp_x} \right)$, respectively, subscript $u$ refers to upward motion and $d$, downward motion.
+
+The global mean, climatological components $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ are provided in Figure~\ref{fig:4panel}a,b for the aqua-planet simulations.
+The magnitude of both $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increase monotonically with resolution, and are both equal and opposite, which is a requirement of mass conservation in the model and a convenient check of the calculation.
+The individual component $\langle f_{x} \rangle$ varies by only few percent in the simulations, and so the monotonic behavior of $\langle f_{x} \rangle \, \langle \omega_{x} \rangle$ is primarily due to $ \langle \omega_{x} \rangle$ (not shown).
+
+\begin{figure}[t]
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_diags_4panel.pdf}\\
+\end{center}
+\caption{(a,b,c,e,f,g) Components of the global mean vertical pressure velocity $\overline{\langle \omega \rangle}$ (d) ratio of global mean ZM precipitation rate to the total precipitation rate versus grid spacing $\Delta x$ and (e) scatter plot of $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ and FREQZM, and the fitted linear regression which has a Pearson's R-value = 0.99.
+(a) $\overline{\langle \omega_{u} \rangle}$ (b) $\overline{\langle f_{u} \rangle}$ (c) $\overline{\langle f_{u} \rangle} \, \overline{\langle \omega_{u} \rangle}$ (e) $\overline{\langle \omega_{d} \rangle}$ (f) $ \overline{\langle f_{d} \rangle}$ (g) $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$.
+Colors are as in Figure~\ref{fig:2pdf}.
+Grey crosses are for the perturbed parameter runs.}
+\label{fig:4panel}
+\end{figure}
+
+\subsection{Convective Precipitation and Vertical Velocities}
+
+The \cite{ZM1995AO} deep convection scheme (referred to as the $ZM$ scheme, hereafter) is modulated by the dilute CAPE calculation, which itself is intertwined with the vertical velocity field \citep{SZ2018JCLIM}.
+The CAPE budget can be separated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through advection of dry static energy and moisture by the environment, i.e., the resolved flow.
+The latter term is defined as,
+\begin{equation}
+-R_d \int_{p_t}^{p_b} \frac{\partial T_{ve}}{\partial t} dlnp \label{eq:cape}
+\end{equation}
+where $T_{ve}$ is the virtual temperature of the environment, $R_d$ the gas constant for dry air and subscripts $b$ and $t$ refer to the parcel launch level (typically in the boundary layer) and the level of neutral buoyancy, respectively \citep{Z2002JGR}.
+Equation~\ref{eq:cape} contains a term, the vertical advection of potential energy, which simplifies to $w \frac{\partial gz}{\partial z} = gw$, and directly proportional to the vertical velocity $w$ by the factor $g$, the acceleration of gravity (F.
+Song, personal communication).
+In subsiding regions, $w$ is negative and opposes the generation of CAPE through adiabatic warming of the environment.
+
+There is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the ZM scheme in the simulations, $FREQZM$ (Figure~\ref{fig:4panel}c).
+At any given grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.
+Time-mean $FREQZM$ therefore indicates the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg.
+It is hypothesized that the greater subsiding motion with resolution opposes the generation of dilute CAPE through subsidence warming, reducing the frequency that the ZM scheme is triggered, and leading to the strong negative correlation in Figure~\ref{fig:4panel}c.
 

### 2019-08-08 results.tex
edits

@@
 The simulations use the CAM6 physics package.
-The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed PDF high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
+The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed probability distribution function (PDF) high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
 The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with a three mode Modular Aerosol Model \citep{MAM}.
 The combined macrophysics/microphysics routines generate grid-scale clouds with stratiform precipitation.
-Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} and a dilute form of the convective available potential energy (CAPE) is computed to assess the convective trigger ($\geq 70$ J/kg), and for closing the mass fluxes in the cloud ensemble \citep{NRJ2008JC}.
+Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} and a dilute form of the convective available potential energy (CAPE) is computed and used as the convective trigger (convection occurs if dilute CAPE $\geq 70$ J/kg), and for closing the mass fluxes in the cloud ensemble \citep{NRJ2008JC}.
 The deep convection scheme also parameterizes convective momentum transport \citep{RR2008JC}.
@@
    \hline
+   $\Delta x$ (km) & 166.8 & 111.2 & 83.4 & 55.6 & 41.7 & 27.8 \\
    $\nu$ ($m^4/s$) & $1.5 \times 10^{15}$ & $4.0 \times 10^{14}$ & $1.5 \times 10^{14}$ & $4.0 \times 10^{13}$  & $1.5 \times 10^{13}$ & $4.0 \times 10^{12}$\\
@@
 
-Table~\ref{tbl:table6-1} provides a list of some relevant globally averaged metrics that are typically published in CAM-lineage convergence studies.
+Table~\ref{tbl:table1} provides globally averaged, climatological (omitting the first month) metrics for the CAM6 simulations which are typically published in CAM/CCM convergence studies.
 Total precipitable water, total cloud fraction and deep convective precipitation rate decreases, while stratiform precipitation increases, monotonically with resolution.
-Resolution sensitivity in CAM6 is similar to all prior model versions within the models' lineage.
+Resolution sensitivity in CAM6 is similar to all prior versions of the model.
 
-The probability density function (PDF) of negative, or upward vertical pressure velocities $\omega$ in the aqua-planets is shown in Figure~\ref{fig:2pdf}a.
-The magnitude of upward $\omega$ increases in a monotonic way with resolution, with positive, or downward $\omega$ behaving similarly (not shown).
-The PDF's may be scaled to the high-resolution $ne120$ resolution through $P(\omega)_s = \alpha P (\omega / \alpha)$, where $\alpha$ is the scale factor equation~\ref{eq:eq6-1}.
-In general, the scaled PDF's all line up on top of the high-resolution reference (Figure~\ref{fig:2pdf}b); equation~\ref{eq:eq6-1} explains the variation in vertical velocity across resolutions to first order.
+\subsection{Vertical Velocities and Resolution}
 
@@
 \end{center}
-\caption{Probability density distribution of the upward vertical pressure velocities $\omega$ computed everywhere in the model from 1 year of 6-hourly data (a) raw values (b) scaled to the $ne120$ resolution using $P(\omega)_s = \alpha P (\omega / \alpha)$, where $\alpha$ is from equation~\ref{eq:eq6-1}.}
+\caption{Probability density distribution of the upward vertical pressure velocities $\omega$ computed everywhere in the model from 1 year of 6-hourly data (a) raw values (solid) and values remapped to the $ne20$ grid (dotted), (b) values scaled to the $ne120$ resolution.}
 \label{fig:2pdf}
@@
 
-Changes to the time-mean vertical velocity field can be understood through decomposing the global mean $\omega$ into upward and downward components,
+The PDF of negative, or upward vertical pressure velocities $\omega$ in the aqua-planets is shown in Figure~\ref{fig:2pdf}a.
+The magnitude of upward $\omega$ increases monotonicaly with resolution, with positive, or downward $\omega$ behaving similarly (not shown).
+This monotonic increase in the magnitude of $\omega$ is evident even after remapping all the model output to a common grid ($ne20$; dotted curves Figure~\ref{fig:2pdf}a).
+
+The PDF's may be scaled to the highest-resolution resolution grid through $P(\omega)_s = \alpha P (\omega / \alpha)$, where $\alpha$ is the scale factor from equation~\ref{eq:alpha}, and setting $\Delta x_0$ to the $ne120$ grid-spacing.
+Figure~\ref{fig:2pdf}b shows the scaled PDFs using $n=-1$ in equation~\ref{eq:alpha}.
+The scaled PDF's all collapse onto the high-resolution reference, indicating the a power-law with $n=-1$ explains to first-order the variation in vertical velocity with resolution in the aqua-planet simulations.
+
+Changes to the vertical velocity field can be further understood through decomposing the mass weighted vertical mean $\omega$ into upward and downward components,
 \begin{equation}
-\overline{\langle \omega \rangle} = \overline{\langle f_{u} \rangle} \, \overline{\langle \omega_{u} \rangle} + \overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}, \label{eq:eq6-2}
+\langle \omega \rangle =\langle f_{u} \rangle \, \langle \omega_{u} \rangle + \langle f_{d} \rangle \, \langle \omega_{d} \rangle, \label{eq:omega}
 \end{equation}
-where $\langle f_x \rangle$ and $\langle \omega_x \rangle$ refers to the vertical mass fraction and contribution to mass weighted vertical mean $\omega$, respectively, subscript $u$ refers to upward motion and $d$, downward motion; overbars indicate time-mean.
+where $\langle f_x \rangle$ and $\langle \omega_x \rangle$ refers to the vertical mass fraction $ \left( \frac{\int dp_x}{\int dp} \right)$ and the $x$ component of the mass weighted vertical mean of $\omega$ $ \left( \frac{\int \omega_x dp_x}{\int dp_x} \right)$, respectively, subscript $u$ refers to upward motion and $d$, downward motion.
 
-The components of equation~\ref{eq:eq6-2} are shown in Figure~\ref{fig:8panel}a,b,e,f for the six aqua-planet simulations.
-The magnitude of both $\overline{\langle \omega_{u} \rangle}$ and $\overline{\langle \omega_{d} \rangle}$ increase monotonically with resolution, the upward component increasing more than the downward component.
-In contrast, $\overline{\langle f_{u} \rangle}$ decreases with resolution while $\overline{\langle f_{d} \rangle}$ increases, although in both cases there is a reversal in this trend in the highest resolution simulation, $ne120$.
-Note that the magnitude of the change in $\overline{\langle f_{u} \rangle}$ and $\overline{\langle f_{d} \rangle}$ is small, a range of about $0.015$ across all simulations.
-Figure~\ref{fig:8panel}c,g shows that the products  $\overline{\langle f_{u} \rangle} \, \overline{\langle \omega_{u} \rangle}$ and $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ are equal and opposite, which is a requirement of mass conservation in the model and a convenient check of the calculation.
-Besides the six aqua-planets described in Section 6.2, 23 additional year-long aqua-planet experiments were carried out at various resolutions, with modified parameters in the dynamical core and the physics.
-The spread in the components of equation~\ref{eq:eq6-2} in the simulations with perturbed parameters is relatively large for $\overline{\langle f_{x} \rangle}$, compared with $\overline{\langle \omega_{x} \rangle}$.
+The global mean, climatological components $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ are provided in Figure~\ref{fig:4panel}a,b for the aqua-planet simulations.
+The magnitude of both $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increase monotonically with resolution, and are both equal and opposite, which is a requirement of mass conservation in the model and a convenient check of the calculation.
+The individual component $\langle f_{x} \rangle$ varies by only few percent in the simulations, and so the monotonic behavior of $\langle f_{x} \rangle \, \langle \omega_{x} \rangle$ is primarily due to $ \langle \omega_{x} \rangle$ (not shown).
 
@@
 Grey crosses are for the perturbed parameter runs.}
-\label{fig:8panel}
+\label{fig:4panel}
 \end{figure}
+
+\subsection{Convective Precipitation and Vertical Velocities}
+
+The \cite{ZM1995AO} deep convection scheme (referred to as the $ZM$ scheme, hereafter) is modulated by the dilute CAPE calculation, which itself is intertwined with the vertical velocity field \citep{SZ2018JCLIM}.
+The CAPE budget can be separated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through advection of dry static energy and moisture by the environment, i.e., the resolved flow.
+The latter term is defined as,
+\begin{equation}
+-R_d \int_{p_t}^{p_b} \frac{\partial T_{ve}}{\partial t} dlnp \label{eq:cape}
+\end{equation}
+where $T_{ve}$ is the virtual temperature of the environment, $R_d$ the gas constant for dry air and subscripts $b$ and $t$ refer to the parcel launch level (typically in the boundary layer) and the level of neutral buoyancy, respectively \citep{Z2002JGR}.
+Equation~\ref{eq:cape} contains a term, the vertical advection of potential energy, which simplifies to $w \frac{\partial gz}{\partial z} = gw$, and directly proportional to the vertical velocity $w$ by the factor $g$, the acceleration of gravity (F.
+Song, personal communication).
+In subsiding regions, $w$ is negative and opposes the generation of CAPE through adiabatic warming of the environment.
+
+There is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the ZM scheme in the simulations, $FREQZM$ (Figure~\ref{fig:4panel}c).
+At any given grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, and zero if it is not.
+Time-mean $FREQZM$ therefore indicates the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg.
+It is hypothesized that the greater subsiding motion with resolution opposes the generation of dilute CAPE through subsidence warming, reducing the frequency that the ZM scheme is triggered, and leading to the strong negative correlation in Figure~\ref{fig:4panel}c.
 
@@
 \end{figure}
-
-The ZM trigger uses a dilute form of CAPE, which is consistent with the negative relation between $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ and FREQZM in space, and across resolutions.
-In the classical, non-dilute case, CAPE can be seperated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through advection of dry static energy and moisture by the environment, i.e., the resolved flow.
-The latter term generally contributes positively to CAPE in regions of ascent and negatively in regions of subsidence \citep{SZ2018JCLIM}, consistent with the negative correlation of $\langle f_{d} \rangle \langle \omega_{d} \rangle$ in space, and across resolutions.
 

### 2019-08-08 resolution.tex
edits

@@
 
-The global mean, climatological components $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ are provided in Figure~\ref{fig:4panel}a,b for the aqua-planet simulations.
+The global mean, climatological components $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ are provided in Figure~\ref{fig:2panel}a,b for the aqua-planet simulations.
 The magnitude of both $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increase monotonically with resolution, and are both equal and opposite, which is a requirement of mass conservation in the model and a convenient check of the calculation.
@@
 \begin{center}
-\noindent\includegraphics[width=20pc,angle=0]{figs/temp_diags_4panel.pdf}\\
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_diags_2panel.pdf}\\
 \end{center}
@@
 Grey crosses are for the perturbed parameter runs.}
-\label{fig:4panel}
+\label{fig:2panel}
 \end{figure}
@@
 
+\begin{figure}[t]
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_diags_corr.pdf}\\
+\end{center}
+\caption{Scatter plot of $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ and FREQZM, and the fitted linear regression which has a Pearson's R-value = 0.99.
+Colors are as in Figure~\ref{fig:2pdf}.
+Grey crosses are for the perturbed parameter ensemble.}
+\label{fig:corr}
+\end{figure}
+
 There is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the ZM scheme in the simulations, $FREQZM$ (Figure~\ref{fig:4panel}c).

### 2019-08-08 resolution.tex
edits

@@
 \end{center}
-\caption{(a,b,c,e,f,g) Components of the global mean vertical pressure velocity $\overline{\langle \omega \rangle}$ (d) ratio of global mean ZM precipitation rate to the total precipitation rate versus grid spacing $\Delta x$ and (e) scatter plot of $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ and FREQZM, and the fitted linear regression which has a Pearson's R-value = 0.99.
-(a) $\overline{\langle \omega_{u} \rangle}$ (b) $\overline{\langle f_{u} \rangle}$ (c) $\overline{\langle f_{u} \rangle} \, \overline{\langle \omega_{u} \rangle}$ (e) $\overline{\langle \omega_{d} \rangle}$ (f) $ \overline{\langle f_{d} \rangle}$ (g) $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$.
+\caption{(a,b,c,e,f,g) Components of the global mean vertical pressure velocity, (a) $\overline{\langle f_{u} \rangle} \, \overline{\langle \omega_{u} \rangle}$ (b) $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$.
 Colors are as in Figure~\ref{fig:2pdf}.
@@
 
-There is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the ZM scheme in the simulations, $FREQZM$ (Figure~\ref{fig:4panel}c).
+There is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the ZM scheme in the simulations, $FREQZM$ (Figure~\ref{fig:corr}).
 At any given grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.
 Time-mean $FREQZM$ therefore indicates the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg.
-It is hypothesized that the greater subsiding motion with resolution opposes the generation of dilute CAPE through subsidence warming, reducing the frequency that the ZM scheme is triggered, and leading to the strong negative correlation in Figure~\ref{fig:4panel}c.
+It is hypothesized that the greater subsiding motion with resolution opposes the generation of dilute CAPE through subsidence warming, reducing the frequency the ZM scheme is triggered, and creating the strong negative correlation in Figure~\ref{fig:corr}.
+
+\begin{figure}[t]
+\begin{center}
+\noindent\includegraphics[width=14pc,angle=0]{figs/temp_4zonal.pdf}\\
+\end{center}
+\caption{Zonal mean (a) R-values and (b) sensitivity parameter in the logistic regression, (d) time mean surface latent heat fluxes and (c) drizzle fraction. Colors are as in Figure~\ref{fig:2pdf}.}
+\label{fig:4zonal}
+\end{figure}
+
+To test the hypothesis, a logistic regression between subsiding motion and $FREQZM$ is performed for each grid column within each of the simulations.
+Logistic regression uses an iterative method to fit a continuous variable predictor, $x$ to a binary predictand $p$ \citep{WILKSBOOK},
+\begin{equation}
+p = \frac{exp{[b_0 + b_1 x]}}{1 + exp{[b_0 + b_1 x]}}, \label{eq:eq6-3}
+\end{equation}
+where $b_0$ and $b_1$ are the shape parameters of the exponential.
+The predictor is then the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
+Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and parameter $b_1$ (hereafter referred to as the sensitivity parameter; Figure~\ref{fig:4zonal}a,b).
+
+The zonal mean R-values indicates the greatest goodness of fit in the deep tropics ($\pm 10^{circ}$ latitude).
+Figure~\ref{fig:4zonal}c shows the time mean zonal mean latent heat flux in the simulations, which is expected to contribute positively to CAPE through the component associated with the thermodynamic state of boundary layer parcels.
+In the deep tropics, the latent heat flux is small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), which is consistent with our hypothesis that the subsiding motion actively depresses CAPE and the activity of the ZM scheme in the simulations.
+The sensitivity parameter becomes less negative in the deep tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary of whether the ZM scheme is active.
+The R-values generally decrease with resolution indicating that there is degradation in the relationship with resolution.
+All values shown are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK}.
+
+\begin{figure}[t]
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_cape.pdf}\\
+\end{center}
+\caption{(a) time mean fraction of the deep tropics in the simulations with upward $\langle \omega \rangle$ (red) and downward $\langle \omega \rangle$ (blue), (b) CAPE computed from the mean temperature and moisture profiles of upward regions and downward regions.
+Black is for CAPE computed from the mean temperature and moisture profiles for the entire deep tropics, grey is the approximate discussed in the text.}
+\label{fig:cape}
+\end{figure}
+
+To understand the degree to which atmospheric stability in the deep tropics is dependent on ascending and descending motion, temperature and moisture profiles are conditionally sampled based on whether $\langle \omega \rangle$ is positive or negative, indicating predominantly subsiding or ascending grid columns.
+The mean temperature and moisture profiles of subsiding and ascending regions in the deep tropics are used to compute the dilute-CAPE used by the ZM scheme, offline.
+Figure~\ref{fig:cape}a indicates that the mean profile of ascending regions are associated with large values of CAPE ($>170$ J/kg), and low values in subsiding regions ($<110$ J/kg).
+Figure~\ref{fig:profiles} shows the climatological temperature and specific humidity profiles of subsiding grid cells, averaged over the deep tropics and presented as anomalies from the mean temperature and specific humidity of the entire deep Tropics.
+The mean profiles of subsiding regions have an anomalous warming layer in the $600-800$ hPa layer and an anomalous moisture deficit throughout the entire column.
+This warming of the environment in subsiding regions opposes the generation of CAPE through equation~\ref{eq:cape}.
+
+\begin{figure}[t]
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_profiles.pdf}\\
+\end{center}
+\caption{}
+\label{fig:profiles}
+\end{figure}
+
+Figure~\ref{fig:cape}b shows that fractional area of air columns in the deep tropics that are predominantly subsiding (ascending) changes drastically with resolution, from $0.32$ ($0.68$) in the $ne20$ ($\Delta x = 166.8$ km) run, and monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in the $ne120$ ($\Delta x = 27.8$ km) run.
+Interestingly, the sum of the product of the fractional areas with their corresponding CAPE values gives approximately the same CAPE values computed from the mean temperature and moisture profiles over the entire deep tropics (Figure~\ref{fig:cape}a).
+This provides strong evidence that CAPE values in the deep tropics are decreasing with resolution because a larger (smaller) space-time fraction of the deep tropics are made up of predominantly subsiding (ascending) grid columns with increasing resolution.
+
+Poleward of the deep tropics and within the subtropics, the logistic regression indicates that subsidence is a poor predictor of $FREQZM$ (Figure~\ref{fig:4reg}a,b).
+The R-value decreases to a local minimum between $15^{\circ} - 20^{\circ}$ latitude, and the magnitude of the sensitivity parameter steeply declines.
+The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4reg}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
+The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4reg}d).
+The predominance of drizzle in this region is probably a result of the large subsiding motion in the subtropics (not shown) constraining CAPE from becoming too large.
+AGCMs are known to suffer from an excess drizzle bias in precisely this region \citep{D2006JCLIM}, and this analysis indicates that this bias is due to the use of a CAPE trigger function.
 

### 2019-08-08 resolution.tex
edits

@@
 
-Poleward of the deep tropics and within the subtropics, the logistic regression indicates that subsidence is a poor predictor of $FREQZM$ (Figure~\ref{fig:4reg}a,b).
+Poleward of the deep tropics and within the subtropics, the logistic regression indicates that subsidence is a poor predictor of $FREQZM$ (Figure~\ref{fig:4zonal}a,b).
 The R-value decreases to a local minimum between $15^{\circ} - 20^{\circ}$ latitude, and the magnitude of the sensitivity parameter steeply declines.
-The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4reg}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
-The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4reg}d).
+The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4rzonal}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
+The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d).
 The predominance of drizzle in this region is probably a result of the large subsiding motion in the subtropics (not shown) constraining CAPE from becoming too large.

### 2019-08-08 resolution.tex
edits

@@
 Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum for mesoscale motions \citep{NG1985JAS}.
-\cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope is common in both observations and models, and provides an emergent constraint for $n= -\frac{2}{3}$.
+\cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope is common in both observations and models, and provides an emergent constraint for $b=\frac{1}{3}$ and $n= -\frac{2}{3}$.
 
@@
 The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4rzonal}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
-The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d).
+The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d), discussed in \cite{TETAL2018CD}.
 The predominance of drizzle in this region is probably a result of the large subsiding motion in the subtropics (not shown) constraining CAPE from becoming too large.

### 2019-08-08 resolution.tex
edits

@@
 The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4rzonal}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
-The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d), discussed in \cite{TETAL2018CD}.
+The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d), also discussed in \cite{TETAL2018CD}.
 The predominance of drizzle in this region is probably a result of the large subsiding motion in the subtropics (not shown) constraining CAPE from becoming too large.

### 2019-08-08 resolution.tex
edits

@@
 The magnitude of both $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increase monotonically with resolution, and are both equal and opposite, which is a requirement of mass conservation in the model and a convenient check of the calculation.
-The individual component $\langle f_{x} \rangle$ varies by only few percent in the simulations, and so the monotonic behavior of $\langle f_{x} \rangle \, \langle \omega_{x} \rangle$ is primarily due to $ \langle \omega_{x} \rangle$ (not shown).
+While $\langle f_{d} \rangle$ is about 25\% larger than $\langle f_{u} \rangle$ in all simulations, the vertical mass fractions varies by only few percent with resolution, and so the monotonic behavior of $\langle f_{x} \rangle \, \langle \omega_{x} \rangle$ with resolution is primarily from variations in $ \langle \omega_{x} \rangle$ (not shown).
 
@@
 
-\subsection{Convective Precipitation and Vertical Velocities}
+\subsection{Vertical Velocities and Convective Precipitation}
 
@@
 
-The zonal mean R-values indicates the greatest goodness of fit in the deep tropics ($\pm 10^{circ}$ latitude).
-Figure~\ref{fig:4zonal}c shows the time mean zonal mean latent heat flux in the simulations, which is expected to contribute positively to CAPE through the component associated with the thermodynamic state of boundary layer parcels.
+The zonal mean R-values indicates the greatest goodness of fit in the deep tropics ($\pm 10^{\circ}$ latitude).
+Figure~\ref{fig:4zonal}c shows the climatological, zonal-mean latent heat flux in the simulations, which is expected to contribute positively to CAPE through the component associated with the thermodynamic state of boundary layer parcels.
 In the deep tropics, the latent heat flux is small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), which is consistent with our hypothesis that the subsiding motion actively depresses CAPE and the activity of the ZM scheme in the simulations.
@@
 The R-value decreases to a local minimum between $15^{\circ} - 20^{\circ}$ latitude, and the magnitude of the sensitivity parameter steeply declines.
-The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4rzonal}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
+The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4zonal}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
 The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d), also discussed in \cite{TETAL2018CD}.

### 2019-08-08 resolution.tex
edits

@@
 Strictly speaking, $\Delta x$ here refers to the distance between two points for which the velocity increment is computed in the structure function, but with this distance set to the model grid-spacing.
-Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum for mesoscale motions \citep{NG1985JAS}.
-\cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope is common in both observations and models, and provides an emergent constraint for $b=\frac{1}{3}$ and $n= -\frac{2}{3}$.
+Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum, which is true for observations of mesoscale flow \citep{NG1985JAS}.
+\cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope being common in both observations and models, provides an emergent constraint for $b=\frac{1}{3}$ and $n= -\frac{2}{3}$.
 

### 2019-08-09 resolution.tex
edits

@@
 An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x = 5$ km and less) and climate ($\Delta x = 50$ km and less) simulations \citep{MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
-These models are built using unstructured meshes that while allows for substantial grid flexibility, require physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
+These models are built using unstructured meshes that while allows for substantial grid flexibility, would require physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
 The most common approach towards developing scale-aware physics is through the lens of limited area, large-eddy simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.

### 2019-08-10 resolution.tex
edits

@@
 Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum, which is true for observations of mesoscale flow \citep{NG1985JAS}.
-\cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope being common in both observations and models, provides an emergent constraint for $b=\frac{1}{3}$ and $n= -\frac{2}{3}$.
+\cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope being common in both observations and models provides an emergent constraint for $b=\frac{1}{3}$ and $n= -\frac{2}{3}$.
 
@@
 This behavior is summarized in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies.
+The propensity for precipitation rates to shift from the convective to the stratiform scheme with resolution has been documented in other AGCMs as well \citep{PS2002CD,TETAL2018CD}, but none have given a satisfactory explanation for this sensitivity.
 The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the tendency to reduce $\Delta t_{phys}$ with resolution would by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convective precipitation rates with resolution.
@@
 \end{equation}
-where $\Delta t_{phys,0}$ is taken to be the standard $1800 s$ used in CAM-SE-CSLAM at the standard resolution, $n_{e,0} = ne30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km).
+where $\Delta t_{phys,0}$ is taken to be the standard $1800$ s used in CAM-SE-CSLAM at the standard resolution, $n_{e,0} = ne30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km).
 This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test \citep[Appendix A in][]{HETAL2019JAMES}, and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS}.
@@
 Figure~\ref{fig:2pdf}b shows the scaled PDFs using $n=-1$ in equation~\ref{eq:alpha}.
-The scaled PDF's all collapse onto the high-resolution reference, indicating the a power-law with $n=-1$ explains to first-order the variation in vertical velocity with resolution in the aqua-planet simulations.
+The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law with $n=-1$ explains to first-order the variation in vertical velocity with resolution in the aqua-planet simulations.
 
@@
 The global mean, climatological components $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ are provided in Figure~\ref{fig:2panel}a,b for the aqua-planet simulations.
-The magnitude of both $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increase monotonically with resolution, and are both equal and opposite, which is a requirement of mass conservation in the model and a convenient check of the calculation.
+The magnitude of both $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increase monotonically with resolution, and are equal and opposite, which is a requirement of mass conservation in the model and a convenient check of the calculation.
 While $\langle f_{d} \rangle$ is about 25\% larger than $\langle f_{u} \rangle$ in all simulations, the vertical mass fractions varies by only few percent with resolution, and so the monotonic behavior of $\langle f_{x} \rangle \, \langle \omega_{x} \rangle$ with resolution is primarily from variations in $ \langle \omega_{x} \rangle$ (not shown).
@@
 where $T_{ve}$ is the virtual temperature of the environment, $R_d$ the gas constant for dry air and subscripts $b$ and $t$ refer to the parcel launch level (typically in the boundary layer) and the level of neutral buoyancy, respectively \citep{Z2002JGR}.
-Equation~\ref{eq:cape} contains a term, the vertical advection of potential energy, which simplifies to $w \frac{\partial gz}{\partial z} = gw$, and directly proportional to the vertical velocity $w$ by the factor $g$, the acceleration of gravity (F.
+Equation~\ref{eq:cape} shows that warming the environment reduces CAPE though reducing the buoyancy of lifted air parcels.
+The $T_{ve}$ budget contains a term; the vertical advection of potential energy, which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
 Song, personal communication).
@@
 
-There is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the ZM scheme in the simulations, $FREQZM$ (Figure~\ref{fig:corr}).
+There is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the ZM scheme in the simulations, global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
 At any given grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.
@@
 To test the hypothesis, a logistic regression between subsiding motion and $FREQZM$ is performed for each grid column within each of the simulations.
-Logistic regression uses an iterative method to fit a continuous variable predictor, $x$ to a binary predictand $p$ \citep{WILKSBOOK},
+Logistic regression uses an iterative method to fit a continuous variable predictor, $x$ to a binary predictand $p$ through the exponential \citep{WILKSBOOK},
 \begin{equation}
@@
 The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4zonal}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
-The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d), also discussed in \cite{TETAL2018CD}.
+The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d).
 The predominance of drizzle in this region is probably a result of the large subsiding motion in the subtropics (not shown) constraining CAPE from becoming too large.

### 2019-08-14 resolution.tex
edits

@@
 
+\subsection{Convection Activity and Atmospheric Stability}
+
 \begin{figure}[t]
@@
 
+ \begin{table}
+ \caption{Fractional contribution of the deep Tropics ($\pm 15^{\circ}$) to changes in global mean precipitation with resolution.
+The deep Tropics accounts for a majority of the resolution sensitivity despite being only about $25\%$ of the area of the planet.
+The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc...}
+ \centering
+ \scriptsize
+ \begin{tabular}{lcccccc}
+   \hline
+   Variable & $ne30$ & $ne40$ & $ne60$ & $ne80$ & $ne120$ \\ 
+   \hline
+   Convective Precipitation & 0.22 & 0.75 & 0.73 & 0.79 & 0.72 \\
+   Stratiform Precipitation & 0.46 & 0.64 & 0.71 & 0.70 & 0.49 \\      
+ \hline
+ \end{tabular}
+ \label{tbl:table2}
+ \end{table}
+
 \begin{figure}[t]
@@
 The mean temperature and moisture profiles of subsiding and ascending regions in the deep tropics are used to compute the dilute-CAPE used by the ZM scheme, offline.
-Figure~\ref{fig:cape}a indicates that the mean profile of ascending regions are associated with large values of CAPE ($>170$ J/kg), and low values in subsiding regions ($<110$ J/kg).
+Figure~\ref{fig:cape}a indicates that the mean profile of ascending regions are associated with large values of CAPE ($>180$ J/kg), and low values in subsiding regions ($<110$ J/kg).
 Figure~\ref{fig:profiles} shows the climatological temperature and specific humidity profiles of subsiding grid cells, averaged over the deep tropics and presented as anomalies from the mean temperature and specific humidity of the entire deep Tropics.
@@
 
+\subsection{Vertical Velocities and Stratiform Precipitation}
+
+\subsection{Biases Unrelated to Resolution}
+
 Poleward of the deep tropics and within the subtropics, the logistic regression indicates that subsidence is a poor predictor of $FREQZM$ (Figure~\ref{fig:4zonal}a,b).

### 2019-08-14 resolution.tex
edits

@@
 Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
+All analyses exclude the first month of the simulations, and are computed on their native grids unless otherwise stated.
+
 The horizontal hyper-viscosity operators $\nu$ vary with resolution after \cite{HETAL2019JAMES}, also provided in Table~\ref{tbl:table1}.
@@
 
- \begin{table}
+  \begin{table*}
  \caption{Experimental design and global means.}
@@
  \label{tbl:table1}
- \end{table}
+ \end{table*}
 
-Table~\ref{tbl:table1} provides globally averaged, climatological (omitting the first month) metrics for the CAM6 simulations which are typically published in CAM/CCM convergence studies.
+Table~\ref{tbl:table1} provides globally averaged, climatological metrics for the CAM6 simulations which are typically published in CAM/CCM convergence studies.
 Total precipitable water, total cloud fraction and deep convective precipitation rate decreases, while stratiform precipitation increases, monotonically with resolution.
@@
 
-\begin{figure}[t]
+\begin{figure}
 \begin{center}
@@
 \end{center}
-\caption{Probability density distribution of the upward vertical pressure velocities $\omega$ computed everywhere in the model from 1 year of 6-hourly data (a) raw values (solid) and values remapped to the $ne20$ grid (dotted), (b) values scaled to the $ne120$ resolution.}
+\caption{Probability density distribution of the upward vertical pressure velocities $\omega$ computed everywhere in the model from six-hourly output over the entirety of the year-long simulations.
+(a) Values on their native grid (solid) and values remapped to the $ne20$ grid (dotted), (b) values on their native grid, scaled to the $ne120$ resolution.}
 \label{fig:2pdf}
@@
 The deep Tropics accounts for a majority of the resolution sensitivity despite being only about $25\%$ of the area of the planet.
-The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc...}
+The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc..., which are computed through remapping all data to the $ne20$ grid.}
  \centering

### 2019-08-14 resolution.tex
edits

@@
  \begin{table}
- \caption{Fractional contribution of the deep Tropics ($\pm 15^{\circ}$) to changes in global mean precipitation with resolution.
-The deep Tropics accounts for a majority of the resolution sensitivity despite being only about $25\%$ of the area of the planet.
+ \caption{Fractional contribution of latitude bands $\pm 10^{\circ}$ and $\pm 15^{\circ}$ to changes in global mean precipitation with resolution.
 The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc..., which are computed through remapping all data to the $ne20$ grid.}
@@
    \hline
+   $\pm 10^{\circ}$ ($17.6\%$ of global area) \\
+   Convective Precipitation & -0.58 & 0.62 & 0.66 & 0.72 & 0.70 \\
+   Stratiform Precipitation & 0.55 & 0.63 & 0.69 & 0.67 & 0.41 \\ 
+   \hline
+   $\pm 15^{\circ}$ ($25.8\%$ of global area) \\
    Convective Precipitation & 0.22 & 0.75 & 0.73 & 0.79 & 0.72 \\
@@
 
+\begin{figure}[t]
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_pdecomp.pdf}\\
+\end{center}
+\caption{}
+\label{fig:profiles}
+\end{figure}
+
 \subsection{Biases Unrelated to Resolution}

### 2019-08-15 resolution.tex
edits

@@
 
-\author{Adam R. Herrington\corrauth, Kevin A. Reed}
+\author{Adam R. Herrington\corrauth \& Kevin A. Reed}
 \address{School of Marine and Atmospheric Sciences, Stony Brook University, Stony Brook, NY 11794}
@@
 
-\keywords{Climate models, physical parameterizations, physics-dynamics coupling}
+\keywords{Climate models, physical parameterizations, resolution sensitivity}
 
@@
 
-An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x = 5$ km and less) and climate ($\Delta x = 50$ km and less) simulations \citep{MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
+An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x \leq 5$ km) and climate ($\Delta x \leq 50$ km) simulations \citep{SMTMN2008JCP,MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
 These models are built using unstructured meshes that while allows for substantial grid flexibility, would require physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
@@
 The sensitivity of the Community Atmosphere Model \citep[CAM;][]{CAM5}, and its predecessor, the Community Climate Model (CCM) to resolution ({\em{resolution}} refers to {\em{horizontal resolution}}, hereafter) is well documented through convergence studies \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}.
-Despite thirty years of continual model development, there are robust sensitivities to resolution that have persisted in all versions of the model.
+{\color{red}{CAM/CCM is a well supported climate model,}} but despite thirty years of continual model development, there are robust sensitivities to resolution that have persisted in all versions of the model.
 This study argues that a unifying cause, the inherent scale sensitivities of the underlying dynamical equations, can explain the robust responses to resolution that occur in CAM/CCM, {\color{red}{since it is difficult to conceive that inevitable responses to native grid resolution could be ignored in the pursuit of scale-aware physics.}}
@@
 Strictly speaking, $\Delta x$ here refers to the distance between two points for which the velocity increment is computed in the structure function, but with this distance set to the model grid-spacing.
-Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum, which is true for observations of mesoscale flow \citep{NG1985JAS}.
+Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum, which is supported by observations of mesoscale flow \citep{NG1985JAS}.
 \cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope being common in both observations and models provides an emergent constraint for $b=\frac{1}{3}$ and $n= -\frac{2}{3}$.
@@
 In CAM aqua-planet simulations, the largest source of buoyancy is from grid-scale cloud formation, whose horizontal extents are set by the effective resolution of the model (i.e., some multiple of $\Delta x$), indicating $n=-1$ \citep{HR2018JAMES}.
-\cite{HR2017JCLIM} has shown that the $n=-1$ scaling does not explain the behavior of CAM4 in a convergence experiment, but follow-up studies \citep{HR2018JAMES,HETAL2019JAMES} indicate that the inadequacy of the $n=-1$ scaling is not definitive due to time-truncation errors associated with fixing the physics time-step ($\Delta t_{phys}$) across resolutions in that study.
+\cite{HR2017JCLIM} has shown that the $n=-1$ scaling does not explain the behavior of CAM4 in a convergence experiment, but follow-up studies \citep{HR2018JAMES,HETAL2019JAMES} indicate that the inadequacy of the $n=-1$ scaling is not definitive, due to time-truncation errors associated with fixing the physics time-step ($\Delta t_{phys}$) across resolutions in that study.
 
-Another robust response of the CAM/CCM lineage to resolution is an increase in stratiform precipitation rates (i.e., the precipitation from grid-scale clouds), at the expense of parameterized convective precipitation rates.
-This behavior is summarized in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies.
-The propensity for precipitation rates to shift from the convective to the stratiform scheme with resolution has been documented in other AGCMs as well \citep{PS2002CD,TETAL2018CD}, but none have given a satisfactory explanation for this sensitivity.
+Another robust response of the CAM/CCM lineage to resolution is an increase in grid-scale precipitation rates at the expense of parameterized convective precipitation rates.
+Grid-scale precipitation is formed from grid-scale clouds, but is often referred to in the literature as stratiform precipitaiton and stratiform clouds, and used hereafter.
+The behavior of the difference precipitation routines with resolution is shown in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies.
+The propensity for precipitation rates to shift from the convection scheme to the stratiform scheme with resolution has been documented in other AGCMs as well \citep{PS2002CD,TETAL2018CD}, but none have given a satisfactory explanation for this sensitivity.
 The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the tendency to reduce $\Delta t_{phys}$ with resolution would by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convective precipitation rates with resolution.
@@
 \end{center}
-\caption{Bar-graph of the convective (solid) and grid-scale (white) climatological precipitation rates in prior CAM/CCM convergence studies.
+\caption{Bar-graph of the convective (solid) and stratiform (white) climatological precipitation rates in prior CAM/CCM convergence studies.
 Each window contains a single convergence study, with identical x-axis; the approximate grid resolution.
@@
 The notation for the horizontal grid resolution is an `$ne$' followed by the number of elements making up an edge of one cubed-sphere face, e.g., $ne30$.
-Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds.
+Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds \citep{LetAl2018JAMES}.
 CSLAM tracer transport uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.
@@
 The dynamics time-step is subcycled within a longer physics time-step $\Delta t_{phys}$, and the temperature and momentum increments from the physics are divided by the number of subcycles and added to the dynamical core at the beginning of each subcycle.
-The full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LetAl2018JAMES}.
+The full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LW2019JAMES}.
 
@@
 The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with a three mode Modular Aerosol Model \citep{MAM}.
-The combined macrophysics/microphysics routines generate grid-scale clouds with stratiform precipitation.
-Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} and a dilute form of the convective available potential energy (CAPE) is computed and used as the convective trigger (convection occurs if dilute CAPE $\geq 70$ J/kg), and for closing the mass fluxes in the cloud ensemble \citep{NRJ2008JC}.
+The combined macrophysics/microphysics routines generate stratiform precipitation from stratiform clouds.
+Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} and an entraining plume model \citep[referred to as the dilute convective available potential energy, or {\em{dilute CAPE}} hereafter;][]{RB1992JAS, NRJ2008JC} is used as a convective trigger (convection occurs if dilute CAPE $\geq 70$ J/kg), and for closing the mass fluxes in the cloud ensemble.
 The deep convection scheme also parameterizes convective momentum transport \citep{RR2008JC}.
@@
 \end{equation}
-where $\Delta t_{phys,0}$ is taken to be the standard $1800$ s used in CAM-SE-CSLAM at the standard resolution, $n_{e,0} = ne30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km).
+where $\Delta t_{phys,0}$ is taken to be the standard $1800$ s used in CAM-SE-CSLAM for the standard climate resolution, $n_{e,0} = 30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km).
 This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test \citep[Appendix A in][]{HETAL2019JAMES}, and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS}.
@@
   \begin{table*}
- \caption{Experimental design and global means.}
+ \caption{Experimental design and global mean climatologies.}
  \centering
@@
 Table~\ref{tbl:table1} provides globally averaged, climatological metrics for the CAM6 simulations which are typically published in CAM/CCM convergence studies.
-Total precipitable water, total cloud fraction and deep convective precipitation rate decreases, while stratiform precipitation increases, monotonically with resolution.
+Total precipitable water, total cloud fraction and deep convective precipitation rate decreases, while stratiform precipitation increases, monotonically with resolution (also shown in Figure~\ref{fig:cam-history}).
 Resolution sensitivity in CAM6 is similar to all prior versions of the model.
@@
 The PDF of negative, or upward vertical pressure velocities $\omega$ in the aqua-planets is shown in Figure~\ref{fig:2pdf}a.
-The magnitude of upward $\omega$ increases monotonicaly with resolution, with positive, or downward $\omega$ behaving similarly (not shown).
-This monotonic increase in the magnitude of $\omega$ is evident even after remapping all the model output to a common grid ($ne20$; dotted curves Figure~\ref{fig:2pdf}a).
+The magnitude of upward $\omega$ increases monotonically with resolution, with positive, or downward $\omega$ behaving similarly (not shown).
+This monotonic increase in the magnitude of $\omega$ is evident even after remapping all the model output to a common grid ($ne20$; dotted curves in Figure~\ref{fig:2pdf}a).
 
 The PDF's may be scaled to the highest-resolution resolution grid through $P(\omega)_s = \alpha P (\omega / \alpha)$, where $\alpha$ is the scale factor from equation~\ref{eq:alpha}, and setting $\Delta x_0$ to the $ne120$ grid-spacing.
-Figure~\ref{fig:2pdf}b shows the scaled PDFs using $n=-1$ in equation~\ref{eq:alpha}.
+Figure~\ref{fig:2pdf}b shows the scaled PDF's using $n=-1$ in equation~\ref{eq:alpha}.
 The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law with $n=-1$ explains to first-order the variation in vertical velocity with resolution in the aqua-planet simulations.
@@
 \end{center}
-\caption{(a,b,c,e,f,g) Components of the global mean vertical pressure velocity, (a) $\overline{\langle f_{u} \rangle} \, \overline{\langle \omega_{u} \rangle}$ (b) $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$.
-Colors are as in Figure~\ref{fig:2pdf}.
-Grey crosses are for the perturbed parameter runs.}
+\caption{Components of the climatological, global mean vertical pressure velocity, (a) $\langle f_{u} \rangle \langle \omega_{u} \rangle$ and (b) $\langle f_{d} \rangle \langle \omega_{d} \rangle$.
+Grey crosses are for the perturbed parameter ensemble runs.}
 \label{fig:2panel}
@@
 
-The \cite{ZM1995AO} deep convection scheme (referred to as the $ZM$ scheme, hereafter) is modulated by the dilute CAPE calculation, which itself is intertwined with the vertical velocity field \citep{SZ2018JCLIM}.
-The CAPE budget can be separated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through advection of dry static energy and moisture by the environment, i.e., the resolved flow.
+The large increase in magnitude of the upward and downward vertical velocities with resolution may be expected to impact the behavior of other model components.
+The \cite{ZM1995AO} deep convection scheme (referred to as the $ZM$ scheme hereafter) is modulated by the dilute CAPE calculation, a modified form of the traditional CAPE, which is shown to be sensitive to the vertical velocity field in an analysis of a large eddy simulation \citep{SZ2018JCLIM}.
+
+The relationship between vertical velocities and CAPE falls out of the CAPE equation itself (F.
+Song, personal communication).
+The CAPE budget can be separated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through radiative cooling and advection of dry static energy and moisture by the environment, i.e., the resolved flow.
 The latter term is defined as,
@@
 \end{center}
-\caption{Scatter plot of $\overline{\langle f_{d} \rangle} \, \overline{\langle \omega_{d} \rangle}$ and FREQZM, and the fitted linear regression which has a Pearson's R-value = 0.99.
-Colors are as in Figure~\ref{fig:2pdf}.
-Grey crosses are for the perturbed parameter ensemble.}
+\caption{Scatter plot of global mean, climatological $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and FREQZM, and the fitted linear regression which has a Pearson's R-value = 0.99.
+Grey crosses are for the perturbed parameter ensemble runs.}
 \label{fig:corr}
@@
 
-\subsection{Convection Activity and Atmospheric Stability}
+\subsection{Hypothesis Testing}
 
@@
 The predictor is then the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
+Grid column regressions that are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK} are retained for analysis.
 Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and parameter $b_1$ (hereafter referred to as the sensitivity parameter; Figure~\ref{fig:4zonal}a,b).
 
-The zonal mean R-values indicates the greatest goodness of fit in the deep tropics ($\pm 10^{\circ}$ latitude).
+The zonal mean R-values indicate the greatest goodness of fit in the $\pm 10^{\circ}$ latitude band, referred to as the deep Tropics hereafter.
 Figure~\ref{fig:4zonal}c shows the climatological, zonal-mean latent heat flux in the simulations, which is expected to contribute positively to CAPE through the component associated with the thermodynamic state of boundary layer parcels.
-In the deep tropics, the latent heat flux is small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), which is consistent with our hypothesis that the subsiding motion actively depresses CAPE and the activity of the ZM scheme in the simulations.
-The sensitivity parameter becomes less negative in the deep tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary of whether the ZM scheme is active.
+In the deep Tropics, latent heat fluxes are small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), which is consistent with our hypothesis that the subsiding motion actively depresses CAPE and the activity of the ZM scheme in the simulations.
+The sensitivity parameter becomes less negative in the deep Tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary of whether the ZM scheme is active.
 The R-values generally decrease with resolution indicating that there is degradation in the relationship with resolution.
-All values shown are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK}.
+
+Table~\ref{tbl:table2} shows that fractional contribution of the deep Tropics to the climatological, global mean change in convective precipitation with resolution.
+The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep Tropics (except in going from $ne20$ to $ne30$, where deep Tropical convective precipitation rates increase with resolution due to an extremely wide double-ITCZ in the $ne20$ simulation that spans well outside of $\pm 10^{\circ}$ latitude).
+Expanding the latitudes of consideration marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne20$).
+Taken together, the region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme, providing further evidence in support of the hypothesis.
 
@@
  \caption{Fractional contribution of latitude bands $\pm 10^{\circ}$ and $\pm 15^{\circ}$ to changes in global mean precipitation with resolution.
-The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc..., which are computed through remapping all data to the $ne20$ grid.}
+The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc..., and computed through remapping all data to the $ne20$ grid.}
  \centering
@@
 
-\begin{figure}[t]
+To understand how subsiding motion and dilute CAPE vary with resolution in the deep Tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$, the mass-weighted vertical integral of $\omega$, is positive or negative, indicating predominantly subsiding or ascending grid columns.
+The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
+Figure~\ref{fig:cape}a shows the dilute CAPE values associated with mean conditions for ascending, descending and all grid columns in the deep Tropics, with resolution.
+The ascending regions are associated with larger values of CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the CAPE value computed for mean conditions over the entire deep Tropics decrease monotonically with resolution, consistent with the reduction in convective activity with resolution.
+The space-time weights associated with ascending and descending grid columns in the deep Tropics vary drastically with resolution (Figure~\ref{fig:cape}b).
+The subsiding (ascending) space-time weights vary from $0.32$ ($0.68$) at $ne20$, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in $ne120$.
+The increasing occurrence of stable, subsiding grid columns with resolution results in a reduction in CAPE for the entire deep Tropics, verified by the similar CAPE values produced through taking the weighted sum of the ascending/descending CAPE values (grey crosses in Figure~\ref{fig:cape}a).
+
+\begin{figure}
 \begin{center}
-\noindent\includegraphics[width=20pc,angle=0]{figs/temp_cape.pdf}\\
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_cape-ne20.pdf}\\
 \end{center}
@@
 
-To understand the degree to which atmospheric stability in the deep tropics is dependent on ascending and descending motion, temperature and moisture profiles are conditionally sampled based on whether $\langle \omega \rangle$ is positive or negative, indicating predominantly subsiding or ascending grid columns.
-The mean temperature and moisture profiles of subsiding and ascending regions in the deep tropics are used to compute the dilute-CAPE used by the ZM scheme, offline.
-Figure~\ref{fig:cape}a indicates that the mean profile of ascending regions are associated with large values of CAPE ($>180$ J/kg), and low values in subsiding regions ($<110$ J/kg).
-Figure~\ref{fig:profiles} shows the climatological temperature and specific humidity profiles of subsiding grid cells, averaged over the deep tropics and presented as anomalies from the mean temperature and specific humidity of the entire deep Tropics.
+Figure~\ref{fig:profiles} shows the climatological temperature and specific humidity profiles of subsiding grid cells in the deep Tropics, presented as anomalies from the mean temperature and specific humidity of the entire deep Tropics.
 The mean profiles of subsiding regions have an anomalous warming layer in the $600-800$ hPa layer and an anomalous moisture deficit throughout the entire column.
-This warming of the environment in subsiding regions opposes the generation of CAPE through equation~\ref{eq:cape}.
+To test the hypothesis, one needs to establish the relative importance of drying and warming of the environment on dilute CAPE.
+While warming the environment clearly opposes the growth of dilute CAPE (equation~\ref{eq:cape}), the impact of drying the environment on dilute CAPE is less clear.
+One may speculate that their are competing influences; increasing dilute CAPE through reducing $T_{ve}$ and opposing dilute CAPE through mixing drier environmental air into a rising air parcel, reducing the moisture available to condense and warming the parcel through latent heating.
+An experiment is then carried out, in which dilute CAPE is computed using the moisture profile derived from the entire of deep Tropics and temperature profile associated with descending regions, and vice versa.
 
-\begin{figure}[t]
+\begin{figure}
 \begin{center}
@@
 
-Figure~\ref{fig:cape}b shows that fractional area of air columns in the deep tropics that are predominantly subsiding (ascending) changes drastically with resolution, from $0.32$ ($0.68$) in the $ne20$ ($\Delta x = 166.8$ km) run, and monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in the $ne120$ ($\Delta x = 27.8$ km) run.
-Interestingly, the sum of the product of the fractional areas with their corresponding CAPE values gives approximately the same CAPE values computed from the mean temperature and moisture profiles over the entire deep tropics (Figure~\ref{fig:cape}a).
-This provides strong evidence that CAPE values in the deep tropics are decreasing with resolution because a larger (smaller) space-time fraction of the deep tropics are made up of predominantly subsiding (ascending) grid columns with increasing resolution.
+\subsection{Convection Biases Independent of Resolution}
+
+Poleward of the deep tropics and within the subtropics, the logistic regression indicates that subsidence is a poor predictor of $FREQZM$ (Figure~\ref{fig:4zonal}a,b).
+The R-value decreases to a local minimum between $15^{\circ} - 20^{\circ}$ latitude, and the magnitude of the sensitivity parameter steeply declines.
+The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4zonal}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
+The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d).
+The predominance of drizzle in this region is probably a result of the large subsiding motion in the subtropics (not shown) constraining CAPE from becoming too large.
+AGCMs are known to suffer from an excess drizzle bias in precisely this region \citep{D2006JCLIM}, and this analysis indicates that this bias is due to the use of a CAPE trigger function.
 
@@
 
-\subsection{Biases Unrelated to Resolution}
-
-Poleward of the deep tropics and within the subtropics, the logistic regression indicates that subsidence is a poor predictor of $FREQZM$ (Figure~\ref{fig:4zonal}a,b).
-The R-value decreases to a local minimum between $15^{\circ} - 20^{\circ}$ latitude, and the magnitude of the sensitivity parameter steeply declines.
-The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4zonal}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
-The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d).
-The predominance of drizzle in this region is probably a result of the large subsiding motion in the subtropics (not shown) constraining CAPE from becoming too large.
-AGCMs are known to suffer from an excess drizzle bias in precisely this region \citep{D2006JCLIM}, and this analysis indicates that this bias is due to the use of a CAPE trigger function.
-
-\ack This class file was developed by Sunrise Setting Ltd,
-Paignton, Devon, UK. Website:\\
-\href{http://www.sunrise-setting.co.uk}{\texttt{www.sunrise-setting.co.uk}}
+\ack 
+Funding support for this work was in part provided by the U.S. Department of Energy Office of Science (DE-SC0019459) and the National Science Foundation (AGS1648629).
 

### 2019-08-16 resolution.tex
edits

@@
 To generalize the relationship between vertical velocity and resolution, let $\alpha$ refer to the ratio of $W_0$, the vertical velocity scale of some reference grid spacing $\Delta x_0$, to $W$, the vertical velocity scale of any $\Delta x$.
-A power law for $\alpha$ in $\Delta x$ is then,
+A power law for $\alpha^{-1}$ in $\Delta x$ is then,
 \begin{equation}
-\alpha = \frac{W_0}{W} = \left( \frac{\Delta x_0}{\Delta x} \right)^n, \label{eq:alpha}
+\alpha^{-1} = \frac{W}{W_0} = \left( \frac{\Delta x}{\Delta x_0} \right)^n, \label{eq:alpha}
 \end{equation}
@@
 Strictly speaking, $\Delta x$ here refers to the distance between two points for which the velocity increment is computed in the structure function, but with this distance set to the model grid-spacing.
-Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum, which is supported by observations of mesoscale flow \citep{NG1985JAS}.
+Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum, and supported by observations of mesoscale flow \citep{NG1985JAS}.
 \cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope being common in both observations and models provides an emergent constraint for $b=\frac{1}{3}$ and $n= -\frac{2}{3}$.
@@
 The large increase in magnitude of the upward and downward vertical velocities with resolution may be expected to impact the behavior of other model components.
-The \cite{ZM1995AO} deep convection scheme (referred to as the $ZM$ scheme hereafter) is modulated by the dilute CAPE calculation, a modified form of the traditional CAPE, which is shown to be sensitive to the vertical velocity field in an analysis of a large eddy simulation \citep{SZ2018JCLIM}.
-
-The relationship between vertical velocities and CAPE falls out of the CAPE equation itself (F.
-Song, personal communication).
-The CAPE budget can be separated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through radiative cooling and advection of dry static energy and moisture by the environment, i.e., the resolved flow.
-The latter term is defined as,
-\begin{equation}
--R_d \int_{p_t}^{p_b} \frac{\partial T_{ve}}{\partial t} dlnp \label{eq:cape}
-\end{equation}
-where $T_{ve}$ is the virtual temperature of the environment, $R_d$ the gas constant for dry air and subscripts $b$ and $t$ refer to the parcel launch level (typically in the boundary layer) and the level of neutral buoyancy, respectively \citep{Z2002JGR}.
-Equation~\ref{eq:cape} shows that warming the environment reduces CAPE though reducing the buoyancy of lifted air parcels.
-The $T_{ve}$ budget contains a term; the vertical advection of potential energy, which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
-Song, personal communication).
-In subsiding regions, $w$ is negative and opposes the generation of CAPE through adiabatic warming of the environment.
+Curiously, there is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the $ZM$ scheme hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
+At any given grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.
+Time mean $FREQZM$ is therefore the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg.
+In addition to the six simulations used in the convergence experiment, an ensemble of 24 additional simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are used, increasing confidence in the relationship between subsidence and convective activity.
 
@@
 
-There is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the ZM scheme in the simulations, global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
-At any given grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.
-Time-mean $FREQZM$ therefore indicates the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg.
-It is hypothesized that the greater subsiding motion with resolution opposes the generation of dilute CAPE through subsidence warming, reducing the frequency the ZM scheme is triggered, and creating the strong negative correlation in Figure~\ref{fig:corr}.
-
-\subsection{Hypothesis Testing}
+To further understand this relationship, a logistic regression between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ is performed for each grid column within each of the simulations.
+Logistic regression uses an iterative method to fit a continuous variable predictor, $x$ to a binary predictand $p$ through the exponential \citep{WILKSBOOK},
+\begin{equation}
+p = \frac{exp{[b_0 + b_1 x]}}{1 + exp{[b_0 + b_1 x]}}, \label{eq:eq6-3}
+\end{equation}
+where $b_0$ and $b_1$ are the shape parameters of the exponential.
+The predictor is then the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
+The assumption is then that subsidence is the independent variable, which the authors believe is reasonable given subsiding regions are generally more stable, and so dilute CAPE is likely smaller compared with ascending regions.
+Grid column regressions that are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK} are retained for analysis.
+Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and parameter $b_1$ (hereafter referred to as the sensitivity parameter; Figure~\ref{fig:4zonal}a,b).
 
@@
 
-To test the hypothesis, a logistic regression between subsiding motion and $FREQZM$ is performed for each grid column within each of the simulations.
-Logistic regression uses an iterative method to fit a continuous variable predictor, $x$ to a binary predictand $p$ through the exponential \citep{WILKSBOOK},
-\begin{equation}
-p = \frac{exp{[b_0 + b_1 x]}}{1 + exp{[b_0 + b_1 x]}}, \label{eq:eq6-3}
-\end{equation}
-where $b_0$ and $b_1$ are the shape parameters of the exponential.
-The predictor is then the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
-Grid column regressions that are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK} are retained for analysis.
-Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and parameter $b_1$ (hereafter referred to as the sensitivity parameter; Figure~\ref{fig:4zonal}a,b).
-
-The zonal mean R-values indicate the greatest goodness of fit in the $\pm 10^{\circ}$ latitude band, referred to as the deep Tropics hereafter.
-Figure~\ref{fig:4zonal}c shows the climatological, zonal-mean latent heat flux in the simulations, which is expected to contribute positively to CAPE through the component associated with the thermodynamic state of boundary layer parcels.
-In the deep Tropics, latent heat fluxes are small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), which is consistent with our hypothesis that the subsiding motion actively depresses CAPE and the activity of the ZM scheme in the simulations.
-The sensitivity parameter becomes less negative in the deep Tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary of whether the ZM scheme is active.
+The zonal mean R-values indicate the greatest goodness of fit in the $\pm 10^{\circ}$ latitude band, hereafter referred to as the deep Tropics.
+Figure~\ref{fig:4zonal}c shows the climatological, zonal-mean latent heat flux in the simulations, which is expected to contribute positively to the component of dilute CAPE associated with the thermodynamic state of boundary layer parcels \citep{Z2002JGR}.
+In the deep Tropics, latent heat fluxes are small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), consistent with the idea that subsiding motion actively depresses dilute CAPE and the activity of the ZM scheme in the simulations.
+The sensitivity parameter becomes less negative in the deep Tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary $FREQZM$.
 The R-values generally decrease with resolution indicating that there is degradation in the relationship with resolution.
 
-Table~\ref{tbl:table2} shows that fractional contribution of the deep Tropics to the climatological, global mean change in convective precipitation with resolution.
+Table~\ref{tbl:table2} shows the fractional contribution of the deep Tropics to the climatological, global mean change in convective precipitation with resolution.
+This is a reflection of changes in the partitioning of the Intertropical Convergence Zone (ITCZ) between convective and stratiform precipitation with resolution.
 The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep Tropics (except in going from $ne20$ to $ne30$, where deep Tropical convective precipitation rates increase with resolution due to an extremely wide double-ITCZ in the $ne20$ simulation that spans well outside of $\pm 10^{\circ}$ latitude).
-Expanding the latitudes of consideration marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne20$).
-Taken together, the region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme, providing further evidence in support of the hypothesis.
+Expanding the latitudes of consideration marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne30-ne20$).
+Taken together, the region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.
 
@@
 
-To understand how subsiding motion and dilute CAPE vary with resolution in the deep Tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$, the mass-weighted vertical integral of $\omega$, is positive or negative, indicating predominantly subsiding or ascending grid columns.
+To estimate the dilute CAPE values associated with subsiding motion in the deep Tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$, the mass-weighted vertical integral of $\omega$, is positive or negative, indicating predominantly subsiding or ascending grid columns.
 The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
@@
 \begin{center}
-\noindent\includegraphics[width=20pc,angle=0]{figs/temp_cape-ne20.pdf}\\
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_cape.pdf}\\
 \end{center}
@@
 The mean profiles of subsiding regions have an anomalous warming layer in the $600-800$ hPa layer and an anomalous moisture deficit throughout the entire column.
-To test the hypothesis, one needs to establish the relative importance of drying and warming of the environment on dilute CAPE.
-While warming the environment clearly opposes the growth of dilute CAPE (equation~\ref{eq:cape}), the impact of drying the environment on dilute CAPE is less clear.
-One may speculate that their are competing influences; increasing dilute CAPE through reducing $T_{ve}$ and opposing dilute CAPE through mixing drier environmental air into a rising air parcel, reducing the moisture available to condense and warming the parcel through latent heating.
-An experiment is then carried out, in which dilute CAPE is computed using the moisture profile derived from the entire of deep Tropics and temperature profile associated with descending regions, and vice versa.
+This warming and drying patterns is consistent with the effects of subsidence, whose motion results in adiabatic warming and the advection of drier, upper-troposphere air downward.
+Both warming and drying the environment oppose the growth of dilute CAPE through reducing parcel buoyancy; warming the environment relative to the parcel temperature reduces parcel buoyancy \citep{Z2002JGR}, while mixing drier environmental air into a rising air parcel reduces the moisture available to warm the parcel through latent heating \citep{RB1992JAS}.
 
@@
 \end{figure}
+
+An experiment is carried out to unravel the contributions of warming or drying to dilute CAPE values of subsiding grid columns.
+Dilute CAPE is re-computed for descending/ascending regions, but using the mean specific humidity profile derived from the entire deep Tropics.
+Then the opposite is performed, computing dilute CAPE using the ascending/descending regions specific humidity but with the temperature profile from the entire deep Tropics.
+
+%The dilute CAPE calculation is a modified form of the traditional CAPE, which is shown to be sensitive to the vertical velocity field in an analysis of a large eddy simulation \citep{SZ2018JCLIM}.
+A relationship between vertical velocities and CAPE falls out of the CAPE equation itself (F.
+Song, personal communication).
+The CAPE budget can be separated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through radiative cooling and advection of dry static energy and moisture by the environment, i.e., the resolved flow.
+The latter term is defined as,
+%\begin{equation}
+%-R_d \int_{p_t}^{p_b} \frac{\partial T_{ve}}{\partial t} dlnp \label{eq:cape}
+%\end{equation}
+%where $T_{ve}$ is the virtual temperature of the environment, $R_d$ the gas constant for dry air and subscripts $b$ and $t$ refer to the parcel launch level (typically in the boundary layer) and the level of neutral buoyancy, respectively \citep{Z2002JGR}.
+Equation~\ref{eq:cape} shows that warming the environment reduces CAPE though reducing the buoyancy of lifted air parcels.
+The $T_{ve}$ budget contains a term; the vertical advection of potential energy, which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air.
+In subsiding regions, $w$ is negative and opposes the generation of CAPE through adiabatic warming of the environment.
 

### 2019-08-16 resolution.tex
edits

@@
 The large increase in magnitude of the upward and downward vertical velocities with resolution may be expected to impact the behavior of other model components.
-Curiously, there is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the $ZM$ scheme hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
+Curiously, there is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
 At any given grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.

### 2019-08-16 resolution.tex
edits

--- 2019-08-16 resolution.tex
@@
 Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
+In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for year in order to increase confidence in the assessment of resolution sensitivity in this study.
 All analyses exclude the first month of the simulations, and are computed on their native grids unless otherwise stated.
@@
 \caption{Components of the climatological, global mean vertical pressure velocity, (a) $\langle f_{u} \rangle \langle \omega_{u} \rangle$ and (b) $\langle f_{d} \rangle \langle \omega_{d} \rangle$.
-Grey crosses are for the perturbed parameter ensemble runs.}
+Grey crosses are for the 24 member perturbed parameter ensemble.}
 \label{fig:2panel}
@@
 The large increase in magnitude of the upward and downward vertical velocities with resolution may be expected to impact the behavior of other model components.
-Curiously, there is an excellent negative correlation (Pearson's R-value = 0.99) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the $ZM$ scheme hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
+Curiously, there is an excellent negative correlation (Pearson's R-value = 0.99, N = 27) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
 At any given grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.
 Time mean $FREQZM$ is therefore the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg.
-In addition to the six simulations used in the convergence experiment, an ensemble of 24 additional simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are used, increasing confidence in the relationship between subsidence and convective activity.
+The relationship indicates that model simulations with greater subsidence also have less convective activity.
 
@@
 \end{center}
-\caption{Scatter plot of global mean, climatological $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and FREQZM, and the fitted linear regression which has a Pearson's R-value = 0.99.
-Grey crosses are for the perturbed parameter ensemble runs.}
+\caption{Scatter plot of global mean, climatological $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and FREQZM, and the fitted linear regression which has a Pearson's R-value = 0.99, using all 27 simulations.
+Grey crosses are for the 24 member perturbed parameter ensemble runs.}
 \label{fig:corr}
@@
 The predictor is then the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
-The assumption is then that subsidence is the independent variable, which the authors believe is reasonable given subsiding regions are generally more stable, and so dilute CAPE is likely smaller compared with ascending regions.
+The assumption is then that subsidence is the independent variable, which the authors believe is reasonable since the environment of subsiding regions is generally stable, and the ZM scheme is modulated by a stability calculation, dilute CAPE.
 Grid column regressions that are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK} are retained for analysis.
@@
 \end{center}
-\caption{Zonal mean (a) R-values and (b) sensitivity parameter in the logistic regression, (d) time mean surface latent heat fluxes and (c) drizzle fraction. Colors are as in Figure~\ref{fig:2pdf}.}
+\caption{Zonal mean (a) R-values and (b) sensitivity parameter in the logistic regression, (c) climatological surface latent heat fluxes and (d) drizzle fraction.}
 \label{fig:4zonal}
@@
 The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep Tropics (except in going from $ne20$ to $ne30$, where deep Tropical convective precipitation rates increase with resolution due to an extremely wide double-ITCZ in the $ne20$ simulation that spans well outside of $\pm 10^{\circ}$ latitude).
-Expanding the latitudes of consideration marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne30-ne20$).
+Expanding the latitudes boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne30-ne20$).
 Taken together, the region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.
@@
 The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
+
 Figure~\ref{fig:cape}a shows the dilute CAPE values associated with mean conditions for ascending, descending and all grid columns in the deep Tropics, with resolution.
-The ascending regions are associated with larger values of CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the CAPE value computed for mean conditions over the entire deep Tropics decrease monotonically with resolution, consistent with the reduction in convective activity with resolution.
+The ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE value computed for mean conditions over the entire deep Tropics decrease monotonically with resolution, consistent with the reduction in convective activity with resolution.
 The space-time weights associated with ascending and descending grid columns in the deep Tropics vary drastically with resolution (Figure~\ref{fig:cape}b).
-The subsiding (ascending) space-time weights vary from $0.32$ ($0.68$) at $ne20$, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in $ne120$.
-The increasing occurrence of stable, subsiding grid columns with resolution results in a reduction in CAPE for the entire deep Tropics, verified by the similar CAPE values produced through taking the weighted sum of the ascending/descending CAPE values (grey crosses in Figure~\ref{fig:cape}a).
+The subsiding (ascending) space-time weights change from $0.32$ ($0.68$) at $ne20$, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in $ne120$.
+The increasing occurrence of stable, subsiding grid columns with resolution results in a reduction in CAPE for the entire deep Tropics, verified by the similar CAPE values derived through taking the weighted sum of the ascending/descending CAPE values (grey crosses in Figure~\ref{fig:cape}a).
 
@@
 \end{center}
-\caption{(a) time mean fraction of the deep tropics in the simulations with upward $\langle \omega \rangle$ (red) and downward $\langle \omega \rangle$ (blue), (b) CAPE computed from the mean temperature and moisture profiles of upward regions and downward regions.
-Black is for CAPE computed from the mean temperature and moisture profiles for the entire deep tropics, grey is the approximate discussed in the text.}
+\caption{(a) Dilute CAPE computed from time mean temperature and moisture profiles of ascending (red), subsiding (blue) and all grid columns (black) in the deep Tropics ($\pm 10^{\circ}$ latitude), and (b) space-time weights of ascending (red) and descending (blue) grid columns in the deep Tropics.
+(c) Dilute CAPE computed for ascending/descending grid columns, but using the mean temperature profile for the entire deep Tropics, (d) Same as (c) but using the mean moisture profile for the entire deep Tropics.
+Grey crosses in (a) are dilute CAPE derived from the sum of the products of space-time weights with the dilute CAPE values of ascending/descending grid columns.}
 \label{fig:cape}
@@
 
-Figure~\ref{fig:profiles} shows the climatological temperature and specific humidity profiles of subsiding grid cells in the deep Tropics, presented as anomalies from the mean temperature and specific humidity of the entire deep Tropics.
+Figure~\ref{fig:profiles} shows the time mean temperature and specific humidity profiles of subsiding grid cells in the deep Tropics, expressed as anomalies from the mean temperature and specific humidity of the entire deep Tropics.
 The mean profiles of subsiding regions have an anomalous warming layer in the $600-800$ hPa layer and an anomalous moisture deficit throughout the entire column.
-This warming and drying patterns is consistent with the effects of subsidence, whose motion results in adiabatic warming and the advection of drier, upper-troposphere air downward.
-Both warming and drying the environment oppose the growth of dilute CAPE through reducing parcel buoyancy; warming the environment relative to the parcel temperature reduces parcel buoyancy \citep{Z2002JGR}, while mixing drier environmental air into a rising air parcel reduces the moisture available to warm the parcel through latent heating \citep{RB1992JAS}.
+This warming and drying patterns is consistent with the effects of subsidence, whose motion produces adiabatic warming and advection of drier, upper-troposphere air downward.
+Both warming and drying the environment oppose the growth of dilute CAPE through reducing parcel buoyancy; warming the environment relative to the parcel temperature reduces parcel buoyancy \citep{Z2002JGR}, and mixing drier environmental air into a rising air parcel reduces the moisture available to warm the parcel through latent heating \citep{RB1992JAS}.
 
@@
 \end{center}
-\caption{}
+\caption{Time mean (a) temperature and (b) specific humidity profiles of subsiding grid cells in the deep Tropics ($\pm 10^{\circ}$ latitude) in the convergence experiment, presented as anomalies from the mean temperature and specific humidity of the entire deep Tropics in each simulation.}
 \label{fig:profiles}
@@
 
-An experiment is carried out to unravel the contributions of warming or drying to dilute CAPE values of subsiding grid columns.
-Dilute CAPE is re-computed for descending/ascending regions, but using the mean specific humidity profile derived from the entire deep Tropics.
-Then the opposite is performed, computing dilute CAPE using the ascending/descending regions specific humidity but with the temperature profile from the entire deep Tropics.
+An experiment is carried out to unravel the contributions of warming and drying to dilute CAPE values of subsiding grid columns.
+Through computing dilute CAPE using the mean specific humidity of subsiding regions in the deep Tropics, but setting the temperature profile to the mean profile for the entire deep Tropics, one can isolate the role of changes in moisture in subsiding regions on the dilute CAPE calculation, while doing the reverse isolates the influence of changes in temperature.
+Figure~\ref{fig:profiles}c shows the influence of moisture of subsiding/ascending regions on dilute CAPE, and Figure~\ref{fig:profiles}d the influence of temperature.
+The first order cause for the reduction of dilute CAPE is clearly do to the reduction (increase) in moisture in subsiding (ascending) region, while temperature has a smaller, second order influence on the dilute CAPE changes.
+This analysis indicates that the drying of the deep Tropics with resolution manifests through an increase in occurrence of subsiding motion with resolution, which reduces the frequency the ZM scheme is active through reducing the frequency the dilute CAPE threshold for convection occurs.
 
@@
 
-\subsection{Convection Biases Independent of Resolution}
+\subsection{Biases Unaffected by Resolution}
 
@@
 \caption{}
-\label{fig:profiles}
+\label{fig:prdecomp}
 \end{figure}

### 2019-08-16 resolution.tex
edits

@@
 
-\subsection{Biases Unaffected by Resolution}
+\subsection{Vertical Velocities and the Drizzle Bias}
 
-Poleward of the deep tropics and within the subtropics, the logistic regression indicates that subsidence is a poor predictor of $FREQZM$ (Figure~\ref{fig:4zonal}a,b).
+AGCMs are known to suffer from producing too much light rain, or drizzle, relative to observations, predominantly in the subtropics \citep{D2006JCLIM}.
+Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation that is less $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% of ZM precipitation in the region between $10^{\circ}-20^{\circ}$ latitude at all resolutions.
+In this region, there are opposing influences on dilute CAPE, a global maximum in latent heat fluxes (Figure~\ref{fig:4zonal}c) which increases dilute CAPE from below, and abundant subsidence, opposing dilute CAPE from above.
+
+%Poleward of the deep tropics and within the subtropics, the logistic regression indicates that subsidence is a poor predictor of $FREQZM$ (Figure~\ref{fig:4zonal}a,b).
 The R-value decreases to a local minimum between $15^{\circ} - 20^{\circ}$ latitude, and the magnitude of the sensitivity parameter steeply declines.

### 2019-08-16 resolution.tex
edits

@@
 
-The zonal mean R-values indicate the greatest goodness of fit in the $\pm 10^{\circ}$ latitude band, hereafter referred to as the deep Tropics.
+The zonal mean R-values indicate the greatest goodness of fit in the $\pm 10^{\circ}$ latitude band, hereafter referred to as the deep tropics.
 Figure~\ref{fig:4zonal}c shows the climatological, zonal-mean latent heat flux in the simulations, which is expected to contribute positively to the component of dilute CAPE associated with the thermodynamic state of boundary layer parcels \citep{Z2002JGR}.
-In the deep Tropics, latent heat fluxes are small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), consistent with the idea that subsiding motion actively depresses dilute CAPE and the activity of the ZM scheme in the simulations.
-The sensitivity parameter becomes less negative in the deep Tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary $FREQZM$.
+In the deep tropics, latent heat fluxes are small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), consistent with the idea that subsiding motion actively depresses dilute CAPE and the activity of the ZM scheme in the simulations.
+The sensitivity parameter becomes less negative in the deep tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary $FREQZM$.
 The R-values generally decrease with resolution indicating that there is degradation in the relationship with resolution.
 
-Table~\ref{tbl:table2} shows the fractional contribution of the deep Tropics to the climatological, global mean change in convective precipitation with resolution.
+Table~\ref{tbl:table2} shows the fractional contribution of the deep tropics to the climatological, global mean change in convective precipitation with resolution.
 This is a reflection of changes in the partitioning of the Intertropical Convergence Zone (ITCZ) between convective and stratiform precipitation with resolution.
-The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep Tropics (except in going from $ne20$ to $ne30$, where deep Tropical convective precipitation rates increase with resolution due to an extremely wide double-ITCZ in the $ne20$ simulation that spans well outside of $\pm 10^{\circ}$ latitude).
+The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where deep Tropical convective precipitation rates increase with resolution due to an extremely wide double-ITCZ in the $ne20$ simulation that spans well outside of $\pm 10^{\circ}$ latitude).
 Expanding the latitudes boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne30-ne20$).
@@
 
-To estimate the dilute CAPE values associated with subsiding motion in the deep Tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$, the mass-weighted vertical integral of $\omega$, is positive or negative, indicating predominantly subsiding or ascending grid columns.
+To estimate the dilute CAPE values associated with subsiding motion in the deep tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$, the mass-weighted vertical integral of $\omega$, is positive or negative, indicating predominantly subsiding or ascending grid columns.
 The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
 
-Figure~\ref{fig:cape}a shows the dilute CAPE values associated with mean conditions for ascending, descending and all grid columns in the deep Tropics, with resolution.
-The ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE value computed for mean conditions over the entire deep Tropics decrease monotonically with resolution, consistent with the reduction in convective activity with resolution.
-The space-time weights associated with ascending and descending grid columns in the deep Tropics vary drastically with resolution (Figure~\ref{fig:cape}b).
+Figure~\ref{fig:cape}a shows the dilute CAPE values associated with mean conditions for ascending, descending and all grid columns in the deep tropics, with resolution.
+The ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE value computed for mean conditions over the entire deep tropics decrease monotonically with resolution, consistent with the reduction in convective activity with resolution.
+The space-time weights associated with ascending and descending grid columns in the deep tropics vary drastically with resolution (Figure~\ref{fig:cape}b).
 The subsiding (ascending) space-time weights change from $0.32$ ($0.68$) at $ne20$, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in $ne120$.
-The increasing occurrence of stable, subsiding grid columns with resolution results in a reduction in CAPE for the entire deep Tropics, verified by the similar CAPE values derived through taking the weighted sum of the ascending/descending CAPE values (grey crosses in Figure~\ref{fig:cape}a).
+The increasing occurrence of stable, subsiding grid columns with resolution results in a reduction in CAPE for the entire deep tropics, verified by the similar CAPE values derived through taking the weighted sum of the ascending/descending CAPE values (grey crosses in Figure~\ref{fig:cape}a).
 
@@
 \end{center}
-\caption{(a) Dilute CAPE computed from time mean temperature and moisture profiles of ascending (red), subsiding (blue) and all grid columns (black) in the deep Tropics ($\pm 10^{\circ}$ latitude), and (b) space-time weights of ascending (red) and descending (blue) grid columns in the deep Tropics.
-(c) Dilute CAPE computed for ascending/descending grid columns, but using the mean temperature profile for the entire deep Tropics, (d) Same as (c) but using the mean moisture profile for the entire deep Tropics.
+\caption{(a) Dilute CAPE computed from time mean temperature and moisture profiles of ascending (red), subsiding (blue) and all grid columns (black) in the deep tropics ($\pm 10^{\circ}$ latitude), and (b) space-time weights of ascending (red) and descending (blue) grid columns in the deep tropics.
+(c) Dilute CAPE computed for ascending/descending grid columns, but using the mean temperature profile for the entire deep tropics, (d) Same as (c) but using the mean moisture profile for the entire deep tropics.
 Grey crosses in (a) are dilute CAPE derived from the sum of the products of space-time weights with the dilute CAPE values of ascending/descending grid columns.}
@@
 
-Figure~\ref{fig:profiles} shows the time mean temperature and specific humidity profiles of subsiding grid cells in the deep Tropics, expressed as anomalies from the mean temperature and specific humidity of the entire deep Tropics.
+Figure~\ref{fig:profiles} shows the time mean temperature and specific humidity profiles of subsiding grid cells in the deep tropics, expressed as anomalies from the mean temperature and specific humidity of the entire deep tropics.
 The mean profiles of subsiding regions have an anomalous warming layer in the $600-800$ hPa layer and an anomalous moisture deficit throughout the entire column.
@@
 \end{center}
-\caption{Time mean (a) temperature and (b) specific humidity profiles of subsiding grid cells in the deep Tropics ($\pm 10^{\circ}$ latitude) in the convergence experiment, presented as anomalies from the mean temperature and specific humidity of the entire deep Tropics in each simulation.}
+\caption{Time mean (a) temperature and (b) specific humidity profiles of subsiding grid cells in the deep tropics ($\pm 10^{\circ}$ latitude) in the convergence experiment, presented as anomalies from the mean temperature and specific humidity of the entire deep tropics in each simulation.}
 \label{fig:profiles}
@@
 An experiment is carried out to unravel the contributions of warming and drying to dilute CAPE values of subsiding grid columns.
-Through computing dilute CAPE using the mean specific humidity of subsiding regions in the deep Tropics, but setting the temperature profile to the mean profile for the entire deep Tropics, one can isolate the role of changes in moisture in subsiding regions on the dilute CAPE calculation, while doing the reverse isolates the influence of changes in temperature.
+Through computing dilute CAPE using the mean specific humidity of subsiding regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics, one can isolate the role of changes in moisture in subsiding regions on the dilute CAPE calculation, while doing the reverse isolates the influence of changes in temperature.
 Figure~\ref{fig:profiles}c shows the influence of moisture of subsiding/ascending regions on dilute CAPE, and Figure~\ref{fig:profiles}d the influence of temperature.
 The first order cause for the reduction of dilute CAPE is clearly do to the reduction (increase) in moisture in subsiding (ascending) region, while temperature has a smaller, second order influence on the dilute CAPE changes.
-This analysis indicates that the drying of the deep Tropics with resolution manifests through an increase in occurrence of subsiding motion with resolution, which reduces the frequency the ZM scheme is active through reducing the frequency the dilute CAPE threshold for convection occurs.
-
-%The dilute CAPE calculation is a modified form of the traditional CAPE, which is shown to be sensitive to the vertical velocity field in an analysis of a large eddy simulation \citep{SZ2018JCLIM}.
-A relationship between vertical velocities and CAPE falls out of the CAPE equation itself (F.
-Song, personal communication).
-The CAPE budget can be separated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through radiative cooling and advection of dry static energy and moisture by the environment, i.e., the resolved flow.
-The latter term is defined as,
-%\begin{equation}
-%-R_d \int_{p_t}^{p_b} \frac{\partial T_{ve}}{\partial t} dlnp \label{eq:cape}
-%\end{equation}
-%where $T_{ve}$ is the virtual temperature of the environment, $R_d$ the gas constant for dry air and subscripts $b$ and $t$ refer to the parcel launch level (typically in the boundary layer) and the level of neutral buoyancy, respectively \citep{Z2002JGR}.
-Equation~\ref{eq:cape} shows that warming the environment reduces CAPE though reducing the buoyancy of lifted air parcels.
-The $T_{ve}$ budget contains a term; the vertical advection of potential energy, which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air.
-In subsiding regions, $w$ is negative and opposes the generation of CAPE through adiabatic warming of the environment.
+This analysis indicates that the drying of the deep tropics with resolution manifests through an increase in occurrence of subsiding motion with resolution, which reduces the frequency the ZM scheme is active through reducing the frequency the dilute CAPE threshold for convection occurs.
 

### 2019-08-17 resolution.tex
edits

@@
 The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where deep Tropical convective precipitation rates increase with resolution due to an extremely wide double-ITCZ in the $ne20$ simulation that spans well outside of $\pm 10^{\circ}$ latitude).
-Expanding the latitudes boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne30-ne20$).
+Expanding the latitude boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne30-ne20$; Table~\ref{tbl:table2}).
 Taken together, the region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.

### 2019-08-17 resolution.tex
edits

@@
 
-\begin{figure}[t]
+In contrast to the impact of vertical motion on the ZM scheme, stratiform precipitation rates are more intuitively connected to vertical velocities.
+\cite{RETAL2016CD} proposed that total precipitation rates in models ($R$, sum of convective and stratiform precipitation rates) are determined by the upward moisture flux through cloud base,
+\begin{equation}
+R \approx -\frac{1}{g\rho_{w}} \omega^+ q^+ \label{eq:mflux}
+\end{equation}
+where $\omega^+$ and $q^+$ are the upward vertical pressure velocity and specific humidity at cloud base, respectively, and $\rho_w$ is the density of rainwater.
+
+Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to be a good approximation to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics \citep{OETAL2016JAMES}.
+Figure~\ref{fig:mflux} shows the moisture flux at the $850$ hPa level conditionally sampled in $1$ mm/day bins of precipitation rate, and separated into the components of total precipitation, stratiform and convective precipitation, in the CAM6 $ne30$ simulation.
+The figure shows that the increase in precipitation rates with moisture flux is due primarily to the stratiform precipitation rates.
+The increase in stratiform precipitation rates with resolution is likely due to the increase in magnitude of $\omega$ with resolution.
+
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_mflux.pdf}\\
+\end{center}
+\caption{Precipitation rates vs.
+upward moisture flux at the 850 hPa level.
+Solid lines refer to median moisture flux conditioned on the corresponding stratiform (green), convective (blue) and total (magenta) precipitation rates, and shaded regions refer to the conditional interquartile ranges.}
+\label{fig:mflux}
+\end{figure}
+
+To unravel the contributions of $\omega$ and $q$ at the 850 hPa level, $\omega_{850}$ and $q_{850}$ to the increase in climatological stratiform precipitation with resolution, the area averaged, climatological stratiform precipitation rate $\overline{R_{s}}$ is decomposed into a time mean magnitude term $M$ and space-time frequency term $f$,
+\begin{equation}
+\overline{R_{s}} = \sum_i \sum_j f \left( \omega_i , q_j \right) M \left( \omega_i , q_j \right) \label{eq:pdecomp}
+\end{equation}
+after \cite{TETAL2018CD}, and the subscript $850$ is dropped from $\omega$ and $q$ for brevity.
+Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution, and so Figure~\ref{fig:pdecomp} is a plot of the terms $M\left( \omega_i , q_j \right)$ and $f\left( \omega_i , q_j \right) \times M\left( \omega_i , q_j \right)$ for all resolutions, and in the $\pm 15^{\circ}$ latitude region.
+The plots are computed using 6-hourly instantaneous output of $\omega_{i}$ and $q_{j}$, with 0.05 Pa/s and 0.4 g/kg bins.
+Values in  $\left( \omega_{i} , q_{j} \right)$ space are only shown for bins with $f \geq1 \times 10^{-5}$, a reasonable cut-off to a bins' contribution to $\overline{R_s}$.
+Larger $\omega$ values correspond to larger magnitude stratiform precipitation rates, which have larger space-time frequencies $f\left( \omega_i , q_j \right)$ at higher resolutions, explaining the increase in $\overline{R_{s}}$ with resolution in Table~\ref{tbl:table2}.
+Changes in the space-time frequency of the $q$ field contributes comparatively less to changes in stratiform precipitation.
+
+\begin{figure}
 \begin{center}
@@
 \end{center}
-\caption{}
-\label{fig:prdecomp}
+\caption{Decomposition of the climatological stratiform precipitation rates, averaged over the $\pm 15^{\circ}$ latitude band into $\omega_{850}$ and $q_{850}$ environmental conditions.
+Left column shows the time mean magnitude term $M\left( \omega_i , q_j \right)$ and the right column is the magnitude term multiplied by the space-time frequency term $f\left( \omega_i , q_j \right) \times M\left( \omega_i , q_j \right)$.
+Integrals over $f \times M$ gives the climatological, area averaged stratiform precipitation rate.
+Panel labels denote the grid resolution of the model run.}
+\label{fig:pdecomp}
 \end{figure}

### 2019-08-17 resolution.tex
edits

@@
 after \cite{TETAL2018CD}, and the subscript $850$ is dropped from $\omega$ and $q$ for brevity.
-Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution, and so Figure~\ref{fig:pdecomp} is a plot of the terms $M\left( \omega_i , q_j \right)$ and $f\left( \omega_i , q_j \right) \times M\left( \omega_i , q_j \right)$ for all resolutions, and in the $\pm 15^{\circ}$ latitude region.
+
+Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution.
+Figure~\ref{fig:pdecomp} is a plot of the terms $M\left( \omega_i , q_j \right)$ and $f\left( \omega_i , q_j \right) \times M\left( \omega_i , q_j \right)$ for all resolutions, and in the $\pm 15^{\circ}$ latitude region.
 The plots are computed using 6-hourly instantaneous output of $\omega_{i}$ and $q_{j}$, with 0.05 Pa/s and 0.4 g/kg bins.

### 2019-08-17 resolution.tex
edits

@@
 
-\subsection{Vertical Velocities and Convective Precipitation}
+\subsection{Vertical Velocities and Deep Convective Precipitation}
 
@@
 Table~\ref{tbl:table2} shows the fractional contribution of the deep tropics to the climatological, global mean change in convective precipitation with resolution.
-This is a reflection of changes in the partitioning of the Intertropical Convergence Zone (ITCZ) between convective and stratiform precipitation with resolution.
-The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where deep Tropical convective precipitation rates increase with resolution due to an extremely wide double-ITCZ in the $ne20$ simulation that spans well outside of $\pm 10^{\circ}$ latitude).
+This is a reflection of changes in the partitioning of the Intertropical Convergence Zone (ITCZ) from convective to stratiform precipitation with resolution.
+The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where deep tropical convective precipitation rates increase with resolution due to an extremely wide double-ITCZ in the $ne20$ simulation that spans well outside of $\pm 10^{\circ}$ latitude).
 Expanding the latitude boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne30-ne20$; Table~\ref{tbl:table2}).
@@
 Figure~\ref{fig:cape}a shows the dilute CAPE values associated with mean conditions for ascending, descending and all grid columns in the deep tropics, with resolution.
-The ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE value computed for mean conditions over the entire deep tropics decrease monotonically with resolution, consistent with the reduction in convective activity with resolution.
+The ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE value computed for mean conditions over the entire deep tropics decreases monotonically with resolution, consistent with the reduction in convective activity with resolution.
 The space-time weights associated with ascending and descending grid columns in the deep tropics vary drastically with resolution (Figure~\ref{fig:cape}b).
@@
 
-\subsection{Vertical Velocities and the Drizzle Bias}
+\subsection{Vertical Velocities and Shallow Convective Precipitation}
 
-AGCMs are known to suffer from producing too much light rain, or drizzle, relative to observations, predominantly in the subtropics \citep{D2006JCLIM}.
-Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation that is less $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% of ZM precipitation in the region between $10^{\circ}-20^{\circ}$ latitude at all resolutions.
-In this region, there are opposing influences on dilute CAPE, a global maximum in latent heat fluxes (Figure~\ref{fig:4zonal}c) which increases dilute CAPE from below, and abundant subsidence, opposing dilute CAPE from above.
+AGCMs are known to suffer from producing too much light rain, or drizzle, relative to observations \citep{D2006JCLIM}.
+Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation that is less $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the region between $10^{\circ}-20^{\circ}$ latitude at all resolutions.
+Figure~\ref{fig:transect} shows a snapshot of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, and overlain by an isoline delineating where the ZM mass fluxes are fairly active.
+The ZM mass fluxes rarely extend above the $800$ hPa level, and often occurs with abundant subsiding motion aloft.
+In this subtropical region, there are opposing influences on dilute CAPE, a global maximum in latent heat fluxes (Figure~\ref{fig:4zonal}c) which increases dilute CAPE from below, and subsidence, opposing dilute CAPE from above.
+It seems plausible that these two opposing influences on stability translates into small values of dilute CAPE, but still large enough to trigger the ZM scheme.
+The ZM scheme, properly a deep convection scheme, is operating as a shallow convection scheme in the subtropics and produces drizzle.
 
-%Poleward of the deep tropics and within the subtropics, the logistic regression indicates that subsidence is a poor predictor of $FREQZM$ (Figure~\ref{fig:4zonal}a,b).
-The R-value decreases to a local minimum between $15^{\circ} - 20^{\circ}$ latitude, and the magnitude of the sensitivity parameter steeply declines.
-The local minimum in the R-value corresponds with a local maximum in the latent heat fluxes (Figure~\ref{fig:4zonal}c), indicating that the boundary layer is being driven unstable by large surface latent heat fluxes.
-The CAPE values in the $10^{\circ} - 15^{\circ}$ latitude region are likely to be small, since the ZM precipitation rate consists primarily of drizzle (Figure~\ref{fig:4zonal}d).
-The predominance of drizzle in this region is probably a result of the large subsiding motion in the subtropics (not shown) constraining CAPE from becoming too large.
-AGCMs are known to suffer from an excess drizzle bias in precisely this region \citep{D2006JCLIM}, and this analysis indicates that this bias is due to the use of a CAPE trigger function.
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_transect.png}\\
+\end{center}
+\caption{}
+\label{fig:transect}
+\end{figure}
+
+Figure~\ref{fig:cape-subt} shows the dilute CAPE values computed from mean temperature and moisture of subsiding and ascending regions in the $\pm \left( 10^{\circ}-30^{\circ} \right)$ latitude bands, hereafter referred to as the subtropics.
+The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), the dilute CAPE computed from the mean temperature and moisture over the entire subtropics varies much less with resolution compared with the deep tropics ($\sim 20$ J/kg across al resolutions, compared with $\sim 80$ J/kg in the deep tropics).
+Unlike the deep tropics, the reduction in dilute CAPE with resolution is not from an increase in occurrence of subsiding grid columns, since the space-time weights of ascending/descending motion are generally invariant with resolution (Figure~\ref{fig:cape-subt}b).
+The decline in dilute CAPE with resolution in the subtropics is a result of a greater magnitude of downward motion, but this sensitivity is not sufficient to suppress the occurrence of drizzling in the subtropics.
+
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_cape-subtropics.pdf}\\
+\end{center}
+\caption{(a) Dilute CAPE computed from time mean temperature and moisture profiles of ascending (red), subsiding (blue) and all grid columns (black) in the subtropics ($\pm \left(10^{\circ}-30^{\circ} \right)$ latitude band), and (b) space-time weights of ascending (red) and descending (blue) grid columns in the subtropics.
+Grey crosses in (a) are dilute CAPE derived from the sum of the products of space-time weights with the dilute CAPE values of ascending/descending grid columns.}
+\label{fig:cape-subt}
+\end{figure}
 
@@
 Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to be a good approximation to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics \citep{OETAL2016JAMES}.
-Figure~\ref{fig:mflux} shows the moisture flux at the $850$ hPa level conditionally sampled in $1$ mm/day bins of precipitation rate, and separated into the components of total precipitation, stratiform and convective precipitation, in the CAM6 $ne30$ simulation.
-The figure shows that the increase in precipitation rates with moisture flux is due primarily to the stratiform precipitation rates.
-The increase in stratiform precipitation rates with resolution is likely due to the increase in magnitude of $\omega$ with resolution.
+Figure~\ref{fig:mflux} shows the median convective, stratiform and total precipitation rates conditioned on the $850$ hPa moisture flux using $1$ mm/day bins, and in the CAM6 $ne30$ simulation.
+The figure shows that the moisture flux is a good approximation to the median total precipitation rate, and that this increase in precipitation rates with moisture flux is primarily due to stratiform precipitation rates.
 
@@
 upward moisture flux at the 850 hPa level.
-Solid lines refer to median moisture flux conditioned on the corresponding stratiform (green), convective (blue) and total (magenta) precipitation rates, and shaded regions refer to the conditional interquartile ranges.}
+Solid lines refer to median stratiform (green), convective (blue) and total (magenta) precipitation rates conditional on bins of the moisture flux, and shaded regions refer to the conditional interquartile ranges.}
 \label{fig:mflux}
@@
 
+The increase in stratiform precipitation rates with resolution is likely a due to the increase in magnitude of $\omega$ with resolution.
 To unravel the contributions of $\omega$ and $q$ at the 850 hPa level, $\omega_{850}$ and $q_{850}$ to the increase in climatological stratiform precipitation with resolution, the area averaged, climatological stratiform precipitation rate $\overline{R_{s}}$ is decomposed into a time mean magnitude term $M$ and space-time frequency term $f$,

### 2019-08-17 resolution.tex
edits

@@
 Figure~\ref{fig:profiles}c shows the influence of moisture of subsiding/ascending regions on dilute CAPE, and Figure~\ref{fig:profiles}d the influence of temperature.
-The first order cause for the reduction of dilute CAPE is clearly do to the reduction (increase) in moisture in subsiding (ascending) region, while temperature has a smaller, second order influence on the dilute CAPE changes.
+The first order cause for the reduction of dilute CAPE is clearly do to the decrease (increase) in moisture in subsiding (ascending) regions, while temperature has a smaller, second order influence on the dilute CAPE changes.
 This analysis indicates that the drying of the deep tropics with resolution manifests through an increase in occurrence of subsiding motion with resolution, which reduces the frequency the ZM scheme is active through reducing the frequency the dilute CAPE threshold for convection occurs.
@@
 AGCMs are known to suffer from producing too much light rain, or drizzle, relative to observations \citep{D2006JCLIM}.
-Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation that is less $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the region between $10^{\circ}-20^{\circ}$ latitude at all resolutions.
+Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation that is less $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the region between $10^{\circ}-20^{\circ}$ latitude irrespective of resolution.
 Figure~\ref{fig:transect} shows a snapshot of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, and overlain by an isoline delineating where the ZM mass fluxes are fairly active.
@@
 \end{equation}
-where $\omega^+$ and $q^+$ are the upward vertical pressure velocity and specific humidity at cloud base, respectively, and $\rho_w$ is the density of rainwater.
+where $\omega^+$ and $q^+$ are the upward $\omega$ and specific humidity at cloud base, respectively, and $\rho_w$ is the density of rainwater.
 
@@
 Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution.
-Figure~\ref{fig:pdecomp} is a plot of the terms $M\left( \omega_i , q_j \right)$ and $f\left( \omega_i , q_j \right) \times M\left( \omega_i , q_j \right)$ for all resolutions, and in the $\pm 15^{\circ}$ latitude region.
+And so Figure~\ref{fig:pdecomp} is a plot of the terms $M\left( \omega_i , q_j \right)$ and $f\left( \omega_i , q_j \right) \times M\left( \omega_i , q_j \right)$ for all resolutions, in the $\pm 15^{\circ}$ latitude region.
 The plots are computed using 6-hourly instantaneous output of $\omega_{i}$ and $q_{j}$, with 0.05 Pa/s and 0.4 g/kg bins.
 Values in  $\left( \omega_{i} , q_{j} \right)$ space are only shown for bins with $f \geq1 \times 10^{-5}$, a reasonable cut-off to a bins' contribution to $\overline{R_s}$.
-Larger $\omega$ values correspond to larger magnitude stratiform precipitation rates, which have larger space-time frequencies $f\left( \omega_i , q_j \right)$ at higher resolutions, explaining the increase in $\overline{R_{s}}$ with resolution in Table~\ref{tbl:table2}.
-Changes in the space-time frequency of the $q$ field contributes comparatively less to changes in stratiform precipitation.
+Larger $\omega$ values correspond to larger magnitude stratiform precipitation rates, of which have larger space-time frequencies $f\left( \omega_i , q_j \right)$ at higher resolutions, increasingly contributing to the increase in $\overline{R_{s}}$ with resolution.
+Changes in the space-time frequency of the $q$ field contributes comparatively much less to changes in stratiform precipitation, and so the increase in global mean stratiform precipitation with resolution is primarily a result of the increase in magnitude of $\omega$ with resolution.
 

### 2019-08-18 resolution.tex
edits

@@
 An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x \leq 5$ km) and climate ($\Delta x \leq 50$ km) simulations \citep{SMTMN2008JCP,MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
-These models are built using unstructured meshes that while allows for substantial grid flexibility, would require physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
+These models are built using unstructured meshes that allows for substantial grid flexibility, but this flexibility is limited by the need for physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
 The most common approach towards developing scale-aware physics is through the lens of limited area, large-eddy simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
-Through subsequently filtering large-eddy solutions to lower-resolution grids, a relationship between first- and higher-order moments \citep{G1992JFM} may be understood and ultimately parameterized as a function of grid resolution.
+By subsequently filtering large-eddy solutions to lower-resolution grids, a relationship between first- and higher-order moments \citep{G1992JFM} may be understood and ultimately parameterized as a function of grid resolution.
 While this approach is likely necessary for developing scale-aware physics, it is not sufficient.
-The equations of motions have inherent scale dependencies, with the properties of dynamical modes a function of native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
+The equations of motions have inherent scale dependencies, and the properties of dynamical modes are a function of native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
 Scale-aware physics should also recognize these native grid dependencies.
@@
 The sensitivity of the Community Atmosphere Model \citep[CAM;][]{CAM5}, and its predecessor, the Community Climate Model (CCM) to resolution ({\em{resolution}} refers to {\em{horizontal resolution}}, hereafter) is well documented through convergence studies \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}.
-{\color{red}{CAM/CCM is a well supported climate model,}} but despite thirty years of continual model development, there are robust sensitivities to resolution that have persisted in all versions of the model.
+CAM/CCM is a fully supported, well-funded climate model, but despite thirty years of continual model development, there are robust sensitivities to resolution that have persisted in all versions of the model.
 This study argues that a unifying cause, the inherent scale sensitivities of the underlying dynamical equations, can explain the robust responses to resolution that occur in CAM/CCM, {\color{red}{since it is difficult to conceive that inevitable responses to native grid resolution could be ignored in the pursuit of scale-aware physics.}}
@@
 While the cause of this sensitivity has been established for large-eddy simulations \citep[see][and references therein]{J2017JAMES}, only recently has the vertical velocity field in AGCMs and their sensitivity to resolution received attention \citep{DETALA2016ACP,OETAL2016JAMES}, albeit with conflicting explanations \citep{RETAL2016CD,HR2018JAMES}.
-To generalize the relationship between vertical velocity and resolution, let $\alpha$ refer to the ratio of $W_0$, the vertical velocity scale of some reference grid spacing $\Delta x_0$, to $W$, the vertical velocity scale of any $\Delta x$.
+To generalize the relationship between vertical velocity and resolution, let $\alpha$ refer to the ratio of $W_0$, the vertical velocity scale of some reference grid spacing $\Delta x_0$, to $W$, the vertical velocity scale for any $\Delta x$.
 A power law for $\alpha^{-1}$ in $\Delta x$ is then,
@@
 Strictly speaking, $\Delta x$ here refers to the distance between two points for which the velocity increment is computed in the structure function, but with this distance set to the model grid-spacing.
-Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which according to the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum, and supported by observations of mesoscale flow \citep{NG1985JAS}.
+Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which through the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum, and supported by observations of mesoscale flow \citep{NG1985JAS}.
 \cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope being common in both observations and models provides an emergent constraint for $b=\frac{1}{3}$ and $n= -\frac{2}{3}$.
@@
 Grid-scale precipitation is formed from grid-scale clouds, but is often referred to in the literature as stratiform precipitaiton and stratiform clouds, and used hereafter.
-The behavior of the difference precipitation routines with resolution is shown in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies.
-The propensity for precipitation rates to shift from the convection scheme to the stratiform scheme with resolution has been documented in other AGCMs as well \citep{PS2002CD,TETAL2018CD}, but none have given a satisfactory explanation for this sensitivity.
-The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the tendency to reduce $\Delta t_{phys}$ with resolution would by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convective precipitation rates with resolution.
+The resolution dependent partitioning between the two different precipitation routines is shown in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies.
+The tendency for precipitation rates to shift from the convection scheme to the stratiform scheme with resolution has been documented in other models \citep{PS2002CD,RETAL2016CD,TETAL2018CD}, but none have provided a satisfactory explanation for this sensitivity.
+The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the practice of reducing $\Delta t_{phys}$ with resolution should by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convective precipitation rates with resolution.
 
@@
 The reduction in convective precipitation rates with resolution in CAM6 is shown to result from the greater magnitude subsiding motion, creating a more stable atmosphere in which the criterion for parameterized convection occurs less often.
+The increase in stratiform precipitation rates with resolution is shown to result more directly from the increase in vertical velocities, by increasing moisture fluxes through the cloud base level.
 The feedback of the resolved vertical motion on the physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution.
-Section 2 describes CAM6 and details the convergence experiment.
+Section 2 describes CAM6 and the details of the convergence experiment.
 Section 3 contains a thorough analysis of the CAM6 simulations and Section 4 provides some discussion and conclusions.
@@
 The simulations use the CAM6 physics package.
-The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed probability distribution function (PDF) high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
+The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed filtered density function \citep{G1992JFM} high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
 The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with a three mode Modular Aerosol Model \citep{MAM}.
@@
 Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
-In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for year in order to increase confidence in the assessment of resolution sensitivity in this study.
+In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in the assessment of resolution sensitivity in this study.
 All analyses exclude the first month of the simulations, and are computed on their native grids unless otherwise stated.
 
-The horizontal hyper-viscosity operators $\nu$ vary with resolution after \cite{HETAL2019JAMES}, also provided in Table~\ref{tbl:table1}.
-The values of $\nu$ are a factor 2.5 greater for divergence damping and are not shown.
-$\Delta t_{phys}$ is chosen to scale with resolution, in proportion to the grid spacing,
-\begin{equation}
-\Delta t_{phys} = \Delta t_{phys,0} \times \frac{n_{e,0}}{n_e}~s,\label{eq:dt-scale}
-\end{equation}
-where $\Delta t_{phys,0}$ is taken to be the standard $1800$ s used in CAM-SE-CSLAM for the standard climate resolution, $n_{e,0} = 30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km).
-This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test \citep[Appendix A in][]{HETAL2019JAMES}, and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS}.
-The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.
-
- \section{Results}
-
  \begin{table*}
- \caption{Experimental design and global mean climatologies.}
+ \caption{Experimental design and global mean climatologies. $\Delta x$ refers to the average equatorial grid-spacing.}
  \centering
@@
 
-Table~\ref{tbl:table1} provides globally averaged, climatological metrics for the CAM6 simulations which are typically published in CAM/CCM convergence studies.
+The horizontal hyper-viscosity operators $\nu$ vary with resolution after \cite{HETAL2019JAMES}, also provided in Table~\ref{tbl:table1}.
+The values of $\nu$ are a factor 2.5 greater for divergence damping and are not shown.
+$\Delta t_{phys}$ is chosen to scale with resolution, in proportion to the grid spacing,
+\begin{equation}
+\Delta t_{phys} = \Delta t_{phys,0} \times \frac{n_{e,0}}{n_e}~s,\label{eq:dt-scale}
+\end{equation}
+where $\Delta t_{phys,0}$ is taken to be the standard $1800$ s used in CAM-SE-CSLAM for the standard climate resolution, $n_{e,0} = 30$ (equivalent to an average equatorial grid spacing $\Delta x = 111.2$km).
+This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test \citep[Appendix A in][]{HETAL2019JAMES}, and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS}.
+The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.
+
+ \section{Results}
+
+Table~\ref{tbl:table1} provides some globally averaged, climatological metrics for the CAM6 convergence experiment, commonly published in CAM/CCM convergence studies.
 Total precipitable water, total cloud fraction and deep convective precipitation rate decreases, while stratiform precipitation increases, monotonically with resolution (also shown in Figure~\ref{fig:cam-history}).
@@
 \caption{Probability density distribution of the upward vertical pressure velocities $\omega$ computed everywhere in the model from six-hourly output over the entirety of the year-long simulations.
-(a) Values on their native grid (solid) and values remapped to the $ne20$ grid (dotted), (b) values on their native grid, scaled to the $ne120$ resolution.}
+(a) Values on their native grid (solid) and values remapped to the $ne20$ grid (dotted), (b) values on their native grid, scaled to the $ne120$ resolution using a power law exponent $n=-1$ in equation~\ref{eq:alpha}.}
 \label{fig:2pdf}
@@
 
-The PDF of negative, or upward vertical pressure velocities $\omega$ in the aqua-planets is shown in Figure~\ref{fig:2pdf}a.
+The probability density function (PDF) of negative, or upward vertical pressure velocities $\omega$ in the aqua-planets is provided in Figure~\ref{fig:2pdf}a.
 The magnitude of upward $\omega$ increases monotonically with resolution, with positive, or downward $\omega$ behaving similarly (not shown).
@@
 The PDF's may be scaled to the highest-resolution resolution grid through $P(\omega)_s = \alpha P (\omega / \alpha)$, where $\alpha$ is the scale factor from equation~\ref{eq:alpha}, and setting $\Delta x_0$ to the $ne120$ grid-spacing.
-Figure~\ref{fig:2pdf}b shows the scaled PDF's using $n=-1$ in equation~\ref{eq:alpha}.
-The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law with $n=-1$ explains to first-order the variation in vertical velocity with resolution in the aqua-planet simulations.
+Figure~\ref{fig:2pdf}b shows the scaled PDF's for power law exponent $n=-1$ in equation~\ref{eq:alpha}.
+The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law exponent $n=-1$ explains to first-order the variation in vertical velocity with resolution in the aqua-planet simulations.
 
-Changes to the vertical velocity field can be further understood through decomposing the mass weighted vertical mean $\omega$ into upward and downward components,
+Changes to the vertical velocity field can be further understood through decomposing the mass weighted vertical mean $ \langle \omega \rangle$ into upward and downward components,
 \begin{equation}
@@
 The magnitude of both $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ and $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increase monotonically with resolution, and are equal and opposite, which is a requirement of mass conservation in the model and a convenient check of the calculation.
-While $\langle f_{d} \rangle$ is about 25\% larger than $\langle f_{u} \rangle$ in all simulations, the vertical mass fractions varies by only few percent with resolution, and so the monotonic behavior of $\langle f_{x} \rangle \, \langle \omega_{x} \rangle$ with resolution is primarily from variations in $ \langle \omega_{x} \rangle$ (not shown).
+While $\langle f_{d} \rangle$ is about 25\% larger than $\langle f_{u} \rangle$ in all simulations, the vertical mass fractions vary by only few percent with resolution, and so the monotonic behavior of $\langle f_{x} \rangle \, \langle \omega_{x} \rangle$ with resolution is primarily from variations in $ \langle \omega_{x} \rangle$ (not shown).
 
@@
 Curiously, there is an excellent negative correlation (Pearson's R-value = 0.99, N = 27) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
-At any given grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.
+At any grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.
 Time mean $FREQZM$ is therefore the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg.
-The relationship indicates that model simulations with greater subsidence also have less convective activity.
+The regression indicates that model simulations with greater subsidence also have less convective activity.
 
@@
 The predictor is then the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
-The assumption is then that subsidence is the independent variable, which the authors believe is reasonable since the environment of subsiding regions is generally stable, and the ZM scheme is modulated by a stability calculation, dilute CAPE.
+The assumption is then that subsidence is the independent variable, which the authors believe is reasonable since the environment of subsiding regions is generally more stable than its surroundings, and the ZM scheme is modulated by a stability calculation, dilute CAPE.
 Grid column regressions that are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK} are retained for analysis.
@@
 The zonal mean R-values indicate the greatest goodness of fit in the $\pm 10^{\circ}$ latitude band, hereafter referred to as the deep tropics.
-Figure~\ref{fig:4zonal}c shows the climatological, zonal-mean latent heat flux in the simulations, which is expected to contribute positively to the component of dilute CAPE associated with the thermodynamic state of boundary layer parcels \citep{Z2002JGR}.
-In the deep tropics, latent heat fluxes are small, and the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), consistent with the idea that subsiding motion actively depresses dilute CAPE and the activity of the ZM scheme in the simulations.
+In this region, the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), consistent with the idea that subsiding motion stabilizes the environment and actively depresses dilute CAPE and the activity of the ZM scheme in the simulations.
 The sensitivity parameter becomes less negative in the deep tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary $FREQZM$.
@@
 This is a reflection of changes in the partitioning of the Intertropical Convergence Zone (ITCZ) from convective to stratiform precipitation with resolution.
-The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where deep tropical convective precipitation rates increase with resolution due to an extremely wide double-ITCZ in the $ne20$ simulation that spans well outside of $\pm 10^{\circ}$ latitude).
-Expanding the latitude boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occur in this region (again, ignoring $ne30-ne20$; Table~\ref{tbl:table2}).
+The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where convective precipitation rates increase in the $ne30$ run due to an extremely wide double-ITCZ in the $ne20$ run that spans well outside of $\pm 10^{\circ}$ latitude).
+Expanding the latitude boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occurs in this region (again, ignoring $ne30-ne20$; Table~\ref{tbl:table2}).
 Taken together, the region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.
@@
 
-To estimate the dilute CAPE values associated with subsiding motion in the deep tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$, the mass-weighted vertical integral of $\omega$, is positive or negative, indicating predominantly subsiding or ascending grid columns.
+To estimate the dilute CAPE values associated with subsiding motion in the deep tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$ is positive or negative, indicating predominantly subsiding or ascending grid columns.
 The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
@@
 The subsiding (ascending) space-time weights change from $0.32$ ($0.68$) at $ne20$, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in $ne120$.
-The increasing occurrence of stable, subsiding grid columns with resolution results in a reduction in CAPE for the entire deep tropics, verified by the similar CAPE values derived through taking the weighted sum of the ascending/descending CAPE values (grey crosses in Figure~\ref{fig:cape}a).
+The increasing occurrence of stable, subsiding grid columns with resolution results in a reduction in dilute CAPE for the entire deep tropics, verified by the similar values derived through taking the weighted sum of the ascending/descending CAPE values (grey crosses in Figure~\ref{fig:cape}a).
 
@@
 The mean profiles of subsiding regions have an anomalous warming layer in the $600-800$ hPa layer and an anomalous moisture deficit throughout the entire column.
-This warming and drying patterns is consistent with the effects of subsidence, whose motion produces adiabatic warming and advection of drier, upper-troposphere air downward.
-Both warming and drying the environment oppose the growth of dilute CAPE through reducing parcel buoyancy; warming the environment relative to the parcel temperature reduces parcel buoyancy \citep{Z2002JGR}, and mixing drier environmental air into a rising air parcel reduces the moisture available to warm the parcel through latent heating \citep{RB1992JAS}.
+This warming and drying patterns is consistent with the effects of subsidence, whose motion adiabatically warms the environment while simultaneously advecting drier conditions aloft downward.
+Both warming and drying the environment oppose the growth of dilute CAPE through reducing parcel buoyancy; warming the environment relative to the temperature of rising air parcels reduces parcel buoyancy \citep{Z2002JGR}, and mixing drier environmental air into rising air parcels reduces the moisture available to warm parcels through latent heating \citep{RB1992JAS}.
 
@@
 An experiment is carried out to unravel the contributions of warming and drying to dilute CAPE values of subsiding grid columns.
-Through computing dilute CAPE using the mean specific humidity of subsiding regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics, one can isolate the role of changes in moisture in subsiding regions on the dilute CAPE calculation, while doing the reverse isolates the influence of changes in temperature.
-Figure~\ref{fig:profiles}c shows the influence of moisture of subsiding/ascending regions on dilute CAPE, and Figure~\ref{fig:profiles}d the influence of temperature.
-The first order cause for the reduction of dilute CAPE is clearly do to the decrease (increase) in moisture in subsiding (ascending) regions, while temperature has a smaller, second order influence on the dilute CAPE changes.
-This analysis indicates that the drying of the deep tropics with resolution manifests through an increase in occurrence of subsiding motion with resolution, which reduces the frequency the ZM scheme is active through reducing the frequency the dilute CAPE threshold for convection occurs.
+Through recomputing dilute CAPE using the mean specific humidity of subsiding regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics, one can isolate the role of changes in moisture in subsiding regions on the dilute CAPE calculation.
+Doing the reverse then isolates the role of temperature on dilute CAPE.
+Figure~\ref{fig:profiles}c shows the influence of moisture on the dilute CAPE of subsiding/ascending regions in the deep tropics, and Figure~\ref{fig:profiles}d the influence of temperature.
+To first order, the large spread in dilute CAPE between subsiding and ascending regions is due to the difference in moisture in subsiding and ascending regions, while temperature differences between the two regimes has a smaller, second order influence on the spread of dilute CAPE.
+
+This analysis provides the following reasons that subsidence and the activity of the ZM scheme are so tightly intertwined.
+The negative correlation between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ within the simulations, and determined by the logistic regression, is predominantly due to the effectiveness of drier conditions in depressing dilute CAPE below the threshold for convection.
+The strong negative correlation between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ across resolutions (Figure~\ref{fig:corr}) is primarily the result of the increase in occurrence of dry, subsiding regions in the deep tropics with resolution.
 
@@
 
-AGCMs are known to suffer from producing too much light rain, or drizzle, relative to observations \citep{D2006JCLIM}.
+AGCMs are known to suffer from a drizzle bias, producing too much light rain relative to observations \citep{D2006JCLIM}.
 Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation that is less $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the region between $10^{\circ}-20^{\circ}$ latitude irrespective of resolution.
-Figure~\ref{fig:transect} shows a snapshot of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, and overlain by an isoline delineating where the ZM mass fluxes are fairly active.
-The ZM mass fluxes rarely extend above the $800$ hPa level, and often occurs with abundant subsiding motion aloft.
-In this subtropical region, there are opposing influences on dilute CAPE, a global maximum in latent heat fluxes (Figure~\ref{fig:4zonal}c) which increases dilute CAPE from below, and subsidence, opposing dilute CAPE from above.
+Focusing in on this drizzle region, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
+The ZM mass fluxes typically only extend up to about the $800$ hPa level, which often occurs with abundant subsiding motion aloft.
+
+In drizzling regions, there are opposing influences on dilute CAPE, a global maximum in latent heat fluxes (Figure~\ref{fig:4zonal}c), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and subsidence, which opposes dilute CAPE from above.
 It seems plausible that these two opposing influences on stability translates into small values of dilute CAPE, but still large enough to trigger the ZM scheme.
-The ZM scheme, properly a deep convection scheme, is operating as a shallow convection scheme in the subtropics and produces drizzle.
+The ZM scheme, properly a deep convection scheme, is operating as a shallow convection scheme in the $10^{\circ}-20^{\circ}$ latitude region, and produces drizzle.
 
@@
 \begin{center}
-\noindent\includegraphics[width=20pc,angle=0]{figs/temp_transect.png}\\
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_trans.pdf}\\
 \end{center}
-\caption{}
+\caption{(a,b) Two snapshots of $\omega$ for a longitude-pressure transect at $\sim 18^{\circ}$ latitude in the $ne30$ simulation, overlain by the $0.0075$ kg/m$^2$/s contour of the ZM mass flux (white) delineating the region where the ZM scheme is active, and the 15 K/day contour of the total physics tendencies (black), indicating stratiform cloud formation.}
 \label{fig:transect}
@@
 Figure~\ref{fig:cape-subt} shows the dilute CAPE values computed from mean temperature and moisture of subsiding and ascending regions in the $\pm \left( 10^{\circ}-30^{\circ} \right)$ latitude bands, hereafter referred to as the subtropics.
-The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), the dilute CAPE computed from the mean temperature and moisture over the entire subtropics varies much less with resolution compared with the deep tropics ($\sim 20$ J/kg across al resolutions, compared with $\sim 80$ J/kg in the deep tropics).
-Unlike the deep tropics, the reduction in dilute CAPE with resolution is not from an increase in occurrence of subsiding grid columns, since the space-time weights of ascending/descending motion are generally invariant with resolution (Figure~\ref{fig:cape-subt}b).
-The decline in dilute CAPE with resolution in the subtropics is a result of a greater magnitude of downward motion, but this sensitivity is not sufficient to suppress the occurrence of drizzling in the subtropics.
+The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), and the dilute CAPE computed from the mean temperature and moisture over the entire subtropics varies much less with resolution compared with the deep tropics ($\sim 20$ J/kg across all resolutions, compared with $\sim 80$ J/kg across resolutions in the deep tropics).
+Also unlike the deep tropics, the reduction in dilute CAPE with resolution is not from an increase in occurrence of subsiding grid columns, since the space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
+The decline in dilute CAPE with resolution in the subtropics is a result of a greater magnitude of downward motion (not shown), but this sensitivity is not sufficient to suppress the occurrence of drizzling in the subtropics.
 
@@
 \end{center}
-\caption{(a) Dilute CAPE computed from time mean temperature and moisture profiles of ascending (red), subsiding (blue) and all grid columns (black) in the subtropics ($\pm \left(10^{\circ}-30^{\circ} \right)$ latitude band), and (b) space-time weights of ascending (red) and descending (blue) grid columns in the subtropics.
+\caption{(a) Dilute CAPE computed from time mean temperature and moisture profiles of ascending (red), subsiding (blue) and all grid columns (black) in the subtropics ($\pm \left(10^{\circ}-30^{\circ} \right)$ latitude bands), and (b) space-time weights of ascending (red) and descending (blue) grid columns in the subtropics.
 Grey crosses in (a) are dilute CAPE derived from the sum of the products of space-time weights with the dilute CAPE values of ascending/descending grid columns.}
@@
 
-In contrast to the impact of vertical motion on the ZM scheme, stratiform precipitation rates are more intuitively connected to vertical velocities.
+In contrast to the impact of vertical motion on the ZM scheme, the stratiform scheme is more intuitively connected to vertical velocities.
 \cite{RETAL2016CD} proposed that total precipitation rates in models ($R$, sum of convective and stratiform precipitation rates) are determined by the upward moisture flux through cloud base,
@@
 \end{equation}
-where $\omega^+$ and $q^+$ are the upward $\omega$ and specific humidity at cloud base, respectively, and $\rho_w$ is the density of rainwater.
+where $\omega^+$ and $q^+$ are the upward $\omega$ and specific humidity $q$ at cloud base, respectively, and $\rho_w$ is the density of rainwater.
 
 Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to be a good approximation to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics \citep{OETAL2016JAMES}.
-Figure~\ref{fig:mflux} shows the median convective, stratiform and total precipitation rates conditioned on the $850$ hPa moisture flux using $1$ mm/day bins, and in the CAM6 $ne30$ simulation.
-The figure shows that the moisture flux is a good approximation to the median total precipitation rate, and that this increase in precipitation rates with moisture flux is primarily due to stratiform precipitation rates.
+Figure~\ref{fig:mflux} shows the median convective, stratiform and total precipitation rates conditioned on the $850$ hPa moisture flux using $1$ mm/day bins, in the CAM6 $ne30$ simulation.
+The figure shows that the moisture flux is a good approximation to the median total precipitation rate, and that the increase in precipitation rates with moisture flux is primarily from the stratiform precipitation scheme.
 
@@
 
-The increase in stratiform precipitation rates with resolution is likely a due to the increase in magnitude of $\omega$ with resolution.
-To unravel the contributions of $\omega$ and $q$ at the 850 hPa level, $\omega_{850}$ and $q_{850}$ to the increase in climatological stratiform precipitation with resolution, the area averaged, climatological stratiform precipitation rate $\overline{R_{s}}$ is decomposed into a time mean magnitude term $M$ and space-time frequency term $f$,
+The increase in stratiform precipitation rates with resolution is likely due to the increase in magnitude of $\omega$ with resolution.
+To unravel the contributions of $\omega$ and $q$ at the 850 hPa level, $\omega_{850}$ and $q_{850}$ to the increase in climatological stratiform precipitation with resolution, the area averaged, climatological stratiform precipitation rate $\overline{R_{s}}$ can be decomposed into a double sum of the product of the time mean magnitude term $M_s$ and time mean spatial frequency term $f_s$, over $\omega_{850}$ and $q_{850}$ space,
 \begin{equation}
@@
 
-Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution.
-And so Figure~\ref{fig:pdecomp} is a plot of the terms $M\left( \omega_i , q_j \right)$ and $f\left( \omega_i , q_j \right) \times M\left( \omega_i , q_j \right)$ for all resolutions, in the $\pm 15^{\circ}$ latitude region.
+Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution, and so $\overline{R_s}$ is defined as the mean over the $\pm 15^{\circ}$ latitude region.
+Figure~\ref{fig:pdecomp} is a plot of the terms $M_s \left( \omega_i , q_j \right)$ and $f_s \left( \omega_i , q_j \right) M\left( \omega_i , q_j \right)$ for all resolutions, in the $\pm 15^{\circ}$ latitude region.
 The plots are computed using 6-hourly instantaneous output of $\omega_{i}$ and $q_{j}$, with 0.05 Pa/s and 0.4 g/kg bins.
-Values in  $\left( \omega_{i} , q_{j} \right)$ space are only shown for bins with $f \geq1 \times 10^{-5}$, a reasonable cut-off to a bins' contribution to $\overline{R_s}$.
-Larger $\omega$ values correspond to larger magnitude stratiform precipitation rates, of which have larger space-time frequencies $f\left( \omega_i , q_j \right)$ at higher resolutions, increasingly contributing to the increase in $\overline{R_{s}}$ with resolution.
-Changes in the space-time frequency of the $q$ field contributes comparatively much less to changes in stratiform precipitation, and so the increase in global mean stratiform precipitation with resolution is primarily a result of the increase in magnitude of $\omega$ with resolution.
+Values in  $\left( \omega_{i} , q_{j} \right)$ space are only shown for bins with $f_s \geq1 \times 10^{-5}$, a reasonable cut-off to a bins' contribution to $\overline{R_s}$.
+The $M_s$ plots show that larger $\omega$ values correspond to larger magnitude time-mean stratiform precipitation rates, and these larger precipitation rates are associated with larger spatial frequencies $f_s \left( \omega_i , q_j \right)$ at higher resolutions, increasingly contributing to the increase in $\overline{R_{s}}$ with resolution.
+Changes to the time mean spatial frequency of the $q$ field contributes comparatively much less to changes in stratiform precipitation, and so the increase in global mean stratiform precipitation with resolution is primarily a result of the increase in magnitude of $\omega$ with resolution.
 
@@
 
+\section{Conclusions}
+
+In the $\pm 10^{\circ}$ latitude band, the deep tropics, there is a reduction in the occurrence of ascending grid columns with resolution, which drives an increase in occurrence of subsiding regions.
+The reduction in occurrence of ascending motion in the deep tropics with resolution likely manifests as a reduction in the area of resolved updrafts with resolution, since the horizontal scale of large buoyancy stratiform clouds collapse to smallest resolvable feature in the model \citep{HR2018JAMES}, i.e., the models effective resolution.
+The PDF's of vertical velocities indicate that it is this resolution dependence of stratiform clouds that sets a power law relationship of the magnitude of the vertical velocities with resolution $\Delta x^{n}$, with $n=-1$.
+This scaling also indicates that for a doubling of the resolution, the same upward mass flux can be achieved in the half the area, and so it seems plausible that the reduction in the occurrence of ascending motion with resolution is also a result of the greater efficiency of the resolved mass fluxes with resolution.
+
 \ack 
-Funding support for this work was in part provided by the U.S. Department of Energy Office of Science (DE-SC0019459) and the National Science Foundation (AGS1648629).
+Funding support for this work was in part provided by the U.S.
+Department of Energy Office of Science (DE-SC0019459) and the National Science Foundation (AGS1648629).
+Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at the National Center for Atmospheric Research (NCAR).
+Herrington would like to thank Dr.
+Sultan Hameed for his assistance with the statistical analysis used in this study.
 

### 2019-08-18 resolution.tex
edits

@@
 The increase in stratiform precipitation rates with resolution is likely due to the increase in magnitude of $\omega$ with resolution.
-To unravel the contributions of $\omega$ and $q$ at the 850 hPa level, $\omega_{850}$ and $q_{850}$ to the increase in climatological stratiform precipitation with resolution, the area averaged, climatological stratiform precipitation rate $\overline{R_{s}}$ can be decomposed into a double sum of the product of the time mean magnitude term $M_s$ and time mean spatial frequency term $f_s$, over $\omega_{850}$ and $q_{850}$ space,
+To unravel the contributions of $\omega$ and $q$ at the 850 hPa level, $\omega_{850}$ and $q_{850}$ to the increase in climatological stratiform precipitation with resolution, the area averaged, climatological stratiform precipitation rate $\overline{R_{s}}$ can be decomposed into a double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
 \begin{equation}
-\overline{R_{s}} = \sum_i \sum_j f \left( \omega_i , q_j \right) M \left( \omega_i , q_j \right) \label{eq:pdecomp}
+\overline{R_{s}} = \sum_i \sum_j f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right) \label{eq:pdecomp}
 \end{equation}
@@
 Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution, and so $\overline{R_s}$ is defined as the mean over the $\pm 15^{\circ}$ latitude region.
-Figure~\ref{fig:pdecomp} is a plot of the terms $M_s \left( \omega_i , q_j \right)$ and $f_s \left( \omega_i , q_j \right) M\left( \omega_i , q_j \right)$ for all resolutions, in the $\pm 15^{\circ}$ latitude region.
+Figure~\ref{fig:pdecomp} is a plot of the terms $M_s \left( \omega_i , q_j \right)$ and $f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right)$ for all resolutions, in the $\pm 15^{\circ}$ latitude region.
 The plots are computed using 6-hourly instantaneous output of $\omega_{i}$ and $q_{j}$, with 0.05 Pa/s and 0.4 g/kg bins.
 Values in  $\left( \omega_{i} , q_{j} \right)$ space are only shown for bins with $f_s \geq1 \times 10^{-5}$, a reasonable cut-off to a bins' contribution to $\overline{R_s}$.
-The $M_s$ plots show that larger $\omega$ values correspond to larger magnitude time-mean stratiform precipitation rates, and these larger precipitation rates are associated with larger spatial frequencies $f_s \left( \omega_i , q_j \right)$ at higher resolutions, increasingly contributing to the increase in $\overline{R_{s}}$ with resolution.
-Changes to the time mean spatial frequency of the $q$ field contributes comparatively much less to changes in stratiform precipitation, and so the increase in global mean stratiform precipitation with resolution is primarily a result of the increase in magnitude of $\omega$ with resolution.
+The $M_s$ plots show that larger magnitude $\omega_{850}$ correspond to larger magnitude time-mean stratiform precipitation rates, and these larger precipitation rates are associated with larger spatial frequencies $f_s \left( \omega_i , q_j \right)$ at higher resolutions.
+These large magnitude stratiform precipitation rates increasingly contribute to the increase in $\overline{R_{s}}$ with resolution, as seen by the expansion of $f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right)$ to include larger magnitude $\omega_{850}$.
+Changes to the time mean spatial frequency of the $q_{850}$ field contributes comparatively much less to changes in stratiform precipitation.
 

### 2019-08-18 resolution.tex
edits

@@
 To further understand this relationship, a logistic regression between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ is performed for each grid column within each of the simulations.
-Logistic regression uses an iterative method to fit a continuous variable predictor, $x$ to a binary predictand $p$ through the exponential \citep{WILKSBOOK},
+Logistic regression uses an iterative method to fit a continuous variable predictor, $x$ to a binary predictand $p$ using the exponential \citep{WILKSBOOK},
 \begin{equation}
@@
 where $b_0$ and $b_1$ are the shape parameters of the exponential.
-The predictor is then the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
+The predictor is the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
 The assumption is then that subsidence is the independent variable, which the authors believe is reasonable since the environment of subsiding regions is generally more stable than its surroundings, and the ZM scheme is modulated by a stability calculation, dilute CAPE.
@@
 This is a reflection of changes in the partitioning of the Intertropical Convergence Zone (ITCZ) from convective to stratiform precipitation with resolution.
-The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where convective precipitation rates increase in the $ne30$ run due to an extremely wide double-ITCZ in the $ne20$ run that spans well outside of $\pm 10^{\circ}$ latitude).
+The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where convective precipitation rates increase due to a wide double-ITCZ in the $ne20$ run that spans well outside of $\pm 10^{\circ}$ latitude, but which is contained within $\pm 10^{\circ}$ latitude in the $ne30$ run).
 Expanding the latitude boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occurs in this region (again, ignoring $ne30-ne20$; Table~\ref{tbl:table2}).
@@
  \caption{Fractional contribution of latitude bands $\pm 10^{\circ}$ and $\pm 15^{\circ}$ to changes in global mean precipitation with resolution.
-The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc..., and computed through remapping all data to the $ne20$ grid.}
+The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc...
+All differences are computed after remapping the data to the $ne20$ grid.}
  \centering
@@
 The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
+Figure~\ref{fig:cape}a shows the dilute CAPE values associated with mean conditions for ascending, descending and all grid columns in the deep tropics, with resolution.
+Ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE value computed for mean conditions over the entire deep tropics decreases monotonically with resolution, consistent with the reduction in convective activity with resolution.
 
-Figure~\ref{fig:cape}a shows the dilute CAPE values associated with mean conditions for ascending, descending and all grid columns in the deep tropics, with resolution.
-The ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE value computed for mean conditions over the entire deep tropics decreases monotonically with resolution, consistent with the reduction in convective activity with resolution.
 The space-time weights associated with ascending and descending grid columns in the deep tropics vary drastically with resolution (Figure~\ref{fig:cape}b).
@@
 An experiment is carried out to unravel the contributions of warming and drying to dilute CAPE values of subsiding grid columns.
-Through recomputing dilute CAPE using the mean specific humidity of subsiding regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics, one can isolate the role of changes in moisture in subsiding regions on the dilute CAPE calculation.
-Doing the reverse then isolates the role of temperature on dilute CAPE.
-Figure~\ref{fig:profiles}c shows the influence of moisture on the dilute CAPE of subsiding/ascending regions in the deep tropics, and Figure~\ref{fig:profiles}d the influence of temperature.
+Through recomputing dilute CAPE using the mean specific humidity of subsiding regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics, one can isolate the role of moisture in subsiding regions on the dilute CAPE calculation.
+Doing the reverse then isolates the role of temperature in subsiding regions on dilute CAPE.
+
+Figure~\ref{fig:cape}c shows the influence of moisture on the dilute CAPE of subsiding/ascending regions in the deep tropics, and Figure~\ref{fig:cape}d the influence of temperature.
 To first order, the large spread in dilute CAPE between subsiding and ascending regions is due to the difference in moisture in subsiding and ascending regions, while temperature differences between the two regimes has a smaller, second order influence on the spread of dilute CAPE.
+A large spread in dilute CAPE between ascending/descending regions is important for maintaining a strong relationship between subsidence and the activity of the ZM scheme.
+If the model operated more similar to the fixed moisture experiment (Figure~\ref{fig:cape}d), then subsidence would be associated with larger values of dilute CAPE, and less effective at depressing dilute CAPE below the threshold for convection.
 
-This analysis provides the following reasons that subsidence and the activity of the ZM scheme are so tightly intertwined.
+The large spread in dilute CAPE of subsiding/ascending regions multiplied by the large change in space-time fraction of subsiding/ascending are responsible for about half of the change in the dilute CAPE computed from mean temperature and moisture profiles over the entire deep tropics.
+There is a systematic offset of dilute CAPE values in both subsiding/ascending regions with resolution (Figure~\ref{fig:cape}a) that accounts for another half of the resolution sensitivity.
+To understand whether moisture or temperature changes with resolution are responsible for this behavior, a similar experiment is performed in which dilute CAPE is recomputed at all resolutions, but fixing the moisture profiles to the lowest resolution $ne20$ profile, and then again through fixing the temperature profile to the $ne20 profile$.
+
+%This analysis indicates the following reasons that subsidence and the activity of the ZM scheme are so tightly intertwined.
 The negative correlation between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ within the simulations, and determined by the logistic regression, is predominantly due to the effectiveness of drier conditions in depressing dilute CAPE below the threshold for convection.
@@
 AGCMs are known to suffer from a drizzle bias, producing too much light rain relative to observations \citep{D2006JCLIM}.
-Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation that is less $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the region between $10^{\circ}-20^{\circ}$ latitude irrespective of resolution.
+Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the region between $10^{\circ}-20^{\circ}$ latitude irrespective of resolution.
 Focusing in on this drizzle region, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
@@
 
-In the $\pm 10^{\circ}$ latitude band, the deep tropics, there is a reduction in the occurrence of ascending grid columns with resolution, which drives an increase in occurrence of subsiding regions.
+%In the $\pm 10^{\circ}$ latitude band, the deep tropics, there is a reduction in the occurrence of ascending grid columns with resolution, which drives an increase in occurrence of subsiding regions.
 The reduction in occurrence of ascending motion in the deep tropics with resolution likely manifests as a reduction in the area of resolved updrafts with resolution, since the horizontal scale of large buoyancy stratiform clouds collapse to smallest resolvable feature in the model \citep{HR2018JAMES}, i.e., the models effective resolution.

### 2019-08-18 resolution.tex
edits

@@
 (c) Dilute CAPE computed for ascending/descending grid columns, but using the mean temperature profile for the entire deep tropics, (d) Same as (c) but using the mean moisture profile for the entire deep tropics.
+(e) Same as (c) but using the fixing the temperature to the $ne20$ profiles, and (f) same as (d) but fixing the moisture to the $ne20$ profiles.
 Grey crosses in (a) are dilute CAPE derived from the sum of the products of space-time weights with the dilute CAPE values of ascending/descending grid columns.}
@@
 
-The large spread in dilute CAPE of subsiding/ascending regions multiplied by the large change in space-time fraction of subsiding/ascending are responsible for about half of the change in the dilute CAPE computed from mean temperature and moisture profiles over the entire deep tropics.
+The large spread in dilute CAPE of subsiding/ascending regions multiplied by the large change in space-time fraction of subsiding/ascending are responsible for about half of the change in the dilute CAPE computed from the mean temperature and moisture profiles over the entire deep tropics.
 There is a systematic offset of dilute CAPE values in both subsiding/ascending regions with resolution (Figure~\ref{fig:cape}a) that accounts for another half of the resolution sensitivity.
 To understand whether moisture or temperature changes with resolution are responsible for this behavior, a similar experiment is performed in which dilute CAPE is recomputed at all resolutions, but fixing the moisture profiles to the lowest resolution $ne20$ profile, and then again through fixing the temperature profile to the $ne20 profile$.
-
-%This analysis indicates the following reasons that subsidence and the activity of the ZM scheme are so tightly intertwined.
-The negative correlation between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ within the simulations, and determined by the logistic regression, is predominantly due to the effectiveness of drier conditions in depressing dilute CAPE below the threshold for convection.
-The strong negative correlation between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ across resolutions (Figure~\ref{fig:corr}) is primarily the result of the increase in occurrence of dry, subsiding regions in the deep tropics with resolution.
+Figure~\ref{fig:cape}e,d shows the influence of the change of resolution on moisture, and the change due to temperature, and indicates that temperature changes with resolution are primarily responsible for the other half of the decrease in dilute CAPE in deep tropics with resolution.
 

### 2019-08-19 resolution.tex
edits

@@
 
-The large spread in dilute CAPE of subsiding/ascending regions multiplied by the large change in space-time fraction of subsiding/ascending are responsible for about half of the change in the dilute CAPE computed from the mean temperature and moisture profiles over the entire deep tropics.
-There is a systematic offset of dilute CAPE values in both subsiding/ascending regions with resolution (Figure~\ref{fig:cape}a) that accounts for another half of the resolution sensitivity.
-To understand whether moisture or temperature changes with resolution are responsible for this behavior, a similar experiment is performed in which dilute CAPE is recomputed at all resolutions, but fixing the moisture profiles to the lowest resolution $ne20$ profile, and then again through fixing the temperature profile to the $ne20 profile$.
-Figure~\ref{fig:cape}e,d shows the influence of the change of resolution on moisture, and the change due to temperature, and indicates that temperature changes with resolution are primarily responsible for the other half of the decrease in dilute CAPE in deep tropics with resolution.
+The large spread in dilute CAPE of subsiding/ascending regions combined with the large changes in space-time fraction of subsiding/ascending with resolution are responsible for about half of the change in the dilute CAPE computed for the entire deep tropics.
+There is also a systematic offset of dilute CAPE values in both subsiding/ascending regions with resolution (Figure~\ref{fig:cape}a) that accounts for another half of the resolution sensitivity.
+To understand whether moisture or temperature changes with resolution are responsible for this behavior, a similar experiment is performed in which dilute CAPE is recomputed at all resolutions, but fixing the moisture profiles to the lowest resolution $ne20$ profile, and then repeated through fixing only the temperature profile to $ne20$ values.
+Figure~\ref{fig:cape}e,d shows the influence of the change of resolution on moisture, and the change due to temperature, respectively.
+The calculation shows that temperature changes with resolution are primarily responsible for the other half of the decrease in dilute CAPE in the deep tropics with resolution.
 
@@
 Focusing in on this drizzle region, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
-The ZM mass fluxes typically only extend up to about the $800$ hPa level, which often occurs with abundant subsiding motion aloft.
+The ZM mass fluxes typically only extend up to about the $800$ hPa level, which often occurs with subsiding motion aloft.
 
 In drizzling regions, there are opposing influences on dilute CAPE, a global maximum in latent heat fluxes (Figure~\ref{fig:4zonal}c), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and subsidence, which opposes dilute CAPE from above.
-It seems plausible that these two opposing influences on stability translates into small values of dilute CAPE, but still large enough to trigger the ZM scheme.
-The ZM scheme, properly a deep convection scheme, is operating as a shallow convection scheme in the $10^{\circ}-20^{\circ}$ latitude region, and produces drizzle.
+It seems plausible that these two opposing influences on stability translates into small values of dilute CAPE, but for drizzle to occur, still large enough to trigger the ZM scheme.
+The ZM scheme, properly a deep convection scheme, is operating as a shallow convection scheme in the $10^{\circ}-20^{\circ}$ latitude region.
 
@@
 Also unlike the deep tropics, the reduction in dilute CAPE with resolution is not from an increase in occurrence of subsiding grid columns, since the space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
-The decline in dilute CAPE with resolution in the subtropics is a result of a greater magnitude of downward motion (not shown), but this sensitivity is not sufficient to suppress the occurrence of drizzling in the subtropics.
+The decline in dilute CAPE with resolution in the subtropics is a result of increased stability due to changes in environmental temperature from the larger magnitude downward motion (not shown), but this sensitivity is not sufficient to suppress the occurrence of drizzling in the subtropics.
 
@@
 Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution, and so $\overline{R_s}$ is defined as the mean over the $\pm 15^{\circ}$ latitude region.
-Figure~\ref{fig:pdecomp} is a plot of the terms $M_s \left( \omega_i , q_j \right)$ and $f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right)$ for all resolutions, in the $\pm 15^{\circ}$ latitude region.
+Figure~\ref{fig:pdecomp} is a plot of the terms $M_s \left( \omega_i , q_j \right)$ and $f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right)$ for all resolutions,.
 The plots are computed using 6-hourly instantaneous output of $\omega_{i}$ and $q_{j}$, with 0.05 Pa/s and 0.4 g/kg bins.

### 2019-08-19 resolution.tex
edits

@@
 Also unlike the deep tropics, the reduction in dilute CAPE with resolution is not from an increase in occurrence of subsiding grid columns, since the space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
-The decline in dilute CAPE with resolution in the subtropics is a result of increased stability due to changes in environmental temperature from the larger magnitude downward motion (not shown), but this sensitivity is not sufficient to suppress the occurrence of drizzling in the subtropics.
+The decline in dilute CAPE with resolution in the subtropics is a result of increased stability due to changes in environmental temperature from the larger magnitude downward motion (not shown), but this sensitivity is not sufficient oppose the instability generated by the global maximum in latent heat fluxes.
 

### 2019-08-19 resolution.tex
edits

--- 2019-08-19 resolution.tex
@@
 
+{\color{red}{The next three paragraphs need to be condensed into one or two. Working on it.}}
+
 An experiment is carried out to unravel the contributions of warming and drying to dilute CAPE values of subsiding grid columns.
@@
 Also unlike the deep tropics, the reduction in dilute CAPE with resolution is not from an increase in occurrence of subsiding grid columns, since the space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
-The decline in dilute CAPE with resolution in the subtropics is a result of increased stability due to changes in environmental temperature from the larger magnitude downward motion (not shown), but this sensitivity is not sufficient to suppress the occurrence of drizzling in the subtropics.
+The decline in dilute CAPE with resolution in the subtropics is a result of increased stability due to changes in environmental temperature from the larger magnitude downward motion (not shown).
+This increased stability with resolution is not sufficient to overcome the large latent heat fluxes and mitigate the shallow convection regime of the ZM scheme, consistent with the local minimum in the R-values from the logistic regression in drizzling regions (Figure~\ref{fig:4zonal}a).
 

### 2019-08-19 resolution.tex
edits

@@
 (c) Dilute CAPE computed for ascending/descending grid columns, but using the mean temperature profile for the entire deep tropics, (d) Same as (c) but using the mean moisture profile for the entire deep tropics.
-(e) Same as (c) but using the fixing the temperature to the $ne20$ profiles, and (f) same as (d) but fixing the moisture to the $ne20$ profiles.
+(e) Same as (c) but fixing temperature to the $ne20$ profiles, and (f) same as (d) but fixing moisture to the $ne20$ profile.
 Grey crosses in (a) are dilute CAPE derived from the sum of the products of space-time weights with the dilute CAPE values of ascending/descending grid columns.}
@@
 The $M_s$ plots show that larger magnitude $\omega_{850}$ correspond to larger magnitude time-mean stratiform precipitation rates, and these larger precipitation rates are associated with larger spatial frequencies $f_s \left( \omega_i , q_j \right)$ at higher resolutions.
-These large magnitude stratiform precipitation rates increasingly contribute to the increase in $\overline{R_{s}}$ with resolution, as seen by the expansion of $f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right)$ to include larger magnitude $\omega_{850}$.
+These large magnitude stratiform precipitation rates increasingly contribute to the increase in $\overline{R_{s}}$ with resolution, as seen by the increase in $f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right)$ in the larger magnitude $\omega_{850}$ space with resolution.
 Changes to the time mean spatial frequency of the $q_{850}$ field contributes comparatively much less to changes in stratiform precipitation.

### 2019-08-19 resolution.tex
edits

@@
 Figure~\ref{fig:cape}a shows the dilute CAPE values associated with mean conditions for ascending, descending and all grid columns in the deep tropics, with resolution.
-Ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE value computed for mean conditions over the entire deep tropics decreases monotonically with resolution, consistent with the reduction in convective activity with resolution.
+Ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE in both regimes decreases monotonically with resolution.
 
@@
 The subsiding (ascending) space-time weights change from $0.32$ ($0.68$) at $ne20$, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in $ne120$.
-The increasing occurrence of stable, subsiding grid columns with resolution results in a reduction in dilute CAPE for the entire deep tropics, verified by the similar values derived through taking the weighted sum of the ascending/descending CAPE values (grey crosses in Figure~\ref{fig:cape}a).
+It is the increasing occurrence of stable, subsiding grid columns with resolution in addition to the reduction in dilute CAPE for both ascending and descending grid columns with resolution (Figure~\ref{fig:cape}a), that produces a reduction in dilute CAPE for the entire deep tropics, verified by the similar values derived through taking the weighted sum of the ascending/descending dilute CAPE values (grey crosses in Figure~\ref{fig:cape}a).
 
@@
 \caption{(a) Dilute CAPE computed from time mean temperature and moisture profiles of ascending (red), subsiding (blue) and all grid columns (black) in the deep tropics ($\pm 10^{\circ}$ latitude), and (b) space-time weights of ascending (red) and descending (blue) grid columns in the deep tropics.
-(c) Dilute CAPE computed for ascending/descending grid columns, but using the mean temperature profile for the entire deep tropics, (d) Same as (c) but using the mean moisture profile for the entire deep tropics.
-(e) Same as (c) but fixing temperature to the $ne20$ profiles, and (f) same as (d) but fixing moisture to the $ne20$ profile.
+(c) Dilute CAPE computed for ascending/descending grid columns, but using the mean temperature profile for the entire deep tropics, and (d) Dilute CAPE for ascending/descending regions but fixing moisture to the $ne20$ profile.
 Grey crosses in (a) are dilute CAPE derived from the sum of the products of space-time weights with the dilute CAPE values of ascending/descending grid columns.}
@@
 The mean profiles of subsiding regions have an anomalous warming layer in the $600-800$ hPa layer and an anomalous moisture deficit throughout the entire column.
-This warming and drying patterns is consistent with the effects of subsidence, whose motion adiabatically warms the environment while simultaneously advecting drier conditions aloft downward.
+This warming and drying patterns is consistent with the effects of subsidence, whose motion adiabatically warms the environment while simultaneously advecting drier conditions aloft, downward.
 Both warming and drying the environment oppose the growth of dilute CAPE through reducing parcel buoyancy; warming the environment relative to the temperature of rising air parcels reduces parcel buoyancy \citep{Z2002JGR}, and mixing drier environmental air into rising air parcels reduces the moisture available to warm parcels through latent heating \citep{RB1992JAS}.
@@
 
-{\color{red}{The next three paragraphs need to be condensed into one or two. Working on it.}}
+An experiment is carried out to unravel the contributions of warming and drying to the large spread in dilute CAPE between ascending/descending grid columns in the deep tropics.
+Through recomputing dilute CAPE using the mean specific humidity of ascending/descending regions, but setting the temperature profile to the mean profile for the entire deep tropics, one can isolate the contribution of changing moisture to the spread in dilute CAPE in the two regimes, and for each resolution.
+Doing the reverse then isolates the role of changing temperature profiles.
+Figure~\ref{fig:cape}c shows the influence of changing moisture on the dilute CAPE of subsiding/ascending regions in the deep tropics.
+To first order, the large spread in dilute CAPE between subsiding and ascending regions is due to the difference in moisture in subsiding and ascending regions, while temperature differences between the two regimes has a smaller, second order influence (not shown).
+The large spread in dilute CAPE between ascending/descending regions is crucial for maintaining the relationship in the deep tropics illustrated by the logistic regression, since the much smaller values of dilute CAPE of subsiding grid columns is required to depress dilute CAPE below the threshold for convection.
 
-An experiment is carried out to unravel the contributions of warming and drying to dilute CAPE values of subsiding grid columns.
-Through recomputing dilute CAPE using the mean specific humidity of subsiding regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics, one can isolate the role of moisture in subsiding regions on the dilute CAPE calculation.
-Doing the reverse then isolates the role of temperature in subsiding regions on dilute CAPE.
-
-Figure~\ref{fig:cape}c shows the influence of moisture on the dilute CAPE of subsiding/ascending regions in the deep tropics, and Figure~\ref{fig:cape}d the influence of temperature.
-To first order, the large spread in dilute CAPE between subsiding and ascending regions is due to the difference in moisture in subsiding and ascending regions, while temperature differences between the two regimes has a smaller, second order influence on the spread of dilute CAPE.
-A large spread in dilute CAPE between ascending/descending regions is important for maintaining a strong relationship between subsidence and the activity of the ZM scheme.
-If the model operated more similar to the fixed moisture experiment (Figure~\ref{fig:cape}d), then subsidence would be associated with larger values of dilute CAPE, and less effective at depressing dilute CAPE below the threshold for convection.
-
-The large spread in dilute CAPE of subsiding/ascending regions combined with the large changes in space-time fraction of subsiding/ascending with resolution are responsible for about half of the change in the dilute CAPE computed for the entire deep tropics.
-There is also a systematic offset of dilute CAPE values in both subsiding/ascending regions with resolution (Figure~\ref{fig:cape}a) that accounts for another half of the resolution sensitivity.
-To understand whether moisture or temperature changes with resolution are responsible for this behavior, a similar experiment is performed in which dilute CAPE is recomputed at all resolutions, but fixing the moisture profiles to the lowest resolution $ne20$ profile, and then repeated through fixing only the temperature profile to $ne20$ values.
-Figure~\ref{fig:cape}e,d shows the influence of the change of resolution on moisture, and the change due to temperature, respectively.
-The calculation shows that temperature changes with resolution are primarily responsible for the other half of the decrease in dilute CAPE in the deep tropics with resolution.
+The increasing occurrence of subsiding grid columns accounts for about half of the changes in dilute CAPE in the deep tropics with resolution, the other half being due to the systematic reduction in dilute CAPE for both ascending and descending regions.
+To isolate whether temperature or moisture contributes to the latter contribution to the changes in dilute CAPE, a similar experiment is performed in which dilute CAPE is recomputed at all resolutions, but fixing the moisture profiles to the lowest resolution $ne20$ profile, and then repeated through fixing only the temperature profile to $ne20$ values.
+Figure~\ref{fig:cape}d shows the influence of the changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution is primarily from changes to the temperature field.
+Moisture changes with resolution has a smaller influence, only appreciably impacting the dilute CAPE of ascending regions at higher resolutions (not shown).
 

### 2019-08-19 resolution.tex
edits

@@
 The predictor is the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
-The assumption is then that subsidence is the independent variable, which the authors believe is reasonable since the environment of subsiding regions is generally more stable than its surroundings, and the ZM scheme is modulated by a stability calculation, dilute CAPE.
+The assumption is then that subsidence is the independent variable, which the authors believe is reasonable since the environment of subsiding regions is generally more stable than its surroundings, and the ZM scheme is modulated by the dilute CAPE stability metric.
 Grid column regressions that are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK} are retained for analysis.
@@
 
+\subsubsection{Deep Tropics}
+
 Table~\ref{tbl:table2} shows the fractional contribution of the deep tropics to the climatological, global mean change in convective precipitation with resolution.
-This is a reflection of changes in the partitioning of the Intertropical Convergence Zone (ITCZ) from convective to stratiform precipitation with resolution.
 The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where convective precipitation rates increase due to a wide double-ITCZ in the $ne20$ run that spans well outside of $\pm 10^{\circ}$ latitude, but which is contained within $\pm 10^{\circ}$ latitude in the $ne30$ run).
 Expanding the latitude boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occurs in this region (again, ignoring $ne30-ne20$; Table~\ref{tbl:table2}).
+This trend reflects changes in the partitioning of the Intertropical Convergence Zone (ITCZ) from convective to stratiform precipitation with resolution.
 Taken together, the region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.
@@
 The space-time weights associated with ascending and descending grid columns in the deep tropics vary drastically with resolution (Figure~\ref{fig:cape}b).
-The subsiding (ascending) space-time weights change from $0.32$ ($0.68$) at $ne20$, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in $ne120$.
+The subsiding (ascending) space-time weights change from $0.32$ ($0.68$) in the $ne20$ run, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in the $ne120$ run.
 It is the increasing occurrence of stable, subsiding grid columns with resolution in addition to the reduction in dilute CAPE for both ascending and descending grid columns with resolution (Figure~\ref{fig:cape}a), that produces a reduction in dilute CAPE for the entire deep tropics, verified by the similar values derived through taking the weighted sum of the ascending/descending dilute CAPE values (grey crosses in Figure~\ref{fig:cape}a).
@@
 The mean profiles of subsiding regions have an anomalous warming layer in the $600-800$ hPa layer and an anomalous moisture deficit throughout the entire column.
-This warming and drying patterns is consistent with the effects of subsidence, whose motion adiabatically warms the environment while simultaneously advecting drier conditions aloft, downward.
+This warming and drying pattern is consistent with the effects of subsidence, whose motion adiabatically warms the environment while simultaneously advecting drier conditions aloft, downward.
 Both warming and drying the environment oppose the growth of dilute CAPE through reducing parcel buoyancy; warming the environment relative to the temperature of rising air parcels reduces parcel buoyancy \citep{Z2002JGR}, and mixing drier environmental air into rising air parcels reduces the moisture available to warm parcels through latent heating \citep{RB1992JAS}.
@@
 
-An experiment is carried out to unravel the contributions of warming and drying to the large spread in dilute CAPE between ascending/descending grid columns in the deep tropics.
-Through recomputing dilute CAPE using the mean specific humidity of ascending/descending regions, but setting the temperature profile to the mean profile for the entire deep tropics, one can isolate the contribution of changing moisture to the spread in dilute CAPE in the two regimes, and for each resolution.
-Doing the reverse then isolates the role of changing temperature profiles.
-Figure~\ref{fig:cape}c shows the influence of changing moisture on the dilute CAPE of subsiding/ascending regions in the deep tropics.
-To first order, the large spread in dilute CAPE between subsiding and ascending regions is due to the difference in moisture in subsiding and ascending regions, while temperature differences between the two regimes has a smaller, second order influence (not shown).
-The large spread in dilute CAPE between ascending/descending regions is crucial for maintaining the relationship in the deep tropics illustrated by the logistic regression, since the much smaller values of dilute CAPE of subsiding grid columns is required to depress dilute CAPE below the threshold for convection.
+The large spread in dilute CAPE between ascending/descending regions is crucial for maintaining the relationship shown by the logistic regression, since the much smaller values of dilute CAPE of subsiding grid columns is required to depress dilute CAPE below the threshold for convection.
+To unravel the contributions of warming and drying to this large spread in dilute CAPE between the two regimes, dilute CAPE is recomputed using the mean specific humidity of ascending/descending regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics.
+Figure~\ref{fig:cape}c shows this influence of changing moisture on the dilute CAPE of subsiding/ascending regions in the deep tropics, which to first order, explains the large spread in dilute CAPE between the two regimes.
+Temperature differences between ascending/descending regimes has a smaller, second order influence on the spread in dilute CAPE (not shown).
 
@@
 
-\subsection{Vertical Velocities and Shallow Convective Precipitation}
+\subsubsection{Subtropics}
 
-AGCMs are known to suffer from a drizzle bias, producing too much light rain relative to observations \citep{D2006JCLIM}.
-Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the region between $10^{\circ}-20^{\circ}$ latitude irrespective of resolution.
-Focusing in on this drizzle region, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
+Figure~\ref{fig:cape-subt}a shows the dilute CAPE values computed from mean temperature and moisture of subsiding and ascending regions in the $\pm \left( 10^{\circ}-30^{\circ} \right) $ latitude bands, hereafter referred to as the subtropics.
+The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), and their dilute CAPE  values vary much less with resolution ($\sim 20$ J/kg, compared with $\sim 80$ J/kg across all resolutions in the deep tropics).
+The lack of spread in dilute CAPE between ascending/descending regions explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}a), since without a clear distinction between dilute CAPE of ascending/descending grid columns, subsidence cannot be effective at depressing dilute CAPE below the threshold for convection.
+There is also no significant changes to the occurrence of subsiding grid columns with resolution; the space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
+
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_cape-subtropics.pdf}\\
+\end{center}
+\caption{(a) Dilute CAPE computed from time mean temperature and moisture profiles of ascending (red), subsiding (blue) and all grid columns (black) in the subtropics ($\pm \left(10^{\circ}-30^{\circ} \right)$ latitude bands), and (b) space-time weights of ascending (red) and descending (blue) grid columns in the subtropics.
+Grey crosses in (a) are dilute CAPE derived from the sum of the products of space-time weights with the dilute CAPE values of ascending/descending grid columns.}
+\label{fig:cape-subt}
+\end{figure}
+
+To understand why dilute CAPE is so insensitive to subsidence, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
 The ZM mass fluxes typically only extend up to about the $800$ hPa level, which often occurs with subsiding motion aloft.
-
-In drizzling regions, there are opposing influences on dilute CAPE, a global maximum in latent heat fluxes (Figure~\ref{fig:4zonal}c), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and subsidence, which opposes dilute CAPE from above.
-It seems plausible that these two opposing influences on stability translates into small values of dilute CAPE, but for drizzle to occur, still large enough to trigger the ZM scheme.
-The ZM scheme, properly a deep convection scheme, is operating as a shallow convection scheme in the $10^{\circ}-20^{\circ}$ latitude region.
+In the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands there are opposing influences on dilute CAPE; a global maximum in surface latent heat fluxes (Figure~\ref{fig:4zonal}c), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and subsidence, which opposes dilute CAPE from above.
+The shallow convection regime of the ZM scheme depicted in Figure~\ref{fig:transect} is likely a result of these two opposing influences on CAPE, with large the latent heat fluxes increasing dilute CAPE just above the threshold for convection, but with the subsidence restricting dilute CAPE from becoming much larger.
 
@@
 
-Figure~\ref{fig:cape-subt} shows the dilute CAPE values computed from mean temperature and moisture of subsiding and ascending regions in the $\pm \left( 10^{\circ}-30^{\circ} \right)$ latitude bands, hereafter referred to as the subtropics.
-The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), and the dilute CAPE computed from the mean temperature and moisture over the entire subtropics varies much less with resolution compared with the deep tropics ($\sim 20$ J/kg across all resolutions, compared with $\sim 80$ J/kg across resolutions in the deep tropics).
-Also unlike the deep tropics, the reduction in dilute CAPE with resolution is not from an increase in occurrence of subsiding grid columns, since the space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
-The decline in dilute CAPE with resolution in the subtropics is a result of increased stability due to changes in environmental temperature from the larger magnitude downward motion (not shown).
-This increased stability with resolution is not sufficient to overcome the large latent heat fluxes and mitigate the shallow convection regime of the ZM scheme, consistent with the local minimum in the R-values from the logistic regression in drizzling regions (Figure~\ref{fig:4zonal}a).
-
-\begin{figure}
-\begin{center}
-\noindent\includegraphics[width=20pc,angle=0]{figs/temp_cape-subtropics.pdf}\\
-\end{center}
-\caption{(a) Dilute CAPE computed from time mean temperature and moisture profiles of ascending (red), subsiding (blue) and all grid columns (black) in the subtropics ($\pm \left(10^{\circ}-30^{\circ} \right)$ latitude bands), and (b) space-time weights of ascending (red) and descending (blue) grid columns in the subtropics.
-Grey crosses in (a) are dilute CAPE derived from the sum of the products of space-time weights with the dilute CAPE values of ascending/descending grid columns.}
-\label{fig:cape-subt}
-\end{figure}
+AGCMs are known to suffer from a drizzle bias, producing too much light rain relative to observations \citep{D2006JCLIM}.
+Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands, irrespective of resolution, and where the ZM scheme is producing shallow convection.
+The analysis of dilute CAPE in the subtropics indicates that the prevalence of drizzle is a result of this shallow convection regime of the ZM scheme.
+While $FREQZM$ does decrease with resolution in drizzling regions (not shown) consistent with the reduction in dilute CAPE in the subtropics (Figure~\ref{fig:cape-subt}a), the larger magnitude subsidence with resolution is unable to impact the impact the drizzling regime of the ZM scheme with resolution.
 

### 2019-08-19 resolution.tex
edits

@@
 To unravel the contributions of warming and drying to this large spread in dilute CAPE between the two regimes, dilute CAPE is recomputed using the mean specific humidity of ascending/descending regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics.
-Figure~\ref{fig:cape}c shows this influence of changing moisture on the dilute CAPE of subsiding/ascending regions in the deep tropics, which to first order, explains the large spread in dilute CAPE between the two regimes.
+Figure~\ref{fig:cape}c shows this influence of changing moisture on dilute CAPE, which to first order, explains the large spread in dilute CAPE between the ascending/descending regions in the deep tropics.
 Temperature differences between ascending/descending regimes has a smaller, second order influence on the spread in dilute CAPE (not shown).
@@
 The increasing occurrence of subsiding grid columns accounts for about half of the changes in dilute CAPE in the deep tropics with resolution, the other half being due to the systematic reduction in dilute CAPE for both ascending and descending regions.
-To isolate whether temperature or moisture contributes to the latter contribution to the changes in dilute CAPE, a similar experiment is performed in which dilute CAPE is recomputed at all resolutions, but fixing the moisture profiles to the lowest resolution $ne20$ profile, and then repeated through fixing only the temperature profile to $ne20$ values.
+To isolate whether temperature or moisture contributes to the latter contributions change in dilute CAPE, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then repeated through fixing only the temperature profile to $ne20$ values.
 Figure~\ref{fig:cape}d shows the influence of the changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution is primarily from changes to the temperature field.
@@
 
-To understand why dilute CAPE is so insensitive to subsidence, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
+To understand why dilute CAPE is so insensitive to subsidence in the subtropics, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
 The ZM mass fluxes typically only extend up to about the $800$ hPa level, which often occurs with subsiding motion aloft.
 In the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands there are opposing influences on dilute CAPE; a global maximum in surface latent heat fluxes (Figure~\ref{fig:4zonal}c), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and subsidence, which opposes dilute CAPE from above.
-The shallow convection regime of the ZM scheme depicted in Figure~\ref{fig:transect} is likely a result of these two opposing influences on CAPE, with large the latent heat fluxes increasing dilute CAPE just above the threshold for convection, but with the subsidence restricting dilute CAPE from becoming much larger.
+The shallow convection regime of the ZM scheme depicted in Figure~\ref{fig:transect} is likely a result of these two opposing influences on CAPE, with large latent heat fluxes increasing dilute CAPE just above the threshold for convection, but with subsidence restricting dilute CAPE from becoming much larger.
 
@@
 The analysis of dilute CAPE in the subtropics indicates that the prevalence of drizzle is a result of this shallow convection regime of the ZM scheme.
-While $FREQZM$ does decrease with resolution in drizzling regions (not shown) consistent with the reduction in dilute CAPE in the subtropics (Figure~\ref{fig:cape-subt}a), the larger magnitude subsidence with resolution is unable to impact the impact the drizzling regime of the ZM scheme with resolution.
+While $FREQZM$ does decrease with resolution in drizzling regions (not shown) consistent with the reduction in dilute CAPE in the subtropics (Figure~\ref{fig:cape-subt}a), the larger magnitude subsidence with resolution is unable to change the drizzling regime of the ZM scheme with resolution.
 
@@
 \end{equation}
-where $\omega^+$ and $q^+$ are the upward $\omega$ and specific humidity $q$ at cloud base, respectively, and $\rho_w$ is the density of rainwater.
+where $\omega^+$ and $q^+$ are the upward $\omega$ and specific humidity $q$ at cloud base, respectively, with $g$ the acceleration of gravity and $\rho_w$ the density of rainwater.
 
-Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to be a good approximation to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics \citep{OETAL2016JAMES}.
+Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to be a good approximation to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics, across multiple resolutions \citep{OETAL2016JAMES}.
 Figure~\ref{fig:mflux} shows the median convective, stratiform and total precipitation rates conditioned on the $850$ hPa moisture flux using $1$ mm/day bins, in the CAM6 $ne30$ simulation.
@@
 The increase in stratiform precipitation rates with resolution is likely due to the increase in magnitude of $\omega$ with resolution.
-To unravel the contributions of $\omega$ and $q$ at the 850 hPa level, $\omega_{850}$ and $q_{850}$ to the increase in climatological stratiform precipitation with resolution, the area averaged, climatological stratiform precipitation rate $\overline{R_{s}}$ can be decomposed into a double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
+To unravel the contributions of $\omega$ and $q$ at the 850 hPa level $\omega_{850}$ and $q_{850}$, to the increase in climatological stratiform precipitation with resolution, the area averaged, time mean stratiform precipitation rate $\overline{R_{s}}$ can be decomposed into a double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
 \begin{equation}
@@
 \end{equation}
-after \cite{TETAL2018CD}, and the subscript $850$ is dropped from $\omega$ and $q$ for brevity.
+after \cite{TETAL2018CD}. The subscript $850$ is dropped from $\omega$ and $q$ for brevity.
 

### 2019-08-19 resolution.tex
edits

@@
 
-\title{Parameterized convection, grid-scale clouds and resolution sensitivity in the Community Atmosphere Model}
+\title{Parameterized convection, stratiform clouds and resolution sensitivity in the Community Atmosphere Model}
 
@@
 The most common approach towards developing scale-aware physics is through the lens of limited area, large-eddy simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
-By subsequently filtering large-eddy solutions to lower-resolution grids, a relationship between first- and higher-order moments \citep{G1992JFM} may be understood and ultimately parameterized as a function of grid resolution.
+By subsequently filtering large-eddy solutions to lower-resolution grids, a relationship between first- and higher-order moments may be understood and ultimately parameterized as a function of grid resolution.
 While this approach is likely necessary for developing scale-aware physics, it is not sufficient.
@@
 
-Another robust response of the CAM/CCM lineage to resolution is an increase in grid-scale precipitation rates at the expense of parameterized convective precipitation rates.
-Grid-scale precipitation is formed from grid-scale clouds, but is often referred to in the literature as stratiform precipitaiton and stratiform clouds, and used hereafter.
+Another robust response of the CAM/CCM lineage to resolution is an increase in stratiform precipitation rates at the expense of parameterized convective precipitation rates.
 The resolution dependent partitioning between the two different precipitation routines is shown in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies.
@@
 The reduction in convective precipitation rates with resolution in CAM6 is shown to result from the greater magnitude subsiding motion, creating a more stable atmosphere in which the criterion for parameterized convection occurs less often.
-The increase in stratiform precipitation rates with resolution is shown to result more directly from the increase in vertical velocities, by increasing moisture fluxes through the cloud base level.
+The increase in stratiform precipitation rates with resolution is shown to result more directly from the increase in vertical velocities, by increasing moisture fluxes through cloud base.
 The feedback of the resolved vertical motion on the physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution.
@@
 The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed filtered density function \citep{G1992JFM} high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
-The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with a three mode Modular Aerosol Model \citep{MAM}.
+The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with the three mode Modular Aerosol Model \citep{MAM}.
 The combined macrophysics/microphysics routines generate stratiform precipitation from stratiform clouds.
@@
 Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
-In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in the assessment of resolution sensitivity in this study.
+In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing resolution sensitivity in this study.
 All analyses exclude the first month of the simulations, and are computed on their native grids unless otherwise stated.
@@
 The magnitude of upward $\omega$ increases monotonically with resolution, with positive, or downward $\omega$ behaving similarly (not shown).
-This monotonic increase in the magnitude of $\omega$ is evident even after remapping all the model output to a common grid ($ne20$; dotted curves in Figure~\ref{fig:2pdf}a).
+This monotonic increase in the magnitude of $\omega$ is evident even after remapping all model output to a common grid ($ne20$; dotted curves in Figure~\ref{fig:2pdf}a).
 
-The PDF's may be scaled to the highest-resolution resolution grid through $P(\omega)_s = \alpha P (\omega / \alpha)$, where $\alpha$ is the scale factor from equation~\ref{eq:alpha}, and setting $\Delta x_0$ to the $ne120$ grid-spacing.
-Figure~\ref{fig:2pdf}b shows the scaled PDF's for power law exponent $n=-1$ in equation~\ref{eq:alpha}.
+The PDF's may be scaled to the highest-resolution resolution grid through $ \alpha P (\omega / \alpha)$, where $\alpha$ is the scale factor from equation~\ref{eq:alpha}, $P$ the original PDF and setting $\Delta x_0$ to the $ne120$ grid-spacing.
+Figure~\ref{fig:2pdf}b shows the scaled PDF's for a power law exponent $n=-1$ in $\Delta x$.
 The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law exponent $n=-1$ explains to first-order the variation in vertical velocity with resolution in the aqua-planet simulations.
@@
 The predictor is the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
-The assumption is then that subsidence is the independent variable, which the authors believe is reasonable since the environment of subsiding regions is generally more stable than its surroundings, and the ZM scheme is modulated by the dilute CAPE stability metric.
+The assumption is then that subsidence is the independent variable, which is reasonable considering the environment of subsiding regions is generally more stable than its surroundings, and the ZM scheme is modulated by the dilute CAPE stability metric.
 Grid column regressions that are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK} are retained for analysis.
@@
 To isolate whether temperature or moisture contributes to the latter contributions change in dilute CAPE, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then repeated through fixing only the temperature profile to $ne20$ values.
-Figure~\ref{fig:cape}d shows the influence of the changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution is primarily from changes to the temperature field.
+Figure~\ref{fig:cape}d shows the influence of the changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
 Moisture changes with resolution has a smaller influence, only appreciably impacting the dilute CAPE of ascending regions at higher resolutions (not shown).

### 2019-08-19 resolution.tex
edits

@@
 The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), and their dilute CAPE  values vary much less with resolution ($\sim 20$ J/kg, compared with $\sim 80$ J/kg across all resolutions in the deep tropics).
-The lack of spread in dilute CAPE between ascending/descending regions explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}a), since without a clear distinction between dilute CAPE of ascending/descending grid columns, subsidence cannot be effective at depressing dilute CAPE below the threshold for convection.
+The lack of spread in dilute CAPE between ascending/descending regions explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}a), since with no strong dependence of dilute CAPE on ascending/descending regimes, subsidence cannot be a skillful predictor of depressing dilute CAPE below the threshold for convection.
 There is also no significant changes to the occurrence of subsiding grid columns with resolution; the space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
@@
 To understand why dilute CAPE is so insensitive to subsidence in the subtropics, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
-The ZM mass fluxes typically only extend up to about the $800$ hPa level, which often occurs with subsiding motion aloft.
+The ZM mass fluxes typically only extend up to about the $800$ hPa level in this region, which often occurs with appreciable subsiding motion aloft.
 In the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands there are opposing influences on dilute CAPE; a global maximum in surface latent heat fluxes (Figure~\ref{fig:4zonal}c), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and subsidence, which opposes dilute CAPE from above.
@@
 The analysis of dilute CAPE in the subtropics indicates that the prevalence of drizzle is a result of this shallow convection regime of the ZM scheme.
-While $FREQZM$ does decrease with resolution in drizzling regions (not shown) consistent with the reduction in dilute CAPE in the subtropics (Figure~\ref{fig:cape-subt}a), the larger magnitude subsidence with resolution is unable to change the drizzling regime of the ZM scheme with resolution.
+While $FREQZM$ does decrease with resolution in drizzling regions as a result of larger magnitude subsidence and its influence on dilute CAPE (not shown), these resolution sensitivities are not an effective means to eliminate the shallow convection regime of the ZM scheme in the subtropics.
 

### 2019-08-19 resolution.tex
edits

@@
 Curiously, there is an excellent negative correlation (Pearson's R-value = 0.99, N = 27) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
-At any grid-point and time-step, $FREQZM$ is a binary variable: one if the ZM scheme is active, zero if it is not.
+At any grid-point and time-step, $FREQZM$ is a binary variable: 1 if the ZM scheme is active, 0 if it is not.
 Time mean $FREQZM$ is therefore the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg.
@@
 AGCMs are known to suffer from a drizzle bias, producing too much light rain relative to observations \citep{D2006JCLIM}.
-Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands, irrespective of resolution, and where the ZM scheme is producing shallow convection.
-The analysis of dilute CAPE in the subtropics indicates that the prevalence of drizzle is a result of this shallow convection regime of the ZM scheme.
-While $FREQZM$ does decrease with resolution in drizzling regions as a result of larger magnitude subsidence and its influence on dilute CAPE (not shown), these resolution sensitivities are not an effective means to eliminate the shallow convection regime of the ZM scheme in the subtropics.
+Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands, irrespective of resolution.
+The analysis of dilute CAPE in the subtropics indicates that the predominance of drizzle is a result of this shallow convection regime of the ZM scheme.
+While $FREQZM$ does decrease with resolution in drizzling regions as a result of larger magnitude subsidence and its influence on dilute CAPE (not shown), these resolution sensitivities are not an effective means to eliminate the shallow convection regime of the ZM scheme.
 

### 2019-08-19 resolution.tex
edits

@@
 
-The PDF's may be scaled to the highest-resolution resolution grid through $ \alpha P (\omega / \alpha)$, where $\alpha$ is the scale factor from equation~\ref{eq:alpha}, $P$ the original PDF and setting $\Delta x_0$ to the $ne120$ grid-spacing.
+The PDF's may be scaled to the highest-resolution resolution grid through $ \alpha P (\omega / \alpha)$, where $P$ the original PDF, $\alpha$ is the scale factor from equation~\ref{eq:alpha} and setting $\Delta x_0$ to the $ne120$ grid-spacing.
 Figure~\ref{fig:2pdf}b shows the scaled PDF's for a power law exponent $n=-1$ in $\Delta x$.
@@
 In contrast to the impact of vertical motion on the ZM scheme, the stratiform scheme is more intuitively connected to vertical velocities.
-\cite{RETAL2016CD} proposed that total precipitation rates in models ($R$, sum of convective and stratiform precipitation rates) are determined by the upward moisture flux through cloud base,
+\cite{RETAL2016CD} proposed that total precipitation rates in models ($R$, sum of convective and stratiform precipitation rates) can be approximated by the upward moisture flux through cloud base,
 \begin{equation}

### 2019-08-23 resolution.tex
edits

@@
 
-Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to be a good approximation to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics, across multiple resolutions \citep{OETAL2016JAMES}.
+Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to provide to provide a good fit to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics, across multiple resolutions \citep{OETAL2016JAMES}.
 Figure~\ref{fig:mflux} shows the median convective, stratiform and total precipitation rates conditioned on the $850$ hPa moisture flux using $1$ mm/day bins, in the CAM6 $ne30$ simulation.

### 2019-08-23 resolution.tex
edits

@@
 \end{center}
-\caption{Zonal mean (a) R-values and (b) sensitivity parameter in the logistic regression, (c) climatological surface latent heat fluxes and (d) drizzle fraction.}
+\caption{Zonal mean (a) R-values and (b) sensitivity parameter in the logistic regression, (c) climatological $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and (d) climatological $FREQZM$.}
 \label{fig:4zonal}
@@
 
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=14pc,angle=0]{figs/temp_2zonal.pdf}\\
+\end{center}
+\caption{Climatological, zonal mean (a) drizzle fraction and (b) surface latent heat fluxes in the convergence experiment.
+Drizzle fraction refers to the fractional contribution of $\leq 5$ mm/day convective precipitation rates to all convection precipitation rates, for the duration of the simulations and computed from 6-hourly instantaneous fields.}
+\label{fig:2zonal}
+\end{figure}
+
 \subsection{Vertical Velocities and Stratiform Precipitation}

### 2019-08-23 resolution.tex
edits

@@
 \caption{Decomposition of the climatological stratiform precipitation rates, averaged over the $\pm 15^{\circ}$ latitude band into $\omega_{850}$ and $q_{850}$ environmental conditions.
-Left column shows the time mean magnitude term $M\left( \omega_i , q_j \right)$ and the right column is the magnitude term multiplied by the space-time frequency term $f\left( \omega_i , q_j \right) \times M\left( \omega_i , q_j \right)$.
-Integrals over $f \times M$ gives the climatological, area averaged stratiform precipitation rate.
+Left column shows the time mean magnitude term $M\left( \omega_i , q_j \right)$ and the right column is the magnitude term multiplied by the space-time frequency term $f\left( \omega_i , q_j \right) M\left( \omega_i , q_j \right)$.
+Integrals over $f M$ gives the climatological, area averaged stratiform precipitation rate.
 Panel labels denote the grid resolution of the model run.}

### 2019-08-26 resolution.tex
edits

@@
 An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x \leq 5$ km) and climate ($\Delta x \leq 50$ km) simulations \citep{SMTMN2008JCP,MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
-These models are built using unstructured meshes that allows for substantial grid flexibility, but this flexibility is limited by the need for physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
+These models are built using unstructured meshes that allows for substantial grid flexibility, this newfound freedom is restrained by the need for physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
 The most common approach towards developing scale-aware physics is through the lens of limited area, large-eddy simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
@@
 CAM/CCM is a fully supported, well-funded climate model, but despite thirty years of continual model development, there are robust sensitivities to resolution that have persisted in all versions of the model.
-This study argues that a unifying cause, the inherent scale sensitivities of the underlying dynamical equations, can explain the robust responses to resolution that occur in CAM/CCM, {\color{red}{since it is difficult to conceive that inevitable responses to native grid resolution could be ignored in the pursuit of scale-aware physics.}}
+This study argues that a unifying cause, the inherent scale sensitivities of the underlying dynamical equations, can explain the robust responses to resolution that occur in CAM/CCM.
+{\color{red}{Establishing a complete understanding of resolution sensitivity in CAM/CCM and AGCMs alike is crucial, since it is difficult to conceive that inevitable responses to native grid resolution could be ignored in the pursuit of scale-aware physics.}}
 
@@
 It is well known that the magnitude of vertical velocities increase with resolution in atmospheric models.
-While the cause of this sensitivity has been established for large-eddy simulations \citep[see][and references therein]{J2017JAMES}, only recently has the vertical velocity field in AGCMs and their sensitivity to resolution received attention \citep{DETALA2016ACP,OETAL2016JAMES}, albeit with conflicting explanations \citep{RETAL2016CD,HR2018JAMES}.
+While the cause of this sensitivity has been established for large-eddy simulations \citep[see][and references therein]{J2017JAMES}, only recently has the vertical velocity field in AGCMs and their sensitivity to resolution received attention \citep{DETALA2016ACP,OETAL2016JAMES}, albeit with seemingly conflicting explanations \citep{RETAL2016CD,HR2018JAMES}.
 To generalize the relationship between vertical velocity and resolution, let $\alpha$ refer to the ratio of $W_0$, the vertical velocity scale of some reference grid spacing $\Delta x_0$, to $W$, the vertical velocity scale for any $\Delta x$.
@@
 
-\cite{RETAL2016CD} derive an estimate $n= b-1$ by combining a scale analysis of the continuity equation with a power law representation $\Delta x^{2b}$ of the second-order structure function of the horizontal wind.
+\cite{RETAL2016CD} derive an estimate $n= h-1$ in equation~\ref{eq:alpha} by combining a scale analysis of the continuity equation with a power law representation $\Delta x^{2h}$ of the second-order structure function of the horizontal wind.
 Strictly speaking, $\Delta x$ here refers to the distance between two points for which the velocity increment is computed in the structure function, but with this distance set to the model grid-spacing.
-Observations indicate that $b=\frac{1}{3}$ for scales less than about $1000$ km \citep{CETAL1999JGR}, which through the Weiner–Khinchin theorem $- \left( 2b+1 \right) = -\frac{5}{3}$ is equal to the slope of the kinetic energy spectrum, and supported by observations of mesoscale flow \citep{NG1985JAS}.
-\cite{RETAL2016CD} argue that the $-\frac{5}{3}$ slope being common in both observations and models provides an emergent constraint for $b=\frac{1}{3}$ and $n= -\frac{2}{3}$.
+Regional models analyzed in \cite{RETAL2016CD} provide evidence for $h<1$, implying $n<0$ and consistent with the increase in $W$ with resolution in models.
+
+Observations show that $h=\frac{1}{3}$ for scales less than about $1000$ km \citep[hereafter referred to as the {\em{mesoscale}};][]{L1999JFM,CL2001JGR}.
+This value for $h$ is also supported by the slope of the kinetic energy spectrum $\beta$, which can be related to the second-order structure function through the Weiner–Khinchine theorem, $\beta = - \left( 2h+1 \right)$, valid for the range $0<h<1$ \citep{DETAL1996JAS}.
+For $h=\frac{1}{3}$, $\beta=-\frac{5}{3}$, which is true for the kinetic energy spectrum in observations \citep{NG1985JAS,CETAL1999JGR} and models \citep[e.g.,][]{THO2006GRL,SPKS2014JAS} at the mesoscale.
+\cite{RETAL2016CD} proposed that the robust support for $h=\frac{1}{3}$ at the mesoscale provides an emergent constraint for $n=-\frac{2}{3}$ in equation~\ref{eq:alpha}.
 
 In large-eddy simulations, the sensitivity of vertical velocities to resolution is adequately explained by a scale analysis of the dynamical equations \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
-For hydrostatic scales relevant to AGCMs, a scale analysis of the Poisson equation gives $W \propto D^{-1}$, where $D$ is the horizontal scale of a buoyancy perturbation driving vertical motion \citep{HR2018JAMES}.
-In CAM aqua-planet simulations, the largest source of buoyancy is from grid-scale cloud formation, whose horizontal extents are set by the effective resolution of the model (i.e., some multiple of $\Delta x$), indicating $n=-1$ \citep{HR2018JAMES}.
-\cite{HR2017JCLIM} has shown that the $n=-1$ scaling does not explain the behavior of CAM4 in a convergence experiment, but follow-up studies \citep{HR2018JAMES,HETAL2019JAMES} indicate that the inadequacy of the $n=-1$ scaling is not definitive, due to time-truncation errors associated with fixing the physics time-step ($\Delta t_{phys}$) across resolutions in that study.
+For hydrostatic scales relevant to AGCMs, a scale analysis of the Poisson equation gives $W \propto D^{-1}$, where $D$ is the horizontal scale of buoyancy perturbations driving vertical motion \citep{HR2018JAMES}.
+In CAM aqua-planet simulations, the largest source of buoyancy is from stratifrom cloud formation, which are grid-limited with horizontal scales set by the effective resolution of the model \citep[i.e., some multiple of $\Delta x$;][]{S2011LNCSE}, indicating $n=-1$ in equation~\ref{eq:alpha} \citep{HR2018JAMES}.
+\cite{HR2017JCLIM} has shown that the $n=-1$ scaling does not explain the behavior of CAM4 in a convergence experiment, but follow-up studies \citep{HR2018JAMES,HETAL2019JAMES} indicate that the inadequacy of the $n=-1$ scaling is not definitive, due to large time-truncation errors associated with fixing the physics time-step ($\Delta t_{phys}$) across resolutions in that study.
 
@@
 Each window contains a single convergence study, with identical x-axis; the approximate grid resolution.
-Colors indicate the model configuration; January ensemble (black) and aqua-planet configurations with SST profiles $QOBS$ (blue) and $CNTL$ (red) after \cite{NH2000ASL}.
+Colors indicate the model configuration; January ensemble (black) and aqua-planet configurations with SST profiles $QOBS$ (blue) and $CNTL$ (red) in \cite{NH2000ASL}.
 Studies included in this figure are \cite{KW1991JGR} (CCM1), \cite{WETAL1995CD} (CCM2), \cite{W2008TELLUS} (CAM3), \cite{RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM} (CAM4), \cite{ZetAl2014JCb} (CAM5) and this study (CAM6).
@@
 Section 3 contains a thorough analysis of the CAM6 simulations and Section 4 provides some discussion and conclusions.
-
-\section{Methods}
-
-\subsection{Dynamical Core}
-
-This study uses the spectral-element dynamical core option of Community Atmosphere Model \citep[CAM-SE;][]{DetAl2012IJHPCA}, coupled with a mass conserving, semi-Lagrangian advection method for accelerated multi-tracer transport \citep[CSLAM;][]{LTOUNGK2017MWR}, and dry-mass vertical coordinate with comprehensive treatment of moisture and energy \citep{LetAl2018JAMES}.
-The dry dynamics are solved using the high-order, momentum, mass and energy conserving spectral element method \citep{TF2010JCP}, with the elements defined by a cubed-sphere grid.
-The notation for the horizontal grid resolution is an `$ne$' followed by the number of elements making up an edge of one cubed-sphere face, e.g., $ne30$.
-Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds \citep{LetAl2018JAMES}.
-CSLAM tracer transport uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.
-
-\subsection{Physical Parameterizations}
-
-The physics are evaluated on the finite-volume CSLAM grid, and the tendencies mapped back to the spectral element grid.
-The coupled system, referred to as CAM-SE-CSLAM, conserves energy, mass and preserves linear correlations between two reactive species to within machine precision \citep{HL2018MWR}.
-A coarser physics grid, containing $\frac{5}{9}$ fewer degrees of freedom than the dynamical core grid is also available as part of the CAM-SE-CSLAM package \citep{HETAL2019JAMES}.
-This lower-resolution physics grid is used in this study, but only as a member of a perturbed parameter ensemble and not in the default convergence experiment.
-The dynamics time-step is subcycled within a longer physics time-step $\Delta t_{phys}$, and the temperature and momentum increments from the physics are divided by the number of subcycles and added to the dynamical core at the beginning of each subcycle.
-The full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LW2019JAMES}.
-
-The simulations use the CAM6 physics package.
-The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed filtered density function \citep{G1992JFM} high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
-The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with the three mode Modular Aerosol Model \citep{MAM}.
-The combined macrophysics/microphysics routines generate stratiform precipitation from stratiform clouds.
-Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} and an entraining plume model \citep[referred to as the dilute convective available potential energy, or {\em{dilute CAPE}} hereafter;][]{RB1992JAS, NRJ2008JC} is used as a convective trigger (convection occurs if dilute CAPE $\geq 70$ J/kg), and for closing the mass fluxes in the cloud ensemble.
-The deep convection scheme also parameterizes convective momentum transport \citep{RR2008JC}.
-
-\subsection{Experimental Design}
-
-The convergence experiment is performed in an aqua-planet configuration \citep{NH2000ASL,MWO2016JAMES}, an all ocean planet with fixed, zonally symmetric sea surface temperatures modeled after present day Earth \citep[$QOBS$ in][]{NH2000ASL}.
-The aqua-planets are in a perpetual equinox, and aerosols are largely absent from the simulations.
-Each simulation is ran for one simulated year.
-Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
-In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing resolution sensitivity in this study.
-All analyses exclude the first month of the simulations, and are computed on their native grids unless otherwise stated.
 
@@
 
+\section{Methods}
+
+\subsection{Dynamical Core}
+
+This study uses the spectral-element dynamical core option of Community Atmosphere Model \citep[CAM-SE;][]{DetAl2012IJHPCA}, coupled with a mass conserving, semi-Lagrangian advection method for accelerated multi-tracer transport \citep[CSLAM;][]{LTOUNGK2017MWR}, and dry-mass vertical coordinate with comprehensive treatment of moisture and energy \citep{LetAl2018JAMES}.
+The dry dynamics are solved using the high-order, momentum, mass and energy conserving spectral element method \citep{TF2010JCP}, with the elements defined by a cubed-sphere grid.
+The notation for the horizontal grid resolution is an `$ne$' followed by the number of elements making up an edge of one cubed-sphere face, e.g., $ne30$.
+Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds \citep{LetAl2018JAMES}.
+CSLAM tracer transport uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.
+
+\subsection{Physical Parameterizations}
+
+The physics are evaluated on the finite-volume CSLAM grid, and the tendencies mapped back to the spectral element grid.
+The coupled system, referred to as CAM-SE-CSLAM, conserves energy, mass and preserves linear correlations between two reactive species to within machine precision \citep{HL2018MWR}.
+A coarser physics grid, containing $\frac{5}{9}$ fewer degrees of freedom than the dynamical core grid is also available as part of the CAM-SE-CSLAM package \citep{HETAL2019JAMES}.
+This lower-resolution physics grid is used in this study, but only as a member of a perturbed parameter ensemble and not in the control convergence experiment.
+The dynamics time-step is subcycled within a longer physics time-step $\Delta t_{phys}$, and the temperature and momentum increments from the physics are divided by the number of subcycles and added to the dynamical core at the beginning of each subcycle.
+The full moisture increment from the physics is applied only at the start of the first subcycle to conserve tracer mass \citep[$ftype=2$ option in][]{LW2019JAMES}.
+
+The simulations use the CAM6 physics package.
+The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed filtered density function \citep{G1992JFM} high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
+The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with the three mode Modular Aerosol Model \citep{MAM}.
+The combined macrophysics/microphysics routines generate stratiform clouds and stratiform precipitation.
+Deep convection is parameterized using a quasi-equilibrium mass flux scheme \citep{ZM1995AO} and an entraining plume calculation \citep[referred to as the dilute convective available potential energy, or {\em{dilute CAPE}} hereafter;][]{RB1992JAS, NRJ2008JC} is used as a convective trigger (convection occurs if dilute CAPE $\geq 70$ J/kg), and for closing the mass fluxes in the cloud ensemble.
+The deep convection scheme also parameterizes convective momentum transport \citep{RR2008JC}.
+
+\subsection{Experimental Design}
+
+The convergence experiment is performed in an aqua-planet configuration \citep{NH2000ASL,MWO2016JAMES}, an all ocean planet with fixed, zonally symmetric sea surface temperatures modeled after present day Earth \citep[$QOBS$ in][]{NH2000ASL}.
+The aqua-planets are in a perpetual equinox, and aerosols are largely absent from the simulations.
+Each simulation is ran for one simulated year.
+Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
+In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing resolution sensitivity in this study.
+All analyses exclude the first month of the simulations, and are computed on their native grids unless otherwise stated.
+
 The horizontal hyper-viscosity operators $\nu$ vary with resolution after \cite{HETAL2019JAMES}, also provided in Table~\ref{tbl:table1}.
@@
 \begin{equation}
-\Delta t_{phys} = \Delta t_{phys,0} \times \frac{n_{e,0}}{n_e}~s,\label{eq:dt-scale}
+\Delta t_{phys} = \Delta t_{phys,0} \times \frac{n_{e,0}}{n_e},\label{eq:dt-scale}
 \end{equation}
@@
 \caption{Probability density distribution of the upward vertical pressure velocities $\omega$ computed everywhere in the model from six-hourly output over the entirety of the year-long simulations.
-(a) Values on their native grid (solid) and values remapped to the $ne20$ grid (dotted), (b) values on their native grid, scaled to the $ne120$ resolution using a power law exponent $n=-1$ in equation~\ref{eq:alpha}.}
+(a) Values on their native grid (solid) and values bilinearly remapped to the $ne20$ grid (dotted), (b) values on their native grid, scaled to the $ne120$ resolution using a power law exponent $n=-1$ in equation~\ref{eq:alpha}.}
 \label{fig:2pdf}
@@
 
-The PDF's may be scaled to the highest-resolution resolution grid through $ \alpha P (\omega / \alpha)$, where $P$ the original PDF, $\alpha$ is the scale factor from equation~\ref{eq:alpha} and setting $\Delta x_0$ to the $ne120$ grid-spacing.
+The PDF's may be scaled to the highest-resolution resolution grid through $ \alpha P (\omega / \alpha)$, where $P(x)$ is the PDF, and $\alpha$ the scale factor from equation~\ref{eq:alpha} with $\Delta x_0$ set to the $ne120$ grid-spacing.
 Figure~\ref{fig:2pdf}b shows the scaled PDF's for a power law exponent $n=-1$ in $\Delta x$.
-The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law exponent $n=-1$ explains to first-order the variation in vertical velocity with resolution in the aqua-planet simulations.
+The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law exponent $n=-1$ explains to first-order the variation in vertical velocity with resolution as shown by the aqua-planet simulations.
 
@@
 \begin{equation}
-p = \frac{exp{[b_0 + b_1 x]}}{1 + exp{[b_0 + b_1 x]}}, \label{eq:eq6-3}
+p = \frac{exp{[b_0 + b_1 x]}}{1 + exp{[b_0 + b_1 x]}}, \label{eq:logreg}
 \end{equation}
@@
 Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and parameter $b_1$ (hereafter referred to as the sensitivity parameter; Figure~\ref{fig:4zonal}a,b).
+{\color{red}{The large latitudinal dependence in the logistic regression skill across all resolutions however, is not mirrored by the changes in the climatological zonal mean $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ with resolution, shown in Figures~\ref{fig:4zonal}c,d., being much less dependent on latitude.}}
 
@@
 This trend reflects changes in the partitioning of the Intertropical Convergence Zone (ITCZ) from convective to stratiform precipitation with resolution.
-Taken together, the region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.
+Taken together, the
+region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.
+
+%The change in  activity of the convection scheme $FREQZM$ with resolution is not substantially different in the deeper tropical regions compared with other regions (Figure~\ref{fig:4zonal}d), but rather, this region is unique for its substantially larger change in convective precipitation per change in $FREQZM$ with resolution.
 
@@
 The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc...
-All differences are computed after remapping the data to the $ne20$ grid.}
+All differences are computed after conservative remapping to a common $ne20$ grid.}
  \centering
@@
 
-To estimate the dilute CAPE values associated with subsiding motion in the deep tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$ is positive or negative, indicating predominantly subsiding or ascending grid columns.
+To characterize the dilute CAPE and frequency of subsiding regions in the deep tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$ is positive or negative, indicating predominantly subsiding or ascending grid columns.
 The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
@@
 The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), and their dilute CAPE  values vary much less with resolution ($\sim 20$ J/kg, compared with $\sim 80$ J/kg across all resolutions in the deep tropics).
+Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but with temperature changes playing a larger role (not shown).
+
 The lack of spread in dilute CAPE between ascending/descending regions explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}a), since with no strong dependence of dilute CAPE on ascending/descending regimes, subsidence cannot be a skillful predictor of depressing dilute CAPE below the threshold for convection.
-There is also no significant changes to the occurrence of subsiding grid columns with resolution; the space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
+There is also no significant changes to the occurrence of subsiding grid columns with resolution.
+The space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
 
@@
 To understand why dilute CAPE is so insensitive to subsidence in the subtropics, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
-The ZM mass fluxes typically only extend up to about the $800$ hPa level in this region, which often occurs with appreciable subsiding motion aloft.
-In the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands there are opposing influences on dilute CAPE; a global maximum in surface latent heat fluxes (Figure~\ref{fig:4zonal}c), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and subsidence, which opposes dilute CAPE from above.
+The ZM mass fluxes typically only extend up to about the 800 hPa level in this region, which often occurs with appreciable subsiding motion aloft.
+In the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands there are opposing influences on dilute CAPE; a global maximum in surface latent heat fluxes (Figure~\ref{fig:2zonal}b), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and subsidence, which opposes dilute CAPE from above.
 The shallow convection regime of the ZM scheme depicted in Figure~\ref{fig:transect} is likely a result of these two opposing influences on CAPE, with large latent heat fluxes increasing dilute CAPE just above the threshold for convection, but with subsidence restricting dilute CAPE from becoming much larger.
@@
 AGCMs are known to suffer from a drizzle bias, producing too much light rain relative to observations \citep{D2006JCLIM}.
-Figure~\ref{fig:4zonal}d shows the fraction of ZM precipitation $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands, irrespective of resolution.
+Figure~\ref{fig:2zonal}a shows the fraction of ZM precipitation $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands, irrespective of resolution.
 The analysis of dilute CAPE in the subtropics indicates that the predominance of drizzle is a result of this shallow convection regime of the ZM scheme.
-While $FREQZM$ does decrease with resolution in drizzling regions as a result of larger magnitude subsidence and its influence on dilute CAPE (not shown), these resolution sensitivities are not an effective means to eliminate the shallow convection regime of the ZM scheme.
+While $FREQZM$ does decrease with resolution in drizzling regions (Figure~\ref{fig:4zonal}d), this resolution sensitivity is not an effective means to eliminate the shallow convection regime of the ZM scheme.
 
@@
 In contrast to the impact of vertical motion on the ZM scheme, the stratiform scheme is more intuitively connected to vertical velocities.
-\cite{RETAL2016CD} proposed that total precipitation rates in models ($R$, sum of convective and stratiform precipitation rates) can be approximated by the upward moisture flux through cloud base,
+\cite{RETAL2016CD} proposed an approximate scaling for the total precipitation rate in models $P_{tot}$, proportional to the upward moisture flux through cloud base,
 \begin{equation}
-R \approx -\frac{1}{g\rho_{w}} \omega^+ q^+ \label{eq:mflux}
+P_{tot} \approx -\frac{1}{g\rho_{w}} \omega^{cb}_u q^{cb}_u \label{eq:mflux}
 \end{equation}
-where $\omega^+$ and $q^+$ are the upward $\omega$ and specific humidity $q$ at cloud base, respectively, with $g$ the acceleration of gravity and $\rho_w$ the density of rainwater.
+where $\omega^{cb}_u$ and $q^{cb}_u$ refers to $\omega$ and specific humidity $q$ for ascending motion at cloud base, respectively, with $g=9.80616$ m/s$^2$ the acceleration of gravity and $\rho_w=1000$ kg/m$^3$ the density of rainwater used in CAM.
+\cite{OETAL2016JAMES} deem this the ``what goes up, must go down" model, which highlights the physically intuitive nature of the scaling, but beware that other potentially important processes are ignored in the scaling, such as detrainment of cloudy updrafts into the environment or updrafts embedded within circulations with large gross moist stability.
 
-Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to provide to provide a good fit to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics, across multiple resolutions \citep{OETAL2016JAMES}.
-Figure~\ref{fig:mflux} shows the median convective, stratiform and total precipitation rates conditioned on the $850$ hPa moisture flux using $1$ mm/day bins, in the CAM6 $ne30$ simulation.
-The figure shows that the moisture flux is a good approximation to the median total precipitation rate, and that the increase in precipitation rates with moisture flux is primarily from the stratiform precipitation scheme.
+Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to provide a good fit to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics, across multiple resolutions \citep{OETAL2016JAMES}.
+As noted by \cite{RETAL2016CD}, using only resolved quantities in equation~\ref{eq:mflux} omits the sub-grid scale contribution to moisture fluxes, and so it is instructive to analyze the relative skill of this approximation for the individual components of $P_{tot}$, being the sum of stratiform $P_s$ and parameterized convective $P_c$ precipitation rates.
+Figure~\ref{fig:mflux} shows the median $P_c$, $P_s$ and $P_{tot}$ conditioned on the $850$ hPa resolved moisture flux scaling using $1$ mm/day bins, in the CAM6 $ne30$ simulation.
+The figure shows that the moisture flux scaling is a good first-order approximation to the median total precipitation rate, and that the increase in precipitation rates with moisture flux is primarily from the stratiform precipitation scheme.
+The scaling is just as skillful for different resolutions \citep[not shown;][]{OETAL2016JAMES} and so changes to the global mean, climatological $P_s$ with resolution can be understood through quantifying changes in moisture fluxes with resolution.
 
@@
 
-The increase in stratiform precipitation rates with resolution is likely due to the increase in magnitude of $\omega$ with resolution.
-To unravel the contributions of $\omega$ and $q$ at the 850 hPa level $\omega_{850}$ and $q_{850}$, to the increase in climatological stratiform precipitation with resolution, the area averaged, time mean stratiform precipitation rate $\overline{R_{s}}$ can be decomposed into a double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
+To unravel the contributions of changes in $\omega$ and $q$ at the 850 hPa level $\omega_{850}$ and $q_{850}$, to the increase in global mean, climatological stratiform precipitation rates $\overline{P_s}$, $\overline{P_s}$ is decomposed in $\omega_{850}$ and $q_{850}$ space following \cite{TETAL2018CD}.
+Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution.
+To simplify the analysis, the decomposition is restricted to stratiform precipitation rates and moisture fluxes in the $\pm 15^{\circ}$ latitude region, and $\overline{P_s}$ is redefined as the average over $\pm 15^{\circ}$ latitude.
+$\overline{P_s}$ can then be expressed as the double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
 \begin{equation}
-\overline{R_{s}} = \sum_i \sum_j f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right) \label{eq:pdecomp}
+\overline{P_{s}} = \sum_i \sum_j f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right) \label{eq:pdecomp}
 \end{equation}
-after \cite{TETAL2018CD}. The subscript $850$ is dropped from $\omega$ and $q$ for brevity.
+where the subscript $850$ is dropped from $\omega$ and $q$ for brevity.
+$f_s$ is a measure of the occurrence of a particular combination $\left( \omega_i , q_j \right)$ in a simulation, and $M_s$ the mean stratiform precipitation rate associated with the combination $\left( \omega_i , q_j \right)$.
+Their product $f_s M_s$ is then the contribution of a particular combination of $\left( \omega_i , q_j \right)$ to $\overline{P_s}$.
 
-Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution, and so $\overline{R_s}$ is defined as the mean over the $\pm 15^{\circ}$ latitude region.
-Figure~\ref{fig:pdecomp} is a plot of the terms $M_s \left( \omega_i , q_j \right)$ and $f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right)$ for all resolutions,.
-The plots are computed using 6-hourly instantaneous output of $\omega_{i}$ and $q_{j}$, with 0.05 Pa/s and 0.4 g/kg bins.
-Values in  $\left( \omega_{i} , q_{j} \right)$ space are only shown for bins with $f_s \geq1 \times 10^{-5}$, a reasonable cut-off to a bins' contribution to $\overline{R_s}$.
-The $M_s$ plots show that larger magnitude $\omega_{850}$ correspond to larger magnitude time-mean stratiform precipitation rates, and these larger precipitation rates are associated with larger spatial frequencies $f_s \left( \omega_i , q_j \right)$ at higher resolutions.
-These large magnitude stratiform precipitation rates increasingly contribute to the increase in $\overline{R_{s}}$ with resolution, as seen by the increase in $f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right)$ in the larger magnitude $\omega_{850}$ space with resolution.
-Changes to the time mean spatial frequency of the $q_{850}$ field contributes comparatively much less to changes in stratiform precipitation.
+Figure~\ref{fig:pdecomp} shows plots of the terms $M_s$ and $f_s M_s$ for all resolutions.
+The plots are computed using 6-hourly instantaneous output of $\omega_{850}$ and $q_{850}$, with 0.05 Pa/s and 0.4 g/kg bin wdiths, respectively.
+Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but still reasonable cut-off to a bins' contribution to $\overline{P_s}$.
+The $M_s$ plots (left column) show that larger magnitude $\omega_{850}$ correspond to larger magnitude stratiform precipitation rates, while the impact of changing $q_{850}$ is less clear.
+The changes in $M_s$ with resolution are subtle, while the changes in $f_s$ with resolution are large (not shown).
+The changes to $f_s$ can be inferred from the larger space of $\left( \omega_i , q_j \right)$ plotted at higher resolutions, indicating that larger magnitude $\omega_{850}$ values are occurring more frequently at higher resolutions, above the cutoff for plotting ($f_s \geq1 \times 10^{-5}$).
+The $f_s M_s$ plots (right column) clearly shows that larger magnitude $\omega_{850}$, and therefore larger magnitude stratiform precipitation rates are contributing to the increase in $\overline{P_s}$ with resolution.
+In contrast, smaller magnitude $\omega_{850}$ and hence smaller stratiform precipitation rates contribute less to $\overline{P_s}$ at higher resolutions.
 
@@
 Funding support for this work was in part provided by the U.S.
-Department of Energy Office of Science (DE-SC0019459) and the National Science Foundation (AGS1648629).
+Department of Energy Office of Science (DE-SC0019459) and the National Science Foundation (AGS1648629 and AGS1830729).
 Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at the National Center for Atmospheric Research (NCAR).

### 2019-08-26 resolution.tex
edits

@@
 
-To further understand this relationship, a logistic regression between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ is performed for each grid column within each of the simulations.
+Figure~\ref{fig:4zonal}a,b, shows the zonal mean variations in $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and $FREQZM$ in the control convergence experiment.
+$FREQZM$ is largest in the $\pm 10^{\circ}$ latitude region, within the Intertropical Convergence Zone (ITCZ), and rapidly decreasing polewards into the subtropics.
+$\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increases away from the ITCZ region and reaches a maximum at the poleward limit of the Hadley Cell.
+The increase (decrease) in $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ ($FREQZM$) with resolution is to first-order, independent of latitude.
+
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=14pc,angle=0]{figs/temp_4zonal.pdf}\\
+\end{center}
+\caption{Zonal mean (a) climatological $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and (b) climatological $FREQZM$.
+Zonal mean (c) R-values and (d) sensitivity parameter in the logistic regression.}
+\label{fig:4zonal}
+\end{figure}
+
+To further understand the relationship between subsidence and activity of the ZM scheme, a logistic regression between $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ is performed for each grid column within each of the simulations.
 Logistic regression uses an iterative method to fit a continuous variable predictor, $x$ to a binary predictand $p$ using the exponential \citep{WILKSBOOK},
@@
 The predictor is the instantaneous $\langle f_{d} \rangle \langle \omega_{d} \rangle$ of a grid column, and the predictand the binary $FREQZM$.
-The assumption is then that subsidence is the independent variable, which is reasonable considering the environment of subsiding regions is generally more stable than its surroundings, and the ZM scheme is modulated by the dilute CAPE stability metric.
+The assumption is then that subsidence is the independent variable, which is reasonable considering the environment of subsiding regions is generally more stable than its surroundings, and the ZM scheme is modulated by the dilute CAPE stability calculation.
 Grid column regressions that are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK} are retained for analysis.
-Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and parameter $b_1$ (hereafter referred to as the sensitivity parameter; Figure~\ref{fig:4zonal}a,b).
-{\color{red}{The large latitudinal dependence in the logistic regression skill across all resolutions however, is not mirrored by the changes in the climatological zonal mean $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$ with resolution, shown in Figures~\ref{fig:4zonal}c,d., being much less dependent on latitude.}}
-
-\begin{figure}
-\begin{center}
-\noindent\includegraphics[width=14pc,angle=0]{figs/temp_4zonal.pdf}\\
-\end{center}
-\caption{Zonal mean (a) R-values and (b) sensitivity parameter in the logistic regression, (c) climatological $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and (d) climatological $FREQZM$.}
-\label{fig:4zonal}
-\end{figure}
+Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and parameter $b_1$ (hereafter referred to as the sensitivity parameter; Figure~\ref{fig:4zonal}c,d).
 
 The zonal mean R-values indicate the greatest goodness of fit in the $\pm 10^{\circ}$ latitude band, hereafter referred to as the deep tropics.
-In this region, the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}b), consistent with the idea that subsiding motion stabilizes the environment and actively depresses dilute CAPE and the activity of the ZM scheme in the simulations.
+In this region, the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}d), consistent with the idea that subsiding motion stabilizes the environment and actively depresses dilute CAPE and the activity of the ZM scheme in the simulations.
 The sensitivity parameter becomes less negative in the deep tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary $FREQZM$.
@@
 Table~\ref{tbl:table2} shows the fractional contribution of the deep tropics to the climatological, global mean change in convective precipitation with resolution.
-The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where convective precipitation rates increase due to a wide double-ITCZ in the $ne20$ run that spans well outside of $\pm 10^{\circ}$ latitude, but which is contained within $\pm 10^{\circ}$ latitude in the $ne30$ run).
+The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where convective precipitation rates increase, in part due to a wide double-ITCZ in the $ne20$ run that spans outside of $\pm 10^{\circ}$ latitude, but which is contained within $\pm 10^{\circ}$ latitude in the $ne30$ run).
 Expanding the latitude boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occurs in this region (again, ignoring $ne30-ne20$; Table~\ref{tbl:table2}).
-This trend reflects changes in the partitioning of the Intertropical Convergence Zone (ITCZ) from convective to stratiform precipitation with resolution.
-Taken together, the
-region with the largest change in convective precipitation with resolution is also the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.
+This trend reflects changes in the partitioning of the ITCZ from convective to stratiform precipitation with resolution.
 
-%The change in  activity of the convection scheme $FREQZM$ with resolution is not substantially different in the deeper tropical regions compared with other regions (Figure~\ref{fig:4zonal}d), but rather, this region is unique for its substantially larger change in convective precipitation per change in $FREQZM$ with resolution.
+The large reduction in convective precipitation with resolution in the deep tropics also happens to be the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.
+But the change in $FREQZM$ in the deep tropics with resolution is not substantially different from any other region (Figure~\ref{fig:4zonal}d), including regions where the logistic regression indicates a poor relationship between subsidence and $FREQZM$.
+The deep tropics is then only unique for its substantially larger change in ZM precipitation per change in $FREQZM$ with resolution.
 
@@
 
-To characterize the dilute CAPE and frequency of subsiding regions in the deep tropics, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$ is positive or negative, indicating predominantly subsiding or ascending grid columns.
+To characterize the changes to dilute CAPE of subsiding regions in the deep tropics with resolution, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$ is positive or negative, indicating predominantly subsiding or ascending grid columns.
 The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
@@
 Ascending regions are associated with larger values of dilute CAPE ($>180$ J/kg) relative to subsiding regions ($<110$ J/kg), and the dilute CAPE in both regimes decreases monotonically with resolution.
-
-The space-time weights associated with ascending and descending grid columns in the deep tropics vary drastically with resolution (Figure~\ref{fig:cape}b).
-The subsiding (ascending) space-time weights change from $0.32$ ($0.68$) in the $ne20$ run, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in the $ne120$ run.
-It is the increasing occurrence of stable, subsiding grid columns with resolution in addition to the reduction in dilute CAPE for both ascending and descending grid columns with resolution (Figure~\ref{fig:cape}a), that produces a reduction in dilute CAPE for the entire deep tropics, verified by the similar values derived through taking the weighted sum of the ascending/descending dilute CAPE values (grey crosses in Figure~\ref{fig:cape}a).
 
@@
 The large spread in dilute CAPE between ascending/descending regions is crucial for maintaining the relationship shown by the logistic regression, since the much smaller values of dilute CAPE of subsiding grid columns is required to depress dilute CAPE below the threshold for convection.
-To unravel the contributions of warming and drying to this large spread in dilute CAPE between the two regimes, dilute CAPE is recomputed using the mean specific humidity of ascending/descending regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics.
+To unravel the contributions of warming and drying shown in Figure~\ref{fig:profiles} to the large spread in dilute CAPE between the two regimes, dilute CAPE is recomputed using the mean specific humidity of ascending/descending regions in the deep tropics, but setting the temperature profile to the mean profile for the entire deep tropics.
 Figure~\ref{fig:cape}c shows this influence of changing moisture on dilute CAPE, which to first order, explains the large spread in dilute CAPE between the ascending/descending regions in the deep tropics.
@@
 
-The increasing occurrence of subsiding grid columns accounts for about half of the changes in dilute CAPE in the deep tropics with resolution, the other half being due to the systematic reduction in dilute CAPE for both ascending and descending regions.
-To isolate whether temperature or moisture contributes to the latter contributions change in dilute CAPE, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then repeated through fixing only the temperature profile to $ne20$ values.
-Figure~\ref{fig:cape}d shows the influence of the changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
+The space-time weights associated with ascending and descending grid columns in the deep tropics vary drastically with resolution (Figure~\ref{fig:cape}b).
+The subsiding (ascending) space-time weights change from $0.32$ ($0.68$) in the $ne20$ run, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in the $ne120$ run.
+This increasing occurrence of stable, subsiding grid columns accounts for about half of the changes in dilute CAPE in the deep tropics with resolution, the other half being due to the systematic reduction in dilute CAPE for both ascending and descending regions (Figure~\ref{fig:cape}a).
+This is verified through comparing the weighted sum of the ascending/descending dilute CAPE values with the dilute CAPE for the entire deep tropics (compare black and grey crosses in Figure~\ref{fig:cape}a).
+
+To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and repeated through fixing only the temperature to the $ne20$ profile.
+Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
 Moisture changes with resolution has a smaller influence, only appreciably impacting the dilute CAPE of ascending regions at higher resolutions (not shown).
@@
 Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but with temperature changes playing a larger role (not shown).
-
-The lack of spread in dilute CAPE between ascending/descending regions explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}a), since with no strong dependence of dilute CAPE on ascending/descending regimes, subsidence cannot be a skillful predictor of depressing dilute CAPE below the threshold for convection.
-There is also no significant changes to the occurrence of subsiding grid columns with resolution.
-The space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
 
@@
 
-To understand why dilute CAPE is so insensitive to subsidence in the subtropics, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
-The ZM mass fluxes typically only extend up to about the 800 hPa level in this region, which often occurs with appreciable subsiding motion aloft.
-In the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands there are opposing influences on dilute CAPE; a global maximum in surface latent heat fluxes (Figure~\ref{fig:2zonal}b), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and subsidence, which opposes dilute CAPE from above.
-The shallow convection regime of the ZM scheme depicted in Figure~\ref{fig:transect} is likely a result of these two opposing influences on CAPE, with large latent heat fluxes increasing dilute CAPE just above the threshold for convection, but with subsidence restricting dilute CAPE from becoming much larger.
+The lack of spread in dilute CAPE between ascending/descending regions explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}a), since with no strong dependence of dilute CAPE on ascending/descending regimes, subsidence cannot be a skillful predictor of depressing dilute CAPE below the threshold for convection.
+There is also no significant changes to the occurrence of subsiding grid columns with resolution.
+The space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
 
@@
 
-AGCMs are known to suffer from a drizzle bias, producing too much light rain relative to observations \citep{D2006JCLIM}.
-Figure~\ref{fig:2zonal}a shows the fraction of ZM precipitation $\leq 5$ mm/day in the simulations, which stubbornly persists at 70\% in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands, irrespective of resolution.
-The analysis of dilute CAPE in the subtropics indicates that the predominance of drizzle is a result of this shallow convection regime of the ZM scheme.
-While $FREQZM$ does decrease with resolution in drizzling regions (Figure~\ref{fig:4zonal}d), this resolution sensitivity is not an effective means to eliminate the shallow convection regime of the ZM scheme.
+To understand why dilute CAPE is so insensitive to subsidence in the subtropics, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
+The ZM mass fluxes typically only extend up to about the 800 hPa level in this region, which often occurs with appreciable subsiding motion aloft.
+Though properly a deep convection scheme, the ZM scheme is acting as a shallow convection scheme in this region.
+This shallow convection regime tends to produce light rain, or drizzle, which is a common bias in AGCMs \citep{D2006JCLIM}.
+Figure~\ref{fig:2zonal}a shows the fraction of ZM precipitation $\leq 5$ mm/day in the simulations, which stubbornly persists at $\sim$ 70\% in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands, irrespective of resolution.
 
@@
 \end{center}
-\caption{Climatological, zonal mean (a) drizzle fraction and (b) surface latent heat fluxes in the convergence experiment.
-Drizzle fraction refers to the fractional contribution of $\leq 5$ mm/day convective precipitation rates to all convection precipitation rates, for the duration of the simulations and computed from 6-hourly instantaneous fields.}
+\caption{Climatological zonal mean (a) drizzle fraction and (b) surface latent heat fluxes in the convergence experiment.
+Drizzle fraction is defined as sum  convective precipitation rates $\leq 5$ mm/day divided by the sum of all convective precipitation rates, computed from 6-hourly instantaneous fields over the duration of the simulation.}
 \label{fig:2zonal}
 \end{figure}
+
+In the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands there are opposing influences on dilute CAPE; a global maximum in surface latent heat fluxes (Figure~\ref{fig:2zonal}b), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and increasing subsidence (Figure~\ref{fig:4zonal}a), which opposes dilute CAPE from above.
+The shallow convection regime of the ZM scheme is likely a result of these two opposing influences on CAPE, with large latent heat fluxes increasing dilute CAPE just above the threshold for convection, but with subsidence restricting dilute CAPE from becoming much larger.
+This interpretation is supported by the logistic regression, which shows a local minimum in goodness-of-fit in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude region (Figure~\ref{fig:4zonal}c), indicating the elevated subsidence is not that skillful in depressing dilute CAPE below the threshold for convection.
 

### 2019-08-27 resolution.tex
edits

@@
 An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x \leq 5$ km) and climate ($\Delta x \leq 50$ km) simulations \citep{SMTMN2008JCP,MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
-These models are built using unstructured meshes that allows for substantial grid flexibility, this newfound freedom is restrained by the need for physical parameterizations ({\em{physics}}) that behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
+These models are built using unstructured meshes that allows for substantial grid flexibility, but lack physical parameterizations ({\em{physics}}) which behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
 The most common approach towards developing scale-aware physics is through the lens of limited area, large-eddy simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
@@
 This study argues that a unifying cause, the inherent scale sensitivities of the underlying dynamical equations, can explain the robust responses to resolution that occur in CAM/CCM.
-{\color{red}{Establishing a complete understanding of resolution sensitivity in CAM/CCM and AGCMs alike is crucial, since it is difficult to conceive that inevitable responses to native grid resolution could be ignored in the pursuit of scale-aware physics.}}
+{\color{red}{Establishing a complete understanding of resolution sensitivity in CAM/CCM and AGCMs alike is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of scale-aware physics.}}
 
@@
 
-Observations show that $h=\frac{1}{3}$ for scales less than about $1000$ km \citep[hereafter referred to as the {\em{mesoscale}};][]{L1999JFM,CL2001JGR}.
-This value for $h$ is also supported by the slope of the kinetic energy spectrum $\beta$, which can be related to the second-order structure function through the Weiner–Khinchine theorem, $\beta = - \left( 2h+1 \right)$, valid for the range $0<h<1$ \citep{DETAL1996JAS}.
-For $h=\frac{1}{3}$, $\beta=-\frac{5}{3}$, which is true for the kinetic energy spectrum in observations \citep{NG1985JAS,CETAL1999JGR} and models \citep[e.g.,][]{THO2006GRL,SPKS2014JAS} at the mesoscale.
+Observations show that $h=\frac{1}{3}$ for scales less than about $100$ km \citep[hereafter referred to as the {\em{mesoscale}};][]{L1999JFM,CL2001JGR}.
+This value for $h$ is also supported by the slope of the kinetic energy spectrum $-\beta$, which can be related to the second-order structure function through the Weiner–Khinchine theorem, $\beta = \left( 2h+1 \right)$, valid in the range $1<\beta<3$ \citep[satisfies stationarity;][]{DETAL1996JAS}.
+For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is true for the kinetic energy spectrum in observations \citep{NG1985JAS,CETAL1999JGR} and models \citep[e.g.,][]{THO2006GRL,SPKS2014JAS} at the mesoscale.
 \cite{RETAL2016CD} proposed that the robust support for $h=\frac{1}{3}$ at the mesoscale provides an emergent constraint for $n=-\frac{2}{3}$ in equation~\ref{eq:alpha}.
@@
 Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but with temperature changes playing a larger role (not shown).
+The increase in subsidence with resolution is likely responsible for the reduction in $FREQZM$ with resolution in the subtropics (Figures~\ref{fig:4zonal}a,b).
 
@@
 
-The lack of spread in dilute CAPE between ascending/descending regions explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}a), since with no strong dependence of dilute CAPE on ascending/descending regimes, subsidence cannot be a skillful predictor of depressing dilute CAPE below the threshold for convection.
-There is also no significant changes to the occurrence of subsiding grid columns with resolution.
-The space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
+The lack of spread in dilute CAPE between ascending/descending regions explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}a).
+Without a strong dependence of dilute CAPE on ascending/descending regimes, subsidence cannot be a skillful predictor of depressing dilute CAPE below the threshold for convection.
+There is also no significant changes to the occurrence of subsiding grid columns with resolution (Figure~\ref{fig:cape-subt}b); space-time weights of ascending/descending motion are more-or-less invariant with resolution.
 
@@
 
-To understand why dilute CAPE is so insensitive to subsidence in the subtropics, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
+To understand why dilute CAPE is less sensitive to subsidence in the subtropics, Figure~\ref{fig:transect} shows two snapshots of $\omega$ in the longitude-pressure plane at  $\sim 18^{\circ}$ latitude in the $ne30$ simulations, overlain by an isoline delineating where the ZM mass fluxes are quite active.
 The ZM mass fluxes typically only extend up to about the 800 hPa level in this region, which often occurs with appreciable subsiding motion aloft.
@@
 The shallow convection regime of the ZM scheme is likely a result of these two opposing influences on CAPE, with large latent heat fluxes increasing dilute CAPE just above the threshold for convection, but with subsidence restricting dilute CAPE from becoming much larger.
-This interpretation is supported by the logistic regression, which shows a local minimum in goodness-of-fit in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude region (Figure~\ref{fig:4zonal}c), indicating the elevated subsidence is not that skillful in depressing dilute CAPE below the threshold for convection.
+This interpretation is supported by the logistic regression, which shows a local minimum in goodness-of-fit in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude region (Figure~\ref{fig:4zonal}c), indicating the elevated subsidence is not very skillful in depressing dilute CAPE below the threshold for convection.
 
@@
 where $\omega^{cb}_u$ and $q^{cb}_u$ refers to $\omega$ and specific humidity $q$ for ascending motion at cloud base, respectively, with $g=9.80616$ m/s$^2$ the acceleration of gravity and $\rho_w=1000$ kg/m$^3$ the density of rainwater used in CAM.
-\cite{OETAL2016JAMES} deem this the ``what goes up, must go down" model, which highlights the physically intuitive nature of the scaling, but beware that other potentially important processes are ignored in the scaling, such as detrainment of cloudy updrafts into the environment or updrafts embedded within circulations with large gross moist stability.
+\cite{OETAL2016JAMES} deem this the ``what goes up, must go down" model, highlighting the physically intuitive nature of the scaling, but beware the scaling omits other potentially important processes, such as detrainment of cloudy updrafts into the environment or updrafts embedded within efficient circulations, containing high gross moist stabilities.
 
 Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to provide a good fit to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics, across multiple resolutions \citep{OETAL2016JAMES}.
-As noted by \cite{RETAL2016CD}, using only resolved quantities in equation~\ref{eq:mflux} omits the sub-grid scale contribution to moisture fluxes, and so it is instructive to analyze the relative skill of this approximation for the individual components of $P_{tot}$, being the sum of stratiform $P_s$ and parameterized convective $P_c$ precipitation rates.
+As noted by \cite{RETAL2016CD}, using only resolved quantities in equation~\ref{eq:mflux} omits the sub-grid scale contribution to moisture fluxes, and so it is instructive to analyze the relative skill of this approximation for the individual components of $P_{tot}$, being the sum of stratiform $P_s$ and ZM convective $P_c$ precipitation rates.
 Figure~\ref{fig:mflux} shows the median $P_c$, $P_s$ and $P_{tot}$ conditioned on the $850$ hPa resolved moisture flux scaling using $1$ mm/day bins, in the CAM6 $ne30$ simulation.
@@
 Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution.
-To simplify the analysis, the decomposition is restricted to stratiform precipitation rates and moisture fluxes in the $\pm 15^{\circ}$ latitude region, and $\overline{P_s}$ is redefined as the average over $\pm 15^{\circ}$ latitude.
-$\overline{P_s}$ can then be expressed as the double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
+The decomposition is then restricted to stratiform precipitation rates and moisture fluxes in the $\pm 15^{\circ}$ latitude region to simplify the analysis, and $\overline{P_s}$ is redefined as the average over $\pm 15^{\circ}$ latitude.
+$\overline{P_s}$ can be expressed as the double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
 \begin{equation}
@@
 Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but still reasonable cut-off to a bins' contribution to $\overline{P_s}$.
-The $M_s$ plots (left column) show that larger magnitude $\omega_{850}$ correspond to larger magnitude stratiform precipitation rates, while the impact of changing $q_{850}$ is less clear.
+The $M_s$ plots (left column) show that larger magnitude $\omega_{850}$ correspond to larger magnitude stratiform precipitation rates, while the impact of changing $q_{850}$ on $M_s$ is less clear.
 The changes in $M_s$ with resolution are subtle, while the changes in $f_s$ with resolution are large (not shown).
-The changes to $f_s$ can be inferred from the larger space of $\left( \omega_i , q_j \right)$ plotted at higher resolutions, indicating that larger magnitude $\omega_{850}$ values are occurring more frequently at higher resolutions, above the cutoff for plotting ($f_s \geq1 \times 10^{-5}$).
+The changes to $f_s$ can be inferred from the larger space of $\left( \omega_i , q_j \right)$ plotted at higher resolutions, indicating that larger magnitude $\omega_{850}$ values are occurring more frequently at higher resolutions, above the cutoff for plotting ($f_s < 1 \times 10^{-5}$).
 The $f_s M_s$ plots (right column) clearly shows that larger magnitude $\omega_{850}$, and therefore larger magnitude stratiform precipitation rates are contributing to the increase in $\overline{P_s}$ with resolution.

### 2019-08-27 resolution.tex
edits

@@
 This study argues that a unifying cause, the inherent scale sensitivities of the underlying dynamical equations, can explain the robust responses to resolution that occur in CAM/CCM.
-{\color{red}{Establishing a complete understanding of resolution sensitivity in CAM/CCM and AGCMs alike is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of scale-aware physics.}}
 
@@
 
-Observations show that $h=\frac{1}{3}$ for scales less than about $100$ km \citep[hereafter referred to as the {\em{mesoscale}};][]{L1999JFM,CL2001JGR}.
+Observations show that $h=\frac{1}{3}$ for horizontal scales on the order of $10^2$ km and less \citep[hereafter referred to as the {\em{mesoscale}};][]{L1999JFM,CL2001JGR}.
 This value for $h$ is also supported by the slope of the kinetic energy spectrum $-\beta$, which can be related to the second-order structure function through the Weiner–Khinchine theorem, $\beta = \left( 2h+1 \right)$, valid in the range $1<\beta<3$ \citep[satisfies stationarity;][]{DETAL1996JAS}.
 For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is true for the kinetic energy spectrum in observations \citep{NG1985JAS,CETAL1999JGR} and models \citep[e.g.,][]{THO2006GRL,SPKS2014JAS} at the mesoscale.
-\cite{RETAL2016CD} proposed that the robust support for $h=\frac{1}{3}$ at the mesoscale provides an emergent constraint for $n=-\frac{2}{3}$ in equation~\ref{eq:alpha}.
+\cite{RETAL2016CD} propose that the overwhelming support for $h=\frac{1}{3}$ at the mesoscale provides an emergent constraint for $n=-\frac{2}{3}$ in equation~\ref{eq:alpha}.
 
@@
 
-Figure~\ref{fig:4zonal}a,b, shows the zonal mean variations in $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and $FREQZM$ in the control convergence experiment.
+Figure~\ref{fig:4zonal}a,b, shows the zonal mean variations in $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and $FREQZM$ in the convergence experiment.
 $FREQZM$ is largest in the $\pm 10^{\circ}$ latitude region, within the Intertropical Convergence Zone (ITCZ), and rapidly decreasing polewards into the subtropics.
@@
 \caption{Zonal mean (a) climatological $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and (b) climatological $FREQZM$.
-Zonal mean (c) R-values and (d) sensitivity parameter in the logistic regression.}
+Zonal mean (c) R-values and (d) the shape parameter $b_1$ in the logistic regression.}
 \label{fig:4zonal}
@@
 Grid column regressions that are statistically significant at the $95\%$ level using a log-likelihood test \citep{WILKSBOOK} are retained for analysis.
-Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and parameter $b_1$ (hereafter referred to as the sensitivity parameter; Figure~\ref{fig:4zonal}c,d).
+Since the aqua-planets have zonally symmetric boundary conditions, there is a zonally varying structure in the goodness of fit (R-value) and shape parameter $b_1$ (Figure~\ref{fig:4zonal}c,d).
 
 The zonal mean R-values indicate the greatest goodness of fit in the $\pm 10^{\circ}$ latitude band, hereafter referred to as the deep tropics.
-In this region, the sensitivity parameter is large and negative (Figure~\ref{fig:4zonal}d), consistent with the idea that subsiding motion stabilizes the environment and actively depresses dilute CAPE and the activity of the ZM scheme in the simulations.
-The sensitivity parameter becomes less negative in the deep tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower sensitivity parameter to predict the binary $FREQZM$.
+In this region, the shape parameter $b_1$ is large and negative (Figure~\ref{fig:4zonal}d), consistent with the idea that subsiding motion stabilizes the environment and actively depresses dilute CAPE and the activity of the ZM scheme in the simulations.
+The shape parameter becomes less negative in the deep tropics with resolution, likely due to the greater magnitude $\langle f_{d} \rangle \langle \omega_{d} \rangle$ with resolution, which requires a lower $b_1$ to predict the binary $FREQZM$.
 The R-values generally decrease with resolution indicating that there is degradation in the relationship with resolution.
@@
 The large reduction in convective precipitation with resolution in the deep tropics also happens to be the region where the logistic regression indicates that subsiding motion is most skillful at depressing the activity of the convection scheme.
-But the change in $FREQZM$ in the deep tropics with resolution is not substantially different from any other region (Figure~\ref{fig:4zonal}d), including regions where the logistic regression indicates a poor relationship between subsidence and $FREQZM$.
+But the change in $FREQZM$ in the deep tropics with resolution is not substantially different from any other region (Figure~\ref{fig:4zonal}b), such as the midlatitudes where the logistic regression indicates a poor relationship between subsidence and $FREQZM$ (Figure~\ref{fig:4zonal}c).
 The deep tropics is then only unique for its substantially larger change in ZM precipitation per change in $FREQZM$ with resolution.
@@
 
-To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and repeated through fixing only the temperature to the $ne20$ profile.
+To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then through only fixing only the temperature to the $ne20$ profile.
 Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
@@
 Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but with temperature changes playing a larger role (not shown).
-The increase in subsidence with resolution is likely responsible for the reduction in $FREQZM$ with resolution in the subtropics (Figures~\ref{fig:4zonal}a,b).
+This analysis suggests that Increasing subsidence with resolution increases the stability of the mean state, since both ascending/descending regimes become more stable with resolution, and reducing climatological $FREQZM$ despite the poor relationship between instantaneous subsidence and $FREQZM$ in the subtropics, as shown by the logistic regression (Figure~\ref{fig:4zonal}c).
 
@@
 
-The lack of spread in dilute CAPE between ascending/descending regions explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}a).
-Without a strong dependence of dilute CAPE on ascending/descending regimes, subsidence cannot be a skillful predictor of depressing dilute CAPE below the threshold for convection.
-There is also no significant changes to the occurrence of subsiding grid columns with resolution (Figure~\ref{fig:cape-subt}b); space-time weights of ascending/descending motion are more-or-less invariant with resolution.
+It is the lack of spread in dilute CAPE between ascending/descending regions that explains the declining skill in the logistic regression in the subtropics (Figure~\ref{fig:4zonal}c).
+Without a strong dependence of dilute CAPE on ascending/descending regimes, subsidence is not a very skillful predictor of depressing dilute CAPE below the threshold for convection.
+In contrast to the deep tropics, there is also no significant changes to the occurrence of subsiding grid columns with resolution; space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
 
@@
 In the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude bands there are opposing influences on dilute CAPE; a global maximum in surface latent heat fluxes (Figure~\ref{fig:2zonal}b), influencing the thermodynamic state of boundary layer parcels and increasing dilute CAPE from below \citep{Z2002JGR}, and increasing subsidence (Figure~\ref{fig:4zonal}a), which opposes dilute CAPE from above.
-The shallow convection regime of the ZM scheme is likely a result of these two opposing influences on CAPE, with large latent heat fluxes increasing dilute CAPE just above the threshold for convection, but with subsidence restricting dilute CAPE from becoming much larger.
-This interpretation is supported by the logistic regression, which shows a local minimum in goodness-of-fit in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude region (Figure~\ref{fig:4zonal}c), indicating the elevated subsidence is not very skillful in depressing dilute CAPE below the threshold for convection.
+The shallow convection regime of the ZM scheme is likely a result of these two opposing influences on dilute CAPE, with large latent heat fluxes increasing dilute CAPE above the threshold for convection, but with subsidence restricting dilute CAPE from becoming much larger than this threshold.
+This interpretation is supported by the logistic regression, which shows a local minimum in goodness-of-fit in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude region (Figure~\ref{fig:4zonal}c), indicating that it is particularly difficult for subsiding motion to depress dilute CAPE below the threshold for convection where the surface latent heat fluxes are large.
 
@@
 where $\omega^{cb}_u$ and $q^{cb}_u$ refers to $\omega$ and specific humidity $q$ for ascending motion at cloud base, respectively, with $g=9.80616$ m/s$^2$ the acceleration of gravity and $\rho_w=1000$ kg/m$^3$ the density of rainwater used in CAM.
-\cite{OETAL2016JAMES} deem this the ``what goes up, must go down" model, highlighting the physically intuitive nature of the scaling, but beware the scaling omits other potentially important processes, such as detrainment of cloudy updrafts into the environment or updrafts embedded within efficient circulations, containing high gross moist stabilities.
+\cite{OETAL2016JAMES} deem this the ``what goes up, must go down" model, highlighting the physically intuitive nature of the scaling, but note the scaling omits other potentially important processes, such as detrainment of cloudy updrafts into the environment or updrafts embedded within efficient circulations, with large gross moist stabilities.
 
@@
 The figure shows that the moisture flux scaling is a good first-order approximation to the median total precipitation rate, and that the increase in precipitation rates with moisture flux is primarily from the stratiform precipitation scheme.
-The scaling is just as skillful for different resolutions \citep[not shown;][]{OETAL2016JAMES} and so changes to the global mean, climatological $P_s$ with resolution can be understood through quantifying changes in moisture fluxes with resolution.
+The scaling is just as skillful for different resolutions \citep[not shown;][]{OETAL2016JAMES} and so changes to the global mean, climatological $P_s$ with resolution can be understood through quantifying changes in resolved moisture fluxes with resolution.
 
@@
 The decomposition is then restricted to stratiform precipitation rates and moisture fluxes in the $\pm 15^{\circ}$ latitude region to simplify the analysis, and $\overline{P_s}$ is redefined as the average over $\pm 15^{\circ}$ latitude.
+
 $\overline{P_s}$ can be expressed as the double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
@@
 The plots are computed using 6-hourly instantaneous output of $\omega_{850}$ and $q_{850}$, with 0.05 Pa/s and 0.4 g/kg bin wdiths, respectively.
-Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but still reasonable cut-off to a bins' contribution to $\overline{P_s}$.
+Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but still reasonable cut-off for a bins' contribution to $\overline{P_s}$.
 The $M_s$ plots (left column) show that larger magnitude $\omega_{850}$ correspond to larger magnitude stratiform precipitation rates, while the impact of changing $q_{850}$ on $M_s$ is less clear.
 The changes in $M_s$ with resolution are subtle, while the changes in $f_s$ with resolution are large (not shown).
-The changes to $f_s$ can be inferred from the larger space of $\left( \omega_i , q_j \right)$ plotted at higher resolutions, indicating that larger magnitude $\omega_{850}$ values are occurring more frequently at higher resolutions, above the cutoff for plotting ($f_s < 1 \times 10^{-5}$).
+The changes to $f_s$ can be inferred from the larger space of $\left( \omega_i , q_j \right)$ plotted at higher resolutions, indicating that larger magnitude $\omega_{850}$ values are occurring more frequently at higher resolutions, above the cutoff for plotting.
 The $f_s M_s$ plots (right column) clearly shows that larger magnitude $\omega_{850}$, and therefore larger magnitude stratiform precipitation rates are contributing to the increase in $\overline{P_s}$ with resolution.
-In contrast, smaller magnitude $\omega_{850}$ and hence smaller stratiform precipitation rates contribute less to $\overline{P_s}$ at higher resolutions.
+In contrast, smaller magnitude $\omega_{850}$ and hence smaller stratiform precipitation rates contribute less and less to $\overline{P_s}$ with increasing resolution.
 
@@
 \section{Conclusions}
+
+This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception some 30 years ago.
+Establishing a complete understanding of resolution sensitivity in CAM and AGCMs alike is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of scale-aware physics.
+As with most versions within the CAM lineage (see Figure~\ref{fig:cam-history}), the atmosphere becomes drier and less cloudy with increasing resolution, and convective precipitation decreases while stratiform precipitation increases with resolution.
+Prior studies have established that the drying and reduction in cloudiness is a result of increases in magnitude of vertical velocities with resolution, which more efficiently advect dry air downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}.
+This study aims to then understand why and how vertical velocities change with resolution, and why the convection precipitation decreases at the expense of stratiform precipitation with increasing resolution.
+
+The vertical velocities were found to fit a power law scaling $\Delta x^n$ with $n=-1$.
+The scaling is derived from a scale analysis of the Poisson equation in which the pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, facilitating convergence into an air column and driving vertical motion.
+The horizontal scale of buoyancy perturbations in CAM have been previously shown to result from stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, known as the effective resolution \citep{S2011LNCSE}.
+The horizontal scale of stratiform clouds is then proportional to $\Delta x$, which results in the $n=-1$ power law exponent.
+This scaling is consistent with studies describing the resolution sensitivity of large-eddy simulations, which use similar physical arguments, but adapted for non-hydrostatic scales \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
+
+The $n=-1$ power law scaling is potentially in conflict with the scaling proposed in \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second structure function $\Delta x^{2h}$ of the horizontal wind.
+\cite{RETAL2016CD} provide evidence for $h<1$ in their regional climate model simulations, but propose that $h=\frac{1}{3}$ is an emergent constraint for $n=-\frac{2}{3}$, based on observations and models.
+This is intriguing proposal and the authors don't dismiss it's potential value.
+But rather, it is proposed that the overwhelming support for the $h=\frac{1}{3}$ exponent in second-order structure function is only relevant for horizontal scales on the order of $100$ km and less, which is only marginally resolved in even our highest resolution simulation, with an average equatorial grid spacing of $\Delta x=27.8$ km.
+But it is the effective resolution that determines the resolvable scales of motion, which is in the range of $5-10\Delta x$ in CAM-SE-CSLAM \citep{HETAL2019JAMES}.
+
+The Weiner–Khinchine theorem provides a duality between the second-order structure function and the slope of kinetic energy spectrum $-\beta$, in which $h=\frac{1}{3}$ gives $\beta=\frac{5}{3}$, which has robust support from observations and models at the mesoscale.
+The simulations used in this paper are too coarse to resolve the $\beta=\frac{5}{3}$ slope, and has only been observed in the CAM model with $\Delta x=13.9$ km.
+In this study, $\beta=3$ near the effective resolution of the model, and while tempting to apply the Weiner–Khinchine theorem to derive a value for $h$, it is only valid in the range $1<\beta<3$; $\beta=3$ violates stationarity \citep{DETAL1996JAS}.
+One could however, compute the second-order structure function in the simulations directly to estimate a value for $h$ in the simulations, but this is beyond the scope of the current study.
 

### 2019-08-27 resolution.tex
edits

@@
 
-This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception some 30 years ago.
+This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since the models' inception (see Figure~\ref{fig:cam-history}).
 Establishing a complete understanding of resolution sensitivity in CAM and AGCMs alike is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of scale-aware physics.
-As with most versions within the CAM lineage (see Figure~\ref{fig:cam-history}), the atmosphere becomes drier and less cloudy with increasing resolution, and convective precipitation decreases while stratiform precipitation increases with resolution.
-Prior studies have established that the drying and reduction in cloudiness is a result of increases in magnitude of vertical velocities with resolution, which more efficiently advect dry air downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}.
+As with most versions of CAM, in the CAM6 convergence experiment the atmosphere becomes drier and less cloudy with increasing resolution, and convective precipitation decreases while stratiform precipitation increases with resolution.
+Prior studies have established that the drying and reduction in cloudiness is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}.
 This study aims to then understand why and how vertical velocities change with resolution, and why the convection precipitation decreases at the expense of stratiform precipitation with increasing resolution.
 
-The vertical velocities were found to fit a power law scaling $\Delta x^n$ with $n=-1$.
-The scaling is derived from a scale analysis of the Poisson equation in which the pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, facilitating convergence into an air column and driving vertical motion.
+The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$.
+The power law exponent $n=-1$ is derived from a scale analysis of the Poisson equation, indicating that the pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, facilitating convergence into an air column and driving vertical motion.
 The horizontal scale of buoyancy perturbations in CAM have been previously shown to result from stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, known as the effective resolution \citep{S2011LNCSE}.
@@
 
-The $n=-1$ power law scaling is potentially in conflict with the scaling proposed in \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second structure function $\Delta x^{2h}$ of the horizontal wind.
-\cite{RETAL2016CD} provide evidence for $h<1$ in their regional climate model simulations, but propose that $h=\frac{1}{3}$ is an emergent constraint for $n=-\frac{2}{3}$, based on observations and models.
-This is intriguing proposal and the authors don't dismiss it's potential value.
-But rather, it is proposed that the overwhelming support for the $h=\frac{1}{3}$ exponent in second-order structure function is only relevant for horizontal scales on the order of $100$ km and less, which is only marginally resolved in even our highest resolution simulation, with an average equatorial grid spacing of $\Delta x=27.8$ km.
-But it is the effective resolution that determines the resolvable scales of motion, which is in the range of $5-10\Delta x$ in CAM-SE-CSLAM \citep{HETAL2019JAMES}.
+The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind.
+\cite{RETAL2016CD} provide evidence for $h<1$ in their regional climate model simulations, but propose that $h=\frac{1}{3}$ is an emergent constraint for $n=-\frac{2}{3}$, based on observations.
+This is an intriguing proposal and the authors don't seek to dismiss its potential value.
+But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is not resolved in even our highest resolution simulation $\Delta x=27.8$ km since that CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
+The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
+For $h=\frac{1}{3}$, $\beta=-\frac{5}{3}$, which is not observed in the simulation {\color{red}{(Figure~X)}}.
 
-The Weiner–Khinchine theorem provides a duality between the second-order structure function and the slope of kinetic energy spectrum $-\beta$, in which $h=\frac{1}{3}$ gives $\beta=\frac{5}{3}$, which has robust support from observations and models at the mesoscale.
-The simulations used in this paper are too coarse to resolve the $\beta=\frac{5}{3}$ slope, and has only been observed in the CAM model with $\Delta x=13.9$ km.
-In this study, $\beta=3$ near the effective resolution of the model, and while tempting to apply the Weiner–Khinchine theorem to derive a value for $h$, it is only valid in the range $1<\beta<3$; $\beta=3$ violates stationarity \citep{DETAL1996JAS}.
-One could however, compute the second-order structure function in the simulations directly to estimate a value for $h$ in the simulations, but this is beyond the scope of the current study.
+A curious, strong negative correlation is discovered between the climatological, global mean activity of the \cite{ZM1995AO} deep convection scheme (ZM scheme) and global mean subsiding motion in the simulations.
+Since the ZM scheme is modulated by an entraining plume calculation, referred to as the dilute convective available potential energy (dilute CAPE), the authors sought to understand the influence of subsidence on dilute CAPE.
+In the deep tropics ($\pm 10^{\circ}$ latitude), subsiding regions have much smaller dilute CAPE values compared with ascending regions due to the influence of the drier, subsiding environment on dilute CAPE, and subsidence is a good predictor of the activity of the ZM scheme within in the simulations.
+With increasing resolution and in the deep tropics, the occurrence of subsiding motion increases dramatically and at the expense of ascending motion, and the activity of the the ZM scheme decreases overall.
+This is not completely unexpected since e.g., for a doubling of the resolution, and the power-law $\Delta x^{-1}$ scaling of the vertical velocities, one can simulate the same mass fluxes in half the area.
+The increased efficiency of upward mass fluxes with resolution suggests that the occurrence of ascending motion may decrease with resolution, while still maintaining a statistical equilibrium with the large-scale forcing.
 
-%In the $\pm 10^{\circ}$ latitude band, the deep tropics, there is a reduction in the occurrence of ascending grid columns with resolution, which drives an increase in occurrence of subsiding regions.
-The reduction in occurrence of ascending motion in the deep tropics with resolution likely manifests as a reduction in the area of resolved updrafts with resolution, since the horizontal scale of large buoyancy stratiform clouds collapse to smallest resolvable feature in the model \citep{HR2018JAMES}, i.e., the models effective resolution.
-The PDF's of vertical velocities indicate that it is this resolution dependence of stratiform clouds that sets a power law relationship of the magnitude of the vertical velocities with resolution $\Delta x^{n}$, with $n=-1$.
-This scaling also indicates that for a doubling of the resolution, the same upward mass flux can be achieved in the half the area, and so it seems plausible that the reduction in the occurrence of ascending motion with resolution is also a result of the greater efficiency of the resolved mass fluxes with resolution.
+In addition to the change in balance of ascending/descending motion in the deep tropics, there is also a decrease in dilute CAPE of both ascending/descending regimes with increasing resolution.
+A sensitivity analysis of the dilute CAPE calculation has verified that this change in the background stability of the deep tropics is a result temperature changes with resolution.
+A similar change in background stability of the environment occurs everywhere else in the simulations, equator-to-pole.
+
+The relationship between subsidence and temperature falls out of the CAPE equation, which is closely related to dilute CAPE.
+The CAPE budget can be separated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through radiative cooling and advection of dry static energy and moisture by the environment, i.e., the resolved flow.
+The latter term is defined as,
+\begin{equation}
+-R_d \int_{p_t}^{p_b} \frac{\partial T_{ve}}{\partial t} dlnp \label{eq:cape}
+\end{equation}
+where $T_{ve}$ is the virtual temperature of the environment, $R_d$ the gas constant for dry air and subscripts $b$ and $t$ refer to the parcel launch level (typically in the boundary layer) and the level of neutral buoyancy, respectively \citep{Z2002JGR}.
+Equation~\ref{eq:cape} shows that warming the environment reduces CAPE though reducing the buoyancy of lifted air parcels.
+The $T_{ve}$ budget contains a term; the vertical advection of potential energy, which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air.
+In subsiding regions, $w$ is negative and opposes the generation of CAPE through adiabatic warming of the environment.
+
+Unlike the deep tropics, dilute CAPE in the subtropics ($\pm (10^{\circ}-10^{\circ})$ latitude) is less dependent on subsiding motion, and a poor predictor of the activity of the ZM scheme in general.
+This is likely due to the $\pm (10^{\circ}-10^{\circ})$ latitude region coinciding with a global maximum in surface latent heat fluxes, which increases dilute CAPE above the threshold for convection, and despite descending motion aloft opposing the growth of dilute CAPE.
+These opposing influences results in quite small values of dilute CAPE, with ZM scheme acting more like a shallow convection scheme, and producing copious amounts of drizzle.
+\cite{XETAL2019JAMES} recently proposed a new convective trigger the ZM scheme that avoids this excess drizzle in the model.
 

### 2019-08-27 resolution.tex
edits

@@
 
-This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since the models' inception (see Figure~\ref{fig:cam-history}).
-Establishing a complete understanding of resolution sensitivity in CAM and AGCMs alike is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of scale-aware physics.
-As with most versions of CAM, in the CAM6 convergence experiment the atmosphere becomes drier and less cloudy with increasing resolution, and convective precipitation decreases while stratiform precipitation increases with resolution.
+Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of scale-aware physics.
+This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}).
+As with most versions of CAM, in the CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution.
 Prior studies have established that the drying and reduction in cloudiness is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}.
-This study aims to then understand why and how vertical velocities change with resolution, and why the convection precipitation decreases at the expense of stratiform precipitation with increasing resolution.
+This study aims to then understand why and how vertical velocities change with resolution, and why the convectiive precipitation decreases at the expense of stratiform precipitation with increasing resolution.
 
 The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$.
-The power law exponent $n=-1$ is derived from a scale analysis of the Poisson equation, indicating that the pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, facilitating convergence into an air column and driving vertical motion.
-The horizontal scale of buoyancy perturbations in CAM have been previously shown to result from stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, known as the effective resolution \citep{S2011LNCSE}.
-The horizontal scale of stratiform clouds is then proportional to $\Delta x$, which results in the $n=-1$ power law exponent.
+The $n=-1$ power law exponent is derived from a scale analysis of the Poisson equation, which shows pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, facilitating convergence into the air column and the resulting vertical velocities also scale like $D^{-1}$.
+It has been previously shown that $D$ is set by stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, the effective resolution \citep{S2011LNCSE}, and so $D \propto \Delta x$ and $n=-1$.
 This scaling is consistent with studies describing the resolution sensitivity of large-eddy simulations, which use similar physical arguments, but adapted for non-hydrostatic scales \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
@@
 This is an intriguing proposal and the authors don't seek to dismiss its potential value.
-But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is not resolved in even our highest resolution simulation $\Delta x=27.8$ km since that CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
+But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is not resolved in even our highest resolution simulation $\Delta x=27.8$ km since CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
 The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
@@
 Since the ZM scheme is modulated by an entraining plume calculation, referred to as the dilute convective available potential energy (dilute CAPE), the authors sought to understand the influence of subsidence on dilute CAPE.
-In the deep tropics ($\pm 10^{\circ}$ latitude), subsiding regions have much smaller dilute CAPE values compared with ascending regions due to the influence of the drier, subsiding environment on dilute CAPE, and subsidence is a good predictor of the activity of the ZM scheme within in the simulations.
+In the deep tropics ($\pm 10^{\circ}$ latitude), subsiding regions have much smaller dilute CAPE values compared to ascending regions due to the influence of the drier, subsiding environment on dilute CAPE, and subsidence is a good predictor of the activity of the ZM scheme within the simulations.
 With increasing resolution and in the deep tropics, the occurrence of subsiding motion increases dramatically and at the expense of ascending motion, and the activity of the the ZM scheme decreases overall.
-This is not completely unexpected since e.g., for a doubling of the resolution, and the power-law $\Delta x^{-1}$ scaling of the vertical velocities, one can simulate the same mass fluxes in half the area.
-The increased efficiency of upward mass fluxes with resolution suggests that the occurrence of ascending motion may decrease with resolution, while still maintaining a statistical equilibrium with the large-scale forcing.
+This is not completely unexpected due to the $\Delta x^{-1}$ scaling of the vertical velocities, e.g., a doubling of resolution can generate the same mass flux in just half the area.
+The increased efficiency of upward mass fluxes with resolution suggests that the occurrence of ascending motion may decrease with resolution, while still maintaining statistical equilibrium with the large-scale forcing.
+
+{\color{red}{EVERYTHING BELOW HERE NEEDS SOME WORK}}
 
 In addition to the change in balance of ascending/descending motion in the deep tropics, there is also a decrease in dilute CAPE of both ascending/descending regimes with increasing resolution.
-A sensitivity analysis of the dilute CAPE calculation has verified that this change in the background stability of the deep tropics is a result temperature changes with resolution.
-A similar change in background stability of the environment occurs everywhere else in the simulations, equator-to-pole.
+A sensitivity analysis of the dilute CAPE calculation has verified that this change in the background stability of the deep tropics is a result of temperature changes with resolution.
+A similar change in background stability of the environment occurs everywhere else in the simulations, equator-to-pole, and reduces the activity of the ZM scheme globally.
+The increased stability is a result of adiabatic warming due to increasing subsidence with resolution.
+Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, referred to as dCAPE \citep{Z2002JGR}.
+Subsidence warming arises from the vertical advection of dry static energy which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
+Song, personal communication).
+The magnitude of downward $w$ increases everywhere in the model with increasing resolution, reducing dilute CAPE everywhere in the model with increasing resolution.
 
-The relationship between subsidence and temperature falls out of the CAPE equation, which is closely related to dilute CAPE.
-The CAPE budget can be separated into two components \citep{Z2002JGR}; instability due to the thermodynamic state of parcels in the boundary layer and the instability generated through radiative cooling and advection of dry static energy and moisture by the environment, i.e., the resolved flow.
-The latter term is defined as,
-\begin{equation}
--R_d \int_{p_t}^{p_b} \frac{\partial T_{ve}}{\partial t} dlnp \label{eq:cape}
-\end{equation}
-where $T_{ve}$ is the virtual temperature of the environment, $R_d$ the gas constant for dry air and subscripts $b$ and $t$ refer to the parcel launch level (typically in the boundary layer) and the level of neutral buoyancy, respectively \citep{Z2002JGR}.
-Equation~\ref{eq:cape} shows that warming the environment reduces CAPE though reducing the buoyancy of lifted air parcels.
-The $T_{ve}$ budget contains a term; the vertical advection of potential energy, which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air.
-In subsiding regions, $w$ is negative and opposes the generation of CAPE through adiabatic warming of the environment.
+The subtropics ($\pm (10^{\circ}-10^{\circ})$ latitude) are uniquely less sensitive to subsiding motion, and a poor predictor of the activity of the ZM scheme in general.
+This is likely due to the $\pm (10^{\circ}-10^{\circ})$ latitude region coinciding with a global maximum in surface latent heat fluxes, which increases dilute CAPE above the threshold for convection, despite appreciable subsiding motion aloft opposing the growth of dilute CAPE.
+These opposing influences result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing copious amounts of drizzle.
+AGCMs are known to suffer from an excess of drizzle relative to observations \citep{D2006JCLIM}.
+\cite{XETAL2019JAMES} recently proposed a new convective trigger for the ZM scheme containing an additional criterion for convection, that dCAPE must be positive.
+Based on the analysis in this study, it is reasonable to assume that dCAPE would be negative in drizzling regions and convection would not occur.
+The authors suggest that the \cite{XETAL2019JAMES} be explored as a means to mitigate the drizzling bias in CAM.
 
-Unlike the deep tropics, dilute CAPE in the subtropics ($\pm (10^{\circ}-10^{\circ})$ latitude) is less dependent on subsiding motion, and a poor predictor of the activity of the ZM scheme in general.
-This is likely due to the $\pm (10^{\circ}-10^{\circ})$ latitude region coinciding with a global maximum in surface latent heat fluxes, which increases dilute CAPE above the threshold for convection, and despite descending motion aloft opposing the growth of dilute CAPE.
-These opposing influences results in quite small values of dilute CAPE, with ZM scheme acting more like a shallow convection scheme, and producing copious amounts of drizzle.
-\cite{XETAL2019JAMES} recently proposed a new convective trigger the ZM scheme that avoids this excess drizzle in the model.
+The Intertropical Convergence Zone (ITCZ) changes with increasing resolution, with proportionally more precipitation generated by the the stratiform scheme relative to the ZM scheme.
+Energetically, one might assume that stratiform precipitation rates are increasing with resolution to compensate for the reduction in ZM precipitation rates to satisfy globally average radiative-convective equilibrium.
+But from a process point of view, it is shown that resolved upward moisture fluxes are highly correlated with stratiform precipitation rates, and that there is an increase in both upward moisture fluxes and stratiform precipitation with resolution.
+Through decomposing the stratiform precipitation in the ITCZ region into the components of the moisture flux, it is shown that larger magnitude vertical velocities are associated with larger stratiform precipitation rates, and these large precipitation rates increasingly contribute to the increase in the time mean stratiform precipitation rate with resolution.
 

### 2019-08-27 resolution.tex
edits

@@
 This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}).
-As with most versions of CAM, in the CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution.
+As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution.
 Prior studies have established that the drying and reduction in cloudiness is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}.
@@
 The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$.
-The $n=-1$ power law exponent is derived from a scale analysis of the Poisson equation, which shows pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, facilitating convergence into the air column and the resulting vertical velocities also scale like $D^{-1}$.
+The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, which shows pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and the vertical velocities also scale like $D^{-1}$.
 It has been previously shown that $D$ is set by stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, the effective resolution \citep{S2011LNCSE}, and so $D \propto \Delta x$ and $n=-1$.
@@
 This is an intriguing proposal and the authors don't seek to dismiss its potential value.
-But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is not resolved in even our highest resolution simulation $\Delta x=27.8$ km since CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
+But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is not resolved in even our highest resolution simulation $\Delta x=27.8$ km, since CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
 The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
-For $h=\frac{1}{3}$, $\beta=-\frac{5}{3}$, which is not observed in the simulation {\color{red}{(Figure~X)}}.
+For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is significantly less than the $\beta=3$ slope observed in the simulation {\color{red}{(Figure~X)}}.
 
-A curious, strong negative correlation is discovered between the climatological, global mean activity of the \cite{ZM1995AO} deep convection scheme (ZM scheme) and global mean subsiding motion in the simulations.
+A strong negative correlation is discovered between the climatological, global mean activity of the \cite{ZM1995AO} deep convection scheme (ZM scheme) and global mean subsiding motion in the simulations.
 Since the ZM scheme is modulated by an entraining plume calculation, referred to as the dilute convective available potential energy (dilute CAPE), the authors sought to understand the influence of subsidence on dilute CAPE.
@@
 With increasing resolution and in the deep tropics, the occurrence of subsiding motion increases dramatically and at the expense of ascending motion, and the activity of the the ZM scheme decreases overall.
-This is not completely unexpected due to the $\Delta x^{-1}$ scaling of the vertical velocities, e.g., a doubling of resolution can generate the same mass flux in just half the area.
-The increased efficiency of upward mass fluxes with resolution suggests that the occurrence of ascending motion may decrease with resolution, while still maintaining statistical equilibrium with the large-scale forcing.
+This is not completely unexpected due to the $\Delta x^{-1}$ scaling of the vertical velocities, e.g., a doubling of resolution can generate the same upward mass flux in just half the area.
+The increased efficiency of upward mass fluxes with resolution suggests that the occurrence of ascending motion can decrease with resolution while still maintaining statistical equilibrium with the large-scale forcing.
+
+In addition to the change in balance of ascending/descending motion in the deep tropics, there is also a decrease in dilute CAPE of both ascending/descending regimes with increasing resolution.
+This reduction in background dilute CAPE is not confined to the deep tropics, but everywhere in the model and is in large part the reason for the global mean reduction in the activity in the ZM scheme with resolution.
+A sensitivity analysis of the dilute CAPE calculation has verified that this change in background stability is primarily a result of temperature changes with resolution.
+The temperature changes reflect an increase in adiabatic warming from increasing subsidence with resolution.
+Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE, or {\em{dCAPE}} \citep{Z2002JGR}.
+Subsidence warming is represented in dCAPE as the vertical advection of dry static energy, which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
+Song, personal communication).
+The magnitude of downward $w$ increases everywhere in the model with increasing resolution, reducing dilute CAPE and the activity of the ZM scheme everywhere in the model with increasing resolution
+
+The subtropics ($\pm (10^{\circ}-30^{\circ})$ latitude) are uniquely less sensitive to subsiding motion.
+This is likely due to the the large, global maximum in surface latent heat fluxes in the subtropics, which increases dilute CAPE above the threshold for convection, despite appreciable subsiding motion aloft that opposes the growth of dilute CAPE.
+These opposing influences ho however result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing copious amounts of light precipitation or drizzle, irrespective of resolution.
+AGCMs are known to suffer from an excess of drizzle relative to observations \citep{D2006JCLIM}.
+\cite{XETAL2019JAMES} recently proposed a new convective trigger for the ZM scheme that the authors believe may alleviate the excess drizzle bias in CAM.
+In additional to standard requirement that dilute CAPE exceed a threshold, \cite{XETAL2019JAMES} also requires dCAPE to be positive.
+Based on the analysis in this study, and the strong connection between vertical velocities and dCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that dCAPE is often negative in drizzling regions.
+The authors suggest that the \cite{XETAL2019JAMES} trigger be explored as a means to mitigate the drizzling bias in CAM.
 
 {\color{red}{EVERYTHING BELOW HERE NEEDS SOME WORK}}
-
-In addition to the change in balance of ascending/descending motion in the deep tropics, there is also a decrease in dilute CAPE of both ascending/descending regimes with increasing resolution.
-A sensitivity analysis of the dilute CAPE calculation has verified that this change in the background stability of the deep tropics is a result of temperature changes with resolution.
-A similar change in background stability of the environment occurs everywhere else in the simulations, equator-to-pole, and reduces the activity of the ZM scheme globally.
-The increased stability is a result of adiabatic warming due to increasing subsidence with resolution.
-Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, referred to as dCAPE \citep{Z2002JGR}.
-Subsidence warming arises from the vertical advection of dry static energy which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
-Song, personal communication).
-The magnitude of downward $w$ increases everywhere in the model with increasing resolution, reducing dilute CAPE everywhere in the model with increasing resolution.
-
-The subtropics ($\pm (10^{\circ}-10^{\circ})$ latitude) are uniquely less sensitive to subsiding motion, and a poor predictor of the activity of the ZM scheme in general.
-This is likely due to the $\pm (10^{\circ}-10^{\circ})$ latitude region coinciding with a global maximum in surface latent heat fluxes, which increases dilute CAPE above the threshold for convection, despite appreciable subsiding motion aloft opposing the growth of dilute CAPE.
-These opposing influences result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing copious amounts of drizzle.
-AGCMs are known to suffer from an excess of drizzle relative to observations \citep{D2006JCLIM}.
-\cite{XETAL2019JAMES} recently proposed a new convective trigger for the ZM scheme containing an additional criterion for convection, that dCAPE must be positive.
-Based on the analysis in this study, it is reasonable to assume that dCAPE would be negative in drizzling regions and convection would not occur.
-The authors suggest that the \cite{XETAL2019JAMES} be explored as a means to mitigate the drizzling bias in CAM.
 

### 2019-08-27 resolution.tex
edits

@@
 The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind.
-\cite{RETAL2016CD} provide evidence for $h<1$ in their regional climate model simulations, but propose that $h=\frac{1}{3}$ is an emergent constraint for $n=-\frac{2}{3}$, based on observations.
+\cite{RETAL2016CD} provide evidence for $h<1$ in their regional climate model simulations, and propose that $h=\frac{1}{3}$ is an emergent constraint for $n=-\frac{2}{3}$, based on observations.
 This is an intriguing proposal and the authors don't seek to dismiss its potential value.
@@
 The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
-For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is significantly less than the $\beta=3$ slope observed in the simulation {\color{red}{(Figure~X)}}.
+For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is significantly less than the $\beta=3$ slope observed near the effective resolution in the simulation {\color{red}{(Figure~X)}}.
 
@@
 Since the ZM scheme is modulated by an entraining plume calculation, referred to as the dilute convective available potential energy (dilute CAPE), the authors sought to understand the influence of subsidence on dilute CAPE.
-In the deep tropics ($\pm 10^{\circ}$ latitude), subsiding regions have much smaller dilute CAPE values compared to ascending regions due to the influence of the drier, subsiding environment on dilute CAPE, and subsidence is a good predictor of the activity of the ZM scheme within the simulations.
+In the deep tropics ($\pm 10^{\circ}$ latitude), subsiding regions have much smaller dilute CAPE values compared to ascending regions due to the influence of the drier, subsiding environment on dilute CAPE.
 With increasing resolution and in the deep tropics, the occurrence of subsiding motion increases dramatically and at the expense of ascending motion, and the activity of the the ZM scheme decreases overall.
@@
 A sensitivity analysis of the dilute CAPE calculation has verified that this change in background stability is primarily a result of temperature changes with resolution.
-The temperature changes reflect an increase in adiabatic warming from increasing subsidence with resolution.
-Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE, or {\em{dCAPE}} \citep{Z2002JGR}.
+The temperature changes reflect an increase in adiabatic warming due to increasing subsidence with resolution.
+Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE, or {\em{dCAPE}} \citep{XZ2000JGR,Z2002JGR}.
 Subsidence warming is represented in dCAPE as the vertical advection of dry static energy, which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.

### 2019-08-28 resolution.tex
edits

@@
 The subtropics ($\pm (10^{\circ}-30^{\circ})$ latitude) are uniquely less sensitive to subsiding motion.
-This is likely due to the the large, global maximum in surface latent heat fluxes in the subtropics, which increases dilute CAPE above the threshold for convection, despite appreciable subsiding motion aloft that opposes the growth of dilute CAPE.
-These opposing influences ho however result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing copious amounts of light precipitation or drizzle, irrespective of resolution.
+This is likely due to the the large, global maximum in surface latent heat fluxes in the subtropics, which increases dilute CAPE above the threshold for convection, but is restricted from growing too much larger due to substantial subsiding motion aloft.
+These opposing influences result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing copious amounts of light precipitation or drizzle, irrespective of resolution.
 AGCMs are known to suffer from an excess of drizzle relative to observations \citep{D2006JCLIM}.
-\cite{XETAL2019JAMES} recently proposed a new convective trigger for the ZM scheme that the authors believe may alleviate the excess drizzle bias in CAM.
-In additional to standard requirement that dilute CAPE exceed a threshold, \cite{XETAL2019JAMES} also requires dCAPE to be positive.
+\cite{XETAL2019JAMES} recently proposed a new convective trigger for the ZM scheme that the authors suggest may alleviate the excess drizzle bias in CAM.
+As with the ZM scheme used in this study that dilute CAPE exceed a threshold fro convection to ccur, \cite{XETAL2019JAMES} also requires dCAPE to be positive.
 Based on the analysis in this study, and the strong connection between vertical velocities and dCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that dCAPE is often negative in drizzling regions.
@@
 
-{\color{red}{EVERYTHING BELOW HERE NEEDS SOME WORK}}
+The stratiform scheme becomes more active with resolution, compensating for the reduction in activity of the ZM scheme with resolution, particularly as it relates to the Intertropical Convergence Zone (ITCZ).
+Energetically it might be expected that stratiform precipitation rates increase with resolution, compensating for the reduction in ZM precipitation rates to satisfy the global mean radiative convective equilibrium.
+But from a process point of view, it is shown that resolved upward moisture fluxes are highly correlated with stratiform precipitation rates, and that there is an increase in both upward moisture fluxes and stratiform precipitation with resolution.
+Through decomposing the stratiform precipitation in the ITCZ region into the components of the moisture flux, it is shown that larger magnitude vertical velocities are associated with larger stratiform precipitation rates, and these large precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate with resolution.
 
-The Intertropical Convergence Zone (ITCZ) changes with increasing resolution, with proportionally more precipitation generated by the the stratiform scheme relative to the ZM scheme.
-Energetically, one might assume that stratiform precipitation rates are increasing with resolution to compensate for the reduction in ZM precipitation rates to satisfy globally average radiative-convective equilibrium.
-But from a process point of view, it is shown that resolved upward moisture fluxes are highly correlated with stratiform precipitation rates, and that there is an increase in both upward moisture fluxes and stratiform precipitation with resolution.
-Through decomposing the stratiform precipitation in the ITCZ region into the components of the moisture flux, it is shown that larger magnitude vertical velocities are associated with larger stratiform precipitation rates, and these large precipitation rates increasingly contribute to the increase in the time mean stratiform precipitation rate with resolution.
+{\color{red}{Wrap up with a closing paragraph}}
 

### 2019-08-28 resolution.tex
edits

@@
 The stratiform scheme becomes more active with resolution, compensating for the reduction in activity of the ZM scheme with resolution, particularly as it relates to the Intertropical Convergence Zone (ITCZ).
-Energetically it might be expected that stratiform precipitation rates increase with resolution, compensating for the reduction in ZM precipitation rates to satisfy the global mean radiative convective equilibrium.
+Energetically, it is expected that stratiform precipitation rates increase with resolution to compensate for the reduction in convection from the ZM scheme in order satisfy energy balance through global mean radiative convective equilibrium.
 But from a process point of view, it is shown that resolved upward moisture fluxes are highly correlated with stratiform precipitation rates, and that there is an increase in both upward moisture fluxes and stratiform precipitation with resolution.

### 2019-08-28 resolution.tex
edits

@@
 \cite{XETAL2019JAMES} recently proposed a new convective trigger for the ZM scheme that the authors suggest may alleviate the excess drizzle bias in CAM.
-As with the ZM scheme used in this study that dilute CAPE exceed a threshold fro convection to ccur, \cite{XETAL2019JAMES} also requires dCAPE to be positive.
-Based on the analysis in this study, and the strong connection between vertical velocities and dCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that dCAPE is often negative in drizzling regions.
+As with the ZM scheme used in this study, dilute CAPE must exceed a threshold for convection to occur, but \cite{XETAL2019JAMES} also requires dCAPE to be positive.
+Based on the analysis in this study, and the strong connection between vertical velocities and dCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that dCAPE is often negative in drizzling regions, which would prevent the ZM scheme from triggering.
 The authors suggest that the \cite{XETAL2019JAMES} trigger be explored as a means to mitigate the drizzling bias in CAM.
+
+{\color{red}{What follows needs serious work}}
 

### 2019-08-28 resolution.tex
edits

--- 2019-08-28 resolution.tex
@@
 \cite{XETAL2019JAMES} recently proposed a new convective trigger for the ZM scheme that the authors suggest may alleviate the excess drizzle bias in CAM.
-As with the ZM scheme used in this study that dilute CAPE exceed a threshold fro convection to ccur, \cite{XETAL2019JAMES} also requires dCAPE to be positive.
-Based on the analysis in this study, and the strong connection between vertical velocities and dCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that dCAPE is often negative in drizzling regions.
+As with the ZM scheme used in this study, dilute CAPE must exceed a threshold for convection to occur, but \cite{XETAL2019JAMES} also require dCAPE to be positive.
+Based on the analysis in this study, and the strong connection between vertical velocities and dCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that dCAPE is often negative in drizzling regions, which may prevent the ZM scheme from triggering.
 The authors suggest that the \cite{XETAL2019JAMES} trigger be explored as a means to mitigate the drizzling bias in CAM.
+
+{\color{red}{What follows needs serious work}}
 

### 2019-08-28 resolution.tex
edits

@@
 In addition to the change in balance of ascending/descending motion in the deep tropics, there is also a decrease in dilute CAPE of both ascending/descending regimes with increasing resolution.
-This reduction in background dilute CAPE is not confined to the deep tropics, but everywhere in the model and is in large part the reason for the global mean reduction in the activity in the ZM scheme with resolution.
-A sensitivity analysis of the dilute CAPE calculation has verified that this change in background stability is primarily a result of temperature changes with resolution.
+This reduction in background dilute CAPE is not confined to the deep tropics, but everywhere in the model and is in large part the reason for the global mean reduction in activity in the ZM scheme with resolution.
+A sensitivity analysis of the dilute CAPE calculation indicates that this change in background stability is primarily a result of temperature changes with resolution.
 The temperature changes reflect an increase in adiabatic warming due to increasing subsidence with resolution.
+
 Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE, or {\em{dCAPE}} \citep{XZ2000JGR,Z2002JGR}.
-Subsidence warming is represented in dCAPE as the vertical advection of dry static energy, which simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, and directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
+Subsidence warming is represented in the dCAPE budget as negative the vertical advection of dry static energy, since warming the environment reduces dCAPE, and simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
 Song, personal communication).
-The magnitude of downward $w$ increases everywhere in the model with increasing resolution, reducing dilute CAPE and the activity of the ZM scheme everywhere in the model with increasing resolution
+The magnitude of negative, or downward $w$ increases everywhere in the model with increasing resolution, opposing the growth of dilute CAPE and reducing the activity of the ZM scheme with increasing resolution.
 
@@
 These opposing influences result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing copious amounts of light precipitation or drizzle, irrespective of resolution.
+
 AGCMs are known to suffer from an excess of drizzle relative to observations \citep{D2006JCLIM}.

### 2019-08-28 resolution.tex
edits

@@
 
-{\color{red}{What follows needs serious work}}
-
 The stratiform scheme becomes more active with resolution, compensating for the reduction in activity of the ZM scheme with resolution, particularly as it relates to the Intertropical Convergence Zone (ITCZ).
-Energetically, it is expected that stratiform precipitation rates increase with resolution to compensate for the reduction in convection from the ZM scheme in order satisfy energy balance through global mean radiative convective equilibrium.
-But from a process point of view, it is shown that resolved upward moisture fluxes are highly correlated with stratiform precipitation rates, and that there is an increase in both upward moisture fluxes and stratiform precipitation with resolution.
+Energetically, it is expected that stratiform precipitation rates increase with resolution to compensate for the reduction in convection from the ZM scheme in order to balance radiative cooling.
+But this argument requires one to assume the increase in stratiform precipitation is responding to a reduction in ZM activity, whereas this study indicates the reverse, that the ZM scheme is responding to increases in subsidence with resolution, driven by greater ascent associated with stratiform cloud formation.
+The authors instead sought to understand the increase in stratiform precipitation with resolution at the process level.
+It is shown that resolved upward moisture fluxes are highly correlated with stratiform precipitation rates, and that there is an increase in both upward moisture fluxes and stratiform precipitation with resolution.
 Through decomposing the stratiform precipitation in the ITCZ region into the components of the moisture flux, it is shown that larger magnitude vertical velocities are associated with larger stratiform precipitation rates, and these large precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate with resolution.

### 2019-08-28 resolution.tex
edits

@@
 This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}).
+
 As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution.
 Prior studies have established that the drying and reduction in cloudiness is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}.
-This study aims to then understand why and how vertical velocities change with resolution, and why the convectiive precipitation decreases at the expense of stratiform precipitation with increasing resolution.
+This study aims to then understand why and how vertical velocities change with resolution, and why convective precipitation decreases at the expense of stratiform precipitation with increasing resolution.
 

### 2019-08-28 resolution.tex
edits

@@
 
-Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of scale-aware physics.
+Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of physical parameterizations compatible with the highly flexible grid structures now being used to maximize efficiency on modern day supercomputers.
 This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}).
-
 As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution.
-Prior studies have established that the drying and reduction in cloudiness is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}.
+Prior studies have established that the drying and reduction in cloudiness with resolution is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}.
 This study aims to then understand why and how vertical velocities change with resolution, and why convective precipitation decreases at the expense of stratiform precipitation with increasing resolution.
@@
 The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind.
-\cite{RETAL2016CD} provide evidence for $h<1$ in their regional climate model simulations, and propose that $h=\frac{1}{3}$ is an emergent constraint for $n=-\frac{2}{3}$, based on observations.
+\cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations, it is an emergent constraint for $n=-\frac{2}{3}$.
 This is an intriguing proposal and the authors don't seek to dismiss its potential value.
@@
 Since the ZM scheme is modulated by an entraining plume calculation, referred to as the dilute convective available potential energy (dilute CAPE), the authors sought to understand the influence of subsidence on dilute CAPE.
-In the deep tropics ($\pm 10^{\circ}$ latitude), subsiding regions have much smaller dilute CAPE values compared to ascending regions due to the influence of the drier, subsiding environment on dilute CAPE.
+In the deep tropics ($\pm 10^{\circ}$ latitude), subsiding regions have much smaller dilute CAPE values compared to ascending regions due to the influence of the drier subsiding environment on dilute CAPE.
 With increasing resolution and in the deep tropics, the occurrence of subsiding motion increases dramatically and at the expense of ascending motion, and the activity of the the ZM scheme decreases overall.
-This is not completely unexpected due to the $\Delta x^{-1}$ scaling of the vertical velocities, e.g., a doubling of resolution can generate the same upward mass flux in just half the area.
+This is consistent with the $\Delta x^{-1}$ scaling of the vertical velocities, e.g., a doubling of resolution can generate the same upward mass flux in just half the area.
 The increased efficiency of upward mass fluxes with resolution suggests that the occurrence of ascending motion can decrease with resolution while still maintaining statistical equilibrium with the large-scale forcing.
 
-In addition to the change in balance of ascending/descending motion in the deep tropics, there is also a decrease in dilute CAPE of both ascending/descending regimes with increasing resolution.
-This reduction in background dilute CAPE is not confined to the deep tropics, but everywhere in the model and is in large part the reason for the global mean reduction in activity in the ZM scheme with resolution.
+In addition to the change in balance of ascending/descending motion in the deep tropics, the temperature and moisture profiles indicate a decrease in dilute CAPE for both ascending/descending regimes with increasing resolution.
+This reduction in background dilute CAPE is not confined to the deep tropics, but is also evident from the dilute CAPE derived from the time mean temperature and moisture profiles everywhere in the model (not shown) and is in large part the reason for the global mean reduction in activity in the ZM scheme with resolution.
 A sensitivity analysis of the dilute CAPE calculation indicates that this change in background stability is primarily a result of temperature changes with resolution.
@@
 
-Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE, or {\em{dCAPE}} \citep{XZ2000JGR,Z2002JGR}.
-Subsidence warming is represented in the dCAPE budget as negative the vertical advection of dry static energy, since warming the environment reduces dCAPE, and simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
+Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE \citep[DCAPE;][]{XZ2000JGR,Z2002JGR}.
+Subsidence warming is represented in the DCAPE budget as negative the vertical advection of dry static energy term, since warming the environment reduces dCAPE, and simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
 Song, personal communication).
@@
 \cite{XETAL2019JAMES} recently proposed a new convective trigger for the ZM scheme that the authors suggest may alleviate the excess drizzle bias in CAM.
-As with the ZM scheme used in this study, dilute CAPE must exceed a threshold for convection to occur, but \cite{XETAL2019JAMES} also require dCAPE to be positive.
-Based on the analysis in this study, and the strong connection between vertical velocities and dCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that dCAPE is often negative in drizzling regions, which may prevent the ZM scheme from triggering.
+As with the ZM scheme used in this study, dilute CAPE must exceed a threshold for convection to occur, but \cite{XETAL2019JAMES} also require DCAPE to be positive.
+Based on the analysis in this study, and the strong connection between vertical velocities and DCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that DCAPE is often negative in drizzling regions, and would prevent the ZM scheme from triggering.
 The authors suggest that the \cite{XETAL2019JAMES} trigger be explored as a means to mitigate the drizzling bias in CAM.

### 2019-08-29 resolution.tex
edits

@@
 The ZM mass fluxes typically only extend up to about the 800 hPa level in this region, which often occurs with appreciable subsiding motion aloft.
-Though properly a deep convection scheme, the ZM scheme is acting as a shallow convection scheme in this region.
+Though properly a deep convection scheme, the ZM scheme is acting as a shallow convection scheme in this region \citep{TETAL2016AGU}.
 This shallow convection regime tends to produce light rain, or drizzle, which is a common bias in AGCMs \citep{D2006JCLIM}.
@@
 
-In addition to the change in balance of ascending/descending motion in the deep tropics, the temperature and moisture profiles indicate a decrease in dilute CAPE for both ascending/descending regimes with increasing resolution.
+In addition to the change in balance of ascending/descending motion in the deep tropics with resolution, temperature and moisture profiles shows a decrease in dilute CAPE for both ascending/descending regimes with increasing resolution.
 This reduction in background dilute CAPE is not confined to the deep tropics, but is also evident from the dilute CAPE derived from the time mean temperature and moisture profiles everywhere in the model (not shown) and is in large part the reason for the global mean reduction in activity in the ZM scheme with resolution.
 A sensitivity analysis of the dilute CAPE calculation indicates that this change in background stability is primarily a result of temperature changes with resolution.
-The temperature changes reflect an increase in adiabatic warming due to increasing subsidence with resolution.
+Temperature profiles are more stable with increasing resolution, likely due to an increase in adiabatic warming from increasing subsidence with resolution.
 
 Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE \citep[DCAPE;][]{XZ2000JGR,Z2002JGR}.
-Subsidence warming is represented in the DCAPE budget as negative the vertical advection of dry static energy term, since warming the environment reduces dCAPE, and simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
+Subsidence warming is represented in the DCAPE budget as negative the vertical advection of dry static energy term, since warming the environment reduces DCAPE, and simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
 Song, personal communication).
@@
 The subtropics ($\pm (10^{\circ}-30^{\circ})$ latitude) are uniquely less sensitive to subsiding motion.
-This is likely due to the the large, global maximum in surface latent heat fluxes in the subtropics, which increases dilute CAPE above the threshold for convection, but is restricted from growing too much larger due to substantial subsiding motion aloft.
+This is likely due to the the large, global maximum in surface latent heat fluxes in the subtropics, which increases dilute CAPE above the threshold for convection, but which is restricted from growing much larger due to substantial subsiding motion aloft.
 These opposing influences result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing copious amounts of light precipitation or drizzle, irrespective of resolution.

### 2019-08-29 resolution.tex
edits

@@
 The stratiform scheme becomes more active with resolution, compensating for the reduction in activity of the ZM scheme with resolution, particularly as it relates to the Intertropical Convergence Zone (ITCZ).
-Energetically, it is expected that stratiform precipitation rates increase with resolution to compensate for the reduction in convection from the ZM scheme in order to balance radiative cooling.
-But this argument requires one to assume the increase in stratiform precipitation is responding to a reduction in ZM activity, whereas this study indicates the reverse, that the ZM scheme is responding to increases in subsidence with resolution, driven by greater ascent associated with stratiform cloud formation.
+Energetically, it may expected that stratiform precipitation rates increase with resolution to compensate for the reduction in convection from the ZM scheme in order to balance radiative cooling.
+But this argument requires one to assume the increase in stratiform precipitation is responding to a reduction in ZM activity, whereas this study indicates the reverse, that the ZM scheme is responding to increases in subsidence with resolution, driven by greater ascent from stratiform cloud formation.
 The authors instead sought to understand the increase in stratiform precipitation with resolution at the process level.

### 2019-09-10 resolution.tex
edits

@@
 
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_diags_2panel.pdf}\\
+\end{center}
+\caption{Components of the climatological, global mean vertical pressure velocity, (a) $\langle f_{u} \rangle \langle \omega_{u} \rangle$ and (b) $\langle f_{d} \rangle \langle \omega_{d} \rangle$.
+Grey crosses are for the 24 member perturbed parameter ensemble.}
+\label{fig:2panel}
+\end{figure}
+
 The probability density function (PDF) of negative, or upward vertical pressure velocities $\omega$ in the aqua-planets is provided in Figure~\ref{fig:2pdf}a.
@@
 
+\subsection{Vertical Velocities and Stratiform Precipitation}
+
+\cite{RETAL2016CD} proposed an approximate scaling for the total precipitation rate in models $P_{tot}$, proportional to the upward moisture flux through cloud base,
+\begin{equation}
+P_{tot} \approx -\frac{1}{g\rho_{w}} \omega q |_{cb} \label{eq:mflux}
+\end{equation}
+where $\omega q |_{cb}$ refers to the product of $\omega$ and specific humidity $q$ for ascending motion at cloud base, respectively, with $g=9.80616$ m/s$^2$ the acceleration of gravity and $\rho_w=1000$ kg/m$^3$ the density of rainwater used in CAM.
+\cite{OETAL2016JAMES} deem this the ``what goes up, must go down" model, highlighting the physically intuitive nature of the scaling, but note the scaling omits other potentially important processes, such as detrainment of cloudy updrafts into the environment or updrafts embedded within efficient circulations, with large gross moist stabilities.
+
+Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to provide a good fit to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics, across multiple resolutions \citep[recall that CAM-SE is different from CAM-SE-CSLAM, used in this study;][]{OETAL2016JAMES}.
+As noted by \cite{RETAL2016CD}, using only resolved quantities in equation~\ref{eq:mflux} omits the sub-grid scale contribution to moisture fluxes, and so it is instructive to analyze the relative skill of this approximation for the individual components of $P_{tot}$, being the sum of stratiform $P_s$ and parameterized deep convective $P_c$ precipitation rates.
+Figure~\ref{fig:mflux} shows the median $P_c$, $P_s$ and $P_{tot}$ conditioned on the $850$ hPa resolved moisture flux scaling using $1$ mm/day bins, in the CAM6 $ne30$ simulation.
+The figure shows that the moisture flux scaling is a good first-order approximation to the median total precipitation rate, and that the increase in precipitation rates with moisture flux is primarily from the stratiform precipitation scheme.
+The scaling is just as skillful for different resolutions \citep[not shown;][]{OETAL2016JAMES} and so changes in stratiform precipitation rate with resolution can be understood through quantifying changes in resolved moisture fluxes with resolution.
+
 \begin{figure}
 \begin{center}
-\noindent\includegraphics[width=20pc,angle=0]{figs/temp_diags_2panel.pdf}\\
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_mflux.pdf}\\
 \end{center}
-\caption{Components of the climatological, global mean vertical pressure velocity, (a) $\langle f_{u} \rangle \langle \omega_{u} \rangle$ and (b) $\langle f_{d} \rangle \langle \omega_{d} \rangle$.
-Grey crosses are for the 24 member perturbed parameter ensemble.}
-\label{fig:2panel}
+\caption{Precipitation rates vs.
+upward moisture flux at the 850 hPa level.
+Solid lines refer to median stratiform (green), convective (blue) and total (magenta) precipitation rates conditional on bins of the moisture flux, and shaded regions refer to the conditional interquartile ranges.}
+\label{fig:mflux}
+\end{figure}
+
+ \begin{table}
+ \caption{Fractional contribution of latitude bands $\pm 10^{\circ}$ and $\pm 15^{\circ}$ to changes in global mean precipitation with resolution.
+The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc...
+All differences are computed after conservative remapping to a common $ne20$ grid.}
+ \centering
+ \scriptsize
+ \begin{tabular}{lcccccc}
+   \hline
+   Variable & $ne30$ & $ne40$ & $ne60$ & $ne80$ & $ne120$ \\ 
+   \hline
+   $\pm 10^{\circ}$ ($17.6\%$ of global area) \\
+   Convective Precipitation & -0.58 & 0.62 & 0.66 & 0.72 & 0.70 \\
+   Stratiform Precipitation & 0.55 & 0.63 & 0.69 & 0.67 & 0.41 \\ 
+   \hline
+   $\pm 15^{\circ}$ ($25.8\%$ of global area) \\
+   Convective Precipitation & 0.22 & 0.75 & 0.73 & 0.79 & 0.72 \\
+   Stratiform Precipitation & 0.46 & 0.64 & 0.71 & 0.70 & 0.49 \\      
+ \hline
+ \end{tabular}
+ \label{tbl:table2}
+ \end{table}
+
+To unravel the contributions of changes in $\omega$ and $q$ at the 850 hPa level, $\omega_{850}$ and $q_{850}$, to the increase in global mean, climatological stratiform precipitation rates $\overline{P_s}$, $\overline{P_s}$ is decomposed in $\omega_{850}$ and $q_{850}$ space following \cite{TETAL2018CD}.
+Table~\ref{tbl:table2} shows the fractional contribution of deep tropical belts to the global mean change in stratiform precipitation with resolution.
+The table indicates that roughly a majority ($45-70 \%$) of the increase in stratiform precipitation with resolution is from changes within the $\pm 15^{\circ}$ latitude band, and so the moisture flux decomposition is restricted to this region to simplify the analysis; $\overline{P_s}$ is redefined as the average over $\pm 15^{\circ}$ latitude.
+
+$\overline{P_s}$ can be expressed as the double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
+\begin{equation}
+\overline{P_{s}} = \sum_i \sum_j f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right) \label{eq:pdecomp}
+\end{equation}
+where the subscript $850$ is dropped from $\omega$ and $q$ for brevity.
+$f_s$ is a measure of the occurrence of a particular combination $\left( \omega_i , q_j \right)$ in a simulation, and $M_s$ the mean stratiform precipitation rate associated with the combination $\left( \omega_i , q_j \right)$.
+Their product $f_s M_s$ is then the contribution of a particular combination of $\left( \omega_i , q_j \right)$ to $\overline{P_s}$.
+
+Figure~\ref{fig:pdecomp} shows plots of the terms $M_s$ and $f_s M_s$ for all resolutions.
+The plots are computed using 6-hourly instantaneous output of $\omega_{850}$ and $q_{850}$, with 0.05 Pa/s and 0.4 g/kg bin widths, respectively.
+Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but still reasonable cut-off for a bins' contribution to $\overline{P_s}$.
+The $M_s$ plots (left column) show that larger magnitude $\omega_{850}$ correspond to larger magnitude stratiform precipitation rates, while the impact of changing $q_{850}$ on $M_s$ is less clear.
+The changes in $M_s$ with resolution are subtle, while the changes in $f_s$ with resolution are large (not shown).
+The changes to $f_s$ can be inferred from the larger space of $\left( \omega_i , q_j \right)$ plotted at higher resolutions, indicating that larger magnitude $\omega_{850}$ values are occurring more frequently at higher resolutions, above the cutoff for plotting.
+The $f_s M_s$ plots (right column) clearly shows that larger magnitude $\omega_{850}$, and therefore larger magnitude stratiform precipitation rates, are contributing to the increase in $\overline{P_s}$ with resolution.
+In contrast, smaller magnitude $\omega_{850}$ and hence smaller stratiform precipitation rates contribute less and less to $\overline{P_s}$ with increasing resolution.
+
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_pdecomp.pdf}\\
+\end{center}
+\caption{Decomposition of the climatological stratiform precipitation rates, averaged over the $\pm 15^{\circ}$ latitude band into $\omega_{850}$ and $q_{850}$ environmental conditions.
+Left column shows the time mean magnitude term $M\left( \omega_i , q_j \right)$ and the right column is the magnitude term multiplied by the space-time frequency term $f\left( \omega_i , q_j \right) M\left( \omega_i , q_j \right)$.
+Integrals over $f M$ gives the climatological, area averaged stratiform precipitation rate.
+Panel labels denote the grid resolution of the model run.}
+\label{fig:pdecomp}
 \end{figure}
@@
 
-The large increase in magnitude of the upward and downward vertical velocities with resolution may be expected to impact the behavior of other model components.
-Curiously, there is an excellent negative correlation (Pearson's R-value = 0.99, N = 27) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
+While the increase in magnitude of vertical velocities with resolution clearly drives the increase in moisture fluxes and stratiform precipitation rates with resolution, it is not clear why there is a reduction in parameterized deep convective precipitation rates with resolution.
+Curiously though, there is an excellent negative correlation (Pearson's R-value = 0.99, N = 27) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
 At any grid-point and time-step, $FREQZM$ is a binary variable: 1 if the ZM scheme is active, 0 if it is not.
@@
 Figure~\ref{fig:4zonal}a,b, shows the zonal mean variations in $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and $FREQZM$ in the convergence experiment.
-$FREQZM$ is largest in the $\pm 10^{\circ}$ latitude region, within the Intertropical Convergence Zone (ITCZ), and rapidly decreasing polewards into the subtropics.
+$FREQZM$ is largest in the $\pm 10^{\circ}$ latitude region, within the ITCZ, and rapidly decreasing polewards into the subtropics.
 $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increases away from the ITCZ region and reaches a maximum at the poleward limit of the Hadley Cell.
@@
 The R-values generally decrease with resolution indicating that there is degradation in the relationship with resolution.
+The logistic regression is less skillful in the subtropics, with skill declining even further in the mid-latitudes (Figure~\ref{fig:4zonal}c).
 
@@
 
-Table~\ref{tbl:table2} shows the fractional contribution of the deep tropics to the climatological, global mean change in convective precipitation with resolution.
-The table indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where convective precipitation rates increase, in part due to a wide double-ITCZ in the $ne20$ run that spans outside of $\pm 10^{\circ}$ latitude, but which is contained within $\pm 10^{\circ}$ latitude in the $ne30$ run).
+Table~\ref{tbl:table2} indicates that a majority ($60-70 \%$) of the reduction in convective precipitation with resolution is from changes within the deep tropics (except in going from $ne20$ to $ne30$, where convective precipitation rates increase, in part due to a wide double-ITCZ in the $ne20$ run that spans outside of $\pm 10^{\circ}$ latitude, but which is contained within $\pm 10^{\circ}$ latitude in the $ne30$ run).
 Expanding the latitude boundaries marginally to $\pm 15^{\circ}$, roughly $75\%$ of the changes in convective precipitation with resolution occurs in this region (again, ignoring $ne30-ne20$; Table~\ref{tbl:table2}).
@@
 The deep tropics is then only unique for its substantially larger change in ZM precipitation per change in $FREQZM$ with resolution.
-
- \begin{table}
- \caption{Fractional contribution of latitude bands $\pm 10^{\circ}$ and $\pm 15^{\circ}$ to changes in global mean precipitation with resolution.
-The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc...
-All differences are computed after conservative remapping to a common $ne20$ grid.}
- \centering
- \scriptsize
- \begin{tabular}{lcccccc}
-   \hline
-   Variable & $ne30$ & $ne40$ & $ne60$ & $ne80$ & $ne120$ \\ 
-   \hline
-   $\pm 10^{\circ}$ ($17.6\%$ of global area) \\
-   Convective Precipitation & -0.58 & 0.62 & 0.66 & 0.72 & 0.70 \\
-   Stratiform Precipitation & 0.55 & 0.63 & 0.69 & 0.67 & 0.41 \\ 
-   \hline
-   $\pm 15^{\circ}$ ($25.8\%$ of global area) \\
-   Convective Precipitation & 0.22 & 0.75 & 0.73 & 0.79 & 0.72 \\
-   Stratiform Precipitation & 0.46 & 0.64 & 0.71 & 0.70 & 0.49 \\      
- \hline
- \end{tabular}
- \label{tbl:table2}
- \end{table}
 
@@
 The subsiding (ascending) space-time weights change from $0.32$ ($0.68$) in the $ne20$ run, monotonically increasing (decreasing) with resolution to $0.51$ ($0.49$) in the $ne120$ run.
-This increasing occurrence of stable, subsiding grid columns accounts for about half of the changes in dilute CAPE in the deep tropics with resolution, the other half being due to the systematic reduction in dilute CAPE for both ascending and descending regions (Figure~\ref{fig:cape}a).
-This is verified through comparing the weighted sum of the ascending/descending dilute CAPE values with the dilute CAPE for the entire deep tropics (compare black and grey crosses in Figure~\ref{fig:cape}a).
+The increasing occurrence of stable, subsiding grid columns accounts for about half of the changes in dilute CAPE in the deep tropics with resolution, the other half being due to the systematic reduction in dilute CAPE for both ascending and descending regions (Figure~\ref{fig:cape}a).
+This breakdown is straightforward to compute since the sum of the products of the space-time fractions with their associated ascending/descending dilute CAPE values approximate the dilute CAPE computed from the mean temperature/moisture profiles over the entire deep tropics (compare black and grey crosses in Figure~\ref{fig:cape}a) rather well.
 
-To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then through only fixing only the temperature to the $ne20$ profile.
+To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then through fixing only the temperature to the $ne20$ profile.
 Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
@@
 Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but with temperature changes playing a larger role (not shown).
-This analysis suggests that Increasing subsidence with resolution increases the stability of the mean state, since both ascending/descending regimes become more stable with resolution, and reducing climatological $FREQZM$ despite the poor relationship between instantaneous subsidence and $FREQZM$ in the subtropics, as shown by the logistic regression (Figure~\ref{fig:4zonal}c).
+This analysis suggests that increasing subsidence with resolution (\ref{fig:4zonal}a) increases the stability of the mean state, since both ascending/descending regimes become more stable with resolution, and reducing climatological $FREQZM$ (\ref{fig:4zonal}b) despite the poor relationship between instantaneous subsidence and $FREQZM$ in the subtropics, as shown by the logistic regression (Figure~\ref{fig:4zonal}c).
 
@@
 This interpretation is supported by the logistic regression, which shows a local minimum in goodness-of-fit in the $\pm \left(10^{\circ}-20^{\circ} \right)$ latitude region (Figure~\ref{fig:4zonal}c), indicating that it is particularly difficult for subsiding motion to depress dilute CAPE below the threshold for convection where the surface latent heat fluxes are large.
-
-\subsection{Vertical Velocities and Stratiform Precipitation}
-
-In contrast to the impact of vertical motion on the ZM scheme, the stratiform scheme is more intuitively connected to vertical velocities.
-\cite{RETAL2016CD} proposed an approximate scaling for the total precipitation rate in models $P_{tot}$, proportional to the upward moisture flux through cloud base,
-\begin{equation}
-P_{tot} \approx -\frac{1}{g\rho_{w}} \omega^{cb}_u q^{cb}_u \label{eq:mflux}
-\end{equation}
-where $\omega^{cb}_u$ and $q^{cb}_u$ refers to $\omega$ and specific humidity $q$ for ascending motion at cloud base, respectively, with $g=9.80616$ m/s$^2$ the acceleration of gravity and $\rho_w=1000$ kg/m$^3$ the density of rainwater used in CAM.
-\cite{OETAL2016JAMES} deem this the ``what goes up, must go down" model, highlighting the physically intuitive nature of the scaling, but note the scaling omits other potentially important processes, such as detrainment of cloudy updrafts into the environment or updrafts embedded within efficient circulations, with large gross moist stabilities.
-
-Through approximating the cloud base as the $850$ hPa level, equation~\ref{eq:mflux} was found to provide a good fit to total precipitation rates in a regional model \citep{RETAL2016CD} and in the CAM-SE AGCM with CAM5 physics, across multiple resolutions \citep{OETAL2016JAMES}.
-As noted by \cite{RETAL2016CD}, using only resolved quantities in equation~\ref{eq:mflux} omits the sub-grid scale contribution to moisture fluxes, and so it is instructive to analyze the relative skill of this approximation for the individual components of $P_{tot}$, being the sum of stratiform $P_s$ and ZM convective $P_c$ precipitation rates.
-Figure~\ref{fig:mflux} shows the median $P_c$, $P_s$ and $P_{tot}$ conditioned on the $850$ hPa resolved moisture flux scaling using $1$ mm/day bins, in the CAM6 $ne30$ simulation.
-The figure shows that the moisture flux scaling is a good first-order approximation to the median total precipitation rate, and that the increase in precipitation rates with moisture flux is primarily from the stratiform precipitation scheme.
-The scaling is just as skillful for different resolutions \citep[not shown;][]{OETAL2016JAMES} and so changes to the global mean, climatological $P_s$ with resolution can be understood through quantifying changes in resolved moisture fluxes with resolution.
-
-\begin{figure}
-\begin{center}
-\noindent\includegraphics[width=20pc,angle=0]{figs/temp_mflux.pdf}\\
-\end{center}
-\caption{Precipitation rates vs.
-upward moisture flux at the 850 hPa level.
-Solid lines refer to median stratiform (green), convective (blue) and total (magenta) precipitation rates conditional on bins of the moisture flux, and shaded regions refer to the conditional interquartile ranges.}
-\label{fig:mflux}
-\end{figure}
-
-To unravel the contributions of changes in $\omega$ and $q$ at the 850 hPa level $\omega_{850}$ and $q_{850}$, to the increase in global mean, climatological stratiform precipitation rates $\overline{P_s}$, $\overline{P_s}$ is decomposed in $\omega_{850}$ and $q_{850}$ space following \cite{TETAL2018CD}.
-Table~\ref{tbl:table2} indicates that the $\pm 15^{\circ}$ latitude band accounts for most of the change in global mean stratiform precipitation with resolution.
-The decomposition is then restricted to stratiform precipitation rates and moisture fluxes in the $\pm 15^{\circ}$ latitude region to simplify the analysis, and $\overline{P_s}$ is redefined as the average over $\pm 15^{\circ}$ latitude.
-
-$\overline{P_s}$ can be expressed as the double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
-\begin{equation}
-\overline{P_{s}} = \sum_i \sum_j f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right) \label{eq:pdecomp}
-\end{equation}
-where the subscript $850$ is dropped from $\omega$ and $q$ for brevity.
-$f_s$ is a measure of the occurrence of a particular combination $\left( \omega_i , q_j \right)$ in a simulation, and $M_s$ the mean stratiform precipitation rate associated with the combination $\left( \omega_i , q_j \right)$.
-Their product $f_s M_s$ is then the contribution of a particular combination of $\left( \omega_i , q_j \right)$ to $\overline{P_s}$.
-
-Figure~\ref{fig:pdecomp} shows plots of the terms $M_s$ and $f_s M_s$ for all resolutions.
-The plots are computed using 6-hourly instantaneous output of $\omega_{850}$ and $q_{850}$, with 0.05 Pa/s and 0.4 g/kg bin wdiths, respectively.
-Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but still reasonable cut-off for a bins' contribution to $\overline{P_s}$.
-The $M_s$ plots (left column) show that larger magnitude $\omega_{850}$ correspond to larger magnitude stratiform precipitation rates, while the impact of changing $q_{850}$ on $M_s$ is less clear.
-The changes in $M_s$ with resolution are subtle, while the changes in $f_s$ with resolution are large (not shown).
-The changes to $f_s$ can be inferred from the larger space of $\left( \omega_i , q_j \right)$ plotted at higher resolutions, indicating that larger magnitude $\omega_{850}$ values are occurring more frequently at higher resolutions, above the cutoff for plotting.
-The $f_s M_s$ plots (right column) clearly shows that larger magnitude $\omega_{850}$, and therefore larger magnitude stratiform precipitation rates are contributing to the increase in $\overline{P_s}$ with resolution.
-In contrast, smaller magnitude $\omega_{850}$ and hence smaller stratiform precipitation rates contribute less and less to $\overline{P_s}$ with increasing resolution.
-
-\begin{figure}
-\begin{center}
-\noindent\includegraphics[width=20pc,angle=0]{figs/temp_pdecomp.pdf}\\
-\end{center}
-\caption{Decomposition of the climatological stratiform precipitation rates, averaged over the $\pm 15^{\circ}$ latitude band into $\omega_{850}$ and $q_{850}$ environmental conditions.
-Left column shows the time mean magnitude term $M\left( \omega_i , q_j \right)$ and the right column is the magnitude term multiplied by the space-time frequency term $f\left( \omega_i , q_j \right) M\left( \omega_i , q_j \right)$.
-Integrals over $f M$ gives the climatological, area averaged stratiform precipitation rate.
-Panel labels denote the grid resolution of the model run.}
-\label{fig:pdecomp}
-\end{figure}
 

### 2019-09-10 resolution.tex
edits

@@
 \cite{KW1991JGR} and \cite{WETAL1995CD} suggested that the drying of the atmosphere is due to greater magnitude resolved vertical velocities with increasing resolution, with greater subsiding motion increasing the export of dry air from the upper troposphere.
-This mechanism is consistent with an analysis of moisture budgets in CAM, version 4 \citep[CAM4;][]{CAM4} across multiple resolutions \citep{YETAL2014JCLIM,HR2017JCLIM}.
+This mechanism is consistent with an analysis of moisture budgets in CAM, version 4 \citep[CAM4;][]{CAM4} across multiple resolutions \citep{HR2017JCLIM}.
 
@@
 
-Observations show that $h=\frac{1}{3}$ for horizontal scales on the order of $10^2$ km and less \citep[hereafter referred to as the {\em{mesoscale}};][]{L1999JFM,CL2001JGR}.
+Observations show that $h=\frac{1}{3}$ for horizontal scales on the order of $100$ km and less \citep[hereafter referred to as the {\em{mesoscale}};][]{L1999JFM,CL2001JGR}.
 This value for $h$ is also supported by the slope of the kinetic energy spectrum $-\beta$, which can be related to the second-order structure function through the Weiner–Khinchine theorem, $\beta = \left( 2h+1 \right)$, valid in the range $1<\beta<3$ \citep[satisfies stationarity;][]{DETAL1996JAS}.
@@
 It is shown that the resolution sensitivity of vertical velocities are well described with $n=-1$ in equation~\eqref{eq:alpha}, provided $\Delta t_{phys}$ is defined in a way that avoids large truncation errors across resolutions.
-The reduction in convective precipitation rates with resolution in CAM6 is shown to result from the greater magnitude subsiding motion, creating a more stable atmosphere in which the criterion for parameterized convection occurs less often.
-The increase in stratiform precipitation rates with resolution is shown to result more directly from the increase in vertical velocities, by increasing moisture fluxes through cloud base.
+The increase in stratiform precipitation rates with resolution are attributed to an increase in resolved moisture fluxes at cloud base, primarily due to the increase in upward vertical velocities with resolution.
+The corresponding increase in downward vertical velocities with resolution are shown to stabilize the mean state, reducing the activity of the convection scheme and associated precipitation rates with increasing resolution.
 The feedback of the resolved vertical motion on the physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution.
@@
 \end{center}
-\caption{(a,b) Two snapshots of $\omega$ for a longitude-pressure transect at $\sim 18^{\circ}$ latitude in the $ne30$ simulation, overlain by the $0.0075$ kg/m$^2$/s contour of the ZM mass flux (white) delineating the region where the ZM scheme is active, and the 15 K/day contour of the total physics tendencies (black), indicating stratiform cloud formation.}
+\caption{(a,b) Two snapshots of $\omega$ for a longitude-pressure transect at $\sim 18^{\circ}$ latitude in the $ne30$ simulation, overlain by the $0.0075$ kg/m$^2$/s contour of the ZM mass flux delineating the region where the ZM scheme is active.}
 \label{fig:transect}
@@
 As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution.
-Prior studies have established that the drying and reduction in cloudiness with resolution is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}.
+Prior studies have established that the drying and reduction in cloudiness with resolution is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,HR2017JCLIM}.
 This study aims to then understand why and how vertical velocities change with resolution, and why convective precipitation decreases at the expense of stratiform precipitation with increasing resolution.

### 2019-09-11 resolution.tex
done

@@
 By subsequently filtering large-eddy solutions to lower-resolution grids, a relationship between first- and higher-order moments may be understood and ultimately parameterized as a function of grid resolution.
-While this approach is likely necessary for developing scale-aware physics, it is not sufficient.
+While this approach is necessary for developing scale-aware physics, it is not sufficient.
 The equations of motions have inherent scale dependencies, and the properties of dynamical modes are a function of native grid resolution \citep{O1981JAS,WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
@@
 
-The sensitivity of the Community Atmosphere Model \citep[CAM;][]{CAM5}, and its predecessor, the Community Climate Model (CCM) to resolution ({\em{resolution}} refers to {\em{horizontal resolution}}, hereafter) is well documented through convergence studies \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}.
+The sensitivity of the Community Atmosphere Model \citep[CAM;][]{CAM5}, and its predecessor, the Community Climate Model (CCM) to resolution ({\em{resolution}} refers to {\em{horizontal resolution}} hereafter) is well documented through convergence studies \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}.
 CAM/CCM is a fully supported, well-funded climate model, but despite thirty years of continual model development, there are robust sensitivities to resolution that have persisted in all versions of the model.
@@
 The plots are computed using 6-hourly instantaneous output of $\omega_{850}$ and $q_{850}$, with 0.05 Pa/s and 0.4 g/kg bin widths, respectively.
-Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but still reasonable cut-off for a bins' contribution to $\overline{P_s}$.
+Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but reasonable cut-off for a bins' contribution to $\overline{P_s}$.
 The $M_s$ plots (left column) show that larger magnitude $\omega_{850}$ correspond to larger magnitude stratiform precipitation rates, while the impact of changing $q_{850}$ on $M_s$ is less clear.
@@
 
-While the increase in magnitude of vertical velocities with resolution clearly drives the increase in moisture fluxes and stratiform precipitation rates with resolution, it is not clear why there is a reduction in parameterized deep convective precipitation rates with resolution.
+While the increase in magnitude of vertical velocities with resolution clearly drives an increase in moisture fluxes and stratiform precipitation rates with resolution, it is not clear why their is a reduction parameterized deep convective precipitation rates with resolution, in all versions of the model (Figure~\ref{fig:cam-history}).
 Curiously though, there is an excellent negative correlation (Pearson's R-value = 0.99, N = 27) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
@@
 \end{center}
-\caption{Scatter plot of global mean, climatological $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and FREQZM, and the fitted linear regression which has a Pearson's R-value = 0.99, using all 27 simulations.
+\caption{Scatter plot of global mean, climatological $\langle f_{d} \rangle \langle \omega_{d} \rangle$ and $FREQZM$, and the fitted linear regression which has a Pearson's R-value = 0.99, using all 27 simulations.
 Grey crosses are for the 24 member perturbed parameter ensemble runs.}
@@
 Figure~\ref{fig:4zonal}a,b, shows the zonal mean variations in $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and $FREQZM$ in the convergence experiment.
-$FREQZM$ is largest in the $\pm 10^{\circ}$ latitude region, within the ITCZ, and rapidly decreasing polewards into the subtropics.
+$FREQZM$ is largest in the $\pm 10^{\circ}$ latitude region, within the Inter-Tropical Convergence Zone (ITCZ), and rapidly decreasing polewards into the subtropics.
 $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ increases away from the ITCZ region and reaches a maximum at the poleward limit of the Hadley Cell.
@@
 The R-values generally decrease with resolution indicating that there is degradation in the relationship with resolution.
-The logistic regression is less skillful in the subtropics, with skill declining even further in the mid-latitudes (Figure~\ref{fig:4zonal}c).
+The logistic regression is less skillful in the subtropics, with skill declining even further in the mid-latitudes irregardless of resolution (Figure~\ref{fig:4zonal}c).
 
@@
 The increasing occurrence of stable, subsiding grid columns accounts for about half of the changes in dilute CAPE in the deep tropics with resolution, the other half being due to the systematic reduction in dilute CAPE for both ascending and descending regions (Figure~\ref{fig:cape}a).
-This breakdown is straightforward to compute since the sum of the products of the space-time fractions with their associated ascending/descending dilute CAPE values approximate the dilute CAPE computed from the mean temperature/moisture profiles over the entire deep tropics (compare black and grey crosses in Figure~\ref{fig:cape}a) rather well.
+This breakdown is straightforward to compute since the sum of the products of the space-time fractions with their associated ascending/descending dilute CAPE values approximate the dilute CAPE computed from the mean temperature/moisture profiles over the entire deep tropics quite well (compare black and grey crosses in Figure~\ref{fig:cape}a).
 
@@
 Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but with temperature changes playing a larger role (not shown).
-This analysis suggests that increasing subsidence with resolution (\ref{fig:4zonal}a) increases the stability of the mean state, since both ascending/descending regimes become more stable with resolution, and reducing climatological $FREQZM$ (\ref{fig:4zonal}b) despite the poor relationship between instantaneous subsidence and $FREQZM$ in the subtropics, as shown by the logistic regression (Figure~\ref{fig:4zonal}c).
+This analysis suggests that increasing subsidence with resolution (Figure~\ref{fig:4zonal}a) increases the stability of the mean state, since both ascending/descending regimes become more stable with resolution, and reducing climatological $FREQZM$ (Figure~\ref{fig:4zonal}b) despite the poor relationship between instantaneous subsidence and $FREQZM$ in the subtropics, as shown by the logistic regression (Figure~\ref{fig:4zonal}c).
 
@@
 
-\section{Conclusions}
+\section{Discussion and conclusions}
 
-Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of physical parameterizations compatible with the highly flexible grid structures now being used to maximize efficiency on modern day supercomputers.
+Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of physical parameterizations compatible with the highly flexible grid structures now being used to maximize efficiency on modern day supercomputers (scale-aware physics).
 This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}).
 As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution.
-Prior studies have established that the drying and reduction in cloudiness with resolution is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,HR2017JCLIM}.
+Prior studies have established that the drying and reduction in cloudiness with resolution is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advects dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,HR2017JCLIM}.
 This study aims to then understand why and how vertical velocities change with resolution, and why convective precipitation decreases at the expense of stratiform precipitation with increasing resolution.
@@
 The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$.
-The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, which shows pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and the vertical velocities also scale like $D^{-1}$.
-It has been previously shown that $D$ is set by stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, the effective resolution \citep{S2011LNCSE}, and so $D \propto \Delta x$ and $n=-1$.
+The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, indicating that pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and the vertical velocities also scale like $D^{-1}$.
+It has been previously shown that $D$ is set by stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, the effective resolution \citep{S2011LNCSE}, and so $D \propto \Delta x$ giving $n=-1$.
 This scaling is consistent with studies describing the resolution sensitivity of large-eddy simulations, which use similar physical arguments, but adapted for non-hydrostatic scales \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
@@
 This is an intriguing proposal and the authors don't seek to dismiss its potential value.
-But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is not resolved in even our highest resolution simulation $\Delta x=27.8$ km, since CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
+But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km, since CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
 The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
-For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is significantly less than the $\beta=3$ slope observed near the effective resolution in the simulation {\color{red}{(Figure~X)}}.
+For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is significantly shallower than the slope observed near the effective resolution for all grid resolutions in this study (Figure~~\ref{fig:2ke}).
+
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_kespectra.pdf}\\
+\end{center}
+\caption{(a) Kinetic energy spectrum and (b) compensated kinetic energy spectrum at the 200 hPa level in the simulations.
+The diagonal black line in (a) is a reference spectrum with a $-3$ slope.
+Compensated kinetic energy spectrum refers to the kinetic energy spectrum multiplied by $n^{-\frac{5}{3}}$, with $n$ the spherical wavenumber, so that horizontal lines indicate a spectrum with a $-\frac{5}{3}$ slope.
+All resolutions in this study have a slope of the kinetic energy spectrum steeper than $-\frac{5}{3}$ near the effective resolution of the model.}
+\label{fig:2ke}
+\end{figure}
+
+An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, balanced by an increase in stratiform precipitation rates.
+Through decomposing the stratiform precipitation into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved mass fluxes arising from the increase in resolved vertical velocities with resolution.
+At higher resolutions, larger stratiform precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate, with smaller magnitude precipitation rates contributing less.
+The increasing occurrence of larger magnitude vertical velocities and stratiform precipitation rates at higher resolution is at odds with a similar analysis by \cite{TETAL2018CD} using a similar model.
+The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux) which complicated their calculation.
 
@@
 
-In addition to the change in balance of ascending/descending motion in the deep tropics with resolution, temperature and moisture profiles shows a decrease in dilute CAPE for both ascending/descending regimes with increasing resolution.
+In addition to the change in balance of ascending/descending motion in the deep tropics with resolution, temperature and moisture profiles show a decrease in dilute CAPE for both ascending/descending regimes with increasing resolution.
 This reduction in background dilute CAPE is not confined to the deep tropics, but is also evident from the dilute CAPE derived from the time mean temperature and moisture profiles everywhere in the model (not shown) and is in large part the reason for the global mean reduction in activity in the ZM scheme with resolution.
 A sensitivity analysis of the dilute CAPE calculation indicates that this change in background stability is primarily a result of temperature changes with resolution.
-Temperature profiles are more stable with increasing resolution, likely due to an increase in adiabatic warming from increasing subsidence with resolution.
+Temperature profiles become more stable with increasing resolution, likely due to an increase in adiabatic warming due to the increase in subsidence with resolution.
 
@@
 As with the ZM scheme used in this study, dilute CAPE must exceed a threshold for convection to occur, but \cite{XETAL2019JAMES} also require DCAPE to be positive.
-Based on the analysis in this study, and the strong connection between vertical velocities and DCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that DCAPE is often negative in drizzling regions, and would prevent the ZM scheme from triggering.
-The authors suggest that the \cite{XETAL2019JAMES} trigger be explored as a means to mitigate the drizzling bias in CAM.
+Based on the analysis in this study, and the strong connection between vertical velocities and DCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that DCAPE is often negative in drizzling regions, and the \cite{XETAL2019JAMES} trigger would not allow for convection to occur.
+The authors suggest that this trigger be explored as a means to mitigate the drizzling bias in future versions of CAM.
 
-The stratiform scheme becomes more active with resolution, compensating for the reduction in activity of the ZM scheme with resolution, particularly as it relates to the Intertropical Convergence Zone (ITCZ).
-Energetically, it may expected that stratiform precipitation rates increase with resolution to compensate for the reduction in convection from the ZM scheme in order to balance radiative cooling.
-But this argument requires one to assume the increase in stratiform precipitation is responding to a reduction in ZM activity, whereas this study indicates the reverse, that the ZM scheme is responding to increases in subsidence with resolution, driven by greater ascent from stratiform cloud formation.
-The authors instead sought to understand the increase in stratiform precipitation with resolution at the process level.
-It is shown that resolved upward moisture fluxes are highly correlated with stratiform precipitation rates, and that there is an increase in both upward moisture fluxes and stratiform precipitation with resolution.
-Through decomposing the stratiform precipitation in the ITCZ region into the components of the moisture flux, it is shown that larger magnitude vertical velocities are associated with larger stratiform precipitation rates, and these large precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate with resolution.
-
-{\color{red}{Wrap up with a closing paragraph}}
+This study argues that the sensitivity of parameterized convection and stratiform precipitation rates to resolution are a response to the increase in magnitude of vertical velocities with resolution.
+This is essentially the hypothesis put forth by \cite{HR2017JCLIM} to explain resolution sensitivity, that increased vertical velocities with resolution interacts with other model components, steering the model towards a new mean state.
+Here, those interactions are documented in detail.
+Resolution sensitivity therefore arises from the relationship between vertical velocities and resolution, a well-known scale dependency of the equations of motion \citep{O1981JAS}.
+This scale dependency cannot be ignored in the pursuit of scale-aware physics.
 

### 2019-09-11 resolution.tex
edits

@@
 Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
-Moisture changes with resolution has a smaller influence, only appreciably impacting the dilute CAPE of ascending regions at higher resolutions (not shown).
+Moisture changes with resolution do not contribute to the reduction in dilute CAPE with resolution, with the exception that the 20 J/kg reduction in dilute CAPE of ascending regions between the $ne80$ and $ne120$ runs (Figure~\ref{fig:cape}a) is entirely due to changes in moisture (not shown).
 
@@
 This is essentially the hypothesis put forth by \cite{HR2017JCLIM} to explain resolution sensitivity, that increased vertical velocities with resolution interacts with other model components, steering the model towards a new mean state.
-Here, those interactions are documented in detail.
+Here, those interactions are documented explicitly.
 Resolution sensitivity therefore arises from the relationship between vertical velocities and resolution, a well-known scale dependency of the equations of motion \citep{O1981JAS}.

### 2019-09-14 resolution.tex
edits

@@
 
-\title{Parameterized convection, stratiform clouds and resolution sensitivity in the Community Atmosphere Model}
+\title{On resolution sensitivity in the Community Atmosphere Model}
 
@@
 This value for $h$ is also supported by the slope of the kinetic energy spectrum $-\beta$, which can be related to the second-order structure function through the Weiner–Khinchine theorem, $\beta = \left( 2h+1 \right)$, valid in the range $1<\beta<3$ \citep[satisfies stationarity;][]{DETAL1996JAS}.
-For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is true for the kinetic energy spectrum in observations \citep{NG1985JAS,CETAL1999JGR} and models \citep[e.g.,][]{THO2006GRL,SPKS2014JAS} at the mesoscale.
+For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is true for the kinetic energy spectrum in observations \citep{NG1985JAS,CETAL1999JGR} and models at the mesoscale \citep[e.g.,][]{THO2006GRL,SPKS2014JAS}.
 \cite{RETAL2016CD} propose that the overwhelming support for $h=\frac{1}{3}$ at the mesoscale provides an emergent constraint for $n=-\frac{2}{3}$ in equation~\ref{eq:alpha}.
@@
 Another robust response of the CAM/CCM lineage to resolution is an increase in stratiform precipitation rates at the expense of parameterized convective precipitation rates.
+Stratiform precipitation is sometimes referred to as grid-scale precipitation because unlike parameterized convection, it condenses moisture locally without any transport by unresolved eddies.
 The resolution dependent partitioning between the two different precipitation routines is shown in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies.
-The tendency for precipitation rates to shift from the convection scheme to the stratiform scheme with resolution has been documented in other models \citep{PS2002CD,RETAL2016CD,TETAL2018CD}, but none have provided a satisfactory explanation for this sensitivity.
+The tendency for precipitation rates to shift from the convection scheme to the stratiform scheme with resolution has been documented in other models \citep{PS2002CD,RETAL2016CD,TETAL2018CD}, but none have provided a convincing explanation for this sensitivity.
 The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the practice of reducing $\Delta t_{phys}$ with resolution should by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convective precipitation rates with resolution.
@@
 \caption{Bar-graph of the convective (solid) and stratiform (white) climatological precipitation rates in prior CAM/CCM convergence studies.
-Each window contains a single convergence study, with identical x-axis; the approximate grid resolution.
+Each window contains a single convergence experiment, with identical x-axis; the approximate grid resolution.
 Colors indicate the model configuration; January ensemble (black) and aqua-planet configurations with SST profiles $QOBS$ (blue) and $CNTL$ (red) in \cite{NH2000ASL}.
@@
 The corresponding increase in downward vertical velocities with resolution are shown to stabilize the mean state, reducing the activity of the convection scheme and associated precipitation rates with increasing resolution.
-The feedback of the resolved vertical motion on the physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution.
+The feedback of the resolved vertical motion on the model physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution.
 Section 2 describes CAM6 and the details of the convergence experiment.
@@
    Total Precipitable Water (mm) & 23.31& 23.01 & 22.62 & 22.25 & 21.93 & 21.72 \\
-   Convective Precipitation (mm/day) & 1.91 & 1.83 & 1.68 & 1.47 & 1.29 & 1.08 \\
-   Stratiform Precipitation (mm/day) & 1.26 & 1.42 & 1.60 & 1.85 & 2.05 & 2.22 \\      
+   Convective Precipitation, $P_c$ (mm/day) & 1.91 & 1.83 & 1.68 & 1.47 & 1.29 & 1.08 \\
+   Stratiform Precipitation, $P_s$ (mm/day) & 1.26 & 1.42 & 1.60 & 1.85 & 2.05 & 2.22 \\      
  \hline
@@
 The dry dynamics are solved using the high-order, momentum, mass and energy conserving spectral element method \citep{TF2010JCP}, with the elements defined by a cubed-sphere grid.
-The notation for the horizontal grid resolution is an `$ne$' followed by the number of elements making up an edge of one cubed-sphere face, e.g., $ne30$.
-Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds \citep{LetAl2018JAMES}.
+The notation for horizontal grid resolution is an `$ne$' followed by the number of elements making up an edge of one cubed-sphere face, e.g., $ne30$.
+Hyper-viscous $\nabla^{4}$ damping is applied to temperature, dry pressure thickness, rotational and divergent winds \citep{LetAl2018JAMES}.
 CSLAM tracer transport uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.
@@
 Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
-In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing resolution sensitivity in this study.
-All analyses exclude the first month of the simulations, and are computed on their native grids unless otherwise stated.
+In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing the resolution sensitivity in this study.
+All analyses exclude the first month of the simulations, which is sufficient to avoid the spin-up period; the model comes into global energy balance after about ten days (not shown).
+All calculations are performed on their native grids unless otherwise stated.
 
@@
 
-The PDF's may be scaled to the highest-resolution resolution grid through $ \alpha P (\omega / \alpha)$, where $P(x)$ is the PDF, and $\alpha$ the scale factor from equation~\ref{eq:alpha} with $\Delta x_0$ set to the $ne120$ grid-spacing.
+The PDF's may be scaled to the highest-resolution, $ne120$ grid through $ \alpha P (\omega / \alpha)$, where $P(x)$ is the PDF and $\alpha$ the scale factor from equation~\ref{eq:alpha} with $\Delta x_0$ set to the $ne120$ grid-spacing.
 Figure~\ref{fig:2pdf}b shows the scaled PDF's for a power law exponent $n=-1$ in $\Delta x$.
@@
 \end{equation}
-where $\langle f_x \rangle$ and $\langle \omega_x \rangle$ refers to the vertical mass fraction $ \left( \frac{\int dp_x}{\int dp} \right)$ and the $x$ component of the mass weighted vertical mean of $\omega$ $ \left( \frac{\int \omega_x dp_x}{\int dp_x} \right)$, respectively, subscript $u$ refers to upward motion and $d$, downward motion.
+where $\langle f_x \rangle$ and $\langle \omega_x \rangle$ refers to the vertical mass fraction $ \left( \frac{\int dp_x}{\int dp} \textrm{, with pressure } p \right)$ and the $x$ component of the mass weighted vertical mean of $\omega$ $ \left( \frac{\int \omega_x dp_x}{\int dp_x} \right)$, respectively, subscript $u$ refers to upward motion and $d$, downward motion.
 
@@
  \begin{table}
- \caption{Fractional contribution of latitude bands $\pm 10^{\circ}$ and $\pm 15^{\circ}$ to changes in global mean precipitation with resolution.
+ \caption{Fractional contribution of latitude bands $\pm 10^{\circ}$ and $\pm 15^{\circ}$ to changes in global mean stratiform and parameterized convective precipitation with resolution.
 The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc...
@@
    $\pm 10^{\circ}$ ($17.6\%$ of global area) \\
-   Convective Precipitation & -0.58 & 0.62 & 0.66 & 0.72 & 0.70 \\
-   Stratiform Precipitation & 0.55 & 0.63 & 0.69 & 0.67 & 0.41 \\ 
+   Convective Precipitation, $P_c$ & -0.58 & 0.62 & 0.66 & 0.72 & 0.70 \\
+   Stratiform Precipitation, $P_s$ & 0.55 & 0.63 & 0.69 & 0.67 & 0.41 \\ 
    \hline
    $\pm 15^{\circ}$ ($25.8\%$ of global area) \\
-   Convective Precipitation & 0.22 & 0.75 & 0.73 & 0.79 & 0.72 \\
-   Stratiform Precipitation & 0.46 & 0.64 & 0.71 & 0.70 & 0.49 \\      
+   Convective Precipitation, $P_c$ & 0.22 & 0.75 & 0.73 & 0.79 & 0.72 \\
+   Stratiform Precipitation, $P_s$ & 0.46 & 0.64 & 0.71 & 0.70 & 0.49 \\      
  \hline
@@
 $f_s$ is a measure of the occurrence of a particular combination $\left( \omega_i , q_j \right)$ in a simulation, and $M_s$ the mean stratiform precipitation rate associated with the combination $\left( \omega_i , q_j \right)$.
-Their product $f_s M_s$ is then the contribution of a particular combination of $\left( \omega_i , q_j \right)$ to $\overline{P_s}$.
+Their product $f_s M_s$ is then the contribution of the combination $\left( \omega_i , q_j \right)$ to $\overline{P_s}$.
 
@@
 
-While the increase in magnitude of vertical velocities with resolution clearly drives an increase in moisture fluxes and stratiform precipitation rates with resolution, it is not clear why their is a reduction parameterized deep convective precipitation rates with resolution, in all versions of the model (Figure~\ref{fig:cam-history}).
+While the increase in magnitude of vertical velocities with resolution clearly drives an increase in moisture fluxes and stratiform precipitation rates with resolution, it is not clear why there is a reduction in parameterized deep convective precipitation rates with resolution, in all versions of the model (Figure~\ref{fig:cam-history}).
 Curiously though, there is an excellent negative correlation (Pearson's R-value = 0.99, N = 27) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}).
@@
 To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then through fixing only the temperature to the $ne20$ profile.
-Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
+Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution, with moisture fixed to the $ne20$ profile, on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
 Moisture changes with resolution do not contribute to the reduction in dilute CAPE with resolution, with the exception that the 20 J/kg reduction in dilute CAPE of ascending regions between the $ne80$ and $ne120$ runs (Figure~\ref{fig:cape}a) is entirely due to changes in moisture (not shown).
@@
 The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), and their dilute CAPE  values vary much less with resolution ($\sim 20$ J/kg, compared with $\sim 80$ J/kg across all resolutions in the deep tropics).
-Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but with temperature changes playing a larger role (not shown).
+Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but as in the deep tropics, temperature changes have a larger influence (not shown).
 This analysis suggests that increasing subsidence with resolution (Figure~\ref{fig:4zonal}a) increases the stability of the mean state, since both ascending/descending regimes become more stable with resolution, and reducing climatological $FREQZM$ (Figure~\ref{fig:4zonal}b) despite the poor relationship between instantaneous subsidence and $FREQZM$ in the subtropics, as shown by the logistic regression (Figure~\ref{fig:4zonal}c).
@@
 
-Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of physical parameterizations compatible with the highly flexible grid structures now being used to maximize efficiency on modern day supercomputers (scale-aware physics).
+Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of physical parameterizations compatible with the highly flexible grid structures now being used to maximize efficiency on modern day supercomputers (i.e., scale-aware physics).
 This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}).
@@
 This is an intriguing proposal and the authors don't seek to dismiss its potential value.
-But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km, since CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
+But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km ($ne120$), and CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
 The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
@@
 
-An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, balanced by an increase in stratiform precipitation rates.
+An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, and balanced by an increase in stratiform precipitation rates.
 Through decomposing the stratiform precipitation into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved mass fluxes arising from the increase in resolved vertical velocities with resolution.
@@
 The increasing occurrence of larger magnitude vertical velocities and stratiform precipitation rates at higher resolution is at odds with a similar analysis by \cite{TETAL2018CD} using a similar model.
-The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux) which complicated their calculation.
+The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux) and complicated their calculation.
 
@@
 Song, personal communication).
-The magnitude of negative, or downward $w$ increases everywhere in the model with increasing resolution, opposing the growth of dilute CAPE and reducing the activity of the ZM scheme with increasing resolution.
+The magnitude of negative, or downward, $w$ increases everywhere in the model with increasing resolution, opposing the growth of dilute CAPE and reducing the activity of the ZM scheme with increasing resolution.
 
@@
 This is likely due to the the large, global maximum in surface latent heat fluxes in the subtropics, which increases dilute CAPE above the threshold for convection, but which is restricted from growing much larger due to substantial subsiding motion aloft.
-These opposing influences result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing copious amounts of light precipitation or drizzle, irrespective of resolution.
+These opposing influences result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing light precipitation or drizzle, irrespective of resolution.
 
@@
 Resolution sensitivity therefore arises from the relationship between vertical velocities and resolution, a well-known scale dependency of the equations of motion \citep{O1981JAS}.
-This scale dependency cannot be ignored in the pursuit of scale-aware physics.
+This scale dependency needs to be addressed in the pursuit of scale-aware physics.
 

### 2019-09-15 resolution.tex
edits

@@
 The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
-For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is significantly shallower than the slope observed near the effective resolution for all grid resolutions in this study (Figure~~\ref{fig:2ke}).
+For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is shallower than the slope observed near the effective resolution for all grid resolutions in this study (Figure~~\ref{fig:2ke}).
+There is however, a progressive shoaling of the slope $-\beta$ with increasing resolution (Figure~~\ref{fig:2ke}b).
+This suggests a transition towards the $-\frac{5}{3}$ slope in the model, but seeing this process through would require running the model at higher resolutions than considered in this study.
 

### 2019-09-15 resolution.tex
edits

@@
 But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km ($ne120$), and CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
-The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
-For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is shallower than the slope observed near the effective resolution for all grid resolutions in this study (Figure~~\ref{fig:2ke}).
+The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of the kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
+For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is shallower than the slope near where the spectrum approaches the effective resolution, for all grid resolutions (Figure~~\ref{fig:2ke}).
 There is however, a progressive shoaling of the slope $-\beta$ with increasing resolution (Figure~~\ref{fig:2ke}b).
@@
 The diagonal black line in (a) is a reference spectrum with a $-3$ slope.
-Compensated kinetic energy spectrum refers to the kinetic energy spectrum multiplied by $n^{-\frac{5}{3}}$, with $n$ the spherical wavenumber, so that horizontal lines indicate a spectrum with a $-\frac{5}{3}$ slope.
-All resolutions in this study have a slope of the kinetic energy spectrum steeper than $-\frac{5}{3}$ near the effective resolution of the model.}
+Compensated kinetic energy spectrum refers to the kinetic energy spectrum multiplied by $n^{-\frac{5}{3}}$, with spherical wavenumber $n$, so that horizontal lines indicate a spectrum with a $-\frac{5}{3}$ slope.
+In (b), thick ``caterpillars" are overlain on the portion of each spectrum pertaining to the $5-10 \Delta x$ effective resolution of the model.}
 \label{fig:2ke}

### 2019-09-15 resolution.tex
edits

@@
 The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$.
-The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, indicating that pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and the vertical velocities also scale like $D^{-1}$.
+The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, with pressure gradients scaling like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and vertical velocities that also scale like $D^{-1}$.
 It has been previously shown that $D$ is set by stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, the effective resolution \citep{S2011LNCSE}, and so $D \propto \Delta x$ giving $n=-1$.
@@
 The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind.
-\cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations, it is an emergent constraint for $n=-\frac{2}{3}$.
+\cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations, $n=-\frac{2}{3}$ is an emergent constraint on models.
 This is an intriguing proposal and the authors don't seek to dismiss its potential value.
@@
 The diagonal black line in (a) is a reference spectrum with a $-3$ slope.
-Compensated kinetic energy spectrum refers to the kinetic energy spectrum multiplied by $n^{-\frac{5}{3}}$, with spherical wavenumber $n$, so that horizontal lines indicate a spectrum with a $-\frac{5}{3}$ slope.
-In (b), thick ``caterpillars" are overlain on the portion of each spectrum pertaining to the $5-10 \Delta x$ effective resolution of the model.}
+Compensated kinetic energy spectrum refers to the kinetic energy spectrum multiplied by $k^{-\frac{5}{3}}$, with spherical wavenumber $k$, so that horizontal lines indicate a spectrum with a $-\frac{5}{3}$ slope.
+In (b), solid circles are overlain on each spectrum pertaining to $10 \Delta x$, a conservative estimate for the effective resolution of the model.}
 \label{fig:2ke}
@@
 The increasing occurrence of larger magnitude vertical velocities and stratiform precipitation rates at higher resolution is at odds with a similar analysis by \cite{TETAL2018CD} using a similar model.
-The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux) and complicated their calculation.
+The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux), complicating the analysis.
 
@@
 Resolution sensitivity therefore arises from the relationship between vertical velocities and resolution, a well-known scale dependency of the equations of motion \citep{O1981JAS}.
-This scale dependency needs to be addressed in the pursuit of scale-aware physics.
 

### 2019-09-16 resolution.tex
edits

@@
 This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}).
-As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution.
+As with just about every version of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution.
 Prior studies have established that the drying and reduction in cloudiness with resolution is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advects dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,HR2017JCLIM}.
@@
 The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$.
-The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, with pressure gradients scaling like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and vertical velocities that also scale like $D^{-1}$.
+The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, with pressure gradients scaling like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and bringing about vertical velocities that also scale like $D^{-1}$.
 It has been previously shown that $D$ is set by stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, the effective resolution \citep{S2011LNCSE}, and so $D \propto \Delta x$ giving $n=-1$.
@@
 The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind.
-\cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations, $n=-\frac{2}{3}$ is an emergent constraint on models.
+\cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations and models at the mesoscale, $n=-\frac{2}{3}$ is an emergent constraint on models.
 This is an intriguing proposal and the authors don't seek to dismiss its potential value.

### 2019-09-16 resolution.tex
edits

@@
 \begin{abstract}
-This paper describes...
+
 \end{abstract}
@@
 CAM/CCM is a fully supported, well-funded climate model, but despite thirty years of continual model development, there are robust sensitivities to resolution that have persisted in all versions of the model.
-This study argues that a unifying cause, the inherent scale sensitivities of the underlying dynamical equations, can explain the robust responses to resolution that occur in CAM/CCM.
+This study argues that a unifying cause, the inherent scale sensitivities of the underlying dynamical equations, can explain these robust responses to resolution that occur in CAM/CCM.
 
@@
 Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
-In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing the resolution sensitivity in this study.
+In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing resolution sensitivity in this study.
 All analyses exclude the first month of the simulations, which is sufficient to avoid the spin-up period; the model comes into global energy balance after about ten days (not shown).
@@
 The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law exponent $n=-1$ explains to first-order the variation in vertical velocity with resolution as shown by the aqua-planet simulations.
+In contrast, the $n=-\frac{2}{3}$ slope proposed in \cite{RETAL2016CD} greatly underestimates the increasing occurrence of large magnitude vertical velocities with resolution (Supplementary Figure 1).
 
@@
 There is however, a progressive shoaling of the slope $-\beta$ with increasing resolution (Figure~~\ref{fig:2ke}b).
-This suggests a transition towards the $-\frac{5}{3}$ slope in the model, but seeing this process through would require running the model at higher resolutions than considered in this study.
+This suggests a transition towards the $-\frac{5}{3}$ slope in the simulations, but seeing this process through would require running the model at higher resolutions than considered in this study.
 
@@
 \caption{(a) Kinetic energy spectrum and (b) compensated kinetic energy spectrum at the 200 hPa level in the simulations.
-The diagonal black line in (a) is a reference spectrum with a $-3$ slope.
-Compensated kinetic energy spectrum refers to the kinetic energy spectrum multiplied by $k^{-\frac{5}{3}}$, with spherical wavenumber $k$, so that horizontal lines indicate a spectrum with a $-\frac{5}{3}$ slope.
-In (b), solid circles are overlain on each spectrum pertaining to $10 \Delta x$, a conservative estimate for the effective resolution of the model.}
+The diagonal black line in (a) is a reference spectrum $k^{-3}$, with spherical wavenumber $k$.
+Compensated kinetic energy spectrum refers to the kinetic energy spectrum multiplied by $k^{-\frac{5}{3}}$, so that horizontal lines indicate a spectrum with a $-\frac{5}{3}$ slope.
+In (b), thick ``caterpillars" are overlain on each spectrum in the $5-10 \Delta x$ range, encompassing the effective resolution of the model.}
 \label{fig:2ke}
@@
 Sultan Hameed for his assistance with the statistical analysis used in this study.
+Model output used in this study is publicly available from the GitHub repository \url{https://github.com/adamrher/2019-cam-resolution/tree/master/data}.
 

### 2019-09-17 resolution.tex
abstract

@@
 \begin{abstract}
-
+Advances in high performance computing make it possible to run atmospheric general circulation models (AGCMs) over an increasingly wider range of grid resolutions, using either globally uniform or variable-resolution grid structures, but which are incompatible with the non- or weakly- converging solutions with increasing horizontal resolution that have long characterized AGCMs.
+In the the Community Atmosphere Model (CAM), there are robust sensitivities to resolution that have persisted since the model was first introduced over thirty years ago; the atmosphere progressively dries and becomes less cloudy with resolution, and parameterized deep convective precipitation decreases at the expense of stratiform precipitation with resolution.
+This study argues that a unifying cause, the sensitivity of resolved dynamical modes to native grid resolution, feedback onto other model components and explain these robust sensitivities to resolution.
+The increasing magnitude of resolved vertical velocities with resolution are shown to fit the analytical scaling derived from the equations of motion at hydrostatics scales.
+This trend in vertical velocities results in an increase in resolved moisture fluxes at cloud base, balanced through an increase in stratiform precipitation rates with resolution.
+Compensating, greater magnitude subsiding motion is shown to stabilize the mean state and reduce the activity of parameterized convection and convective precipitation rates with resolution.
+This increase in subsiding motion with resolution has previously shown to dry out the atmosphere, reducing cloud cover, and bringing full circle an explanation for resolution sensitivity rooted in the inherent scale-sensitivities of the equations of motion.
+The impact of changing vertical velocity field with native grid resolution cannot be ignored in any effort to recover convergent solutions in AGCMs, and in particular, the development of scale-aware physical parameterizations.
 \end{abstract}
@@
 The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law exponent $n=-1$ explains to first-order the variation in vertical velocity with resolution as shown by the aqua-planet simulations.
-In contrast, the $n=-\frac{2}{3}$ slope proposed in \cite{RETAL2016CD} greatly underestimates the increasing occurrence of large magnitude vertical velocities with resolution (Supplementary Figure 1).
+In contrast, the $n=-\frac{2}{3}$ slope proposed in \cite{RETAL2016CD} greatly underestimates the increasing occurrence of large magnitude vertical velocities with resolution (SFigure 1).
 

### 2019-09-17 resolution.tex
ready for submission

--- 2019-09-16 resolution.tex
@@
 \begin{abstract}
-
+Advances in high performance computing make it possible to run atmospheric general circulation models (AGCMs) over an increasingly wider range of grid resolutions, using either globally uniform or variable-resolution grid structures.
+In principle, this is an exciting opportunity to resolve atmospheric process and scales in a global model with unprecedented detail, but in practice this grid flexibility is not obviously compatible with the non- or weakly- converging solutions with increasing horizontal resolution that have long characterized AGCMs.
+In the the Community Atmosphere Model (CAM), there are robust sensitivities to horizontal resolution that have persisted since the model was first introduced over thirty years ago; the atmosphere progressively dries and becomes less cloudy with resolution, and parameterized deep convective precipitation decreases at the expense of stratiform precipitation with resolution.
+This study documents a convergence experiment using CAM, version 6, to argue that a unifying cause, the sensitivity of resolved dynamical modes to native grid resolution, feedback onto other model components and explain these robust sensitivities to resolution.
+The increasing magnitude of resolved vertical velocities with resolution are shown to fit an analytic scaling derived for the equations of motion at hydrostatic scales.
+This trend in vertical velocities results in an increase in resolved moisture fluxes at cloud base, balanced through an increase in stratiform precipitation rates with resolution.
+Compensating, greater magnitude subsiding motion with resolution has previously been shown to dry out the atmosphere and reduce cloud cover, and here it also shown to stabilize the mean state and reduce the activity of parameterized convection and convective precipitation rates with resolution.
+The impact of changing vertical velocity field with native grid resolution cannot be ignored in any effort to recover convergent solutions in AGCMs, and in particular, the development of scale-aware physical parameterizations.
 \end{abstract}

### 2019-09-17 template-for-submission/resolution-submit.tex
edits

--- 2019-09-17 resolution.tex
@@
-% qjrms4doc.tex V1.10, 4 October 2013
+\documentclass[alpha-refs]{wiley-article}
 
-\documentclass[times]{qjrms4}
-\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
-\usepackage{moreverb}
+\usepackage{graphicx}
+\usepackage[space]{grffile}
+\usepackage{latexsym}
+\usepackage{textcomp}
+\usepackage{longtable}
+\usepackage{tabulary}
+\usepackage{booktabs,array,multirow}
+\usepackage{amsfonts,amsmath,amssymb}
+\usepackage{natbib}
+\usepackage{url}
+\usepackage{hyperref}
+\hypersetup{colorlinks=false,pdfborder={0 0 0}}
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\@combinedblfloats{\box\@outputbox}{\unvbox\@outputbox}{}{%
+  \errmessage{\noexpand\@combinedblfloats could not be patched}%
+}%
+\makeatother
+% You can conditionalize code for latexml or normal latex using this.
+\newif\iflatexml\latexmlfalse
+\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}%
 
-%\def\volumeyear{2013}
-\def\volumenumber{00}
+\AtBeginDocument{\DeclareGraphicsExtensions{.pdf,.PDF,.eps,.EPS,.png,.PNG,.tif,.TIF,.jpg,.JPG,.jpeg,.JPEG}}
+
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+
+% Add any additional LaTeX packages and macros here
+\usepackage{siunitx}
+
+%---
+
+\iflatexml
+% Add any LateXML specific commands here
+
+%---
+
+\else
+% The commands below will only change the exported PDF. Edit or remove as needed
+
+\renewcommand{\sfdefault}{ptm}
+\renewcommand{\familydefault}{\sfdefault}
+
+\paperfield{Climate Modeling}
+\corraddress{School of Marine and Atmospheric Sciences, Stony Brook University, Stony Brook, NY 11794, USA}
+\corremail{\url{adam.herrington@stonybrook.edu}}
+\presentadd{Climate and Global Dynamics Laboratory, National Center for Atmospheric Research, Boulder, CO, 80305, USA}
+
+%---
+
+\papertype{Original Article}
+
+\title{On resolution sensitivity in the Community Atmosphere Model}
+
+\author[1]{Adam R. Herrington}
+\affil[1]{School of Marine and Atmospheric Sciences, Stony Brook University, Stony Brook, NY 11794, USA}
+\author[1]{Kevin A. Reed}
+\runningauthor{Adam R. Herrington}
 
@@
 
-\runningheads{A.~R.~Herrington and K.~A.~Reed}{CAM resolution sensitivity}
-
-\title{On resolution sensitivity in the Community Atmosphere Model}
-
-\author{Adam R. Herrington\corrauth \& Kevin A. Reed}
-\address{School of Marine and Atmospheric Sciences, Stony Brook University, Stony Brook, NY 11794}
-
-\corraddr{\url{adam.herrington@stonybrook.edu}}
-
+\maketitle
+\selectlanguage{english}
 \begin{abstract}
@@
 The impact of changing vertical velocity field with native grid resolution cannot be ignored in any effort to recover convergent solutions in AGCMs, and in particular, the development of scale-aware physical parameterizations.
-\end{abstract}
-
-\keywords{Climate models, physical parameterizations, resolution sensitivity}
-
-\maketitle
+\textbf{Keywords} --- Climate models, physical parameterizations, resolution sensitivity%
+\end{abstract}%
 
@@
 \begin{center}
-\noindent\includegraphics[width=20pc,angle=0]{figs/temp_pdecomp.pdf}\\
+\noindent\includegraphics[width=10pc,angle=0]{figs/temp_pdecomp.pdf}\\
 \end{center}
@@
 
-\ack 
+\section{Acknowledgements}\label{acknowledgements}
+
 Funding support for this work was in part provided by the U.S.
@@
 
-\bibliographystyle{wileyqj}
-\bibliography{bib}
+\selectlanguage{english}
+%\FloatBarrier
+\bibliography{bibliography/bib.bib}
+
 \end{document}

### 2020-02-24 resolution.tex
bring up to date

@@
 
-%\def\volumeyear{2013}
 \def\volumenumber{00}
@@
 \begin{abstract}
-Advances in high performance computing make it possible to run atmospheric general circulation models (AGCMs) over an increasingly wider range of grid resolutions, using either globally uniform or variable-resolution grid structures.
-In principle, this is an exciting opportunity to resolve atmospheric process and scales in a global model with unprecedented detail, but in practice this grid flexibility is not obviously compatible with the non- or weakly- converging solutions with increasing horizontal resolution that have long characterized AGCMs.
+Advances in high performance computing make it possible to run atmospheric general circulation models (AGCMs) over an increasingly wider range of grid resolutions, using either globally uniform or variable-resolution grids.
+In principle, this is an exciting opportunity to resolve atmospheric process and scales in a global model and in unprecedented detail, but in practice this grid flexibility is incompatible with the non- or weakly- converging solutions with increasing horizontal resolution that have long characterized AGCMs.
 In the the Community Atmosphere Model (CAM), there are robust sensitivities to horizontal resolution that have persisted since the model was first introduced over thirty years ago; the atmosphere progressively dries and becomes less cloudy with resolution, and parameterized deep convective precipitation decreases at the expense of stratiform precipitation with resolution.
-This study documents a convergence experiment using CAM, version 6, to argue that a unifying cause, the sensitivity of resolved dynamical modes to native grid resolution, feedback onto other model components and explain these robust sensitivities to resolution.
+This study documents a convergence experiment using CAM, version 6, and argue that a unifying cause, the sensitivity of resolved dynamical modes to native grid resolution feedback onto other model components and explain these robust sensitivities to resolution.
 The increasing magnitude of resolved vertical velocities with resolution are shown to fit an analytic scaling derived for the equations of motion at hydrostatic scales.
-This trend in vertical velocities results in an increase in resolved moisture fluxes at cloud base, balanced through an increase in stratiform precipitation rates with resolution.
+This trend in vertical velocities results in an increase in resolved upward moisture fluxes at cloud base, balanced by an increase in stratiform precipitation rates with resolution.
 Compensating, greater magnitude subsiding motion with resolution has previously been shown to dry out the atmosphere and reduce cloud cover, and here it also shown to stabilize the mean state and reduce the activity of parameterized convection and convective precipitation rates with resolution.

### 2020-05-16 resolution-revised.tex
updates

--- 2020-02-24 resolution.tex
@@
 
-The sensitivity of the Community Atmosphere Model \citep[CAM;][]{CAM5}, and its predecessor, the Community Climate Model (CCM) to resolution ({\em{resolution}} refers to {\em{horizontal resolution}} hereafter) is well documented through convergence studies \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}.
+{\color{red}{Conventional physics packages interact with the adiabatic dynamics to produce non- or weakly- converging solutions with increasing resolution ({\em{resolution}} refers to {\em{horizontal resolution}} hereafter).
+Although some characteristics of resolution sensitivity are model dependent, others are common across different AGCMs and indicate a common mechanism.
+For example, increasing resolution lends itself to greater stratiform precipitation and reduced parameterized convective precipitation in the Hadley Centre climate model \citep[HadAM3;][]{PS2002CD}, the Max Planck Institute for Meteorology model \citep[ECHAM5;][]{HETAL2006JC} and the National Center for Atmospheric Research (NCAR) Community Climate Model \citep[CCM2;][see Figure~\ref{fig:cam-history}]{WETAL1995CD}.
+So while the purpose of this study is to provide a process-level understanding of resolution sensitivity in the NCAR models, these results are useful for understanding resolution sensitivity in models that exhibit similar behavior.}}
+
+The sensitivity of the Community Atmosphere Model \citep[CAM;][]{CAM5} and its predecessor CCM, to resolution is well documented through convergence studies \citep{KW1991JGR,WETAL1995CD,W2008TELLUS,RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM}.
 CAM/CCM is a fully supported, well-funded climate model, but despite thirty years of continual model development, there are robust sensitivities to resolution that have persisted in all versions of the model.
@@
 The resolution dependent partitioning between the two different precipitation routines is shown in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies.
-The tendency for precipitation rates to shift from the convection scheme to the stratiform scheme with resolution has been documented in other models \citep{PS2002CD,RETAL2016CD,TETAL2018CD}, but none have provided a convincing explanation for this sensitivity.
+The tendency for precipitation rates to shift from the convection scheme to the stratiform scheme with resolution has been documented in other models \citep{PS2002CD,HETAL2006JC,RETAL2016CD,TETAL2018CD}, but none have provided a convincing explanation for this sensitivity.
 The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the practice of reducing $\Delta t_{phys}$ with resolution should by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convective precipitation rates with resolution.
@@
 Each window contains a single convergence experiment, with identical x-axis; the approximate grid resolution.
-Colors indicate the model configuration; January ensemble (black) and aqua-planet configurations with SST profiles $QOBS$ (blue) and $CNTL$ (red) in \cite{NH2000ASL}.
+Colors indicate the model configuration; January ensemble (black), aqua-planet configurations with SST profiles $QOBS$ (blue) and $CNTL$ (red) in \cite{NH2000ASL} and annual means from an AMIP configuration \citep[green;][]{AMIP1999}.
 Studies included in this figure are \cite{KW1991JGR} (CCM1), \cite{WETAL1995CD} (CCM2), \cite{W2008TELLUS} (CAM3), \cite{RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM} (CAM4), \cite{ZetAl2014JCb} (CAM5) and this study (CAM6).
-CCM2* refers to the modified parameter experiment of \cite{WETAL1995CD}, where parameters vary with resolution to reduce the dependence of cloud fraction on resolution.}
+The asterisk (*) indicates that parameters are modified with resolution to reduce the dependence of cloud fraction and top-of-atmosphere radiative balance on resolution.
+{\color{red}{Two additional studies are included that do no not use the NCAR model, \cite{PS2002CD} (HadAM3) and \cite{HETAL2006JC} (ECHAM5).}}}
 \label{fig:cam-history}
@@
 
-In this study, a convergence experiment using CAM, version 6 (CAM6; \url{https://ncar.github.io/CAM/doc/build/html/users_guide/index.html}) is carried out and analyzed in detail.
+In this study, a convergence experiment using CAM, version 6 (CAM6; \url{https://github.com/ESCOMP/CAM}) is carried out and analyzed in detail.
 It is shown that the resolution sensitivity of vertical velocities are well described with $n=-1$ in equation~\eqref{eq:alpha}, provided $\Delta t_{phys}$ is defined in a way that avoids large truncation errors across resolutions.
@@
 
-\subsection{Experimental Design}
+\subsection{Experimental Design},\label{sec:ed}
 
@@
 This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test \citep[Appendix A in][]{HETAL2019JAMES}, and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS}.
-The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.
+{\color{red}{Due to the large sensitivity of vertical velocity to $\Delta t_{phys}$ \citep{HR2018JAMES}, the lack of scientific consensus on the numerical and physical justifications for how $\Delta t_{phys}$ should vary with resolution remains a severe limitation to characterizing vertical velocities across resolutions in AGCMs.}} The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.
 
@@
 In contrast, the $n=-\frac{2}{3}$ slope proposed in \cite{RETAL2016CD} greatly underestimates the increasing occurrence of large magnitude vertical velocities with resolution (SFigure 1).
+{\color{red}{But the $n=-1$ scaling exponent cannot be interpreted more broadly owing to the large sensitivity of vertical velocity to $\Delta t_{phys}$, and the use of equation~\ref{eq:dt-scale} to determine $\Delta t_{phys}$ in the convergence experiment.}}
 
@@
 
-The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$.
-The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, with pressure gradients scaling like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and bringing about vertical velocities that also scale like $D^{-1}$.
+The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$, {\color{red}{although the sensitivity of vertical velocities to choice of physics time-step across resolutions \citep{HR2018JAMES} precludes a more definitive statement on the generality of the $n=-1$ exponent in CAM.}} The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, with pressure gradients scaling like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and bringing about vertical velocities that also scale like $D^{-1}$.
 It has been previously shown that $D$ is set by stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, the effective resolution \citep{S2011LNCSE}, and so $D \propto \Delta x$ giving $n=-1$.
@@
 
-The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind.
-\cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations and models at the mesoscale, $n=-\frac{2}{3}$ is an emergent constraint on models.
+The $n=-1$ power law scaling is {\color{red}{not necessarily}} in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind.
+\cite{RETAL2016CD} propose that since $h=\frac{1}{3}$ is supported by observations and models at the mesoscale, $n=-\frac{2}{3}$ is an emergent constraint on models.
 This is an intriguing proposal and the authors don't seek to dismiss its potential value.
-But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km ($ne120$), and CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
+But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}.
+This mesoscale regime is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km ($ne120$), since CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}.
 The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of the kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem.
@@
 
-An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, and balanced by an increase in stratiform precipitation rates.
-Through decomposing the stratiform precipitation into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved mass fluxes arising from the increase in resolved vertical velocities with resolution.
+An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, and are balanced by an increase in stratiform precipitation rates.
+Through decomposing the stratiform precipitation into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved mass fluxes / vertical velocities with resolution.
 At higher resolutions, larger stratiform precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate, with smaller magnitude precipitation rates contributing less.
@@
 The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux), complicating the analysis.
+{\color{red}{As it is generally accepted that vertical velocities increase with resolution in most, if not all AGCMs, models other than CAM that show an increases in stratiform precipitation with resolution (HadAM3 and ECHAM5 in Figure~\ref{fig:cam-history}) are likely driven in a similar way by the increase in resolved moisture flux at cloud base.}}
 
@@
 
+{\color{red}{The HadAM3 and ECHAM5 models both show a reduction in convective precipitation rates with increasing resolution (Figure~\ref{fig:cam-history}).
+As with CAM, both models use a stability based closures in their deep convection schemes \citep{GR1990MWR,N1994ECMWF} and both show an increase in tropospheric temperatures with resolution \citep{PS2002CD,RETAL2006JC}.
+In as much as the increase in stability with resolution in these models are driven by greater subsidence warming, the mechanism suppressing parameterized convective activity with resolution in CAM may also explain the resolution sensitivity in HadAM3 and ECHAM5.}}
+
 The subtropics ($\pm (10^{\circ}-30^{\circ})$ latitude) are uniquely less sensitive to subsiding motion.
@@
 These opposing influences result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing light precipitation or drizzle, irrespective of resolution.
-
-AGCMs are known to suffer from an excess of drizzle relative to observations \citep{D2006JCLIM}.
-\cite{XETAL2019JAMES} recently proposed a new convective trigger for the ZM scheme that the authors suggest may alleviate the excess drizzle bias in CAM.
-As with the ZM scheme used in this study, dilute CAPE must exceed a threshold for convection to occur, but \cite{XETAL2019JAMES} also require DCAPE to be positive.
-Based on the analysis in this study, and the strong connection between vertical velocities and DCAPE reported in \cite{SZ2018JCLIM}, it is reasonable to speculate that DCAPE is often negative in drizzling regions, and the \cite{XETAL2019JAMES} trigger would not allow for convection to occur.
-The authors suggest that this trigger be explored as a means to mitigate the drizzling bias in future versions of CAM.
 

### 2020-05-16 resolution-revised.tex
edits

@@
 An increasing number of Atmospheric General Circulation Models (AGCMs) are being developed to maximize efficiency on massively parallel systems, permitting regionally-refined high-resolution, or even globally high-resolution weather ($\Delta x \leq 5$ km) and climate ($\Delta x \leq 50$ km) simulations \citep{SMTMN2008JCP,MPASatm,Z2014QJRMS,HETAL2016JCLIM,DCMIP16,LetAl2018JAMES}.
-These models are built using unstructured meshes that allows for substantial grid flexibility, but lack physical parameterizations ({\em{physics}}) which behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
+These models are built using unstructured meshes that {\color{red}{allow}} for substantial grid flexibility, but lack physical parameterizations ({\em{physics}}) which behave consistently as the truncation scale of the model changes with different grid resolutions, referred to as scale-aware physics.
 The most common approach towards developing scale-aware physics is through the lens of limited area, large-eddy simulations \citep[e.g.,][]{PC2008JAS,AW2013JAS,SZ2018JCLIM}.
@@
 The aqua-planets are in a perpetual equinox, and aerosols are largely absent from the simulations.
-Each simulation is ran for one simulated year.
+Each simulation is {\color{red}{run}} for one simulated year.
 Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}.
-In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing resolution sensitivity in this study.
+In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are {\color{red}{run}} for one year in order to increase confidence in assessing resolution sensitivity in this study.
 All analyses exclude the first month of the simulations, which is sufficient to avoid the spin-up period; the model comes into global energy balance after about ten days (not shown).
@@
 The R-values generally decrease with resolution indicating that there is degradation in the relationship with resolution.
-The logistic regression is less skillful in the subtropics, with skill declining even further in the mid-latitudes irregardless of resolution (Figure~\ref{fig:4zonal}c).
+The logistic regression is less skillful in the subtropics, with skill declining even further in the mid-latitudes irrespective of resolution (Figure~\ref{fig:4zonal}c).
 
@@
 Without a strong dependence of dilute CAPE on ascending/descending regimes, subsidence is not a very skillful predictor of depressing dilute CAPE below the threshold for convection.
-In contrast to the deep tropics, there is also no significant changes to the occurrence of subsiding grid columns with resolution; space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
+In contrast to the deep tropics, there {\color{red}{are}} also no significant changes to the occurrence of subsiding grid columns with resolution; space-time weights of ascending/descending motion are more-or-less invariant with resolution (Figure~\ref{fig:cape-subt}b).
 
@@
 The increasing occurrence of larger magnitude vertical velocities and stratiform precipitation rates at higher resolution is at odds with a similar analysis by \cite{TETAL2018CD} using a similar model.
-The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux), complicating the analysis.
+The authors {\color{red}{suggest}} that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux), complicating the analysis.
 {\color{red}{As it is generally accepted that vertical velocities increase with resolution in most, if not all AGCMs, models other than CAM that show an increases in stratiform precipitation with resolution (HadAM3 and ECHAM5 in Figure~\ref{fig:cam-history}) are likely driven in a similar way by the increase in resolved moisture flux at cloud base.}}

### 2020-05-17 reviews/response-to-reviewers.tex
edits

\documentclass[12pt,oneside,a4paper]{article}%{amsart}
\usepackage[margin=1in]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{caption}
\setlength\parindent{0pt}
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\renewcommand{\bibsection}{}

\title{\large Response to Reviewers}
\author{\normalsize Adam R. Herrington and Kevin. A. Reed}
\date{} % delete this line to display the current date

\begin{document}

\maketitle

\section*{\large General Remarks}

{\color{red}{We would like to thank both reviewers for taking the time to provide thoughtful feedback on our manuscript.}} \newline

{\color{red}{Both reviewers raise a similar concern, that the manuscript lacks discussion on how the results from this single model may extend to other modeling systems. We agree with the reviewers, and in particular reviewer 1, that doing so would generalize the results more broadly, providing the breadth that often characterizes QJRMS articles. We have added a paragraph right after the first paragraph in the introduction, to more clearly frame the problem as common to most AGCMs, and how understanding a single model would shed light on resolution sensitivities in other modeling systems. Figure 1 has been updated to include convergence experiments from two non-NCAR models, HadAM3 \citep{PS2002CD} and ECHAM5 \citep{HETAL2006JC}, and we have discussed whether those model results are consistent with our findings throughout the conclusions (Lines XXX, etc ...).}}

\section*{\large Reviewer 1}

\subsection*{\small Generalizing results to other modeling systems}

The authors should strongly consider expanding the discussion (and perhaps the
meta-analysis shown in Figure 1) to include resolution dependence issues noted
in other modeling systems. There are some other papers in the literature (e.g.,
Volosciuk et al., 2015) that describe resolution dependences in other modeling
systems. Many of the results described in this manuscript seem like they should
generalize to other hydrostatic models, and it would be worth adding some
discussion on whether these results are consistent with what is shown in other
models. \newline

{\color{red}{See General Remarks.}} \newline

Relatedly, I would be curious to hear the authors’ commentary on what these
results imply for how a scale-aware model should be designed. It seems that
now that the resolution dependence (of CAM at least) is understood, a solution
might be possible, and the authors of this manuscript understand this resolution
dependence better than just about anyone. They are therefore well-poised to
comment on possible solutions: especially those that might generalize to other modeling systems. \newline

{\color{red}{Working on this.}}

\subsection*{\small Inappropriate experimental design for a definitive statement on resolution exponent}

Based on my reading, the manuscript states two key messages: (1) vertical velocity
is inversely proportional to horizontal resolution (and therefore is inconsistent
with other theories of resolution-dependence), and (2) dynamic/thermodynamic
changes caused by intensification of vertical velocities can explain the resolution
dependence of stratiform and convective precipitation. The bulk of the
manuscript focuses on (2), and in my opinion, this aspect of the manuscript is
on very strong footing. \newline

However, the first key message seems to be based on interpretation of results from
an inappropriate experimental design. The results of the authors’ experiments
indeed are consistent with a scaling exponent of -1, however the experimental
design does not isolate horizontal resolution from physics timestep. The authors
explicitly state that changing the physics timestep will “lead to greater resolution
sensitivity.” They later state that “the n=-2/3 slope proposed in Rauscher et
al. (2016) greatly underestimates the increasing occurrence of large magnitude
vertical velocities with resolution)” – in other words, the n=-2/3 slope has too
weak of a resolution sensitivity. It is possible that the changes in the physics timestep explain why the resolution-dependence is stronger than predicted by
Rauscher et al. \newline

The authors state that they change the physics timestep in order to “avoid
time-truncation errors”. Based on my understanding of Herrington and Reed
(2018), this seems like a valid consideration for the experimental design. But
why does the physics timestep need to change: why couldn’t it just be fixed
at a value small enough to avoid time-truncation? This would avoid conflating
timestep-dependence with resolution-dependence. \newline

I recommend that the authors either remove their strong assertions that these
results are inconsistent with the Rauscher et al. theory, or I recommend that the
authors perform a set of tests with the physics timestep held fixed at a small
value. \newline

{\color{red}{Based on the reviewers concern, we have amended any strong and/or definitive statements about the -1 exponent due to its dependence on physics time-step (see Lines XXX, XXX and XXX). To the authors knowledge, there is no authoritative work on the appropriate choice of physics time-steps in AGCMs across different resolutions. And so we concede that such definitive statements cannot be made until there is consensus on both the physical and numerical justification for physics time-steps across resolutions. More recent work on physics time-steps \citep[e.g.,][]{W2013QJRMS,WETAL2015JAMES,HR2018JAMES} are not necessarily consistent with one another, and do not form a consensus.}} \newline

{\color{red}{In response to the suggestion of fixing a small physics time-step across all resolutions, we are hesitant to agree that this is a more appropriate design choice \citep[see also Fig. 10 in][]{HR2018JAMES}. In thinking about numerical truncation errors, we are concerned that using the same physics time-step at low resolution as in high resolution, may place the simulations in different error regimes. That is, if you think about the log(physics time-step) vs. log(error) curve, it is likely the curves look different in low and high resolution. At high resolution, the model is resolving faster time-scales and so the errors are expected to be larger for a given physics time-step. But a given physics time-step may also be in a different regime at high resolution, i.e., not converging at the same rate as in the low resolution run.}} \newline

{\color{red}{Figure~\ref{fig:dt-conv} shows an example of how for a fixed physics time-step, low and high resolution can be in different regimes. The slope of the error curves for a fixed physics time-step is shallower for high resolution compared with low resolution, for time-steps larger than about 100 s. That is, high resolution is in a less stable regime than low resolution for a fixed time-step (less stable, not un-stable; the errors still decrease with decreasing time-step at high resolution). Interestingly, neither high or low resolution achieve first order convergence for time steps greater than 100 s, which is not really defensible from a numerics standpoint \citep{WETAL2015JAMES}. Given these numerical uncertainties, and the large sensitivity of vertical velocity to physics time-step, I don't think we can say for certain that fixing the physics time-step to a reasonably small value (say 225 s) is any better of an approach than the approach we have taken. We have softened our language in the manuscript to reflect this uncertainty.}} \newline

{\color{red}{Finally, we would like to clarify our position on one point. In the conclusions, there is discussion about how the  \cite{RETAL2016CD} scaling may not be in conflict with the $n=-1$ exponent. The paragraph that begins with ``[t]he $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}," has been changed to ``[t]he $n=-1$ power law scaling is not necessarily in conflict with the scaling proposed by \cite{RETAL2016CD}." The wording was changed to more clearly convey what this paragraph actually concludes. Which is, that we do not observe a -5/3 slope of the kinetic energy spectrum near the effective resolution in our simulations. And so we argue that the \cite{RETAL2016CD} may be more appropriate for resolutions greater than 1/4 degree, at least as it pertains to our simulations. }} \newline

{\color{red}{We feel our characterization in the introduction ``...only recently has the vertical velocity field in AGCMs and their sensitivity to resolution received attention (Donner et al. 2016; O’Brien et al. 2016), albeit with seemingly conflicting explanations (Rauscher et al. 2016; Herrington and Reed 2018)" is still accurate, since it is only later in the conclusions that we delve into why the two explanations are not in conflict with one another.}}

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=20pc,angle=0]{temp_dt-convergence.pdf}\\
\end{center}
\caption{\small {\color{red}{I-norms, $\sqrt{(\omega_{min} - \omega_{min,0})^2}$, where $\omega_{min,0}$ in the reference solution taken as the 1 s physics time-step simulation after \cite{WETAL2015JAMES}, in a moist bubble configuration \citep{HR2018JAMES} with Kessler microphysics. $\omega_{min}$ refers to the minimum vertical pressure velocity over the entirety of the 1 day simulations. Blue markers refers to a grid resolution with an average equatorial grid spacing of 111.2 km and the red markers, 27.8 km. Dashed lines are lines with slopes of 1 and 2, indicating first-order and second-order convergence, respectively.}}}
\label{fig:dt-conv}
\end{figure}

\subsection*{\small Issues associated with figures}

Figure 5: what do the colors represent? The colorbars should have labels. Based
on the caption, it seems like they should represent precipitation rates. \newline

{\color{red}{Yes, it should be precipitation rate (mm/day). We have added this title to the figure.}} \newline

Relatedly, the associated text (pg 5, col 1, lines 56-57) state that “changes in M\_s
with resolution are subtle, while the changes in f\_s with resolution are large (not
shown).” Why show M\_s if the f\_s term dominates the resolution dependence?
And if space is a concern, why not put the f\_s figures in supplementary material? \newline

{\color{red}{Part of the reason we don't show f\_s is space, yes. We also thought it was more intuitive to see that larger magnitude precipitation rates are contributing to the global mean, by subsequently unmasking larger regions of M\_s that exceed a frequency tolerance. But we understand that readers may be curious about what f\_s looks like and so we have included it as a supplementary material as the reviewer suggests. We decided to plot both log(M\_s) and log(f\_s ) and extended the x-axis so that all the data are shown.}}

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=20pc,angle=0]{../figs/temp_pdecomp_supp.pdf}\\
\end{center}
\caption{\small}
\label{fig:supp}
\end{figure}

\subsection*{\small Minor issues}

• pg 1, col 1, line 55: ‘that allows’ $->$ ‘that allow’ \newline
• pg 3, col 2, line 16: ‘is ran for’ $->$ ‘is run for’ \newline
• pg 3, col 2, line 21: ‘are ran for’ $->$ ‘are run for’ \newline
• pg 5, col 2, line 63: ‘irregardless’ $->$ ‘regardless’ or ‘irrespective’ (see
https://www.businessinsider.com/irregardless-real-word-regardless-korystamper-
education-dictionary-mean-girls-lexicon-merriam-webster-2017-
6) \newline
• pg 9, col 1, line 49: ‘is also no significant changes’ $->$ ‘are also no signficant
changes’ \newline
• pg 9, col 2, line 26: ‘the authors suggests’ $->$ ‘the authors suggest’ \newline

{\color{red}{Thank you for pointing out these grammatical errors, they have been fixed. Irregardless flew right off my tongue and onto paper ... h/t for catching that. }} \newline

• pg 9, col 1, line 58: ‘which often occurs with appreciable subsiding motion
aloft’. . . is this statement justified? On my examination of Figure 11, it
looks like just as many of the the ZM-active contours have large areas of
green (omega ~ -0.1 Pa/s) directly above as have blue (positive omega)
directly above. Take the region in Fig 11b, between 30W and 0, for example:
there is very little subsidence above the ZM convection. In case my eyes
are having a difficult time interpreting this graph, it might be useful to
draw the contour of constant omega=0. \newline

\section*{\large Reviewer 2}

\subsection*{\small Minor comments}

1. Examining positive or negative omega at the grid box level, rather than at the column level \newline

The authors use <fx><omegax> to estimate column-mean omega for ascending or descending motions within a column. Therefore, my understanding is that there are many cases where a grid column has both <fd><omegad> and <fu><omegau> values. \newline

With this in mind, it was hard to make sense of phrases such as ‘greater magnitude subsiding motion’ in the abstract (L35). I could not decide whether the phrase meant that free tropospheric subsidence rates (such as omega850) tend to be stronger or whether there are larger areas of subsidence, until I realized that this is not actually captured by the <fd><omegad> decomposition. This decomposition also led to a few points of confusion in the latter half of the study where comparisons of subsidence were made with columnar values, such as FREQZM and dilute-CAPE, or with temperature profiles, as in Fig. 9. In these cases, I assumed that the column mean omega was used, but it left me wondering why the separation between downward motions and upward motions at the grid cell level (as opposed to the column level) was necessary in the first place. \newline

Given the additional level of complexity that is added by thinking in terms of the <fd><omegad> and <fu><omegau> of each column, would the authors please provide more text explaining the advantages of the decomposition and why this decomposition was chosen, rather than taking either the omega at a particular pressure level or the column mean omega <omega> to capture upward and downward motions? \newline

{\color{red}{Working on this.}} \newline

2. Importance of global mean precipitation rate (total condensation) on atmospheric stability \newline

The abstract currently reads as if the increase in stabilization with resolution can wholly be attributed to the greater subsidence, but as Fig. 8 shows, ascending areas also exhibit a change in temperature that leads to lower dilute-CAPE. I noticed that the global mean precipitation rate increases with resolution across the simulations (except for the change from ne80 to ne120). Could the increased condensational heating from a greater precipitation rate explain the increase in stability and decrease in Dilute-CAPE that is unaccounted for by changes in the subsidence fraction? Over the deep Tropics, I expect the subsidence rates to adjust to the temperature profiles set up by the condensational heating and radiative cooling, rather than the subsidence rates setting the temperature profiles. The authors hint at this partitioning and importance of the mean state temperature profile on Page 7, second full paragraph of second column, where they note that the change in subsidence area explains about half of the decrease in the average dilute-CAPE, but its importance in reducing dilute-CAPE is not further discussed in the conclusions or mentioned in the abstract. \newline

If the output is available, the condensational heating rates and their differences across the models might help distinguish the role of greater condensation on atmospheric temperature profiles. \newline

Related to this point, perhaps the authors can explain why we would expect an increase in precipitation rate as a result of increases in resolved precipitation rate. In other words, do we expect resolved precipitation to be more efficient at removing water than parameterized convective precipitation? If so, this should be discussed. \newline

{\color{red}{Working on this.}} \newline

3. Comparison with previous studies that used other models \newline

Would the authors comment on how general their study is to other climate models? For example, as the authors note, Pope and Stratton (2002) also see a shift of precipitation from convective precipitation to large-scale precipitation. Do results from Pope and Stratton (2002) also show changes in the circulation, temperature, and humidity that are consistent with the mechanism described in this study? \newline

{\color{red}{See General Remarks.}}

\subsection*{\small Specific comments/Typos}

Page 3 ~L58 of first column: ‘macrophyiscs’ $->$ ‘macrophysics’ \newline

{\color{red}{Fixed typo.}}

\section*{\normalsize References}
\setlength{\bibsep}{0pt}
{\footnotesize
\bibliography{bib}}

\end{document}

### 2020-05-17 resolution-revised.tex
edits

@@
 The Cloud Layers Unified by Binormals \citep[CLUBB][]{GETAL2002JAS,BOG2013JCLIM} is an assumed filtered density function \citep{G1992JFM} high-order closure model that handles shallow convection, planetary boundary layer mixing and cloud macrophysics.
-The macrophyiscs are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with the three mode Modular Aerosol Model \citep{MAM}.
+The {\color{red}{macrophysics}} are coupled with a two-moment bulk cloud microphysics scheme with prognostic precipitation \citep{MG2}, and microphysics are coupled with the three mode Modular Aerosol Model \citep{MAM}.
 The combined macrophysics/microphysics routines generate stratiform clouds and stratiform precipitation.
@@
 The plots are computed using 6-hourly instantaneous output of $\omega_{850}$ and $q_{850}$, with 0.05 Pa/s and 0.4 g/kg bin widths, respectively.
-Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but reasonable cut-off for a bins' contribution to $\overline{P_s}$.
+Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but reasonable cut-off for a bins' contribution to $\overline{P_s}$ ({\color{red}{all unmasked values of $M_s$ and $f_s$ are shown in the  supplementary materials.}}).
 The $M_s$ plots (left column) show that larger magnitude $\omega_{850}$ correspond to larger magnitude stratiform precipitation rates, while the impact of changing $q_{850}$ on $M_s$ is less clear.
@@
 Integrals over $f M$ gives the climatological, area averaged stratiform precipitation rate.
-Panel labels denote the grid resolution of the model run.}
+Panel labels denote the grid resolution of the model run.
+{\color{red}{See supplementary materials for all the unmasked values of both $M_s$ and $f_s$.}}}
 \label{fig:pdecomp}

### 2020-06-19 reviews/response-to-reviewers.tex
updates

@@
 We agree with the reviewers, and in particular reviewer 1, that doing so would generalize the results more broadly, providing the breadth that often characterizes QJRMS articles.
-We have added a paragraph right after the first paragraph in the introduction, to more clearly frame the problem as common to most AGCMs, and how understanding a single model would shed light on resolution sensitivities in other modeling systems.
-Figure 1 has been updated to include convergence experiments from two non-NCAR models, HadAM3 \citep{PS2002CD} and ECHAM5 \citep{HETAL2006JC}, and we have discussed whether those model results are consistent with our findings throughout the conclusions (Lines XXX, etc ...).}}
+We have added a paragraph right after the first paragraph in the introduction, to more clearly frame the problem as common to other AGCMs, and how understanding a single model would shed light on resolution sensitivities in other modeling systems.
+Figure 1 has been updated to include convergence experiments from two non-NCAR models, HadAM3 \citep{PS2002CD} and ECHAM5 \citep{HETAL2006JC}, and we have discussed whether those model results are consistent with our findings in the conclusions.
+All edits are in red font in the revised manuscript.}} \newline
+
+{\color{red}{We would like to make aware upfront that our response to reviewer 2 substantially changes one of our main conclusions.
+We had originally attributed the increase in stability with resolution to greater subsiding motion, but as stated in the original manuscript, this only explains about half the increase in stability in deep tropics.
+Four paragraphs of text and two additional figures (one of them supplemental) has been added to the end of the ``deep tropics" sub-section to decipher the additional mechanism driving the increase in stability with resolution (tl;dr, greater ascending motion with resolution drives greater condensational heating, that stabilizes the free troposphere).}} \newline
+
+{\color{red}{All edits are in red font in the revised manuscript.}}
 
@@
 
-{\color{red}{Working on this.}}
+{\color{red}{Thank you for the question.
+My thinking on this issue, at least as it pertains to convection, has evolved over the course of this study.
+While there are some clear advantages to having the cloud macrophyiscs/microphysics take over for some convection at these higher resolutions \citep[e.g.,][]{OETAL2016JAMES}, I've become more concerned about the massive circulations that results from aliasing convection to these scales.
+For example in global 1/4 degree runs AMIP runs with CAM, the Tropical West Pacific double-ITCZ `doubles-down' at higher resolution, resulting in a drier, degraded climate.
+I'm fairly certain that this arises from the expansive subsidence zone that results from rigorously convecting/ascending regions in the West Pacific at high resolution.
+This West Pacific bias is resilient; it exists CAM4, CAM5 and CAM6 simulations.
+I'm un-sure if E3SM was able to tune away this problem in their high-res runs, but I would think that would be a tough uphill battle.}} \newline
+
+{\color{red}{I think what is needed is a deep convection scheme that has the flexibility to intervene / prevent / limit these types of large, unrealistic convective circulations in the model.
+I'm not sure exactly what that would look like, but I think it may require the ability to transfer energy and mass to neighboring grid columns.
+Dave Randall gave a talk at the Banff workshop last year (\url{http://www.birs.ca/events/2019/5-day-workshops/19w5153}) about a novel framework of communicating between grid columns that may be useful in this regard.
+Other than that, I think it's fair to say that the convection scheme should not be forced to maintain convective quasi-equilibrium at fine-scales.
+And I'll also note that the ZM scheme, in addition to being a QCE model, has a CFL limiter under it's hood that really limits this scheme from ramping up and preventing the resolved dynamics from convecting (see Figure~\ref{fig:before-after}).
+I'm starting a new project scientist position in July to implement an EDMF scheme into CLUBB, and potentially dethrone the ZM scheme from CAM.
+Part of my focus will be making sure it has the flexibility to potentially become scale-aware in the future.}} \newline
 
@@
 
-{\color{red}{Based on the reviewers concern, we have amended any strong and/or definitive statements about the -1 exponent due to its dependence on physics time-step (see Lines XXX, XXX and XXX).
+{\color{red}{Based on the reviewers concern, we have amended any strong and/or definitive statements about the -1 exponent due to its dependence on physics time-step (columns 1 and 2 on Page 4, 2nd paragraph in the conclusions).
 To the authors knowledge, there is no authoritative work on the appropriate choice of physics time-steps in AGCMs across different resolutions.
-And so we concede that such definitive statements cannot be made until there is consensus on both the physical and numerical justification for physics time-steps across resolutions.
+And so we concede that definitive statements on the resolution exponent cannot be made until there is consensus on both the physical and numerical justification for physics time-steps across resolutions.
 More recent work on physics time-steps \citep[e.g.,][]{W2013QJRMS,WETAL2015JAMES,HR2018JAMES} are not necessarily consistent with one another, and do not form a consensus.}} \newline
@@
 The slope of the error curves for a fixed physics time-step is shallower for high resolution compared with low resolution, for time-steps larger than about 100 s.
-That is, high resolution is in a less stable regime than low resolution for a fixed time-step (less stable, not un-stable; the errors still decrease with decreasing time-step at high resolution).
+That is, high resolution is in a less stable regime than low resolution for a fixed time-step (less stable, not unstable; the errors still decrease with decreasing time-step at high resolution).
 Interestingly, neither high or low resolution achieve first order convergence for time steps greater than 100 s, which is not really defensible from a numerics standpoint \citep{WETAL2015JAMES}.
-Given these numerical uncertainties, and the large sensitivity of vertical velocity to physics time-step, I don't think we can say for certain that fixing the physics time-step to a reasonably small value (say 225 s) is any better of an approach than the approach we have taken.
-We have softened our language in the manuscript to reflect this uncertainty.}} \newline
+Given these numerical uncertainties, and the large sensitivity of vertical velocity to physics time-step, I don't think we can say for certain that fixing the physics time-step to a reasonably small value (say 225 s) is any better of an approach than the approach we have taken in this manuscript.
+But based on feedback from the reviewer, we have softened our language.
+We think this is the more appropriate choice as of now, and advocate for future studies to investigate this issue in cheaper (e.g., reduced planetary radius) idealized configurations before jumping into expensive year long aqua-planet simulations with small physics time-steps.}} \newline
 
@@
 2016), albeit with seemingly conflicting explanations (Rauscher et al.
-2016; Herrington and Reed 2018)" is still accurate, since it is only later in the conclusions that we delve into why the two explanations are not in conflict with one another.}}
+2016; Herrington and Reed 2018)" is still accurate, since it is only later in the conclusions that we delve into why the two explanations are not in conflict with one another.
+But we are flexible and open to suggestions from the reviewer on whether this language needs to be tweaked.}}
 
@@
 
-{\color{red}{Yes, it should be precipitation rate (mm/day). We have added this title to the figure.}} \newline
+{\color{red}{Yes, it should be precipitation rate (mm/day). We have added this label to the figure.}} \newline
 
@@
 We also thought it was more intuitive to see that larger magnitude precipitation rates are contributing to the global mean, by subsequently unmasking larger regions of M\_s that exceed a frequency tolerance.
-But we understand that readers may be curious about what f\_s looks like and so we have included it as a supplementary material as the reviewer suggests.
+But we understand that readers may be curious about what f\_s looks like and so we have included it as a supplementary material as the reviewer suggests, and have attached it to this document.
 We decided to plot both log(M\_s) and log(f\_s ) and extended the x-axis so that all the data are shown.}}
@@
 
+{\color{red}{We have replaced the original plot to include the $\omega=0$ isoline. We believe the new plot justifies the original statement.}}
+
 \section*{\large Reviewer 2}
@@
 
-The authors use <fx><omegax> to estimate column-mean omega for ascending or descending motions within a column.
-Therefore, my understanding is that there are many cases where a grid column has both <fd><omegad> and <fu><omegau> values.
+The authors use $<fx><omegax>$ to estimate column-mean omega for ascending or descending motions within a column.
+Therefore, my understanding is that there are many cases where a grid column has both $<fd><omegad>$ and $<fu><omegau>$ values.
 \newline
@@
 With this in mind, it was hard to make sense of phrases such as ‘greater magnitude subsiding motion’ in the abstract (L35).
-I could not decide whether the phrase meant that free tropospheric subsidence rates (such as omega850) tend to be stronger or whether there are larger areas of subsidence, until I realized that this is not actually captured by the <fd><omegad> decomposition.
+I could not decide whether the phrase meant that free tropospheric subsidence rates (such as omega850) tend to be stronger or whether there are larger areas of subsidence, until I realized that this is not actually captured by the $<fd><omegad>$ decomposition.
 This decomposition also led to a few points of confusion in the latter half of the study where comparisons of subsidence were made with columnar values, such as FREQZM and dilute-CAPE, or with temperature profiles, as in Fig.
@@
 
-Given the additional level of complexity that is added by thinking in terms of the <fd><omegad> and <fu><omegau> of each column, would the authors please provide more text explaining the advantages of the decomposition and why this decomposition was chosen, rather than taking either the omega at a particular pressure level or the column mean omega <omega> to capture upward and downward motions?
+Given the additional level of complexity that is added by thinking in terms of the $<fd><omegad>$ and $<fu><omegau>$ of each column, would the authors please provide more text explaining the advantages of the decomposition and why this decomposition was chosen, rather than taking either the omega at a particular pressure level or the column mean omega $<omega>$ to capture upward and downward motions?
 \newline
 
-{\color{red}{Working on this.}} \newline
+{\color{red}{We understand that this construct of breaking up $\omega$ at the column level is not typically done, and needs to be more clearly articulated in the text.
+The goal was to provide a 2D (lat,lon) subsidence metric that could be compared against the activity of the ZM scheme (specifically, how often it is triggered), a 2D field.
+Choosing a particular level to evaluate $\omega$ is an additional degree of freedom we sought to avoid in the comparison.
+This is because we wanted a metric of subsidence that characterizes $\omega$ globally, and there is no guarantee that a particular level is representative of $\omega$ everywhere, especially if there are changes to the vertical structure of $\omega$ across resolutions.
+The advantage of this composition over $<\omega>$ is that the mass weighted sum can mask the magnitudes of the strongly ascending/subsiding motion that may occur in a single column, due to cancellations in the vertical integral.
+We agree that not much text is devoted to explaining our rationale on why we use this particular breakdown of $\omega$, and so we have added clarifying statements in the text at the reviewers suggestion (page 4, column 2).
+\newline
+
+In order to conditionally sample an entire vertical profile, as we do for temperature and moisture in the dilute-CAPE calculations, we needed a way of characterizing whether an entire column is predominantly subsiding or ascending.
+In terms of our breakdown, we would say that a column is predominantly subsiding if  $|<f_d><\omega_d>|$ is larger than $|<f_u><\omega_u>|$ in a grid column.
+That's the same as saying $<\omega>$ is greater than zero, and so this is why there is a switch from thinking in terms of $<f_d><\omega_d>$ to thinking in terms of $<\omega>$.
+The reviewers point is well taken, that this is not articulated clearly in the manuscript and could easily lead to confusion.
+We have added language that makes this switch to $<\omega>$ clear, on page 7, column 1.}} \newline
 
@@
 
-{\color{red}{Working on this.}} \newline
+{\color{red}{We thank the reviewer for this helpful comment, as you've brought up something that we've overlooked.
+We have done some additional analysis and have concluded that the tendency to increase stratiform precipitation at the expense of deep convective precipitation itself leads to greater heating in the free troposphere.
+This can be teased out by considering $ne60$ and $ne120$, which have basically the same total precipitation rate, but $ne120$ has less deep convective precipitation indicating its atmosphere is more stable.
+This suggests stratiform cloud formation responds more forcibly (i.e., larger heating tendencies) to remove a given amount of instability, and the radiative-convective balance in the deep tropics adjusts to a warmer, more stable atmosphere.
+We can think of two reasons why the stratiform scheme responds more forcibly to instability.
+(1) Unlike deep convection, the rate of removal of instability is not bound by convective quasi-equilibrium, and (2) the mass fluxes in the deep convection scheme are limited from simulating extreme events.
+This latter feature is because the ZM mass fluxes can and do hit CFL violations, but when they due, a limiter kicks in to lower the mass fluxes to satisfy stability instead of out-right crashing.
+This is illustrated in Figure~\ref{fig:before-after}.}} \newline
+
+\begin{figure}[t]
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{temp_pdf_log.pdf}\\
+\end{center}
+\caption{\small {\color{red}{PDF of cloud base base mass fluxes (hPa/s) in the ZM scheme in an aqua-planet simulation.
+Red lines are the mass fluxes before the CFL limiter is applied, the red lines are after the limiter is applied.}}}
+\label{fig:before-after}
+\end{figure}
+
+{\color{red}{Based on the reviewers comments, we've identified some erroneous statements in the original manuscript that have been removed or fixed.
+These were assertions that subsidence warming is stabilizing the atmosphere, when in fact it is subsidence drying and condensational heating that stabilize the atmosphere with resolution.}} \newline
 

### 2020-06-19 resolution-revised.tex
updates

@@
 In principle, this is an exciting opportunity to resolve atmospheric process and scales in a global model and in unprecedented detail, but in practice this grid flexibility is incompatible with the non- or weakly- converging solutions with increasing horizontal resolution that have long characterized AGCMs.
-In the the Community Atmosphere Model (CAM), there are robust sensitivities to horizontal resolution that have persisted since the model was first introduced over thirty years ago; the atmosphere progressively dries and becomes less cloudy with resolution, and parameterized deep convective precipitation decreases at the expense of stratiform precipitation with resolution.
+In the the Community Atmosphere Model (CAM), there are robust sensitivities to horizontal resolution that have persisted since the model was first introduced over thirty years ago; the atmosphere progressively dries and becomes less cloudy with resolution, and parameterized deep convective precipitation decreases at the expense of stratiform precipitation.
 This study documents a convergence experiment using CAM, version 6, and argue that a unifying cause, the sensitivity of resolved dynamical modes to native grid resolution feedback onto other model components and explain these robust sensitivities to resolution.
@@
 This trend in vertical velocities results in an increase in resolved upward moisture fluxes at cloud base, balanced by an increase in stratiform precipitation rates with resolution.
-Compensating, greater magnitude subsiding motion with resolution has previously been shown to dry out the atmosphere and reduce cloud cover, and here it also shown to stabilize the mean state and reduce the activity of parameterized convection and convective precipitation rates with resolution.
+Compensating, greater magnitude subsiding motion with resolution has previously been shown to dry out the atmosphere and reduce cloud cover.
+Here it shown that {\color{red}{both the increase in condensational heating from stratiform cloud formation and greater subsidence drying contribute to an increase in atmospheric stability with resolution,}} reducing the activity of parameterized convection.
 The impact of changing vertical velocity field with native grid resolution cannot be ignored in any effort to recover convergent solutions in AGCMs, and in particular, the development of scale-aware physical parameterizations.
@@
 
-{\color{red}{Conventional physics packages interact with the adiabatic dynamics to produce non- or weakly- converging solutions with increasing resolution ({\em{resolution}} refers to {\em{horizontal resolution}} hereafter).
+{\color{red}{Conventional physics packages interact with the adiabatic dynamics to produce non- or weakly- converging solutions with increasing resolution \citep[{\em{resolution}} refers to {\em{horizontal resolution}} hereafter;][]{W2008TELLUS}.
 Although some characteristics of resolution sensitivity are model dependent, others are common across different AGCMs and indicate a common mechanism.
 For example, increasing resolution lends itself to greater stratiform precipitation and reduced parameterized convective precipitation in the Hadley Centre climate model \citep[HadAM3;][]{PS2002CD}, the Max Planck Institute for Meteorology model \citep[ECHAM5;][]{HETAL2006JC} and the National Center for Atmospheric Research (NCAR) Community Climate Model \citep[CCM2;][see Figure~\ref{fig:cam-history}]{WETAL1995CD}.
-So while the purpose of this study is to provide a process-level understanding of resolution sensitivity in the NCAR models, these results are useful for understanding resolution sensitivity in models that exhibit similar behavior.}}
+So while the purpose of this study is to provide a process-level understanding of resolution sensitivity in community developed NCAR models, these results are useful for understanding resolution sensitivity in models that exhibit similar behavior.}}
 
@@
 
-In large-eddy simulations, the sensitivity of vertical velocities to resolution is adequately explained by a scale analysis of the dynamical equations \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
+In large-eddy simulations, the sensitivity of vertical velocities to resolution is adequately explained by a scale analysis of the {\color{red}{dry equations of motion}} \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.
 For hydrostatic scales relevant to AGCMs, a scale analysis of the Poisson equation gives $W \propto D^{-1}$, where $D$ is the horizontal scale of buoyancy perturbations driving vertical motion \citep{HR2018JAMES}.
@@
 Studies included in this figure are \cite{KW1991JGR} (CCM1), \cite{WETAL1995CD} (CCM2), \cite{W2008TELLUS} (CAM3), \cite{RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM} (CAM4), \cite{ZetAl2014JCb} (CAM5) and this study (CAM6).
-The asterisk (*) indicates that parameters are modified with resolution to reduce the dependence of cloud fraction and top-of-atmosphere radiative balance on resolution.
-{\color{red}{Two additional studies are included that do no not use the NCAR model, \cite{PS2002CD} (HadAM3) and \cite{HETAL2006JC} (ECHAM5).}}}
+{\color{red}{Two additional studies are included that do no not use the NCAR models, \cite{PS2002CD} (HadAM3) and \cite{HETAL2006JC} (ECHAM5) in order to illustrate similarities in resolution sensitivity across different AGCMs.}} The asterisk (*) indicates that parameters are modified with resolution to reduce the dependence of cloud fraction and top-of-atmosphere radiative balance on resolution.}
 \label{fig:cam-history}
@@
 The increase in stratiform precipitation rates with resolution are attributed to an increase in resolved moisture fluxes at cloud base, primarily due to the increase in upward vertical velocities with resolution.
-The corresponding increase in downward vertical velocities with resolution are shown to stabilize the mean state, reducing the activity of the convection scheme and associated precipitation rates with increasing resolution.
-The feedback of the resolved vertical motion on the model physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution.
+{\color{red}{This increase in ascending motion with resolution drives both greater condensational heating and more frequent and expansive subsidence drying, stabilizing the atmosphere and reducing the activity of the convection scheme and associated precipitation rates with increasing resolution.}} The feedback of the resolved vertical motion on the model physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution.
 Section 2 describes CAM6 and the details of the convergence experiment.
@@
 This scaling was chosen to avoid large time-truncation errors in a rising moist bubble test \citep[Appendix A in][]{HETAL2019JAMES}, and it is understood that this choice of $\Delta t_{phys}$ will likely lead to greater resolution sensitivity \citep{W2008TELLUS}.
-{\color{red}{Due to the large sensitivity of vertical velocity to $\Delta t_{phys}$ \citep{HR2018JAMES}, the lack of scientific consensus on the numerical and physical justifications for how $\Delta t_{phys}$ should vary with resolution remains a severe limitation to characterizing vertical velocities across resolutions in AGCMs.}} The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.
+The convective time-scale in the deep convection scheme is fixed at 3600 s in all simulations.
+{\color{red}{Due to the large sensitivity of AGCM solutions to $\Delta t_{phys}$ \citep{WO2003QJR,W2013QJRMS,HR2018JAMES}, the lack of scientific consensus on the numerical and physical justifications for how $\Delta t_{phys}$ should vary with resolution remains a severe limitation to characterizing solutions across resolutions in AGCMs.}}
 
@@
 In contrast, the $n=-\frac{2}{3}$ slope proposed in \cite{RETAL2016CD} greatly underestimates the increasing occurrence of large magnitude vertical velocities with resolution (SFigure 1).
-{\color{red}{But the $n=-1$ scaling exponent cannot be interpreted more broadly owing to the large sensitivity of vertical velocity to $\Delta t_{phys}$, and the use of equation~\ref{eq:dt-scale} to determine $\Delta t_{phys}$ in the convergence experiment.}}
+{\color{red}{But the failure of any particular value of $n$ to explain vertical velocities in the simulations should be interpreted cautiously.
+Due to the large sensitivity of vertical velocity to $\Delta t_{phys}$ \citep{HR2018JAMES}, the $n=-1$ exponent may only characterize simulations that use equation~\ref{eq:dt-scale} to define $\Delta t_{phys}$ across resolutions.}}
 
@@
 where $\langle f_x \rangle$ and $\langle \omega_x \rangle$ refers to the vertical mass fraction $ \left( \frac{\int dp_x}{\int dp} \textrm{, with pressure } p \right)$ and the $x$ component of the mass weighted vertical mean of $\omega$ $ \left( \frac{\int \omega_x dp_x}{\int dp_x} \right)$, respectively, subscript $u$ refers to upward motion and $d$, downward motion.
+{\color{red}{ This breakdown of $\langle \omega \rangle$ into subsiding and ascending components is useful for characterizing the strength of subsidence or ascent in the model, since $\langle \omega \rangle$ includes cancellations in the vertical integral if there is both subsiding and ascending motion in a column.}}
 
@@
 The plots are computed using 6-hourly instantaneous output of $\omega_{850}$ and $q_{850}$, with 0.05 Pa/s and 0.4 g/kg bin widths, respectively.
-Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but reasonable cut-off for a bins' contribution to $\overline{P_s}$ ({\color{red}{all unmasked values of $M_s$ and $f_s$ are shown in the  supplementary materials.}}).
+Bins in  $\left( \omega_{i} , q_{j} \right)$ space with $f_s <1 \times 10^{-5}$ are masked out, which is a somewhat arbitrary, but reasonable cut-off for a bins' contribution to $\overline{P_s}$ ({\color{red}{all unmasked values of $M_s$ and $f_s$ are plotted in SFigure~2}}).
 The $M_s$ plots (left column) show that larger magnitude $\omega_{850}$ correspond to larger magnitude stratiform precipitation rates, while the impact of changing $q_{850}$ on $M_s$ is less clear.
@@
 Panel labels denote the grid resolution of the model run.
-{\color{red}{See supplementary materials for all the unmasked values of both $M_s$ and $f_s$.}}}
+{\color{red}{See SFigure~2 for all the unmasked values of $M_s$, and the raw $f_s$ field.}}}
 \label{fig:pdecomp}
@@
 To characterize the changes to dilute CAPE of subsiding regions in the deep tropics with resolution, temperature and moisture profiles are conditionally sampled depending on whether $\langle \omega \rangle$ is positive or negative, indicating predominantly subsiding or ascending grid columns.
-The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
+{\color{red}{In the lagnuage of equation~\ref{eq:omega}, the use of positive $\langle \omega \rangle$ to characterize a grid column as descending is equivalent to saying the magnitude of $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ exceeds the magnitude of $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ in a grid column, and vice versa for grid columns with negative $\langle \omega \rangle$.}} The time mean temperature and moisture profiles of subsiding and ascending regions are then used to compute the dilute CAPE used in the ZM scheme, offline.
 Figure~\ref{fig:cape}a shows the dilute CAPE values associated with mean conditions for ascending, descending and all grid columns in the deep tropics, with resolution.
@@
 The increasing occurrence of stable, subsiding grid columns accounts for about half of the changes in dilute CAPE in the deep tropics with resolution, the other half being due to the systematic reduction in dilute CAPE for both ascending and descending regions (Figure~\ref{fig:cape}a).
-This breakdown is straightforward to compute since the sum of the products of the space-time fractions with their associated ascending/descending dilute CAPE values approximate the dilute CAPE computed from the mean temperature/moisture profiles over the entire deep tropics quite well (compare black and grey crosses in Figure~\ref{fig:cape}a).
+This breakdown is straightforward to compute since the sum of the products of the space-time weights with their associated ascending/descending dilute CAPE values approximate the dilute CAPE computed from the mean temperature/moisture profiles over the entire deep tropics quite well (compare black and grey crosses in Figure~\ref{fig:cape}a).
 
-To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then through fixing only the temperature to the $ne20$ profile.
-Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution, with moisture fixed to the $ne20$ profile, on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
+To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, by first fixing the moisture profiles to the lowest resolution $ne20$ profiles, and then through fixing only temperature to the $ne20$ profiles.
+Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution, with moisture fixed to the $ne20$ profiles, on dilute CAPE of ascending/descending grid columns.
+It illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field.
 Moisture changes with resolution do not contribute to the reduction in dilute CAPE with resolution, with the exception that the 20 J/kg reduction in dilute CAPE of ascending regions between the $ne80$ and $ne120$ runs (Figure~\ref{fig:cape}a) is entirely due to changes in moisture (not shown).
+
+\begin{figure}
+\begin{center}
+\noindent\includegraphics[width=20pc,angle=0]{figs/temp_dhgt_panel.pdf}\\
+\end{center}
+\caption{{\color{red}{Climatological change in temperature (left column) and the physics temperature tendencies reflecting the sum of parameterized cloud macrophysics, shallow convection and PBL mixing (CLUBB; right column) between low and high resolution simulations.
+Panels (a-d) show $ne40 - ne20$ differences for (a,b) ascending grid columns and (c,d) subsiding grid columns, whereas panels (e-h) are the $ne80p - ne40$ differences for (e,f) ascending grid columns and (g,h) subsiding grid columns.
+Thick black contour is the zero isoline in all panels.
+For reference, the thick magenta lines in the right column denotes the location of the tropopause in the higher resolution simulations using the lapse-rate WMO definition.
+Only three of the six simulations are used in this calculation because the CLUBB tendency is not available in the other simulations.}}}
+\label{fig:dzonal-hgt}
+\end{figure}
+
+{\color{red}{This increase in stability with resolution is evident from plots of the zonal mean temperature change with height in going from $ne20$ to $ne40$ ($ne40-ne20$) and from $ne40$ to $ne80$ ($ne80-ne40$; left column of Figure~\ref{fig:dzonal-hgt}).
+Anomalous warming occurs in $ne40-ne20$ and $ne80-ne40$ throughout most of the deep tropical troposphere, and cannot be attributed to increasing subsiding motion with resolution as it exists in both ascending ($\langle \omega \rangle<0$; Figure~\ref{fig:dzonal-hgt}a,e) and descending grid columns ($\langle \omega \rangle>0$; Figure~\ref{fig:dzonal-hgt}c,g).
+This trend of tropospheric warming and greater stability with resolution occurs in the mid-latitudes as well, consistent with the reduction in $FREQZM$ in this region (Figure~\ref{fig:4zonal}b).
+
+The right column of Figure~\ref{fig:dzonal-hgt} shows the change in zonal mean CLUBB temperature tendencies versus height for $ne40-ne20$ and $ne80-ne40$.
+CLUBB parameterizes cloud macrophysics, shallow convection and PBL mixing, and so the sum of these processes are reflected in the tendencies.
+Ascending grid columns (Figure~\ref{fig:dzonal-hgt}b,f) show a pattern of increasing CLUBB tendencies in the 200-800 hPa layer (\textit{free troposphere}) with resolution that is absent in descending grid columns (Figure~\ref{fig:dzonal-hgt}d,h).
+This suggests a relative increase in condensational heating from the increase in upward mass fluxes and stratiform cloud formation (Figure~\ref{fig:mflux}).
+Greater condensational heating is consistent with the anomalous warming that occurs in both ascending and descending regimes, as it would be expected to spread horizontally over the weak temperature gradient regime of the tropics \citep{SNP2001JAS}, and increase the temperature of the mean state more broadly.
+
+The high negative correlation between $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and $FREQZM$ in Figure~\ref{fig:corr} not only reflects the reduction in dilute CAPE due to greater subsiding motion, but also greater ascending motion favoring stronger condensational heating in the free troposphere.
+In the time-mean global integral, any increase in magnitude of $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ is reflected in the magnitude of $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ through mass conservation (Figure~\ref{fig:2panel}), and so Figure~\ref{fig:corr} incorporates both the impact of greater magnitude descending and ascending motion on $FREQZM$.
+
+Using instantaneous zonal mean $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ as a predictor of $FREQZM$ in equation~\ref{eq:logreg} is not expected to be very informative due to gravity wave propagation of heating anomalies over the deep tropics, and a positive correlation between resolved mass flux and ZM precipitation rate (Figure~\ref{fig:mflux}).
+Indeed, doing so provides low R-values and a shape parameter predicting an increase in $FREQZM$ with increasing ascending motion in the deep tropics (SFigure~3).
+Outside the tropics, where the deep convection scheme is much less active, there is an increase in condensational heating centered about the mid-latitude jet maximum in $ne40-ne20$ and $ne80-ne40$ for ascending grid columns (Figure~\ref{fig:dzonal-hgt}b,f).
+The logistic regression shape parameter reverses sign poleward of $\pm 40^{\circ}$ latitude (SFigure~3d), indicating that instantaneous $\langle f_{u} \rangle \, \langle \omega_{u} \rangle$ is negatively correlated with $FREQZM$ in these regions.
+This suggests that the increase in stability and reduction in $FREQZM$ with resolution in mid-latitudes is in part due to the increase in upward vertical velocities and resulting increase in stratiform cloud formation with resolution.}}
 
@@
 The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), and their dilute CAPE  values vary much less with resolution ($\sim 20$ J/kg, compared with $\sim 80$ J/kg across all resolutions in the deep tropics).
-Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but as in the deep tropics, temperature changes have a larger influence (not shown).
-This analysis suggests that increasing subsidence with resolution (Figure~\ref{fig:4zonal}a) increases the stability of the mean state, since both ascending/descending regimes become more stable with resolution, and reducing climatological $FREQZM$ (Figure~\ref{fig:4zonal}b) despite the poor relationship between instantaneous subsidence and $FREQZM$ in the subtropics, as shown by the logistic regression (Figure~\ref{fig:4zonal}c).
+Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but as with the deep tropics, temperature changes have a larger influence (not shown).
 
@@
 \end{center}
-\caption{(a,b) Two snapshots of $\omega$ for a longitude-pressure transect at $\sim 18^{\circ}$ latitude in the $ne30$ simulation, overlain by the $0.0075$ kg/m$^2$/s contour of the ZM mass flux delineating the region where the ZM scheme is active.}
+\caption{(a,b) Two snapshots of $\omega$ for a longitude-pressure transect at $\sim 18^{\circ}$ latitude in the $ne30$ simulation, overlain by the $0.0075$ kg/m$^2$/s contour of the ZM mass flux delineating the region where the ZM scheme is active.
+{\color{red}{Magenta contour is the $\omega=0$ Pa/s isoline.}}}
 \label{fig:transect}
@@
 An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, and are balanced by an increase in stratiform precipitation rates.
-Through decomposing the stratiform precipitation into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved mass fluxes / vertical velocities with resolution.
+Through decomposing stratiform precipitation rates into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved upward vertical velocities with resolution.
 At higher resolutions, larger stratiform precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate, with smaller magnitude precipitation rates contributing less.
@@
 The authors {\color{red}{suggest}} that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux), complicating the analysis.
-{\color{red}{As it is generally accepted that vertical velocities increase with resolution in most, if not all AGCMs, models other than CAM that show an increases in stratiform precipitation with resolution (HadAM3 and ECHAM5 in Figure~\ref{fig:cam-history}) are likely driven in a similar way by the increase in resolved moisture flux at cloud base.}}
 
@@
 
-In addition to the change in balance of ascending/descending motion in the deep tropics with resolution, temperature and moisture profiles show a decrease in dilute CAPE for both ascending/descending regimes with increasing resolution.
-This reduction in background dilute CAPE is not confined to the deep tropics, but is also evident from the dilute CAPE derived from the time mean temperature and moisture profiles everywhere in the model (not shown) and is in large part the reason for the global mean reduction in activity in the ZM scheme with resolution.
-A sensitivity analysis of the dilute CAPE calculation indicates that this change in background stability is primarily a result of temperature changes with resolution.
-Temperature profiles become more stable with increasing resolution, likely due to an increase in adiabatic warming due to the increase in subsidence with resolution.
-
-Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE \citep[DCAPE;][]{XZ2000JGR,Z2002JGR}.
-Subsidence warming is represented in the DCAPE budget as negative the vertical advection of dry static energy term, since warming the environment reduces DCAPE, and simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F.
-Song, personal communication).
-The magnitude of negative, or downward, $w$ increases everywhere in the model with increasing resolution, opposing the growth of dilute CAPE and reducing the activity of the ZM scheme with increasing resolution.
-
-{\color{red}{The HadAM3 and ECHAM5 models both show a reduction in convective precipitation rates with increasing resolution (Figure~\ref{fig:cam-history}).
-As with CAM, both models use a stability based closures in their deep convection schemes \citep{GR1990MWR,N1994ECMWF} and both show an increase in tropospheric temperatures with resolution \citep{PS2002CD,RETAL2006JC}.
-In as much as the increase in stability with resolution in these models are driven by greater subsidence warming, the mechanism suppressing parameterized convective activity with resolution in CAM may also explain the resolution sensitivity in HadAM3 and ECHAM5.}}
-
 The subtropics ($\pm (10^{\circ}-30^{\circ})$ latitude) are uniquely less sensitive to subsiding motion.
@@
 These opposing influences result in small values of dilute CAPE, and the ZM scheme acts more like a shallow convection scheme, producing light precipitation or drizzle, irrespective of resolution.
+
+In addition to the increasing occurrence of subsidence grid columns with resolution in the deep tropics, temperature and moisture profiles show a decrease in dilute CAPE for both ascending and descending regimes across resolutions.
+This reduction in background dilute CAPE is not confined to the deep tropics, but is also evident from the dilute CAPE derived from the time mean temperature and moisture profiles everywhere in the model (not shown) and plays a large role in the global mean reduction in activity of the ZM scheme with resolution.
+A sensitivity analysis of the dilute CAPE calculation indicates that this change in background stability is primarily a result of temperature changes with resolution.
+
+{\color{red}{Temperatures generally increase with resolution in the free troposphere as a result of increases in condensational heating and stratiform cloud formation arising from larger magnitude ascending motion.
+The overall increase in atmospheric stability with resolution then occurs through two mechanisms; the direct impact of larger upward vertical velocities on condensational heating, and the indirect impact of more vigorous ascent driving more frequent and expansive compensating subsidence, drying out the atmopshere.
+Since an increase in global mean ascent is reflected in the global mean subsidence calculation through mass conservation (they must be both equal and opposite), both means of stabilizing the atmosphere contribute to the high correlation between global mean subsidence and activity of the ZM scheme in CAM.}}
+
+{\color{red}{The reduction in convective precipitation rates and increase in stratiform precipitation rates with resolution is not unique to CAM (see HadAM3 and ECHAM5 in Figure~\ref{fig:cam-history}).
+The HadAM3 model also gives way to an increase in vertical velocities with resolution \citep[Figure~2 in][]{PS2002CD}, and it is reasonable to assume this is a common feature among AGCMs \citep[e.g.,][]{DETALA2016ACP}.
+The increase in stratiform precipitation rates with resolution in HadAM3 and ECHAM5 are likely driven in a similar way by the increase in resolved moisture flux at cloud base.
+In as much as the greater condensational heating and subsidence drying feedback onto their stability based deep convection schemes \citep{GR1990MWR,N1994ECMWF}, understanding resolution in CAM may also help explain resolution sensitivity in HadAM3 and ECHAM5.}}
 
